### 2021-04-18 SupMat.Rnw
up to section 5

\documentclass[letterpaper,11pt]{article}

%packages
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage[left=2cm,top=2cm,right=2cm,bottom=2cm,head=.5cm,foot=.5cm]{geometry}
\usepackage{url}
\usepackage{multirow}
\usepackage{longtable}
\usepackage{subfig}
\usepackage{float}
\usepackage{setspace}
\usepackage{lineno}
\usepackage{natbib}
\usepackage{amsmath}
\usepackage{xr}
\usepackage{authblk}
%\usepackage{accents}

\DeclareRobustCommand{\bbone}{\text{\usefont{U}{bbold}{m}{n}1}}

\DeclareMathOperator{\EX}{E}% expected value
\DeclareMathOperator{\VarX}{var}
\DeclareMathOperator{\CovX}{cov}
\DeclareMathOperator{\mean}{mean}
\DeclareMathOperator{\se}{se}
\DeclareMathOperator{\sd}{sd}

%new commands and so on

%external documents
\externaldocument[MT-]{Paper}

%header material for paper
\title{Drivers of spatial synchrony in giant kelp populations}
\date{}

%***Tentative author order
\author{Lawrence W. Sheppard}
\author{Daniel C. Reuman}
\author{Tom Bell}
\author{Jonathan Walter}
\author{Kyle Cavanaugh}
\author{Max Castorani}

\begin{document}
\SweaveOpts{concordance=TRUE}

%The following is where you load in the numeric results that will be embedded in the text
<<eval=T,echo=F,message=F,warning=F>>=
#place R code here for loading in necessary variables from Results
@

\maketitle

%probably want table of contents, figure and table lists

\section{The Model}

We used the lottery model, previously studied, for instance, by REFS. Model equations are

\begin{equation}
N_i(t+1)=(1-\delta)N(t)+\delta N \frac{B_i(t)N_i(t)}{B_1(t)N_1(t)+B_2(t)N_2(t)}  \label{model_eq}
\end{equation}

\noindent for species $i=1,2$. Here, $N_i(t)$ is the population density of species $i$ at time $t$, and $N = N_1(t)+N_2(t)$ is fixed through time. The parameter, $\delta$ is a mortality rate, and $B_i(t)$ is the fecundiity of specise $i$ at time $t$. The model postulates that individuals die at rate $\delta$ at each time step, and are replaced by juveniles in population to the reproductive output of the two species that year. We take, for each $i$, $B_i(t)$ to be $iid$ through time. We let $B_i = exp(b_i)$, where $(b_1, b_2)$ is some distribution such that $b_i \sim N(\mu_i, \sigma^2)$ and such that $cov(b_1, b_2) = \rho$, a fixed quantity we will specify later. However, $(b_1, b_2)$ is not necessarily a bivariate normal distribution. We will consider various distributions of $(b_1, b_2)$ with these properties, corresponding to our symmetric and asymmetric tail associations cases defined in the main text and elaborated below. We assume without loss of generality that $\mu_1 \leq \mu_2$, so that species 1 is the weaker competitor.

\section{Noise} \label{sect:noise}

We created three cases for $(b_1,b_2)$, the case of symmetric tail association and the cases of asymmetrical left- and right-tail association. To generate $M$ points $(b_1^{(i)}, b_2^{(i)})$, $i=1,..,M$, for the left tail association case, we first generated M points $(a_1^{(i)},a_2^{(i)})$, $i=1,...,M$, from a bivariate normal distribution $N(\begin{pmatrix} 0\\0\end{pmatrix}, \Sigma)$, where here $\Sigma = \begin{pmatrix} 1&0\\0&1\end{pmatrix}$. Then, for each index $i$, we randomly, with 50\% probability, either (A) let $\tilde b_1^{(i)} = -|a_1^{(i)}|$ and $\tilde b_2^{(i)} = -a|_1^{(i)}|$, or (B) let $\tilde b_1^{(i)} = |a_1^{(i)}|$ and $\tilde b_2^{(i)} = |a_2^{(i)}|$. We then let $b_1^{(i)} = \sigma \tilde b_1^{(i)} + \mu_1$ and $b_2^{(i)} = \sigma \tilde b_2^{(i)} + \mu_2$. Perfect association in the left tails of the resulting distribution $(b_1,b_2)$ results from the fact that, in case (A) above, both $\tilde b_1^{(i)}$ and $\tilde b_2^{(i)}$ were assigned to the same value. For speed of computation, when samples $(\tilde b_1^{(i)},\tilde b_2^{(i)})$, $i=1,...,M$ were needed for every distinct set of values of the parameters $\mu_1$, $\mu_2$, and $\sigma^2$, we used the same $(\tilde b_1^{(i)},\tilde b_2^{(i)})$, $i=1,...,M$, simply transforming these values differently (using the different valus of $\mu_1$, $\mu_2$, $\sigma$) for each set of parameter values. Right-tail associated noise was generated in an analogous way.

Noise with symmetric tail association was generated by taking a very large sample (1000?) of left-tail associated noise, as described above, and then calculating the Pearson correaltion coefficient, rho. The value obtained was ... . The symmetric noise was then generated once from $N(\begin{pmatrix} 0\\0\end{pmatrix}, \Sigma)$ where $\Sigma = \begin{pmatrix} 1&\rho \\ \rho&1\end{pmatrix}$, denoting the points so generated by $(\tilde b_1^{(i)},\tilde b_2^{(i)})$ for $i=1,...,M$. We then set $b_1^{(i)} = \sigma \tilde b_1^{(i)} + \mu_1$ and $b_2^{(i)} = \sigma \tilde b_2^{(i)}+ \mu_2$.

Thus, overall, our model was completely specified by specifying $\delta$, $\mu_1$, $\mu_2$, $\sigma$ and whether noise was generated from the left-tailed association, right-tailed association or symmetric tail association cases specified above.

\section{Invasion growth rate of species 1, and storage effects} \label{sect:IGR}

We consider the invasion growth rate of species 1, $\bar r_1=\EX(\ln (\frac{N_1(t+1)}{N_1(t)}))$, when $N_1$ is close to 0 and $N_2$ is close to $N$. Hence $\EX(\cdot)$ represents expected value. We have

\begin{equation}
\frac{N_1(t+1)}{N_1(t)}=(1-\delta)+\delta N \frac{B_1(t)}{B_1(t)N_1(t)+B_2(t)N_2(t)},  \label{sp1_Nt+1_over_Nt}
\end{equation}

\noindent but setting $N_1=0$ and $N_2=N$ here gives the result that

\begin{equation}
\bar r_1=\EX(\ln [(1-\delta)+\delta exp(b_1-b_2)]).  \label{rbar1_E}
\end{equation}

\noindent We considered the various cases for the joint distribution $(b_1,b_2)$, specified in the main text and in the previous section.

To study storage effect for the model, we also consider the contribution of EC covariance to the difference between the invader's and the resident's mean growth rates, to be denoted $\Delta I$. Following REF, we define

\begin{equation}
\bar r_1^\#=\EX(\ln [(1-\delta)+\delta exp(b_1^\#-b_2)]),  \label{rbar1sharp_E}
\end{equation}

\noindent where $b_1^\#$ is distributed in the same way as $b_1$, but is independent of it and $b_2$. Then $\bar r_1 - \bar r_1^\#$ is the contribution of EC covariance to the mean growth rate species 1, the invader. We know, a priori, that the mean growth rate of the resident, species 2, must be 0, $\bar r_2 =0$, but equation (\ref{model_eq}) also implies that

\begin{equation}
\frac{N_2(t+1)}{N_2(t)}=(1-\delta)+\delta N \frac{B_2(t)}{B_1(t)N_1(t)+B_2(t)N_2(t)}.  \label{sp2_Nt+1_over_Nt}
\end{equation}

\noindent Setting $N_2=N$ and $N_1=0$ then gives

\begin{equation}
\frac{N_2(t+1)}{N_2(t)}=(1-\delta)+\delta=1  \label{sp2_Nt+1_over_Nt_simp}
\end{equation}

\noindent so

\begin{equation}
r_2=(\ln (\frac{N_2(t+1)}{N_2(t)}))=0, \label{r2_is_zero}
\end{equation}

\noindent as expected. Again, following REF, we also define

\begin{equation}
\bar r_2^\#=\EX(\ln [(1-\delta)+\delta exp(b_2^\#-b_2)]),  \label{rbar2sharp_E}
\end{equation}

\noindent where $b_2^\#$ is distributed in the same way as $b_2$, but is independent of it and of $b_1$. Thus $\bar r_2- \bar r_2^\# = -\bar r_2^\#$ is the contribution of EC covariance to the mean growth rate of species 2, the resident. And so

\begin{equation}
\Delta I=(\bar r_1 - \bar r_1^\#)-(\bar r_2- \bar r_2^\#)=\bar r_1 - \bar r_1^\#+\bar r_2^\#  \label{DeltaI}
\end{equation}

\noindent has been defined in terms of expected values of elementary expressions of the random variables $(b_1,b_2)$, $(b_1^\#,b_2)$, and $(b_2^\#, b_2)$.

Having already specified that $b_1$ (and therefore also $b_1^\#$) is distributed as $N(\mu_1, \sigma^2)$, and that $b_2$ (and therefore also $b_2^\#$) is distributed as $N(\mu_2, \sigma^2)$, we know, using $\Sigma$ as notation for a covariance matrix that

\begin{equation}
(b_1^\#, b_2) \sim N(\begin{pmatrix} \mu_1\\\mu_2\end{pmatrix}, \Sigma) \label{b1sharpb2_distribution}
\end{equation}

\noindent and

\begin{equation}
(b_2^\#, b_2) \sim N(\begin{pmatrix} \mu_2\\\mu_2\end{pmatrix}, \Sigma) \label{b2sharpb2_distribution}
\end{equation}

\noindent where

\begin{equation}
\Sigma = \begin{pmatrix} \sigma^2&0\\0&\sigma^2\end{pmatrix}. \label{Sigmamat_cov0}
\end{equation}

\noindent Thus the difference $b_1^\#-b_2$ that occurs in the expression for $\bar r_1^\#$ (\ref{rbar1sharp_E}) is a normally distributed random variable with mean $\mu_1-\mu_2$ and variance $2\sigma^2$. Thus

\begin{equation}
\bar r_1^\#=\EX(\ln [(1-\delta)+\delta exp(u)]),  \label{rbar1sharp_E_usub}
\end{equation}

\noindent where $u \sim N(\mu_1-\mu_2,2\sigma^2)$. Likewise, the difference $b_2^\#-b_2$ in (\ref{rbar2sharp_E}) is a normally distributed random variable with mean $\mu_2-\mu_2=0$ and variance $2\sigma^2$, so

\begin{equation}
\bar r_2^\#=\EX(\ln [(1-\delta)+\delta exp(u)]),  \label{rbar2sharp_E_usub}
\end{equation}

\noindent where now $u \sim N(0,2\sigma^2)$. The expression (\ref{rbar1_E}) for $\bar r_1$ can also be further simplified, in a similar way, when we are in the symmetric tail association case considered in section \ref{sect:noise}. In that case,

\begin{equation}
(b_1,b_2) \sim  N(\begin{pmatrix} \mu_2\\\mu_2\end{pmatrix}, \Sigma) \label{b1b2_distribution_sym}
\end{equation}

\noindent where now

\begin{equation}
\Sigma = \begin{pmatrix} \sigma^2&\rho\\\rho&\sigma^2\end{pmatrix}. \label{Sigmamat_rho}
\end{equation}

\noindent and $rho=put in value$ was determined as in section \ref{sect:noise}. Therefore, the expression $b_1-b_2$ that occurred in \ref{rbar1_E}, in the symmetric tail association case, is a normally distributed random variable with mean $\mu_1-\mu_2$ and variance

\begin{align}
\VarX(b_1-b_2)&=\CovX(b_1-b_2, b_1-b_2) \\
                        &=\CovX(b_1,b_1)+\CovX(b_2,b_2)-\CovX(b_1,b_2)-\CovX(b_2,b_1) \\
                        &=2\sigma^2-2\rho.
\end{align}

\noindent Thus, again only in the symmetric tail association case,

\begin{equation}
\bar r_1 = \EX(\ln[1-\delta+\delta \exp(u)]), \label{rbar1_E_usub}
\end{equation}

\noindent where here $u \sim N(\mu_1-\mu_2, 2\sigma^2-2\rho)$. As for all we know there is no simplicifaction of the expression \ref{rbar1_E}  for $\bar r_1$ in the left and right-tail association cases.

Thus for in summary,

\begin{equation}
\bar r_1 = \EX(\ln[1-\delta+\delta \exp(u)]), u \sim N(\mu_1-\mu_2, 2\sigma^2-2\rho), \label{sym_rbar1}
\end{equation}

\noindent for the symmetric case, and

\begin{equation}
\bar r_1^{\#} = \EX(\ln[1-\delta+\delta \exp(u)]), u \sim N(\mu_1-\mu_2, 2\sigma^2), \label{sym_rbar1sharp}
\end{equation}

\begin{equation}
\bar r_2^{\#} = \EX(\ln[1-\delta+\delta \exp(u)]), u \sim N(0, 2\sigma^2). \label{sym_rbar2sharp}
\end{equation}

\noindent We next use the expressions to develop methods to estimate $\bar r_1$ and $\Delta I$ for any given paramters $\delta,\mu_1,\mu_2,\sigma$.

\section{Estimating $\bar r_1$ and $\Delta I$}

First, for a large ?artigen? $M$, carry out the following steps once:

\begin{enumerate}
\item Generate left-tail assocated noise $(\tilde b_{l1}^{(i)}, \tilde b_{l2}^{(i)})$, $i=1,...,M$, as described in the section \ref{sect:noise}. Hence $\tilde b_{l1}^{i)}$ and $\tilde b_{l2}^{(i)}$ are distributed as $N(0,1)$. 
\item Generate right-tail associated noise $(\tilde b_{r1}^{(i)}, \tilde b_{r2}^{(i)})$, $i=1,...,M$, via the analogous procedure. Again, $\tilde b_{r1}^{i)}$ and $\tilde b_{r2}^{(i)}$ are distributed as $N(0,1)$. 
\item Generate $M$ points $\tilde u_i$, $i = 1,...,M$ from a standard normal distribution. 
\end{enumerate}

\noindent Next, given values of parameters $\delta, \mu_1, \mu_2$, and $\sigma$, proceed as follows to get estimates of $\bar r_1, \bar r_1^{\#}, \bar r_2^{\#}$, and $\Delta I$.

\begin{enumerate}

\item Estimate $\bar r_1$ for the left-tail associated case as $\hat \bar r_1 = \mean_i (\ln [1-\delta+\delta\exp(b_{l1}^{(i)}-b_{l2}^{(i)})])$ where $b_{l1}^{(i)} = \sigma \tilde b_{l1}^{(i)} + \mu_1$ and $b_{l2}^{(i)} = \sigma \tilde b_{l2}^{(i)} + \mu_2$. The standard error of this estimate is $\se (\hat \bar r_1) =\frac{\sd _i (\ln[1-\delta +\delta \exp(b_{l1}^{(i)}-b_{l2}^{(i)})])}{\sqrt M}$, where $\sd _i(\cdot)$ is standard deviation.

\item Estiate $\bar r_1$ for the left-tail associated case as $\hat \bar r_1 = \mean_i (\ln [1-\delta+\delta\exp(b_{r1}^{(i)}-b_{r2}^{(i)})])$ where $b_{r1}^{(i)} = \sigma \tilde b_{r1}^{(i)} + \mu_1$ and $b_{r2}^{(i)} = \sigma \tilde b_{r2}^{(i)} + \mu_2$. The standard error of this estimate is $\se (\hat \bar r_1) =\frac{\sd _i (\ln[1-\delta +\delta \exp(b_{r1}^{(i)}-b_{r2}^{(i)})])}{\sqrt M}$.

\item Estimate $\bar r_1$ in the symmetric tail association case as $\hat \bar r_1=\mean_i(\ln[1-\delta+\delta \exp(u_i)])$, where $u_i = \sqrt{(2\sigma^2-2\rho)}\tilde u_i + \mu_1 - \mu_2$. The standard error of this estimation is $\se(\hat \bar r_1) = \frac {\sd_i(\ln[1-\delta+\delta \exp(u_i)])}{\sqrt M}$.

\item Estimate $\bar r_1^{\#}$ (for all three tail-assocaition cases) as $\hat \bar r_1^{\#} = \mean_i(\ln[1-\delta+\delta \exp(u_i)])$, where $u_i = \sqrt{2\sigma^2}\tilde u_i + \mu_1 - \mu_2$. The standard error is $\se(\hat \bar r_1^{\#}) = \sd_i \frac{ln[1-\delta+\delta \exp(u_i)]}{\sqrt M}$.

\item Estimate $\bar r_2^{\#}$ (for all three tail-assocaition cases) as $\hat \bar r_2^{\#} = \mean_i(\ln[1-\delta+\delta \exp(u_i)])$, where $u_i = \sqrt{2\sigma^2}\tilde u_i$. The standard error is $\se(\hat \bar r_2^{\#}) = \sd_i \frac{ln[1-\delta+\delta \exp(u_i)]}{\sqrt M}$.

\item $\hat {\Delta I} = \hat \bar r_1 - \hat \bar r_1^{\#} + \hat \bar r_2^{\#}$, estimated ?syparatilly? for the left-tail associated, right-tail associated and symmetric noise cases. We obtained conservative standard error estimates by adding the standard error of the estimates of $\hat \bar r_1, \hat \bar r_1^{\#}$, and $\hat \bar r_2^{\#}$.

\end{enumerate}

\noindent For large enough $M$, standard error should be very small. Since random variables are generated once only, computation of these quantitaties for a large number of parameter sets should be fast.

\section{Additional quantities influencing invasion success} \label{sect: addquant}

\end{document}

### 2021-04-18 Paper.Rnw
configuring packages

\documentclass[letterpaper,11pt]{article}

%packages
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage[left=2cm,top=2cm,right=2cm,bottom=2cm,head=.5cm,foot=.5cm]{geometry}
\usepackage{url}
\usepackage{multirow}
\usepackage{longtable}
\usepackage{subfig}
\usepackage{float}
\usepackage{setspace}
\usepackage{lineno}
\usepackage{natbib}
\usepackage{amsmath}
\usepackage{xr}
\usepackage{authblk}

%new commands and so on
\providecommand{\keywords}[1]
{
  \small	
  \textbf{\textit{Keywords---}} #1
}

%external documents
\externaldocument[SM-]{SupMat}

%header material for paper
\title{Place title here}
\date{}

\author[a]{Jasmin Albert}
\author[a,b]{Daniel C. Reuman}

\affil[a]{Department of Ecology and Evolutionary Biology and Kansas Biological Survey, University of Kansas}
\affil[b]{Laboratory of Populations, Rockefeller University}

%***Need to indicate corresponding authors

\begin{document}
\SweaveOpts{concordance=TRUE}

%The following is where you load in the numeric results that will be embedded in the text
<<eval=T,echo=F,message=F,warning=F>>=
#system(paste("biber", sub("\\.Rnw$", "", current_input())))

#place R code here for loading in necessary variables from Results
#ordinarily this would be a readRDS command, but I'm just defining a variable here for now
allregres<-matrix(rnorm(10),2,5)
@

\maketitle

\begin{abstract}
Place abstract here
\end{abstract}

%***Add to these
\keywords{kw1, kw2,kw3}

\section{Introduction}\label{section:introduction}

\section{Methods}\label{section:methods}

\section{Results}\label{section:results}

\section{Discussion}\label{section:discussion}

\section{Reference code for Jasmin to learn Latex/Sweave}

To get normal text you just type.

To get a paragraph break you leave a blank line, indentation is automatic.

\noindent If you want to remove the indent you use this.

In-line math is always surrounded by dollar signs, $x+1=4$. Even when you just have single variables, like if you say ``the notation we use for synchrony is $x$'', you put the $x$ in dollar signs. The reason is x and $x$ look different, in a way that a mathematician will notice and wonder if they are meant to be different variables.

For quotes, do it ``like this'' not "like this" or else one of the quotes will be backwards.

Some basic math, inline, by example: $x_{11}$, $y^{22}$, $\frac{1+y}{x+3}$, $\sum_{i=1}^N x_x$,
$\alpha$, $\beta$.

There is also display math:
\begin{equation}
x=y+12-z \label{eq_about_x_and_y}
\end{equation}
The ``label'' command allows a textual label for refering to equations later.

The math gets pretty complex. Keep in mind, whatever you want to do, there is a way to do it. We'll save most of that for later.

Now we can refer to equation/expression (\ref{eq_about_x_and_y}). Latex will sort
out equation numbering and insert the correct number labels for you.

You can also refer to any section in the same way: section \ref{section:introduction}, section \ref{section:methods}.

You can include the value of previously loaded R variables in the text, e.g., the value of allregres[1,5] is
\Sexpr{allregres[1,5]}, which can also be given rounded: \Sexpr{round(allregres[1,5],3)}. You can do simple R
on the fly, e.g., allregres[1,5] plus 1 is \Sexpr{allregres[1,5]+1}. 
You should never manually type any results, they should always be autolinks like this.

%Comments are lines starting with %

You make a percent with \%

You can refer to Fig. \ref{fig:random_fig} by its label and latex will number figures and use the correct numbers.
The label is defined below where the figure is placed.

You can refer to Table \ref{tab:random_table} in the same way.

%You can also refer to figures and tables and equations from the sup mat, just prefix the labels
%by SM- in your calls to \ref{}

You can cite like this \cite{Abanetal2006} or this \citep{Abanetal2006} or this \citealp{Abanetal2006}.
You cal also cite books \citep{BurnhamAnderson2002}, of course.

\clearpage
\newpage

%you store your bibliographic info in refs.bib, see that file
\bibliographystyle{ecology_letters2}
\bibliography{refs}

\clearpage
\newpage

\section{Tables}

%You make tables like this:
\begin{table}[H]
\caption[Summary caption, for table of contents, can be skipped.]{Full caption here.}
\begin{tabular}{llll} %The four ls mean four columns, each left justified. Can use r or c, too.
\hline 
Col 1 header & Col 2 header & Col 3 header & Col 4 header \\
\hline
text & or & numbers & 3.2 \\
or & math & $x+1$ & or \\
even & the & values & of \\
loaded & R & variables & \Sexpr{allregres[1,5]} \\
you & can & round & \Sexpr{round(allregres[1,5],3)} \\ 
\hline
\end{tabular}
\label{tab:random_table}
\end{table}

\clearpage
\newpage

\section{Figures}

\begin{figure}[!h]
\includegraphics[width=.5\textwidth]{figures/summaryDocFigures/figure8.png}
\caption{Type the caption here.} \label{fig:random_fig}
\end{figure}

\end{document}

### 2021-05-02 makenoise.R
new function for making noise

@@ -1,0 +1,45 @@
+library(MASS)
+# make into function
+
+# generate noise
+
+makenoise <- function(M){
+  
+  #sigma and mu to make standard normal bivariate noise
+  Sigma_norm <- matrix(c(1,0,0,1),nrow=2)
+  mu_norm <- c(0,0)
+  
+  #standard normal
+  b <- mvrnorm(M, mu_norm, Sigma_norm) #points (a1,a2)
+  a1 <- b[,1]
+  a2 <- b[,2]
+  
+  #A <- sample(ind, M*0.5) #50% chance
+  A <- a1>0
+  
+  #left tail
+  b[A,] <- rep(-abs(a1[A]))
+  b[!A,] <- abs(b[!A,])
+  b_l <- b
+  #plot(b_l[,1],b_l[,2])
+  
+  #right tail
+  b[A,] <- -b[A,]
+  b[!A,] <- -b[!A,]
+  b_r <- b
+  #plot(b_r[,1],b_r[,2])
+  
+  #symmetric 
+  Sigma_sym <- cor(b_l) 
+  b_s <- mvrnorm(M, mu_norm, Sigma_sym)
+  #plot(b_s[,1],b_s[,2])
+  
+  return(list(b_l=b_l,b_r=b_r,b_s=b_s))
+}
+
+#makenoise(100)
+
+
+
+
+

### 2021-05-02 paper/SupMat.tex
paper files

--- 2021-04-18 SupMat.Rnw
@@
 \DeclareMathOperator{\sd}{sd}
+\DeclareMathOperator{\Prob}{P}
 
@@
 
+\usepackage{Sweave}
 \begin{document}
-\SweaveOpts{concordance=TRUE}
+\input{SupMat-concordance}
 
 %The following is where you load in the numeric results that will be embedded in the text
-<<eval=T,echo=F,message=F,warning=F>>=
-#place R code here for loading in necessary variables from Results
-@
 
@@
 
-Noise with symmetric tail association was generated by taking a very large sample (1000?) of left-tail associated noise, as described above, and then calculating the Pearson correaltion coefficient, rho.
+Noise with symmetric tail association was generated by taking a very large sample (100000) of left-tail associated noise, as described above, and then calculating the Pearson correaltion coefficient, rho.
 The value obtained was ...
@@
 
-\section{Additional quantities influencing invasion success} \label{sect: addquant}
+\section{Additional quantities influencing invasion success} \label{sect:addquant}
+
+Not only does the expected value $\bar r_1$ influence the capacity for species 1 to invade where rare, but so do other aspectf of the distribution of
+\begin{equation}
+r_1 = \ln[1-\delta+\delta \exp(b_1-b_2)]. \label{r1_distribution}
+\end{equation}
+
+We here elaborate how some other aspects of the distribution of $r_1$ were quantified, for our three tail association cases and for any of our parameters $\delta, \mu_1, \mu_2, \sigma$.
+
+First, we considered $\sd(r_1)$, the standard deviation.
+If this is large enough, then $r_1$ can occasionally be positive even if $\bar r_1$ is strongly negative, and this invasion can occur under the right environmental circumstances.
+
+We also considered the probability $\Prob[r_1 >0]$, which is, explicitly, the probability that environmental variables that permit invasion. We have
+\begin{align}
+\Prob[r_1>0] &= \Prob [1-\delta+\delta \exp(b_1-b_2)] > 1] \\
+&= \Prob[\exp (b_1-b_2)>1] \\
+&= \Prob[b_1>b_2],
+\end{align}
+?? should the above have $\ln$?
+\noindent so this was computed for each considered parameter set of $\delta, \mu_1, \mu_2, \sigma$, and for each of our three tail association cases.
+
+We also consider the mean value of the possible part of the distribution of $r_1$, $\EX(r_1|r_1>0)$.
+This characterizes how quickly invasion may happen, when it happens, with larger values corresponding to faster invasion.
+If it takes several time steps for the population of the invader to rise from negligible level at which species 1 is a noticeable presence in the community, then environmental conditions must be suitable for invasion for several sequential time steps.
+Thus large values of $\EX(r_1|r_1>0)$ should also correspond to a greater probability of invasion because they should correspond to faster invasions and therefore to reduced need for sequential periods of suitable environmental conditions.
+
+Finally, for a ?low? of species of? parameters $\delta, \mu_1, \mu_2, \sigma$, we displayed the whole distribution of $r_1$, values for each of our tail association cases, to make visual comparisons.
+
+\section{Symmetries and parameter reduction}
+
+The quantities we consider to address invasion prospects for the weaker competitor, species 1 (see sections \ref{sect:IGR} and \ref{sect:addquant}) depend only on $\mu_1-\mu_2$ and not, independtly, on $\mu_1$ and $\mu_2$, and also are the same for our left- and right-tailed association cases.
+We prove these statements in this section.
+
+Equations \ref{sym_rbar1} - \ref{sym_rbar2sharp} obviously depend only on $\mu_1-\mu_2$, and not independently on $\mu_1$ and $\mu_2$, ?some?
+only $\mu_1-\mu_2$ appears in these expressions if $\mu_1$ or $\mu_2$ appear at all.
+Equation \ref{sym_rbar1} is $\bar r_1$ in the asymmetric?
+tail association cases.
+But letting $\beta_1 = b_1 + \eta$ and $\beta_2 = b_2 + \eta$ and substituting into \ref{sym_rbar1}, we get
+
+\begin{align}
+\bar r_1 &= \EX(\ln[1-\delta+\delta \exp(\beta_1-\eta-(\beta_2-\eta))])\\
+&= \EX(\ln[1-\delta+\delta \exp(\beta_1-\beta_2)]).
+\end{align}
+\noindent And so altering the mean of both $b_1$ and $b_2$ distributions by the same amount has no affect of $\bar r_1$, and thus $\bar r_1$ in the asymmetric tail associations also depends only on $\mu_1 - \mu_2$ and not independtly on $\mu_1$ or $\mu_2$.
+
+In section \ref{sect:addquant}, we introduce several other metrics. These are all based on the distribution 
+\begin{equation}
+r_1=\ln[1-\delta+\delta \exp(b_1-b_2)] \label{r1distribution}
+\end{equation}
+
+\noindent But, again making the substitutions $\beta_1 = b_1 + \eta$ and $\beta_2 = b_2 + \eta$, we can see via the same resoning as above that the whole distribution $r_1$ depends only on $\mu_1 - \mu_2$, and not on $\mu_1$ and $\mu_2$ independently, for all three of our tail association cases.
+
+To see that our various metrics are the same for out left- and right-tail associated noise ?parameters?, let $(b_1,b_2)$ denote our left-tail associated random variable with parameters $\mu_1,\mu_2,\sigma$.
+Then define $\beta_1 = -b_2 + \mu_1 +\mu_2$, $\beta_2=-b_1+\mu_1+\mu_2$.
+It is easy to see that $\EX(\beta_1)=\mu_1$, $\EX(\beta_2)=\mu_2$, $\sd(\beta_1)=\sd(\beta_2)=\sigma$, so $\beta_1$ and $\beta_2$ are normally distributed.
+From there it is easy to see that $\beta_1$ and $beta_2$ are distributed in the same wat as our right-tail associated noise with parameter $\mu_1,\mu_2$, and $\sigma$.
+Making the substitutions $b_2 = -\beta_1 + \mu_1 + \mu_2$ and $b_1 = -\beta_2 + \mu_1+\mu_2$ in ref{r1distribution}, we get
+
+\begin{align}
+r_1 &= \ln[1-\delta+\delta \exp(-\beta_2+\mu_1+\mu_2-(\beta_1+\mu_1+\mu_2))]\\
+&= \ln[1-\delta+\delta \exp(\beta_1-\beta_2)].
+\end{align}
+
+\noindent And since $(\beta_1, \beta_2)$ is distributed in the same way as our right-tail associated noise, this proves all metrics based on the distribution of $r_1$ \ref{r1distribution} are the same for left- and right-tail associated noise.
 

### 2021-05-02 paper/SupMat.Rnw
moving files

--- 2021-05-02 paper/SupMat.tex
@@
 
-\usepackage{Sweave}
 \begin{document}
-\input{SupMat-concordance}
+\SweaveOpts{concordance=TRUE}
 
 %The following is where you load in the numeric results that will be embedded in the text
+<<eval=T,echo=F,message=F,warning=F>>=
+#place R code here for loading in necessary variables from Results
+@
 
@@
 
-Noise with symmetric tail association was generated by taking a very large sample (100000) of left-tail associated noise, as described above, and then calculating the Pearson correaltion coefficient, rho.
+Noise with symmetric tail association was generated by taking a very large sample (1000?) of left-tail associated noise, as described above, and then calculating the Pearson correaltion coefficient, rho.
 The value obtained was ...
@@
 \end{align}
-\noindent And so altering the mean of both $b_1$ and $b_2$ distributions by the same amount has no affect of $\bar r_1$, and thus $\bar r_1$ in the asymmetric tail associations also depends only on $\mu_1 - \mu_2$ and not independtly on $\mu_1$ or $\mu_2$.
+\noindent And so altering the mean of both $b_1$ and $b_2$ distributions by the same amount has no affect of $\bar r_1$ in the asymmetric tail associations.
+Also ?
+depends only on $\mu_1 - \mu_2$ and not independtly on $\mu_1$ or $\mu_2$.
 

### 2021-05-02 noise_fig.R
noise fig

@@ -1,0 +1,21 @@
+source("./makenoise.R")
+
+b <- makenoise(1000)
+
+x<-cor(b$b_l)[1,2]
+y<-cor(b$b_s)[1,2]
+z<-cor(b$b_r)[1,2]
+r <- c(x,y,z)
+
+
+
+png("noise_figure.png", 800,325, res=110)
+par(mar=c(5,4,4,0), oma=c(1,2,1,2))
+layout(matrix(c(1,2,3,1,2,3,1,2,3),nrow=3,byrow=T))
+plot(b$b_l[,1],b$b_l[,2], col="darkblue", ylab='b_1', xlab=NA, xlim=c(-4,4), ylim=c(-4,4),main='Left-tailed', cex.lab=1.5)
+plot(b$b_s[,1],b$b_s[,2], col="darkblue", xlab='b_2', ylab=NA, xlim=c(-4,4), ylim=c(-4,4),main='Symmetric',cex.lab=1.5)
+plot(b$b_r[,1],b$b_r[,2], col="darkblue", xlab=NA, ylab=NA, xlim=c(-4,4), ylim=c(-4,4),main='Right-tailed',cex.lab=1.5)
+#mtext(expression(rho==0.8258437), 1, outer=TRUE)
+dev.off()
+
+

### 2021-05-02 parameter_loop.R
new scripts

@@ -1,0 +1,19 @@
+# loop through parameters
+
+#mu1>mu2 
+
+delta <- seq(0,1,0.1)
+mu1 <- seq(.1,.9,.1) 
+mu2 <- mu1 
+sigma <- c(.4,.8,1.6, 3.2,6.4)
+
+tilde <- makenoise(100)
+
+
+for (i in 1:length(sigma)){
+  for (j in 1:length(mu1)){
+    for (k in 1:length(mu2)){
+      b_l1 <- sigma[i]*tilde$b_l[,1] + mu1[j]
+    }
+  }
+}

### 2021-05-05 integration/bounds_check_function.R
mathy scripts

@@ -1,0 +1,106 @@
+# write integral functions to test bounds
+# ten values
+
+
+# rbarsharp function - can use for both rbar1sharp and rbar2sharp
+integrand_rsharp <- function(u, mu1, mu2, sigma, delta){
+	mu <- mu1-mu2
+	sigma <- sqrt(2)*sigma
+	pdf1 <- (1/(sigma*sqrt(2*pi))*exp((-1/2)*((u-mu)/sigma)^2))
+	gr <- log(1-delta+delta*exp(u))
+	return(pdf1*gr)
+}
+
+# rbar1-term1
+integrand_r1T1 <- function(u, mu1, mu2, sigma, delta){
+	pdf1 <- (1/(sigma*sqrt(2*pi))*exp((-1/2)*(u/sigma)^2))
+	gr <- log(1-delta+delta*exp(mu1-mu2))
+	return(pdf1*gr)
+}
+
+# r1T2 - bivariate
+integrand_r1T2 <- function(vars, mu1, mu2, sigma, delta){
+	u1 <- vars[1]
+	u2 <- vars[2]
+	#u1 = b1-mu1; u2 = b2-mu2; so pdf has mean 0
+	pdf2 <- exp((-1/2)*(((u1)/sigma)^2+((u2)/sigma)^2))/(2*pi*sigma*sigma)
+	grt2 <- log(1-delta+delta*exp(u1-u2+mu1-mu2))
+	return(pdf2*grt2)
+}
+
+# runs multiple bounds checks for errors and returns value computed with the widest bounds if the value matches the value of the secound widest bounds
+bounds_uni <- function(integrand,mu1,mu2,sigma,delta, bounds=seq(350,800,50)){
+	
+	res <- matrix(c(bounds,rep(NA,10)), ncol=2)
+	
+	for (b in 1:length(bounds)){
+		lower <- -(bounds[b])
+		upper <- bounds[b]
+		int <- try(integrate(integrand, lower, upper, mu1, mu2, sigma, delta, stop.on.error=FALSE), silent=T)
+		
+		if (class(int)=="try-error"){
+			warning("error at ", bounds[b],"\n")
+		} else if (int$abs.error == 0){
+			warning("absolute error < 0 at ",bounds[b],"\n")
+		} else if (int$message != "OK"){
+			warning("message not OK at ",bounds[b],"\n")
+		} else {res[b,2] <- try(int$value, silent=T)
+			}
[…]

### 2021-05-05 integration/check_mathfunctions.R

@@ -1,0 +1,82 @@
+#source("./math_functions.r")
+
+delta <- 0.5
+mu <- seq(0.1,1,0.1)
+#mu1 <- sample(mu,1)
+#mu2 <- sample(mu,1)
+sdev <- seq(0.4,6.4,.4)
+#sigma <- rep(sample(sdev,1),2)
+
+bnoise <- getnoise2(mu=c(mu1,mu2), sigma=sigma, n=10^6, corval=corval, corRT=corRT)
+
+L1 <- data.frame(b1=bnoise$LT1[bnoise$LT1<mu1], b2=bnoise$LT2[bnoise$LT2<mu2])
+L2 <- data.frame(b1=bnoise$LT1[bnoise$LT1>mu1], b2=bnoise$LT2[bnoise$LT2>mu2])
+
+bnoisesharp <- getnoise2(mu=c(mu1,mu2), sigma=sigma, n=10^6, corval=corval, corRT=corRT)
+Bnoise <- exp(bnoise)
+Bnoisesharp <- exp(bnoisesharp)
+
+layout(matrix(c(1,1,1,2,2,
+				1,1,1,2,2,
+				1,1,1,3,3,
+				1,1,1,3,3), nrow=4, byrow=TRUE))
+plot(bnoise$LT1[1:1000], bnoise$LT2[1:1000])
+title(paste("mu1=",mu1," mu2=",mu2," sigma=", sigma[1]))
+plot(L2$b1[1:1000], L2$b2[1:1000], cex=0.4)
+plot(L1$b1[1:1000], L1$b2[1:1000],cex=0.4)
+
+SEres <- SE_lottery(Bnoise, Bnoisesharp, delta)
+semiana_res <- DeltaI(mu1, mu2, sigma[1], delta, 500, -500)
+names(semiana_res) <- colnames(SEres)
+r1t1 <- log(1-delta+delta*exp(L1$b1-L1$b2)); mean(r1t1)
+r1t2 <- log(1-delta+delta*exp(L2$b1-L2$b2)); mean(r1t2)
+r1 <- log(1-delta+delta*exp(bnoise$LT1-bnoise$LT2)); mean(r1)
+
+plot(r1[1:1000], col='grey40')
+abline(h=mean(r1), col='red', lty=2, lwd=2)
+abline(h=mean(r1t1)+mean(r1t2), col='green', lty=2, lwd=2)
+plot(r1t2[1:1000], col='grey40',cex=0.4)
[…]

### 2021-05-05 coexistence_metrics/dom_co.R
coMets

@@ -1,0 +1,299 @@
+#trying to measure dominance periods lengths 
+# 1. generate noise
+# 2. popsim
+# 3. measure lengths
+# 4. other metrics
+source("./new_functions/getnoise.R")
+source("./new_functions/popsim2.R")
+source("./new_functions/co.periods.R")
+source("./new_functions/measure.co3.R")
+
+#set correlation
+corval <- 0.88
+corRT <- getcorpar(0,.5,corval,numpts=100000,allowneg=FALSE,ploton=FALSE)
+
+#set params
+n <- 10^6
+N <- 50
+N1 <- 25
+delta <- seq(0,1,0.1)
+mu1 <- c(0.1,0.5,0.9); mu2 <- mu1
+sigma <- c(0.4, 0.8, 1.6, 3.2, 6.4)
+tot_sims <- length(sigma)*length(mu1)*length(mu2)*length(delta)
+delta_sims <- rep(delta, tot_sims/length(delta))
+mu2_sims <- sort(rep(mu2, length(delta)))
+mu1_sims <- sort(rep(mu1, length(mu2_sims)))
+sigma_sims <- sort(rep(sigma, length(mu1_sims)))
+paramdf <- data.frame(sigma=sigma_sims, mu1=mu1_sims, mu2=mu2_sims, delta=delta_sims)
+
+index1 <- c(1:11, 12, 14, 17, 20, 22, 23 ,25 ,28 ,31 ,33 ,45 ,47 ,50 ,53 ,55 ,56 ,58 ,61 ,64 ,66 ,89 ,91 ,94 ,97 ,99)
+indexpsued <- c(1, 3, 6, 9, 11, 12, 14, 17, 20, 22, 23 ,25 ,28 ,31 ,33 ,45 ,47 ,50 ,53 ,55 ,56 ,58 ,61 ,64 ,66 ,89 ,91 ,94 ,97 ,99)
+index2 <- c(index1, indexpsued+99)
+index3 <- c(index2, indexpsued+(99*2))
+index4 <- c(index3, indexpsued+(99*3))
+index5 <- c(index4, indexpsued+(99*4))
+
+for (i in 2:156){
+	params <- paramdf[index5[i],]; params
+
+	#generate noise
+	noise_b <- getnoise2(mu = c(params$mu1, params$mu2), sigma = c(params$sigma,params$sigma), n=n, corval=corval, corRT=corRT)
+		
+	#exponentiate noise
+	noise_B <- exp(noise_b)
+			
+	#population simulation
[…]

### 2021-05-05 integration/integrate.R
adding old untracked files

@@ -1,0 +1,88 @@
+? integrate
+
+install.packages("cubature")
+library(cubature)
+
+pdf1 <- function(x, mu=0, sigma=1){
+	pdf1 <- 1/(sigma*sqrt(2*pi))*exp((-1/2)*((x-mu)/sigma)^2)
+	return(pdf1)
+}
+
+gr <- function(u, delta=0.5){
+	gr <- log(1-delta+delta*exp(u))
+	return(gr)
+}
+
+integrand_rsharp1 <- function(u, mu=0, sigma=1, delta=0.5){
+	pdf1 <- (1/(sigma*sqrt(2*pi))*exp((-1/2)*((u-mu)/sigma)^2))
+	gr <- log(1-delta+delta*exp(u))
+	return(pdf1*gr)
+}
+
+curve(integrand_rsharp, from=0, to=250)
+
+x <- seq(-250,250,1)
+y <- integrand_rsharp(x, sigma = 1)
+plot(x,y, type="l")
+
+integrate(integrand_rsharp1, -500,500, mu=0, sigma=1.9)
+integrate(pdf1, -Inf, Inf)
+integrate(gr,-500,500)
+gr(1)
+
+mu1 <- 0.5; mu2 <- 0.4
+b1 <- #normal noise 1 #mean mu1 var sigma^2
+b2 <- #normal noise 2 #mean mu2 var sigma^2
+b1sharp <- #normal noise 1 sharp #mean mu1 var sigma^2
+b2sharp <- #normal noise 2 sharp #mean mu2 var sigma^2
+
+u_r1sharp <- b1sharp - b2 #normal with mean mu1-mu2 and var 2sigma^2
+u_r2sharp <- b2sharp - b2 #normal with mean 0 and var 2sigma^2 
+# use to plug into rsharp
+
+#term1 of rbar1
+integrand_r1t1 <- function(u, mu=0, sigma=1, delta=0.5, mu1=0.5, mu2=0.4){
+	pdf1 <- (1/(sigma*sqrt(2*pi))*exp((-1/2)*((u-mu)/sigma)^2))
+	grt1 <- log(1-delta + delta*exp(mu1-mu2))
+	return(pdf1*grt1)
+}
+
+integrate(integrand_r1t1, -Inf, 0)
+
+
[…]

### 2021-05-05 integration/math.R

@@ -1,0 +1,259 @@
+source("./new_functions/getnoise.r")
+source("./new_functions/SE_lottery.R")
+source("./new_functions/measure.co3.R")
+source("./new_functions/popsim2.R")
+source("./new_functions/co.periods.R")
+
+#solving IGR expected value
+
+# adding and subtracting normals
+x1 <- rnorm(10^5, 0.5, 3.2)
+x2 <- rnorm(10^5, 0.3, 3.2)
+mean(x1);sd(x1)
+mean(x2);sd(x2)
+hist(x1, 50)
+hist(x2,50)
+X <- x1-x2
+hist(X,50)
+mean(X); sd(X)
+xp1 <- X+x2 # X = x1-x2, x1 =? X+x2?
+hist(xp1)
+mean(xp1); sd(xp1)
+var(x2)
+var(X)
+var(xp1)
+cov(x2, X)
+
+xp2 <- x1-X
+mean(xp2);sd(xp2)
+
+# expected value of log normal 
+# if X is a normal random var and Y = exp(X) then Y is lognormal
+# in our equation we have B1/B2 where Bi = exp(bi) and bi is normal bivariate with means mu1 and mu2 and sd sigma 
+# thus B1/B2 = exp(b1)/exp(b2) = exp(b1-b2)
+# subtracting normal distributions gives us a normal distribution
+# let b1-b2 = a where a~N(mu1-mu2, 2sigma^2)
+# let Q = exp(a) so Q = exp(b1-b2)=B1/B2
+
+mu1 <- 0.5
+mu2 <- 0.3
+sigma <- 0.8
+b1 <- rnorm(10^5, mu1, sigma); var(b1)
+b2 <- rnorm(10^5, mu2, sigma); var(b2)
+a <- b1-b2
+mean(a);var(a);sd(a)
+Q <- exp(a)
+
+# if X is a normal random var with mean mu and sd sigma and Y = exp(X) then Y is lognormal with first moment (expected value) e^(mu + (sigma^2)/2)
+# since a=b1-b2 its mean is mu1-mu2 and sd is sqrt(2)sigma
+# so expcted value of Q=exp(a) is e^((mu1-mu2) + sigma^2
+
+Z <- rnorm(10^7,0,1)
[…]

### 2021-05-05 new_functions/getnoise.r

@@ -1,0 +1,163 @@
+library(copula)
+
+#Generates points from certain bivariate copulas with tail dependence
+#
+#Args
+#corpar           For the part of the distribution where the variables are
+#                   not perfectly related they are generated with a bivariate 
+#                   normal copula with this parameter. Values between -1 and 1.
+#beta1, beta2     These specify where the tail dependence starts in each tail.
+#                   Numbers between 0 and 1 with beta1<beta2.
+#numpts           Number of draws from the copula desired
+#
+#Output
+#A matrix of dimensions numpts by 2. The two variables are each uniformly
+#distributed, and are perfectly correlated to the left of beta1 and to the right
+#of beta 2.
+#
+taildepcopdat<-function(corpar,beta1,beta2,numpts)
+{
+  #some error catching
+  if (corpar>1 || corpar<(-1))
+  {
+    stop("Error in taildepcopdat: bad value of corpar")
+  }
+  if (beta1<0 || beta2>1 || beta1>beta2)
+  {
+    stop("Error in taildepcopdat: bad values of beta1 and beta2")
+  }
+  
+  #get the necessary copula object
+  corpar<-matrix(corpar,2,2)
+  diag(corpar)<-1
+  corpar<-P2p(corpar)
+  ncop<-normalCopula(corpar,2,"un")
+  
+  #generate the initial noise
+  res<-rCopula(numpts,ncop)
+  
+  #impose the tail association
+  tdornot<-runif(numpts)
+  inds<-(tdornot<beta1 | tdornot>beta2)
+  res[inds,]<-tdornot[inds] #observations are same at the rows
+  res[!inds,]<-(beta2-beta1)*res[!inds,]+beta1 #reassigning values between betas to be shift mean by beta1, scale variance by beta2-beta1 
+
+  return(res)
+}
+
+#Just like the above, but a transformation is applied to make the marginals be normal
+#with mean vector mu and standard deviation vector std
+#
+taildepnormdat<-function(corpar,beta1,beta2,numpts,mu=c(0,0),std=c(1,1))
+{
+  if (any(std<=0))
+  {
+    stop("Error in taildepnormdat: bad value for std")
+  }
+  
+  res<-taildepcopdat(corpar,beta1,beta2,numpts)
+  res<-qnorm(res)
+  res[,1]<-res[,1]*std[1]+mu[1]
[…]

### 2021-05-05 new_functions/tests_getnoise.r

@@ -1,0 +1,202 @@
+rm(list=ls())
+
+library(testthat)
+
+source("getnoise.R")
+
+#***visual checks of taildepcopdat
+
+set.seed(101)
+res<-taildepcopdat(.5,.25,.75,1000)  
+plot(res[,1],res[,2],type="p")
+res<-taildepcopdat(.9,.25,.75,1000)  
+plot(res[,1],res[,2],type="p")
+res<-taildepcopdat(0,.25,.75,1000)  
+plot(res[,1],res[,2],type="p")
+res<-taildepcopdat(.9,.1,.75,1000)  
+plot(res[,1],res[,2],type="p")
+res<-taildepcopdat(.9,.5,.9,1000)  
+plot(res[,1],res[,2],type="p")
+res<-taildepcopdat(-.9,.5,.9,1000)  
+plot(res[,1],res[,2],type="p") #this last one is weird but not wrong - we would not normally use it
+
+#***other tests of tauldepcopdat
+
+set.seed(101)
+beta1<-0.1
+beta2<-0.75
+
+#test the error catching
+expect_error(taildepcopdat(1.1,beta1,beta2,100),"Error in taildepcopdat: bad value of corpar")
+expect_error(taildepcopdat(.9,-.1,.5,100),"Error in taildepcopdat: bad values of beta1 and beta2")
+expect_error(taildepcopdat(.9,.1,1.1,100),"Error in taildepcopdat: bad values of beta1 and beta2")
+expect_error(taildepcopdat(.9,.2,.1,100),"Error in taildepcopdat: bad values of beta1 and beta2")
+
+#test a boundary case with corpar=1
+res<-taildepcopdat(1,beta1,beta2,100)
+expect_equal(res[,1],res[,2])
+  
+#test a boundary case with beta1 and beta2 equal
+res<-taildepcopdat(.5,.5,.5,100)
+expect_equal(res[,1],res[,2])
+    
+#test a non-boundary example
+res<-taildepcopdat(.5,beta1,beta2,100)  
+  
+#make sure it's the right format
+expect_equal(class(res),"matrix")
+expect_equal(dim(res),c(100,2))
+expect_true(all(res>=0))
+expect_true(all(res<=1))
+  
+#make sure the components are the same when below beta1 or above beta2
[…]

### 2021-05-05 plot_2.R

@@ -1,0 +1,80 @@
+dflist[[1]][dflist[[1]]$sdev == 0.4 & dflist[[1]]$mn1 ==0.6 & dflist[[1]]$delta == 1.0,]
+
+#range mn2 while mn1 = 0.6, sdev = 0.4, delta = 1.0
+
+plot1_list <- vector(mode = 'list', length =10)
+for (i in 1:10) {
+	plot1_list[[i]] <- dflist[[i]][dflist[[i]]$sdev == 0.4 & dflist[[i]]$mn1 ==0.6 & dflist[[i]]$delta == 1.0,]
+}
+
+
+plot(0, ylim = c(-6e-04,4e-04), xlim = c(0, 1), type = 'n', ylab = "SE", xlab = "mn2", main = "Delta I according to changes in mn2", sub = "sdev = 0.4, mn1=0.6, delta=1")
+legend("topleft", legend = c("extreme left tail", "symmetric", "extreme right tail"), col = c("red", "green", "blue"), box.lty =0)
+
+for (i in 1:10) {
+	lines(plot1_list[[i]]$mn2, plot1_list[[i]]$ELT, col = rgb(1,0,0, 0.2))
+	lines(plot1_list[[i]]$mn2, plot1_list[[i]]$sym, col = rgb(0,1,0, 0.2))
+	lines(plot1_list[[i]]$mn2, plot1_list[[i]]$ERT, col = rgb(0,0,1, 0.2))
+	
+}
+
+lines(seq(.1,.9,.1), mean_df[mean_df$sdev == 0.4 & mean_df$mn1 == 0.6 & mean_df$delta == 1,]$ELT, col = 'red', lty = 2)
+lines(seq(.1,.9,.1), mean_df[mean_df$sdev == 0.4 & mean_df$mn1 == 0.6 & mean_df$delta == 1,]$sym, col = 'green', lty = 2)
+lines(seq(.1,.9,.1), mean_df[mean_df$sdev == 0.4 & mean_df$mn1 == 0.6 & mean_df$delta == 1,]$ERT, col = 'blue', lty = 2)
+
+#range delta while mn1= 0.6, mn2= 0.1, sdev= 0.4
+mn2 <- 0.9
+
+plot2_list <- vector(mode = 'list', length =10)
+for (i in 1:10) {
+	plot2_list[[i]] <- dflist[[i]][dflist[[i]]$sdev == 0.4 & dflist[[i]]$mn1 ==0.6 & dflist[[i]]$mn2 == mn2,]
+}
+
+
[…]

### 2021-05-05 plot_mean_df2.R

@@ -1,0 +1,43 @@
+source("./plot_paramDelta.R")
+
+params: vector of parameters for the plot c(sigma, mu1,mu2)
+
+plot_SE_Delta <- function(params, df){
+	delta <- seq(0, 1.0, 0.1)
+	plot(delta, df[df[,1] == params[1] & df[,2] == params[2] & df[,3] == params[3],6], col='green', type='l')
+	lines(delta, df[df[,1] == params[1] & df[,2] == params[2] & df[,3] == params[3],5], col='red')
+	lines(delta, df[df[,1] == params[1] & df[,2] == params[2] & df[,3] == params[3],7], col='blue')
+}
+
+params <- c(6.4, 0.1, 0.9)
+mean_df<- signif(mean_df,4)
+
+plot_SE_Delta(params, mean.df2)
+plot_SE_Delta(params, mean_df)
+
+
+params_mu2 <- rep(c(0.1,0.5,0.9), 3*5)
+params_df <- data.frame(sigma=NA, mu1=NA, mu2=params_mu2)
+params_df$mu1 <- rep(sort(rep(c(0.1,0.5,0.9), 3)),5)
+params_df$sigma <- sort(rep(c(0.4,0.8,1.6,3.2,6.4),9))
+
+pdf('check_graphs.pdf')
+
+for(i in 1:nrow(params_df)){
+	plot_SE_Delta(as.numeric(params_df[i,]), mean.df2)
+	title(paste(params_df[i,]))
+	plot_SE_Delta(as.numeric(params_df[i,]), mean_df)
+	title(paste(params_df[i,]))
+}
+
+dev.off()
+
+mean.df2[mean.df2$sigma==params[1] & mean.df2$mu1==params[2] & mean.df2$mu2==params[3],1:7]
+
+mean_df[mean_df$sdev==params[1] & mean_df$mn1==params[2] & mean_df$mn2==params[3],1:7]
+
+
+
+
+
+

### 2021-05-05 random_unused/Extrme_relDif.R

@@ -1,0 +1,29 @@
+#trying to find relative difference of extreme tails to symmetric 
+#ERT-ELT/(sym - mean(ERT,ELT))
+#difference between tail direction relative to difference between symmetric and avg of tailed noise
+
+#useful? maybe no if already have anova
+
+Edif <- matrix(NA, ncol = length(dflist), nrow = nrow(simdfx))
+
+for (i in 1:length(dflist)){
+	dflist[[i]]$Emean <- rowMeans(cbind(dflist[[i]]$ERT, dflist[[i]]$ELT))
+	dflist[[i]]$Edif <-abs(dflist[[i]]$ERT - dflist[[i]]$ELT)/abs(dflist[[i]]$sym - dflist[[i]]$Emean)
+	
+	Edif[,i] <- dflist[[i]]$Edif
+	
+}
+
+mean_df$Edif <- rowMeans(Edif)
+#----------------------------------------------------
+
+
+head(mean_df)
+
+plot(mean_df$sdev, mean_df$Edif)
+
+hist(mean_df$Edif,500)
+
+b <-mean_df[which(mean_df$Edif==max(mean_df$Edif, na.rm = TRUE)),]
+mean(c(b$ELT, b$ERT))
+(b$ERT-b$ELT)/abs(b$sym-mean(c(b$ELT, b$ERT)))

### 2021-05-05 runs/param_sim.R

@@ -1,0 +1,122 @@
+source("./NoiseAndSE.R")
+#This is code for running through the storage effect calculations for various sets of parameters as set below a number of x times
+#Parameters include delta (death rate), mn1 and mn2 (mean birth rate of sp1 and sp2) and sdev (standard deviation of birth rate; equal for both species)
+
+
+delta <- seq(0,1,0.1) #0 to 1 by 0.1
+mn1 <- seq(.1,.9,.1) #0.1 to 0.9 by 0.1
+mn2 <- seq(.1,.9,.1) #0.1 to 0.9 by 0.1
+sdev <- c(.4,.8,1.6, 3.2,6.4) #doubling from 0.4 to 6.4
+x <- 10 #number of run repeats
+n <- 10^6
+
+#noise_SE(mn1, mn2, sdev, n, delta)
+# ^the function we will use for this run; a function that generates noise and then calculates SE
+
+
+tot_sims <- length(sdev)*length(mn1)*length(mn2)*length(delta)
+# for above paramters, there will be 4455 unique runs
+
+
+#Structure for nest: sdev<mn1<mn2<delta
+
+#organize parameters to follow nest structure from terminals inward and keep with length of tot_sims
+
+delta_sims <- rep(delta, tot_sims/length(delta))
+#repeat delta to fill tot_sims
+
+mn2_sims <- sort(rep(mn2, length(delta)))
+#repeat one mn2 value for every value of delta
+
+mn1_sims <- sort(rep(mn1, length(mn2_sims)))
+#repeat one mn1 value for every unique combination of mn2 and delta
+
+sdev_sims <- sort(rep(sdev, length(mn1_sims)))
+#repeat one sdev value for every unique combination of mn1, mn2, and delta 
+
+#these vectors will be used to fill columns of dataframe tot_sims long
+#vectors that did not reach tot_sims length will automatically repeat to reach that length
+
+
+#for data storage: a list of dataframes, each dataframe is a run repeat
+
+dflist <- vector(mode = 'list', length = x)
+
+simdfx <- data.frame(sdev=sdev_sims, mn1=mn1_sims, mn2=mn2_sims, delta=delta_sims, ELT=NA, sym=NA, ERT=NA)
+#data.frame with columns specifying the parameter values and columns to be filled with storage effect for ELT, sym, ERT
+
[…]

### 2021-05-05 runs/run4.R

@@ -1,0 +1,91 @@
+source("./getnoise.R")
+source("./popsim2.R")
+source("./co.periods.R")
+source("./measure.co3.R")
+
+#parameters for noise
+delta <- seq(0,1,0.1)
+mu1 <- c(0.1,0.5,0.9)
+mu2 <- mu1
+sigma <- c(0.4, 0.8, 1.6, 3.2, 6.4)
+
+corval <- 0.88
+corRT <- getcorpar(0,.5,corval,numpts=100000,allowneg=FALSE,ploton=FALSE)
+
+repeats <- 10
+N <- 50
+N1 <- 25
+n <- 10^6
+
+tot_sims <- length(sigma)*length(mu1)*length(mu2)*length(delta)
+
+delta_sims <- rep(delta, tot_sims/length(delta))
+mu2_sims <- sort(rep(mu2, length(delta)))
+mu1_sims <- sort(rep(mu1, length(mu2_sims)))
+sigma_sims <- sort(rep(sigma, length(mu1_sims)))
+
+params <- data.frame(sigma=sigma_sims, mu1=mu1_sims, mu2=mu2_sims, delta=delta_sims)
+
+poplist <- vector(mode='list', length=tot_sims)
+popreps <- vector(mode='list', length=10)
+CoMres <- matrix(0, nrow=tot_sims, ncol=90)
+Sys.time()
+#rep loop
+set.seed(2727)
+for(rep in 0:(repeats-1)){
+	#param loop
+	for (i in 1:tot_sims){
+		#generate noise
+		noise_b <- getnoise2(mu = c(params$mu1[i], params$mu2[i]), sigma = c(params$sigma[i],params$sigma[i]), n=n, corval=corval, corRT=corRT)
+		
+		#exponentiate noise
+		noise_B <- exp(noise_b)
+			
+		#population simulation
+		popLT <- popsim2(B=cbind(noise_B$LT1, noise_B$LT2), N=N, N1=N1, delta = params$delta[i], n=n)
+		popsym <- popsim2(B=cbind(noise_B$sym1, noise_B$sym2), N=N, N1=N1, delta = params$delta[i], n=n)
+		popRT <-popsim2(B=cbind(noise_B$RT1, noise_B$RT2), N=N, N1=N1, delta = params$delta[i], n=n)
+		
+		pop <- cbind(popLT, popsym, popRT)
+	
+		#coexistence periods
[…]

### 2021-05-05 runs/run5.analysis.R

@@ -1,0 +1,9 @@
+# co and dom metrics from saved population simulations from run5
+
+# in loop - read csv, co.periods, and measure.co
+read.csv(...)
+
+
+
+
+

### 2021-05-05 runs/runningParams3.R

@@ -1,0 +1,128 @@
+#running params, third times a charm
+source("./getnoise.R")
+source("./SE_lottery.R")
+source("./popsim.R")
+source("./co.periods.R")
+source("./measure.co3.R")
+
+#parameters for noise
+delta <- seq(0,1,0.1)
+mu1 <- seq(0.1,0.9,0.1)
+mu2 <- mu1
+sigma <- c(0.4, 0.8, 1.6, 3.2, 6.4)
+
+corval <- 0.88
+corRT <- getcorpar(0,.5,corval,numpts=100000,allowneg=FALSE,ploton=FALSE)
+
+repeats <- 10
+
+n <- 10^6
+
+
+#make parameter data.frame to pull from
+tot_sims <- length(sigma)*length(mu1)*length(mu2)*length(delta)
+
+delta_sims <- rep(delta, tot_sims/length(delta))
+mu2_sims <- sort(rep(mu2, length(delta)))
+mu1_sims <- sort(rep(mu1, length(mu2_sims)))
+sigma_sims <- sort(rep(sigma, length(mu1_sims)))
+
+params <- data.frame(sigma=sigma_sims, mu1=mu1_sims, mu2=mu2_sims, delta=delta_sims)
+
+
+
+#initialize storage
+#big.list <- vector(mode='list', length=tot_sims)
+
+
+#initailize storage
+
+SEres <- matrix(0, nrow=tot_sims, ncol=120)
+CoMres <- matrix(0, nrow=tot_sims, ncol=90)
+
+
+#rep loop
+set.seed(2727)
+for(rep in 0:(repeats-1)){
+	#param loop
+	for (i in 1:tot_sims){
+		#generate noise
+		noise_b <- getnoise2(mu = c(params$mu1[i], params$mu2[i]), sigma = c(params$sigma[i],params$sigma[i]), n=n, corval=corval, corRT=corRT)
+		noise_b_sharp <- getnoise2(mu = c(params$mu1[i], params$mu2[i]), c(params$sigma[i],params$sigma[i]), n=n, corval=corval, corRT=corRT)
+	
+		#exponentiate noise
+		noise_B <- exp(noise_b)
+		noise_B_sharp <- exp(noise_b_sharp)
+	
+		#compute SE
+		SE <- SE_lottery(noise_B, noise_B_sharp, params$delta[i])
+		SEvec <- c(SE, recursive=T)
+	
[…]

### 2021-05-05 simcheck_SEplots.R

@@ -1,0 +1,100 @@
+#checking simulation
+
+res1<-read.csv("generated_data/run3/SEandCoMresults202101.csv")
+res1 <- res1[,-1]
+res1all <- read.csv("generated_data/run3/SE_results.csv")
+res1all <- data.frame(cbind(res1[,1:4],res1all[,2:31]))
+res0 <- read.csv("generated_data/run1/sim_mean")
+res0 <- res0[,-1]
+res0_1 <- read.csv("generated_data/run1/sim_df1")[,6:8]
+res0_2 <- read.csv("generated_data/run1/sim_df2")[,6:8]
+res0_3 <- read.csv("generated_data/run1/sim_df3")[,6:8]
+res0_4 <- read.csv("generated_data/run1/sim_df4")[,6:8]
+res0_5 <- read.csv("generated_data/run1/sim_df5")[,6:8]
+res0_6 <- read.csv("generated_data/run1/sim_df6")[,6:8]
+res0_7 <- read.csv("generated_data/run1/sim_df7")[,6:8]
+res0_8 <- read.csv("generated_data/run1/sim_df8")[,6:8]
+res0_9 <- read.csv("generated_data/run1/sim_df9")[,6:8]
+res0_10 <- read.csv("generated_data/run1/sim_df10")[,6:8]
+res0all <- cbind(res0_1,res0_2,res0_3,res0_4,res0_5,res0_6,res0_7,res0_8,res0_9,res0_10)
+res0all <- data.frame(res0all)
+res0all <- res0all[,order(colnames(res0all))]
+res0all <- cbind(res1[,1:4],res0all)
+
+
+
+res0[,5:7] <- res0[,5:7][,order(colnames(res0[,5:7]))]
+colnames(res0) <- c(colnames(res0[,1:4]),"ELT", "ERT","sym")
+head(res0[,5:7])
+
+rp <- sample(1:4455,1);rp #random parameters
+
+rp <- sample(1:4455,1);res0[rp,1:7];res1[rp,1:7]
+
+delta <- res1$delta[1:11]
+
+plot(0, ylim=c(-.1,.5), xlim=c(0,1), type='n', xlab=expression(delta))
+lines(delta, res1[res1$sigma==0.4 & res1$mu1==0.1 & res1$mu2==0.1,]$Delta_B1ELT)
+
+plotDeltaI <- function(df, params){
+	
+	x <- df[1:11,4]
[…]

### 2021-05-05 t1_mean_sd.R

@@ -1,0 +1,18 @@
+
+mu <- seq(0.1,1,0.1)
+mu1 <- sample(mu,1)
+mu2 <- sample(mu,1)
+sdev <- seq(0.4,6.4,.4)
+sigma <- rep(sample(sdev,1),2)
+
+bnoise <- getnoise2(mu=c(mu1,mu2), sigma=sigma, n=10^6, corval=corval, corRT=corRT)
+
+L1 <- data.frame(b1=bnoise$LT1[bnoise$LT1<mu1], b2=bnoise$LT2[bnoise$LT2<mu2])
+L2 <- data.frame(b1=bnoise$LT1[bnoise$LT1>mu1], b2=bnoise$LT2[bnoise$LT2>mu2])
+D <- L1$b1 - L1$b2
+
+cat("parameters: mu1=",mu1, " mu2=",mu2," sigma=", sigma[1], "\n\n")
+cat("D = b1t1-b2t1 \n\n")
+cat("mean(D):",mean(D), "sd(D):",sd(D) )
+
+#source("./t1_mean_sd.R")

### 2021-05-05 testing_ellner2020.R

@@ -1,0 +1,51 @@
+source("./getnoise.R")
+source("./popsim2.R")
+source("./co.periods.R")
+source("./measure.co3.R")
+source("./SE_lottery.R")
+
+###########################################################################
+
+# we do not quwation any facts reported by PFCS, but we argue that they empahasized part of a larger picture, and therefore overstated implications for past and future research. 
+
+# the two-species lottery model illustrates why higher IGRs do not necessarily yeild more robust persistence.
+# IGR_1 (the IGR of the weaker comepetitor) increases with environmental variability, sigma (the temporal standard deviation of log per capita fecundity).
+		#IGR_2 = 0 in lottery model
+# The weaker competitor quickly goes extinct when IGR_1<0 (sigma=0.2) but not when IGR_1>0.
+
+n <- 10^6
+
+corval <- 0.88
+corRT <- getcorpar(0,.5,corval,numpts=100000,allowneg=FALSE,ploton=FALSE); corRT
+
+mu1 <- 0.45
+mu2 <- 0.5
+sigma <- 1.1
+
+#generate noise
+noise_b <- getnoise2(mu = c(mu1,mu2), sigma = c(sigma, sigma), n=n, corval=corval, corRT=corRT)
+noise_b_sharp <- getnoise2(mu = c(mu1,mu2), sigma = c(sigma, sigma), n=n, corval=corval, corRT=corRT)
+	
+#exponentiate noise
+noise_B <- exp(noise_b)
+noise_B_sharp <- exp(noise_b_sharp)
+		
+SE <- SE_lottery(noise_B, noise_B_sharp, 0.5); SE
+
+popLT <- popsim2(B=cbind(noise_B$LT1, noise_B$LT2), N=50, N1=25, delta = 0.5, n=n)
+popsym <- popsim2(B=cbind(noise_B$sym1, noise_B$sym2), N=50, N1=25, delta = 0.5, n=n)
+popRT <-popsim2(B=cbind(noise_B$RT1, noise_B$RT2), N=50, N1=25, delta = 0.5, n=n)
+
+plot(popLT$N1[1:2000], type = 'l', main="Asymetric (left)"); abline(h=1, col="red")
+plot(popsym$N1[1:2000], type = 'l', main="Symetric"); abline(h=1, col="red")
+plot(popRT$N1[1:2000], type = 'l', main="Asymetric (right)"); abline(h=1, col="red")
+
[…]

### 2021-05-05 paper/SupMat.Rnw
new fig

--- 2021-05-02 paper/SupMat.tex
@@
 
-Finally, for a ?low? of species of? parameters $\delta, \mu_1, \mu_2, \sigma$, we displayed the whole distribution of $r_1$, values for each of our tail association cases, to make visual comparisons.
+Finally, for a few choices of parameters $\delta, \mu_1, \mu_2, \sigma$, we displayed the whole distribution of $r_1$, values for each of our tail association cases, to make visual comparisons.
 
@@
 
-Equations \ref{sym_rbar1} - \ref{sym_rbar2sharp} obviously depend only on $\mu_1-\mu_2$, and not independently on $\mu_1$ and $\mu_2$, ?some?
-only $\mu_1-\mu_2$ appears in these expressions if $\mu_1$ or $\mu_2$ appear at all.
+Equations \ref{sym_rbar1} - \ref{sym_rbar2sharp} obviously depend only on $\mu_1-\mu_2$, and not independently on $\mu_1$ and $\mu_2$, since only $\mu_1-\mu_2$ appears in these expressions if $\mu_1$ or $\mu_2$ appear at all.
 Equation \ref{sym_rbar1} is $\bar r_1$ in the asymmetric?
@@
 
-To see that our various metrics are the same for out left- and right-tail associated noise ?parameters?, let $(b_1,b_2)$ denote our left-tail associated random variable with parameters $\mu_1,\mu_2,\sigma$.
+To see that our various metrics are the same for out left- and right-tail associated noise processes, let $(b_1,b_2)$ denote our left-tail associated random variable with parameters $\mu_1,\mu_2,\sigma$.
 Then define $\beta_1 = -b_2 + \mu_1 +\mu_2$, $\beta_2=-b_1+\mu_1+\mu_2$.

### 2021-05-05 noise_fig.R

@@ -10,10 +10,10 @@ r <- c(x,y,z)
 
 
 png("noise_figure.png", 800,325, res=110)
-par(mar=c(5,4,4,0), oma=c(1,2,1,2))
+par(mar=c(5,4.5,4,0), oma=c(1,2,1,2))
 layout(matrix(c(1,2,3,1,2,3,1,2,3),nrow=3,byrow=T))
-plot(b$b_l[,1],b$b_l[,2], col="darkblue", ylab='b_1', xlab=NA, xlim=c(-4,4), ylim=c(-4,4),main='Left-tailed', cex.lab=1.5)
-plot(b$b_s[,1],b$b_s[,2], col="darkblue", xlab='b_2', ylab=NA, xlim=c(-4,4), ylim=c(-4,4),main='Symmetric',cex.lab=1.5)
+plot(b$b_l[,1],b$b_l[,2], col="darkblue", ylab=expression(b[2]), xlab=NA, xlim=c(-4,4), ylim=c(-4,4),main='Left-tailed', cex.lab=1.5)
+plot(b$b_s[,1],b$b_s[,2], col="darkblue", xlab=expression(b[1]), ylab=NA, xlim=c(-4,4), ylim=c(-4,4),main='Symmetric',cex.lab=1.5)
 plot(b$b_r[,1],b$b_r[,2], col="darkblue", xlab=NA, ylab=NA, xlim=c(-4,4), ylim=c(-4,4),main='Right-tailed',cex.lab=1.5)
 #mtext(expression(rho==0.8258437), 1, outer=TRUE)
 dev.off()

### 2021-05-05 makenoise.R
changed list names

@@ -34,7 +34,7 @@ makenoise <- function(M){
   b_s <- mvrnorm(M, mu_norm, Sigma_sym)
   #plot(b_s[,1],b_s[,2])
   
-  return(list(b_l=b_l,b_r=b_r,b_s=b_s))
+  return(list(l=b_l,r=b_r,s=b_s))
 }
 
 #makenoise(100)

### 2021-05-05 paper/SupMat.Rnw
editted

@@
 We take, for each $i$, $B_i(t)$ to be $iid$ through time.
-We let $B_i = exp(b_i)$, where $(b_1, b_2)$ is some distribution such that $b_i \sim N(\mu_i, \sigma^2)$ and such that $cov(b_1, b_2) = \rho$, a fixed quantity we will specify later.
+We let $B_i = \exp(b_i)$, where $(b_1, b_2)$ is some distribution such that $b_i \sim N(\mu_i, \sigma^2)$ and such that $cov(b_1, b_2) = \rho$, a fixed quantity we will specify later.
 However, $(b_1, b_2)$ is not necessarily a bivariate normal distribution.
@@
 \begin{equation}
-\bar r_1=\EX(\ln [(1-\delta)+\delta exp(b_1-b_2)]).  \label{rbar1_E}
+\bar r_1=\EX(\ln [(1-\delta)+\delta \exp(b_1-b_2)]).  \label{rbar1_E}
 \end{equation}
@@
 \begin{equation}
-\bar r_1^\#=\EX(\ln [(1-\delta)+\delta exp(b_1^\#-b_2)]),  \label{rbar1sharp_E}
+\bar r_1^\#=\EX(\ln [(1-\delta)+\delta \exp(b_1^\#-b_2)]),  \label{rbar1sharp_E}
 \end{equation}
@@
 \begin{equation}
-\bar r_2^\#=\EX(\ln [(1-\delta)+\delta exp(b_2^\#-b_2)]),  \label{rbar2sharp_E}
+\bar r_2^\#=\EX(\ln [(1-\delta)+\delta \exp(b_2^\#-b_2)]),  \label{rbar2sharp_E}
 \end{equation}
@@
 \begin{equation}
-\bar r_1^\#=\EX(\ln [(1-\delta)+\delta exp(u)]),  \label{rbar1sharp_E_usub}
+\bar r_1^\#=\EX(\ln [(1-\delta)+\delta \exp(u)]),  \label{rbar1sharp_E_usub}
 \end{equation}
@@
 \begin{equation}
-\bar r_2^\#=\EX(\ln [(1-\delta)+\delta exp(u)]),  \label{rbar2sharp_E_usub}
+\bar r_2^\#=\EX(\ln [(1-\delta)+\delta \exp(u)]),  \label{rbar2sharp_E_usub}
 \end{equation}
@@
 \begin{equation}
-(b_1,b_2) \sim  N(\begin{pmatrix} \mu_2\\\mu_2\end{pmatrix}, \Sigma) \label{b1b2_distribution_sym}
+(b_1,b_2) \sim  N(\begin{pmatrix} \mu_1\\\mu_2\end{pmatrix}, \Sigma) \label{b1b2_distribution_sym}
 \end{equation}
@@
 tail association cases.
-But letting $\beta_1 = b_1 + \eta$ and $\beta_2 = b_2 + \eta$ and substituting into \ref{sym_rbar1}, we get
+But letting $\beta_1 = b_1 + \eta$ and $\beta_2 = b_2 + \eta$ and substituting into \ref{rbar1_E}, we get
 

### 2021-05-05 parameter_loop.R
completed loop

@@ -1,19 +1,96 @@
 # loop through parameters
 
-#mu1>mu2 
 
-delta <- seq(0,1,0.1)
+
+delta <- seq(0.1,1,0.1)
 mu1 <- seq(.1,.9,.1) 
 mu2 <- mu1 
-sigma <- c(.4,.8,1.6, 3.2,6.4)
+sigma <- c(1.6, 3.2,6.4)
 
-tilde <- makenoise(100)
-
-
-for (i in 1:length(sigma)){
-  for (j in 1:length(mu1)){
-    for (k in 1:length(mu2)){
-      b_l1 <- sigma[i]*tilde$b_l[,1] + mu1[j]
+#mu1>mu2 
+mu <- c(NA,NA)
+for (i in 1:length(mu1)){
+  for (j in 1:length(mu2)){
+    if(mu1[i]<=mu2[j]){
+      mu <- rbind(mu, cbind(mu1[i],mu2[j]))
     }
   }
 }
+mu <- mu[-1,]
+#print(mu) #45x2
+
+M <- 100000
+b_tilde <- makenoise(M)
+rho <- cor(b_tilde$l)[1,2];rho
+u_tilde <- rnorm(M)
+
+rbar <- c(NA,NA,NA,NA,NA)
+se_rbar <- rbar
+DeltaI <- c(NA,NA,NA)
+se_DeltaI <- DeltaI
+params <- rep(NA, 4)
+
+for (s in 1:length(sigma)){
+  for (m in 1:nrow(mu)){
+    #print <- append(print,paste(mu1[b],mu2[b]))
+    b_l1 <- sigma[s]*b_tilde$l[,1] + mu[m,1]
+    b_l2 <- sigma[s]*b_tilde$l[,2] + mu[m,2]
+    
+    b_r1 <- sigma[s]*b_tilde$r[,1] + mu[m,1]
+    b_r2 <- sigma[s]*b_tilde$r[,2] + mu[m,2]
+    
+    u_s <- sqrt(2*sigma[s]^2 - 2*rho)*u_tilde + (mu[m,1]-mu[m,2]) 
+    u_1 <- sqrt(2*sigma[s]^2)*u_tilde + (mu[m,1]-mu[m,2])
+    u_2 <- sqrt(2*sigma[s]^2)*u_tilde
+    
+    for (d in 1:length(delta)){
+      #rbar1
+      #left tail (1)
+      rbar1hat_l <- mean(log(1-delta[d]+delta[d]*exp(b_l1-b_l2)))
+      
[…]

### 2021-05-05 parameter_loop.R
added saving protocol

@@ -1,32 +1,12 @@
 # loop through parameters
 
+load("./params.RData")
 
+load("./noise_etc.RData")
 
-delta <- seq(0.1,1,0.1)
-mu1 <- seq(.1,.9,.1) 
-mu2 <- mu1 
-sigma <- c(1.6, 3.2,6.4)
-
-#mu1>mu2 
-mu <- c(NA,NA)
-for (i in 1:length(mu1)){
-  for (j in 1:length(mu2)){
-    if(mu1[i]<=mu2[j]){
-      mu <- rbind(mu, cbind(mu1[i],mu2[j]))
-    }
-  }
-}
-mu <- mu[-1,]
-#print(mu) #45x2
-
-M <- 100000
-b_tilde <- makenoise(M)
-rho <- cor(b_tilde$l)[1,2];rho
-u_tilde <- rnorm(M)
-
-rbar <- c(NA,NA,NA,NA,NA)
+rbar <- rep(NA, 5)
 se_rbar <- rbar
-DeltaI <- c(NA,NA,NA)
+DeltaI <- rep(NA,3)
 se_DeltaI <- DeltaI
 params <- rep(NA, 4)
 
@@ -83,7 +63,7 @@ for (s in 1:length(sigma)){
       DeltaIhat_s <- rbar1hat_s - rbar1sharphat + rbar2sharphat
       se_DeltaIhat_s <- se_rbar1hat_s + se_rbar1sharphat + se_rbar2sharphat
       
-      #saving
+      #save in matrix
       rbar <- rbind(rbar, cbind(rbar1hat_l,rbar1hat_r,rbar1hat_s,rbar1sharphat,rbar2sharphat))
       se_rbar <- rbind(se_rbar, cbind(se_rbar1hat_l,se_rbar1hat_r,se_rbar1hat_s,se_rbar1sharphat,se_rbar2sharphat))
       DeltaI <- rbind(DeltaI, cbind(DeltaIhat_l, DeltaIhat_r, DeltaIhat_s))
@@ -94,3 +74,13 @@ for (s in 1:length(sigma)){
 }
 res <- list(params=params[-1,], rbar=rbar[-1,], se_rbar=se_rbar[-1,], DeltaI=DeltaI[-1,], se_DeltaI=se_DeltaI[-1,])
 lapply(res,head,10)
+
+#save
+
+if (dir.exists("./results_numeric")==FALSE){
+  dir.create("./results_numeric")
+}
+
[…]

### 2021-05-05 MASTER.R
for master script

@@ -1,0 +1,9 @@
+#MASTER
+
+source("./makenoise.R") #make standard noise, change M here 
+
+source("./parameters.R") #set parameters sigma, mu1, mu2, delta
+
+source("./parameter_loop.R") #loop through parameters to get results
+
+

### 2021-05-13 paper/SupMat.Rnw
changes to supmat:

@@
 \usepackage{xr}
-\usepackage{authblk}
+%\usepackage{authblk}
 %\usepackage{accents}
@@
 We take, for each $i$, $B_i(t)$ to be $iid$ through time.
-We let $B_i = \exp(b_i)$, where $(b_1, b_2)$ is some distribution such that $b_i \sim N(\mu_i, \sigma^2)$ and such that $cov(b_1, b_2) = \rho$, a fixed quantity we will specify later.
+We let $B_i = \exp(b_i)$, where $(b_1, b_2)$ is some distribution such that $b_i \sim N(\mu_i, \sigma^2)$ and such that $\CovX(b_1, b_2) = \rho$, a fixed quantity we will specify later.
 However, $(b_1, b_2)$ is not necessarily a bivariate normal distribution.
@@
 We created three cases for $(b_1,b_2)$, the case of symmetric tail association and the cases of asymmetrical left- and right-tail association.
-To generate $M$ points $(b_1^{(i)}, b_2^{(i)})$, $i=1,..,M$, for the left tail association case, we first generated M points $(a_1^{(i)},a_2^{(i)})$, $i=1,...,M$, from a bivariate normal distribution $N(\begin{pmatrix} 0\\0\end{pmatrix}, \Sigma)$, where here $\Sigma = \begin{pmatrix} 1&0\\0&1\end{pmatrix}$.
+To generate $M$ points $(b_1^{(i)}, b_2^{(i)})$, $i=1,..,M$, for the left tail association case, we first generated M points $(a_1^{(i)},a_2^{(i)})$, $i=1,...,M$, from a bivariate normal distribution $N([0,0], \Sigma)$, where here $\Sigma = \begin{pmatrix} 1&0\\0&1\end{pmatrix}$.
 Then, for each index $i$, we randomly, with 50\% probability, either (A) let $\tilde b_1^{(i)} = -|a_1^{(i)}|$ and $\tilde b_2^{(i)} = -a|_1^{(i)}|$, or (B) let $\tilde b_1^{(i)} = |a_1^{(i)}|$ and $\tilde b_2^{(i)} = |a_2^{(i)}|$.
@@
 .
-The symmetric noise was then generated once from $N(\begin{pmatrix} 0\\0\end{pmatrix}, \Sigma)$ where $\Sigma = \begin{pmatrix} 1&\rho \\ \rho&1\end{pmatrix}$, denoting the points so generated by $(\tilde b_1^{(i)},\tilde b_2^{(i)})$ for $i=1,...,M$.
+The symmetric noise was then generated once from $N([0,0], \Sigma)$ where $\Sigma = \begin{pmatrix} 1&\rho \\ \rho&1\end{pmatrix}$, denoting the points so generated by $(\tilde b_1^{(i)},\tilde b_2^{(i)})$ for $i=1,...,M$.
 We then set $b_1^{(i)} = \sigma \tilde b_1^{(i)} + \mu_1$ and $b_2^{(i)} = \sigma \tilde b_2^{(i)}+ \mu_2$.
@@
 
-First, for a large ?artigen? $M$, carry out the following steps once:
+First, for a large integer $M$, carry out the following steps once:
 
@@
 \end{align}
-?? should the above have $\ln$?
+
 \noindent so this was computed for each considered parameter set of $\delta, \mu_1, \mu_2, \sigma$, and for each of our three tail association cases.
@@
 Equations \ref{sym_rbar1} - \ref{sym_rbar2sharp} obviously depend only on $\mu_1-\mu_2$, and not independently on $\mu_1$ and $\mu_2$, since only $\mu_1-\mu_2$ appears in these expressions if $\mu_1$ or $\mu_2$ appear at all.
-Equation \ref{sym_rbar1} is $\bar r_1$ in the asymmetric?
-tail association cases.
+Equation \ref{rbar1_E} is $\bar r_1$ in the asymmetric tail association cases.
 But letting $\beta_1 = b_1 + \eta$ and $\beta_2 = b_2 + \eta$ and substituting into \ref{rbar1_E}, we get
@@
 From there it is easy to see that $\beta_1$ and $beta_2$ are distributed in the same wat as our right-tail associated noise with parameter $\mu_1,\mu_2$, and $\sigma$.
-Making the substitutions $b_2 = -\beta_1 + \mu_1 + \mu_2$ and $b_1 = -\beta_2 + \mu_1+\mu_2$ in ref{r1distribution}, we get
+Making the substitutions $b_2 = -\beta_1 + \mu_1 + \mu_2$ and $b_1 = -\beta_2 + \mu_1+\mu_2$ in \ref{r1distribution}, we get
 

### 2021-05-13 paper/SupMat.tex
changes to supmat:

@@
 We take, for each $i$, $B_i(t)$ to be $iid$ through time.
-We let $B_i = exp(b_i)$, where $(b_1, b_2)$ is some distribution such that $b_i \sim N(\mu_i, \sigma^2)$ and such that $cov(b_1, b_2) = \rho$, a fixed quantity we will specify later.
+We let $B_i = \exp(b_i)$, where $(b_1, b_2)$ is some distribution such that $b_i \sim N(\mu_i, \sigma^2)$ and such that $\CovX(b_1, b_2) = \rho$, a fixed quantity we will specify later.
 However, $(b_1, b_2)$ is not necessarily a bivariate normal distribution.
@@
 
-Equations \ref{sym_rbar1} - \ref{sym_rbar2sharp} obviously depend only on $\mu_1-\mu_2$, and not independently on $\mu_1$ and $\mu_2$, ?some?
-only $\mu_1-\mu_2$ appears in these expressions if $\mu_1$ or $\mu_2$ appear at all.
-Equation \ref{sym_rbar1} is $\bar r_1$ in the asymmetric?
-tail association cases.
-But letting $\beta_1 = b_1 + \eta$ and $\beta_2 = b_2 + \eta$ and substituting into \ref{sym_rbar1}, we get
+Equations \ref{sym_rbar1} - \ref{sym_rbar2sharp} obviously depend only on $\mu_1-\mu_2$, and not independently on $\mu_1$ and $\mu_2$, since only $\mu_1-\mu_2$ appears in these expressions if $\mu_1$ or $\mu_2$ appear at all.
+Equation \ref{rbar1_E} is $\bar r_1$ in the asymmetric tail association cases.
+But letting $\beta_1 = b_1 + \eta$ and $\beta_2 = b_2 + \eta$ and substituting into \ref{rbar1_E}, we get
 

### 2021-05-14 paper/SupMat.Rnw
changing paranthesis

@@
 We take, for each $i$, $B_i(t)$ to be $iid$ through time. 
-We let $B_i = \exp(b_i)$, where $(b_1, b_2)$ is some distribution such that $b_i \sim N(\mu_i, \sigma^2)$ and such that $\CovX(b_1, b_2) = \rho$, a fixed quantity we will specify later.
-However, $(b_1, b_2)$ is not necessarily a bivariate normal distribution.
-We will consider various distributions of $(b_1, b_2)$ with these properties, corresponding to our symmetric and asymmetric tail associations cases defined in the main text and elaborated below.
+We let $B_i = \exp(b_i)$, where $\left(b_1, b_2\right)$ is some distribution such that $b_i \sim N\left(\mu_i, \sigma^2\right)$ and such that $\CovX(b_1, b_2) = \rho$, a fixed quantity we will specify later.
+However, $\left(b_1, b_2\right)$ is not necessarily a bivariate normal distribution.
+We will consider various distributions of $\left(b_1, b_2\right)$ with these properties, corresponding to our symmetric and asymmetric tail associations cases defined in the main text and elaborated below.
 We assume without loss of generality that $\mu_1 \leq \mu_2$, so that species 1 is the weaker competitor.
@@
 We created three cases for $(b_1,b_2)$, the case of symmetric tail association and the cases of asymmetrical left- and right-tail association. 
-To generate $M$ points $(b_1^{(i)}, b_2^{(i)})$, $i=1,..,M$, for the left tail association case, we first generated M points $(a_1^{(i)},a_2^{(i)})$, $i=1,...,M$, from a bivariate normal distribution $N([0,0], \Sigma)$, where here $\Sigma = \begin{pmatrix} 1&0\\0&1\end{pmatrix}$.
+To generate $M$ points $\left(b_1^{(i)}, b_2^{(i)}\right)$, $i=1,..,M$, for the left tail association case, we first generated M points $(a_1^{(i)},a_2^{(i)})$, $i=1,...,M$, from a bivariate normal distribution $N\left( \vec{0}, \Sigma \right)$, where here $\Sigma = \begin{pmatrix} 1&0\\0&1\end{pmatrix}$.
 Then, for each index $i$, we randomly, with 50\% probability, either (A) let $\tilde b_1^{(i)} = -|a_1^{(i)}|$ and $\tilde b_2^{(i)} = -a|_1^{(i)}|$, or (B) let $\tilde b_1^{(i)} = |a_1^{(i)}|$ and $\tilde b_2^{(i)} = |a_2^{(i)}|$.
 We then let $b_1^{(i)} = \sigma \tilde b_1^{(i)} + \mu_1$ and $b_2^{(i)} = \sigma \tilde b_2^{(i)} + \mu_2$. 
-Perfect association in the left tails of the resulting distribution $(b_1,b_2)$ results from the fact that, in case (A) above, both $\tilde b_1^{(i)}$ and $\tilde b_2^{(i)}$ were assigned to the same value.
-For speed of computation, when samples $(\tilde b_1^{(i)},\tilde b_2^{(i)})$, $i=1,...,M$ were needed for every distinct set of values of the parameters $\mu_1$, $\mu_2$, and $\sigma^2$, we used the same $(\tilde b_1^{(i)},\tilde b_2^{(i)})$, $i=1,...,M$, simply transforming these values differently (using the different valus of $\mu_1$, $\mu_2$, $\sigma$) for each set of parameter values.
+Perfect association in the left tails of the resulting distribution $\left(b_1,b_2\right)$ results from the fact that, in case (A) above, both $\tilde b_1^{(i)}$ and $\tilde b_2^{(i)}$ were assigned to the same value.
+For speed of computation, when samples $(\tilde b_1^{(i)},\tilde b_2^{(i)})$, $i=1,...,M$ were needed for every distinct set of values of the parameters $\mu_1$, $\mu_2$, and $\sigma^2$, we used the same $\left(\tilde b_1^{(i)},\tilde b_2^{(i)}\right)$, $i=1,...,M$, simply transforming these values differently (using the different valus of $\mu_1$, $\mu_2$, $\sigma$) for each set of parameter values.
 Right-tail associated noise was generated in an analogous way.
@@
 Noise with symmetric tail association was generated by taking a very large sample (100000) of left-tail associated noise, as described above, and then calculating the Pearson correaltion coefficient, rho.
-The value obtained was ...
-.
-The symmetric noise was then generated once from $N([0,0], \Sigma)$ where $\Sigma = \begin{pmatrix} 1&\rho \\ \rho&1\end{pmatrix}$, denoting the points so generated by $(\tilde b_1^{(i)},\tilde b_2^{(i)})$ for $i=1,...,M$.
+The value obtained was ... . 
+The symmetric noise was then generated once from $N\left(\vec{0}, \Sigma \right)$ where $\Sigma = \begin{pmatrix} 1&\rho \\ \rho&1\end{pmatrix}$, denoting the points so generated by $\left(\tilde b_1^{(i)},\tilde b_2^{(i)}\right)$ for $i=1,...,M$.
 We then set $b_1^{(i)} = \sigma \tilde b_1^{(i)} + \mu_1$ and $b_2^{(i)} = \sigma \tilde b_2^{(i)}+ \mu_2$.
@@
 \begin{equation}
-\bar r_1^\#=\EX(\ln [(1-\delta)+\delta \exp(b_1^\#-b_2)]),  \label{rbar1sharp_E}
+\bar r_1^\#=\EX(\ln [(1-\delta)+\delta \exp (b_1^\#-b_2)]),  \label{rbar1sharp_E}
 \end{equation}
@@
 \begin{equation}
-r_2=(\ln (\frac{N_2(t+1)}{N_2(t)}))=0, \label{r2_is_zero}
+\EX(\ln (\frac{N_2(t+1)}{N_2(t)}))=0, \label{r2_is_zero}
 \end{equation}
 
-\noindent as expected. Again, following REF, we also define
+\noindent as expected. 
+Again, following REF, we also define
 
@@
 \begin{equation}
-\Delta I=(\bar r_1 - \bar r_1^\#)-(\bar r_2- \bar r_2^\#)=\bar r_1 - \bar r_1^\#+\bar r_2^\#  \label{DeltaI}
+\Delta I=\left(\bar r_1 - \bar r_1^\#\right)-\left(\bar r_2- \bar r_2^\#\right)=\bar r_1 - \bar r_1^\#+\bar r_2^\#  \label{DeltaI}
 \end{equation}
 
-\noindent has been defined in terms of expected values of elementary expressions of the random variables $(b_1,b_2)$, $(b_1^\#,b_2)$, and $(b_2^\#, b_2)$.
+\noindent has been defined in terms of expected values of elementary expressions of the random variables $\left(b_1,b_2\right)$, $\left(b_1^\#,b_2\right)$, and $\left(b_2^\#, b_2\right)$.
 
-Having already specified that $b_1$ (and therefore also $b_1^\#$) is distributed as $N(\mu_1, \sigma^2)$, and that $b_2$ (and therefore also $b_2^\#$) is distributed as $N(\mu_2, \sigma^2)$, we know, using $\Sigma$ as notation for a covariance matrix that
+Having already specified that $b_1$ (and therefore also $b_1^\#$) is distributed as $N\left(\mu_1, \sigma^2\right)$, and that $b_2$ (and therefore also $b_2^\#$) is distributed as $N\left(\mu_2, \sigma^2\right)$, we know, using $\Sigma$ as notation for a covariance matrix that
 
 \begin{equation}
-(b_1^\#, b_2) \sim N(\begin{pmatrix} \mu_1\\\mu_2\end{pmatrix}, \Sigma) \label{b1sharpb2_distribution}
+\left(b_1^\#, b_2\right) \sim N\left((\mu_1,\mu_2),\Sigma \right) \label{b1sharpb2_distribution}
 \end{equation}
@@
 \begin{equation}
-(b_2^\#, b_2) \sim N(\begin{pmatrix} \mu_2\\\mu_2\end{pmatrix}, \Sigma) \label{b2sharpb2_distribution}
+\left(b_2^\#, b_2\right) \sim N(( \mu_2,\mu_2), \Sigma) \label{b2sharpb2_distribution}
 \end{equation}
@@
 \begin{equation}
-(b_1,b_2) \sim  N(\begin{pmatrix} \mu_1\\\mu_2\end{pmatrix}, \Sigma) \label{b1b2_distribution_sym}
+(b_1,b_2) \sim  N(( \mu_1,\mu_2), \Sigma) \label{b1b2_distribution_sym}
 \end{equation}
@@
 \begin{equation}
-\Sigma = \begin{pmatrix} \sigma^2&\rho\\\rho&\sigma^2\end{pmatrix}. \label{Sigmamat_rho}
+\Sigma = \begin{pmatrix} \sigma^2&\rho\\\rho&\sigma^2\end{pmatrix} \label{Sigmamat_rho}
 \end{equation}
@@
 \begin{enumerate}
-\item Generate left-tail assocated noise $(\tilde b_{l1}^{(i)}, \tilde b_{l2}^{(i)})$, $i=1,...,M$, as described in the section \ref{sect:noise}.
+\item Generate left-tail assocated noise $\left(\tilde b_{l1}^{(i)}, \tilde b_{l2}^{(i)}\right)$, $i=1,...,M$, as described in the section \ref{sect:noise}.
 Hence $\tilde b_{l1}^{i)}$ and $\tilde b_{l2}^{(i)}$ are distributed as $N(0,1)$.
-\item Generate right-tail associated noise $(\tilde b_{r1}^{(i)}, \tilde b_{r2}^{(i)})$, $i=1,...,M$, via the analogous procedure.
+\item Generate right-tail associated noise $\left(\tilde b_{r1}^{(i)}, \tilde b_{r2}^{(i)}\right)$, $i=1,...,M$, via the analogous procedure.
 Again, $\tilde b_{r1}^{i)}$ and $\tilde b_{r2}^{(i)}$ are distributed as $N(0,1)$.
@@
 
-\item $\hat {\Delta I} = \hat \bar r_1 - \hat \bar r_1^{\#} + \hat \bar r_2^{\#}$, estimated ?syparatilly?
-for the left-tail associated, right-tail associated and symmetric noise cases.
+\item $\hat {\Delta I} = \hat \bar r_1 - \hat \bar r_1^{\#} + \hat \bar r_2^{\#}$, estimated separately for the left-tail associated, right-tail associated and symmetric noise cases.
 We obtained conservative standard error estimates by adding the standard error of the estimates of $\hat \bar r_1, \hat \bar r_1^{\#}$, and $\hat \bar r_2^{\#}$.
@@
 
-Not only does the expected value $\bar r_1$ influence the capacity for species 1 to invade where rare, but so do other aspectf of the distribution of
+Not only does the expected value $\bar r_1$ influence the capacity for species 1 to invade where rare, but so do other aspects of the distribution of
 \begin{equation}
@@
 \begin{align}
-\Prob[r_1>0] &= \Prob [1-\delta+\delta \exp(b_1-b_2)] > 1] \\
+\Prob[r_1>0] &= \Prob [1-\delta+\delta \exp(b_1-b_2) > 1] \\
 &= \Prob[\exp (b_1-b_2)>1] \\

### 2021-05-14 paper/SupMat.Rnw
last sentence paranthesize ref

@@
 
-\noindent And since $(\beta_1, \beta_2)$ is distributed in the same way as our right-tail associated noise, this proves all metrics based on the distribution of $r_1$ \ref{r1distribution} are the same for left- and right-tail associated noise.
+\noindent And since $(\beta_1, \beta_2)$ is distributed in the same way as our right-tail associated noise, this proves all metrics based on the distribution of $r_1$ (\ref{r1distribution}) are the same for left- and right-tail associated noise.
 

### 2021-05-14 paper/SupMat.tex
added outline of new stuff

@@
 \begin{equation}
-r_1=\ln[1-\delta+\delta \exp(b_1-b_2)] \label{r1distribution}
+r_1=\ln[1-\delta+\delta \exp(b_1-b_2)]. \label{r1distribution}
 \end{equation}
@@
 
-\noindent And since $(\beta_1, \beta_2)$ is distributed in the same way as our right-tail associated noise, this proves all metrics based on the distribution of $r_1$ \ref{r1distribution} are the same for left- and right-tail associated noise.
+\noindent And since $(\beta_1, \beta_2)$ is distributed in the same way as our right-tail associated noise, this proves all metrics based on the distribution of $r_1$ (\ref{r1distribution}) are the same for left- and right-tail associated noise.
+
+\section{New Stuff}
+
+Fixed number, base growth rate
+
+\begin{equation}
+\epsilon_i^0=\ln(1-\delta + \delta \frac{\bar B_i}{\bar B_j}) \label{null_i}
+\end{equation}
+
+\begin{equation}
+\epsilon_i^E=\ln(1-\delta + \delta \frac{B_i}{\bar B_j}) -  \epsilon_i^0 \label{Evar}
+\end{equation}
+
+contribution of variance in E
+
+\begin{equation}
+\bar \epsilon_i^E=\EX(\epsilon_i^E) \label{Evar_mean}
+\end{equation}
+
+\begin{equation}
+\epsilon_i^C=\ln(1-\delta + \delta \frac{\bar B_i}{B_j}) -  \epsilon_i^0 \label{Cvar}
+\end{equation}
+
+contribution of variance in C
+
+\begin{equation}
+\bar \epsilon_i^C=\EX(\epsilon_i^C) \label{Cvar_mean}
+\end{equation}
+
+\begin{equation}
+\epsilon_i^{EC}=\ln(1-\delta + \delta \frac{B_i}{B_j}) -  [\epsilon_i^0 + \epsilon_i^E + \epsilon_i^C] \label{ECvar}
+\end{equation}
+
+interaction effect of both E and C varying
+
+\begin{equation}
+\bar \epsilon_i^{EC}=\EX(\ln(1-\delta + \delta \frac{B_i}{B_j})) -[\epsilon_i^0 +\bar \epsilon_i^E + \bar \epsilon_i^C] \label{ECvar_mean}
+\end{equation}
+
+\begin{equation}
+\bar r_i=\EX(\ln(1-\delta + \delta \frac{B_i}{B_j}))\label{rbari}
+\end{equation}
+
+\begin{equation}
+\bar r_i=\epsilon_i^0 +\bar \epsilon_i^E + \bar \epsilon_i^C + \bar \epsilon_i^{EC} \label{rbari_decomp}
+\end{equation}
+
+But some of $\bar \epsilon_i^{EC}$ comes from covariance between E and C and some comes from both varying.
+
+\begin{equation}
+\bar \epsilon_i^{(E\#C)}=\EX(\ln(1-\delta + \delta \frac{B_i}{B_j^{\#}})) -[\epsilon_i^0 +\bar \epsilon_i^E + \bar \epsilon_i^C] \label{ECvar_sharp}
+\end{equation}
+
+\begin{align}
+\bar \epsilon_i^{(EC)}&=\bar \epsilon_i^{EC}-\bar \epsilon_i^{(E\#C)}\\
+&= \EX(\ln(1-\delta + \delta \frac{B_i}{B_j}))-[\epsilon_i^0 +\bar \epsilon_i^E + \bar \epsilon_i^C]\\
+&- \EX(\ln(1-\delta + \delta \frac{B_i}{B_j^{\#}})) +[\epsilon_i^0 +\bar \epsilon_i^E + \bar \epsilon_i^C]
+\label{r_iSE}
+\end{align}
+
+checks out
 

### 2021-05-31 paper/SupMat.Rnw
added outline of new stuff
adding new decomposition stuff
decomp notes

@@
 
+\section{Decomposition}
+
+To quantify the contribution of different drivers to the difference in growth rate between the invader and the resident, the amount variance due to that driver is considered.
+We consider the contributions of the null growth rate, $\epsilon^0$, the contribution of variance in E (environment), $\bar \epsilon^E$, the contribution of the variance in C (competitive pressure), $\bar \epsilon^C$, the contrubition of E and C co-varying $\bar \epsilon^{(EC)}$, and the the contribtion of E and C varying, but independently, $\bar \epsilon^{(E\#C)}$, for the invader and the resident.
+The difference between invader and resident is computed term by term to get a full decomposition as follows,
+
+\begin{equation}
+\bar r_1-\bar r_2 = \Delta^0+\Delta^E+\Delta^C+\Delta^{(EC)}+\Delta^{(E\#C)}, \label(IGR)
+\end{equation}
+
+\noindent where each $\Delta$ denotes a drivers contribution the the invader growth rate subtracted by the drivers contribution to the resident's groeth rate.
+
+The null growth rate is computed as the growrth rate when both drivers (E and C) are set at their mean. Recalling the growth rate eqaution from equation \ref{rbar1_E},
+
+\begin{align}
+\Delta^0 &= \epsilon_1^0 - \epsilon_2^0\\
+&=\ln(1-\delta + \delta \frac{\bar B_1}{\bar B_2}) - \ln(1-\delta + \delta \frac{\bar B_2}{\bar B_2})\\
+&=\ln(1-\delta + \delta \frac{\bar B_1}{\bar B_2}) - 0. \\
+%&=\ln(1-\delta + \delta \exp(\bar b_1-\bar b_2)
+\end{align}
+
+The contribution of variance in E to a species growth rate is computed by letting E vary and setting C to its mean, subtracted by the null.
+The contribution of variance in C to a species' growth rate is computed in an analogous way, but reverse.
+
+\begin{align}
+\Delta^E &= \bar \epsilon_1^E - \bar \epsilon_2^E\\
+&=[\EX (\ln[1-\delta + \delta \frac{B_1}{\bar B_2}]) - \epsilon_1^0] - [\EX(\ln[1-\delta + \delta \frac{ B_2}{\bar B_2}]) - \epsilon_2^0]\\
+&=[\EX (\ln[1-\delta + \delta \frac{B_1}{\bar B_2}]) - \epsilon_1^0] - \EX(\ln[1-\delta + \delta \frac{ B_2}{\bar B_2}]) \\
+\end{align}
+
+\begin{align}
+\Delta^C &= \bar \epsilon_1^C - \bar \epsilon_2^C\\
+&=[\EX (\ln[1-\delta + \delta \frac{\bar B_1}{B_2}]) - \epsilon_1^0] - [\EX(\ln[1-\delta + \delta \frac{\bar B_2}{B_2}]) - \epsilon_2^0]\\
+&=[\EX (\ln[1-\delta + \delta \frac{\bar B_1}{B_2}]) - \epsilon_1^0] - \EX(\ln[1-\delta + \delta \frac{\bar B_2}{B_2}]) \\
+\end{align}
+
+The contribution of E and C both varying but not together is computed by letting both terms vary but decoupling any association by substituting in a new independent varibable for one of the varibales.
+Here we use C.
+The term is subtracted by the previous contributions.
+
+\begin{align}
+\Delta_1^{(E\#C)} &= \bar \epsilon_1^{(E\#C)} - \bar \epsilon_2^{(E\#C)}\\
+&=[\EX(\ln[1-\delta + \delta \frac{B_1}{B_2^{\#}}]) -[\epsilon_1^0 +\bar \epsilon_1^E + \bar \epsilon_1^C]] - [\EX(\ln[1-\delta + \delta \frac{B_2}{B_2^{\#}}]) -[\epsilon_2^0 +\bar \epsilon_2^E + \bar \epsilon_2^C]]\\
+\end{align}
+
+The contribution of E and C co-varying is the storage effect. Computed by,
+
+\begin{equation}
+\Delta^{(EC)} = \bar \epsilon_1^{(EC)} - \bar \epsilon_2^{(EC)}.
+\end{equation}
+
+\noindent Where,
+
+\begin{align}
+\bar \epsilon_i^{(EC)}&=\bar \epsilon_i^{EC}-\bar \epsilon_i^{(E\#C)}\\
+&= \EX(\ln(1-\delta + \delta \frac{B_i}{B_j}))-[\epsilon_i^0 +\bar \epsilon_i^E + \bar \epsilon_i^C]\\
+&- \EX(\ln(1-\delta + \delta \frac{B_i}{B_j^{\#}})) +[\epsilon_i^0 +\bar \epsilon_i^E + \bar \epsilon_i^C].
+\end{align}
+
+By substituting these expansions into \ref(IGR), we get,
+
+Recall from main text that the goal of this research is to investigate the implications of extreme events and thus asymmetric tail associations on coexistence.
+Hence, we compare IGR and mechanism values for each noise type.
+Because only association structure and not means and standard deviation is differ across noise types, different noise types will not result in differences in any of the decomposition terms outside of the storage effect term.
+All other terms do not consider the associative relationship between the two variables.
+The null growth grate is a constant that depends on the mean of both variables, the contribution of variance in one variable only depends on that one variable scaled by the mean of the other, and the varying but independent term while it does consider both variable they are set to be completely uncorrelated.
+These four terms will not differ across noise types.
+
+Fixed number, base growth rate
+
+\begin{equation}
+\epsilon_i^0=\ln(1-\delta + \delta \frac{\bar B_i}{\bar B_j}) \label{null_i}
+\end{equation}
+
+\begin{equation}
+\epsilon_i^E=\ln(1-\delta + \delta \frac{B_i}{\bar B_j}) -  \epsilon_i^0 \label{Evar}
+\end{equation}
+
+contribution of variance in E
+
+\begin{equation}
+\bar \epsilon_i^E=\EX(\epsilon_i^E) \label{Evar_mean}
+\end{equation}
+
+\begin{equation}
+\epsilon_i^C=\ln(1-\delta + \delta \frac{\bar B_i}{B_j}) -  \epsilon_i^0 \label{Cvar}
+\end{equation}
+
+contribution of variance in C
+
+\begin{equation}
+\bar \epsilon_i^C=\EX(\epsilon_i^C) \label{Cvar_mean}
+\end{equation}
+
+\begin{equation}
+\epsilon_i^{EC}=\ln(1-\delta + \delta \frac{B_i}{B_j}) -  [\epsilon_i^0 + \epsilon_i^E + \epsilon_i^C] \label{ECvar}
+\end{equation}
+
+interaction effect of both E and C varying
+
+\begin{equation}
+\bar \epsilon_i^{EC}=\EX(\ln(1-\delta + \delta \frac{B_i}{B_j})) -[\epsilon_i^0 +\bar \epsilon_i^E + \bar \epsilon_i^C] \label{ECvar_mean}
+\end{equation}
+
+\begin{equation}
+\bar r_i=\EX(\ln(1-\delta + \delta \frac{B_i}{B_j}))\label{rbari}
+\end{equation}
+
+\begin{equation}
+\bar r_i=\epsilon_i^0 +\bar \epsilon_i^E + \bar \epsilon_i^C + \bar \epsilon_i^{EC} \label{rbari_decomp}
+\end{equation}
+
+But some of $\bar \epsilon_i^{EC}$ comes from covariance between E and C and some comes from both varying.
+
+\begin{equation}
+\bar \epsilon_i^{(E\#C)}=\EX(\ln(1-\delta + \delta \frac{B_i}{B_j^{\#}})) -[\epsilon_i^0 +\bar \epsilon_i^E + \bar \epsilon_i^C] \label{ECvar_sharp}
+\end{equation}
+
+\begin{align}
+\bar \epsilon_i^{(EC)}&=\bar \epsilon_i^{EC}-\bar \epsilon_i^{(E\#C)}\\
+&= \EX(\ln(1-\delta + \delta \frac{B_i}{B_j}))-[\epsilon_i^0 +\bar \epsilon_i^E + \bar \epsilon_i^C]\\
+&- \EX(\ln(1-\delta + \delta \frac{B_i}{B_j^{\#}})) +[\epsilon_i^0 +\bar \epsilon_i^E + \bar \epsilon_i^C]
+\label{r_iSE}
+\end{align}
+
+checks out
+
+Need to find $\Delta_1^0, \Delta_1^E, \Delta_1^C, \Delta_1^{(EC)}, \Delta_1^{(E\#C)}$ (where species 1 is the invader and species 2 is the resident)
+
 \end{document}

### 2021-05-31 paper/SupMat.tex
adding new decomposition stuff
decomp notes

@@
 
-\section{New Stuff}
+\section{Decomposition}
+
+To quantify the contribution of different drivers to the difference in growth rate between the invader and the resident, the amount variance due to that driver is considered.
+We consider the contributions of the null growth rate, $\epsilon^0$, the contribution of variance in E (environment), $\bar \epsilon^E$, the contribution of the variance in C (competitive pressure), $\bar \epsilon^C$, the contrubition of E and C co-varying $\bar \epsilon^{(EC)}$, and the the contribtion of E and C varying, but independently, $\bar \epsilon^{(E\#C)}$, for the invader and the resident.
+The difference between invader and resident is computed term by term to get a full decomposition as follows,
+
+\begin{equation}
+\bar r_1-\bar r_2 = \Delta^0+\Delta^E+\Delta^C+\Delta^{(EC)}+\Delta^{(E\#C)}, \label(IGR)
+\end{equation}
+
+\noindent where each $\Delta$ denotes a drivers contribution the the invader growth rate subtracted by the drivers contribution to the resident's groeth rate.
+
+The null growth rate is computed as the growrth rate when both drivers (E and C) are set at their mean. Recalling the growth rate eqaution from equation \ref{rbar1_E},
+
+\begin{align}
+\Delta^0 &= \epsilon_1^0 - \epsilon_2^0\\
+&=\ln(1-\delta + \delta \frac{\bar B_1}{\bar B_2}) - \ln(1-\delta + \delta \frac{\bar B_2}{\bar B_2})\\
+&=\ln(1-\delta + \delta \frac{\bar B_1}{\bar B_2}) - 0. \\
+%&=\ln(1-\delta + \delta \exp(\bar b_1-\bar b_2)
+\end{align}
+
+The contribution of variance in E to a species growth rate is computed by letting E vary and setting C to its mean, subtracted by the null.
+The contribution of variance in C to a species' growth rate is computed in an analogous way, but reverse.
+
+\begin{align}
+\Delta^E &= \bar \epsilon_1^E - \bar \epsilon_2^E\\
+&=[\EX (\ln[1-\delta + \delta \frac{B_1}{\bar B_2}]) - \epsilon_1^0] - [\EX(\ln[1-\delta + \delta \frac{ B_2}{\bar B_2}]) - \epsilon_2^0]\\
+&=[\EX (\ln[1-\delta + \delta \frac{B_1}{\bar B_2}]) - \epsilon_1^0] - \EX(\ln[1-\delta + \delta \frac{ B_2}{\bar B_2}]) \\
+\end{align}
+
+\begin{align}
+\Delta^C &= \bar \epsilon_1^C - \bar \epsilon_2^C\\
+&=[\EX (\ln[1-\delta + \delta \frac{\bar B_1}{B_2}]) - \epsilon_1^0] - [\EX(\ln[1-\delta + \delta \frac{\bar B_2}{B_2}]) - \epsilon_2^0]\\
+&=[\EX (\ln[1-\delta + \delta \frac{\bar B_1}{B_2}]) - \epsilon_1^0] - \EX(\ln[1-\delta + \delta \frac{\bar B_2}{B_2}]) \\
+\end{align}
+
+The contribution of E and C both varying but not together is computed by letting both terms vary but decoupling any association by substituting in a new independent varibable for one of the varibales.
+Here we use C.
+The term is subtracted by the previous contributions.
+
+\begin{align}
+\Delta_1^{(E\#C)} &= \bar \epsilon_1^{(E\#C)} - \bar \epsilon_2^{(E\#C)}\\
+&=[\EX(\ln[1-\delta + \delta \frac{B_1}{B_2^{\#}}]) -[\epsilon_1^0 +\bar \epsilon_1^E + \bar \epsilon_1^C]] - [\EX(\ln[1-\delta + \delta \frac{B_2}{B_2^{\#}}]) -[\epsilon_2^0 +\bar \epsilon_2^E + \bar \epsilon_2^C]]\\
+\end{align}
+
+The contribution of E and C co-varying is the storage effect. Computed by,
+
+\begin{equation}
+\Delta^{(EC)} = \bar \epsilon_1^{(EC)} - \bar \epsilon_2^{(EC)}.
+\end{equation}
+
+\noindent Where,
+
+\begin{align}
+\bar \epsilon_i^{(EC)}&=\bar \epsilon_i^{EC}-\bar \epsilon_i^{(E\#C)}\\
+&= \EX(\ln(1-\delta + \delta \frac{B_i}{B_j}))-[\epsilon_i^0 +\bar \epsilon_i^E + \bar \epsilon_i^C]\\
+&- \EX(\ln(1-\delta + \delta \frac{B_i}{B_j^{\#}})) +[\epsilon_i^0 +\bar \epsilon_i^E + \bar \epsilon_i^C].
+\end{align}
+
+By substituting these expansions into \ref(IGR), we get,
+
+Recall from main text that the goal of this research is to investigate the implications of extreme events and thus asymmetric tail associations on coexistence.
+Hence, we compare IGR and mechanism values for each noise type.
+Because only association structure and not means and standard deviation is differ across noise types, different noise types will not result in differences in any of the decomposition terms outside of the storage effect term.
+All other terms do not consider the associative relationship between the two variables.
+The null growth grate is a constant that depends on the mean of both variables, the contribution of variance in one variable only depends on that one variable scaled by the mean of the other, and the varying but independent term while it does consider both variable they are set to be completely uncorrelated.
+These four terms will not differ across noise types.
 
@@
 
+Need to find $\Delta_1^0, \Delta_1^E, \Delta_1^C, \Delta_1^{(EC)}, \Delta_1^{(E\#C)}$ (where species 1 is the invader and species 2 is the resident)
+
 \end{document}

### 2021-06-05 paper/SupMat.Rnw
adding new decomp

@@
 \author{Daniel C. Reuman}
-\author{Tom Bell}
-\author{Jonathan Walter}
-\author{Kyle Cavanaugh}
-\author{Max Castorani}
 
@@
 
+\section{E and C definitions}
+
+The lottery model is given by
+
+\begin{equation}
+N_i(t+1)=(1-\delta)N(t)+\delta N \frac{B_i(t)N_i(t)}{B_1(t)N_1(t)+B_2(t)N_2(t)}.  
+\end{equation}
+
+REF also defines $E_i = B_i$ and $C_1=C_2=\frac{B_1 N_1 + B_2 N_2}{\delta N}$.
+Then in the invader-resident scenario which defines the invader to be essentially zero and $N_2 = N$, $C_1 = C_2 = \frac{B_2}{\delta}$.
+The growth rate equation relates the two variables of $E$ and $C$ by
+
+\begin{equation}
+\bar r_i=\EX(\ln(1-\delta + \frac{E_i}{C_i})).\label{rbarEC}
+\end{equation}
+
+\noindent By substituting the definitions,
+
+\begin{align}
+\bar r_i &=\EX(\ln(1-\delta + \frac{B_i}{B_2/ \delta})) \\
+&= \EX(\ln(1-\delta +\delta \frac{B_i}{B_2})) 
+\end{align}
+
+The null growth rate of species $i$ is computed by setting both E and C at their means.
+
+\begin{align}
+\epsilon_i^0 &=\ln(1-\delta + \frac{\bar E_i}{\bar C_i}) \\
+&=\ln(1-\delta + \delta \frac{\bar B_i}{\bar B_2}) 
+\end{align}
+
+The contribution of variance E in to the IGR is computed by setting C to its mean and letting E vary subtracted by the null.
+
+\begin{align}
+\epsilon_i^E &=\ln(1-\delta + \frac{E_i}{\bar C_i}) -  \epsilon_i^0 \\
+&=\ln(1-\delta + \delta \frac{B_i}{\bar B_2}) -  \epsilon_i^0 
+\end{align}
+
+\begin{equation}
+\bar \epsilon_i^E=\EX(\epsilon_i^E) \label{Evar_mean}
+\end{equation}
+
+Similarly, the contribution of variance C in to the IGR is computed by setting E to its mean and letting C vary subtracted by the null.
+
+\begin{align}
+\epsilon_i^C &=\ln(1-\delta + \frac{\bar E_i}{C_i}) -  \epsilon_i^0 \\
+ &=\ln(1-\delta + \delta \frac{\bar B_i}{ B_2}) -  \epsilon_i^0 
+\end{align}
+
+\begin{equation}
+\bar \epsilon_i^C=\EX(\epsilon_i^C) \label{Cvar_mean}
+\end{equation}
+
+The contribution of both E and C varying to the IGR is computed by letting both variables vary subtracted by the null and the contributions of only one variable varying.
+
+\begin{equation}
+\epsilon_i^{EC}=\ln(1-\delta + \delta \frac{B_i}{B_j}) -  [\epsilon_i^0 + \epsilon_i^E + \epsilon_i^C] \label{ECvar}
+\end{equation}
+
+\begin{equation}
+\bar \epsilon_i^{EC}=\EX(\ln(1-\delta + \delta \frac{B_i}{B_j})) -[\epsilon_i^0 +\bar \epsilon_i^E + \bar \epsilon_i^C] \label{ECvar_mean}
+\end{equation}
+
+So,
+
+\begin{equation}
+\bar r_i=\epsilon_i^0 +\bar \epsilon_i^E + \bar \epsilon_i^C + \bar \epsilon_i^{EC}. \label{rbari_decomp}
+\end{equation}
+
+But some of $\bar \epsilon_i^{EC}$ comes from covariance between E and C and some comes from both varying independently.
+
+\begin{align}
+\bar \epsilon_i^{(E\#C)} &=\EX(\ln(1-\delta + \frac{E_i}{C_i^{\#}})) -[\epsilon_i^0 +\bar \epsilon_i^E + \bar \epsilon_i^C] \\
+ &=\EX(\ln(1-\delta + \delta \frac{B_i}{B_2^{\#}})) -[\epsilon_i^0 +\bar \epsilon_i^E + \bar \epsilon_i^C] 
+\end{align}
+
+\begin{align}
+\bar \epsilon_i^{(EC)} &=\bar \epsilon_i^{EC}-\bar \epsilon_i^{(E\#C)}\\
+&= \EX(\ln(1-\delta + \frac{E_i}{C_i}))-[\epsilon_i^0 +\bar \epsilon_i^E + \bar \epsilon_i^C]\\
+&- [\EX(\ln(1-\delta + \frac{E_i}{C_i^{\#}})) -[\epsilon_i^0 +\bar \epsilon_i^E + \bar \epsilon_i^C]] \\
+&= \EX(\ln(1-\delta + \delta \frac{B_i}{B_2}))-\epsilon_i^0 -\bar \epsilon_i^E - \bar \epsilon_i^C\\
+&- \EX(\ln(1-\delta + \delta \frac{B_i}{B_2^{\#}})) +\epsilon_i^0 +\bar \epsilon_i^E + \bar \epsilon_i^C \\
+&= \EX(\ln(1-\delta + \delta \frac{B_i}{B_2}))-\EX(\ln(1-\delta + \delta \frac{B_i}{B_2^{\#}}))\\
+\end{align}
+
 \section{Decomposition}
@@
 \begin{equation}
-\bar r_1-\bar r_2 = \Delta^0+\Delta^E+\Delta^C+\Delta^{(EC)}+\Delta^{(E\#C)}, \label(IGR)
+\bar r_1-\bar r_2 = \Delta^0+\Delta^E+\Delta^C+\Delta^{(EC)}+\Delta^{(E\#C)}, \label{IGR}
 \end{equation}
@@
 &=[\EX (\ln[1-\delta + \delta \frac{B_1}{\bar B_2}]) - \epsilon_1^0] - [\EX(\ln[1-\delta + \delta \frac{ B_2}{\bar B_2}]) - \epsilon_2^0]\\
-&=[\EX (\ln[1-\delta + \delta \frac{B_1}{\bar B_2}]) - \epsilon_1^0] - \EX(\ln[1-\delta + \delta \frac{ B_2}{\bar B_2}]) \\
+&=\EX (\ln[1-\delta + \delta \frac{B_1}{\bar B_2}]) - \epsilon_1^0 - \EX(\ln[1-\delta + \delta \frac{ B_2}{\bar B_2}]) \\
 \end{align}
@@
 &=[\EX (\ln[1-\delta + \delta \frac{\bar B_1}{B_2}]) - \epsilon_1^0] - [\EX(\ln[1-\delta + \delta \frac{\bar B_2}{B_2}]) - \epsilon_2^0]\\
-&=[\EX (\ln[1-\delta + \delta \frac{\bar B_1}{B_2}]) - \epsilon_1^0] - \EX(\ln[1-\delta + \delta \frac{\bar B_2}{B_2}]) \\
+&=\EX (\ln[1-\delta + \delta \frac{\bar B_1}{B_2}]) - \epsilon_1^0 - \EX(\ln[1-\delta + \delta \frac{\bar B_2}{B_2}]) \\
 \end{align}
@@
 
-By substituting these expansions into \ref(IGR), we get,
+By substituting these expansions into \ref{IGR}, we get,
 
@@
 
-Fixed number, base growth rate
+\section{new decomposition term}
+
+A new decomposition term was proposed to quantify the effect of asymmetric associations to coexistence.
+Since asymmetry only differs in the storage effect term, $\bar \epsilon^{(EC)}$, the contribution of asymmetry associations between the species will be a further decomposition.
 
 \begin{equation}
-\epsilon_i^0=\ln(1-\delta + \delta \frac{\bar B_i}{\bar B_j}) \label{null_i}
+\bar \epsilon^{(\dot {EC})} = \bar \epsilon^{(EC)} - \bar \epsilon^{(\ddot {EC})}
 \end{equation}
 
-\begin{equation}
-\epsilon_i^E=\ln(1-\delta + \delta \frac{B_i}{\bar B_j}) -  \epsilon_i^0 \label{Evar}
-\end{equation}
-
-contribution of variance in E
-
-\begin{equation}
-\bar \epsilon_i^E=\EX(\epsilon_i^E) \label{Evar_mean}
-\end{equation}
-
-\begin{equation}
-\epsilon_i^C=\ln(1-\delta + \delta \frac{\bar B_i}{B_j}) -  \epsilon_i^0 \label{Cvar}
-\end{equation}
-
-contribution of variance in C
-
-\begin{equation}
-\bar \epsilon_i^C=\EX(\epsilon_i^C) \label{Cvar_mean}
-\end{equation}
-
-\begin{equation}
-\epsilon_i^{EC}=\ln(1-\delta + \delta \frac{B_i}{B_j}) -  [\epsilon_i^0 + \epsilon_i^E + \epsilon_i^C] \label{ECvar}
-\end{equation}
-
-interaction effect of both E and C varying
-
-\begin{equation}
-\bar \epsilon_i^{EC}=\EX(\ln(1-\delta + \delta \frac{B_i}{B_j})) -[\epsilon_i^0 +\bar \epsilon_i^E + \bar \epsilon_i^C] \label{ECvar_mean}
-\end{equation}
-
-\begin{equation}
-\bar r_i=\EX(\ln(1-\delta + \delta \frac{B_i}{B_j}))\label{rbari}
-\end{equation}
-
-\begin{equation}
-\bar r_i=\epsilon_i^0 +\bar \epsilon_i^E + \bar \epsilon_i^C + \bar \epsilon_i^{EC} \label{rbari_decomp}
-\end{equation}
-
-But some of $\bar \epsilon_i^{EC}$ comes from covariance between E and C and some comes from both varying.
-
-\begin{equation}
-\bar \epsilon_i^{(E\#C)}=\EX(\ln(1-\delta + \delta \frac{B_i}{B_j^{\#}})) -[\epsilon_i^0 +\bar \epsilon_i^E + \bar \epsilon_i^C] \label{ECvar_sharp}
-\end{equation}
-
-\begin{align}
-\bar \epsilon_i^{(EC)}&=\bar \epsilon_i^{EC}-\bar \epsilon_i^{(E\#C)}\\
-&= \EX(\ln(1-\delta + \delta \frac{B_i}{B_j}))-[\epsilon_i^0 +\bar \epsilon_i^E + \bar \epsilon_i^C]\\
-&- \EX(\ln(1-\delta + \delta \frac{B_i}{B_j^{\#}})) +[\epsilon_i^0 +\bar \epsilon_i^E + \bar \epsilon_i^C]
-\label{r_iSE}
-\end{align}
-
-checks out
-
-Need to find $\Delta_1^0, \Delta_1^E, \Delta_1^C, \Delta_1^{(EC)}, \Delta_1^{(E\#C)}$ (where species 1 is the invader and species 2 is the resident)
+The contribution of asymmetry to the storage effect is the observed storage effect subtracted by a null storage effect if the association between species was completely symmetrical. 
+The descrition of $\bar \epsilon^{(\ddot {EC})}$ is given in section 3 and equations 9 and 21-23.
 

### 2021-06-05 decomptestplotting.R
explore decomposition

@@ -1,0 +1,40 @@
+list <- vector(mode='list', length=10)
+
+for (i in 1:10){
+  source("./decomp_test.R")
+  list[[i]] <- res
+  
+}
+list
+
+#DECsharp
+plot(0, pch='', ylim = c(0, 0.01), xlim=c(0,3.5), main="DECsharp")
+for (i in 1:10){
+  points(1:3, list[[i]]$DECsharp, col=rainbow(10)[i])
+}
+
+
+#DEC_
+plot(0, pch='', ylim = c(0.13, 0.17), xlim=c(0,3.5), main="DEC_")
+for (i in 1:10){
+  points(1:3, list[[i]]$DEC_, col=rainbow(10)[i])
+}
+
+#DE
+plot(0, pch='', ylim = c(0.05, 0.1), xlim=c(0,3.5), main="DE")
+for (i in 1:10){
+  points(1:3, list[[i]]$DE, col=rainbow(10)[i])
+}
+
+#DC
+plot(0, pch='', ylim = c(-0.1, -0.07), xlim=c(0,3.5), main="DC")
+for (i in 1:10){
+  points(1:3, list[[i]]$DC, col=rainbow(10)[i])
+}
+
+#Dnull
+plot(0, pch='', ylim = c(-0.17, -0.10), xlim=c(0,3.5), main="Dnull")
+for (i in 1:10){
+  points(1:3, list[[i]]$Dnull, col=rainbow(10)[i])
+}
+

### 2021-06-22 SupMat.Rnw
trivial change to get compilation to work on my machine

@@
 \begin{document}
-\SweaveOpts{concordance=TRUE}
+%\SweaveOpts{concordance=TRUE}
 
@@
 
-\section{Additional quantities influencing invasion success} \label{sect: addquant}
+\section{Additional quantities influencing invasion success} \label{sect:addquant}
+
+Not only does the expected value $\bar r_1$ influence the capacity for species 1 to invade where rare, but so do other aspectf of the distribution of
+\begin{equation}
+r_1 = \ln[1-\delta+\delta \exp(b_1-b_2)]. \label{r1_distribution}
+\end{equation}
+
+We here elaborate how some other aspects of the distribution of $r_1$ were quantified, for our three tail association cases and for any of our parameters $\delta, \mu_1, \mu_2, \sigma$.
+
+First, we considered $\sd(r_1)$, the standard deviation.
+If this is large enough, then $r_1$ can occasionally be positive even if $\bar r_1$ is strongly negative, and this invasion can occur under the right environmental circumstances.
+
+We also considered the probability $\Prob[r_1 >0]$, which is, explicitly, the probability that environmental variables that permit invasion. We have
+\begin{align}
+\Prob[r_1>0] &= \Prob [1-\delta+\delta \exp(b_1-b_2)] > 1] \\
+&= \Prob[\exp (b_1-b_2)>1] \\
+&= \Prob[b_1>b_2],
+\end{align}
+?? should the above have $\ln$?
+\noindent so this was computed for each considered parameter set of $\delta, \mu_1, \mu_2, \sigma$, and for each of our three tail association cases.
+
+We also consider the mean value of the possible part of the distribution of $r_1$, $\EX(r_1|r_1>0)$.
+This characterizes how quickly invasion may happen, when it happens, with larger values corresponding to faster invasion.
+If it takes several time steps for the population of the invader to rise from negligible level at which species 1 is a noticeable presence in the community, then environmental conditions must be suitable for invasion for several sequential time steps.
+Thus large values of $\EX(r_1|r_1>0)$ should also correspond to a greater probability of invasion because they should correspond to faster invasions and therefore to reduced need for sequential periods of suitable environmental conditions.
+
+Finally, for a ?low? of species of? parameters $\delta, \mu_1, \mu_2, \sigma$, we displayed the whole distribution of $r_1$, values for each of our tail association cases, to make visual comparisons.
+
+\section{Symmetries and parameter reduction}
+
+The quantities we consider to address invasion prospects for the weaker competitor, species 1 (see sections \ref{sect:IGR} and \ref{sect:addquant}) depend only on $\mu_1-\mu_2$ and not, independtly, on $\mu_1$ and $\mu_2$, and also are the same for our left- and right-tailed association cases.
+We prove these statements in this section.
+
+Equations \ref{sym_rbar1} - \ref{sym_rbar2sharp} obviously depend only on $\mu_1-\mu_2$, and not independently on $\mu_1$ and $\mu_2$, ?some?
+only $\mu_1-\mu_2$ appears in these expressions if $\mu_1$ or $\mu_2$ appear at all.
+Equation \ref{sym_rbar1} is $\bar r_1$ in the asymmetric?
+tail association cases.
+But letting $\beta_1 = b_1 + \eta$ and $\beta_2 = b_2 + \eta$ and substituting into \ref{sym_rbar1}, we get
+
+\begin{align}
+\bar r_1 &= \EX(\ln[1-\delta+\delta \exp(\beta_1-\eta-(\beta_2-\eta))])\\
+&= \EX(\ln[1-\delta+\delta \exp(\beta_1-\beta_2)]).
+\end{align}
+\noindent And so altering the mean of both $b_1$ and $b_2$ distributions by the same amount has no affect of $\bar r_1$ in the asymmetric tail associations.
+Also ?
+depends only on $\mu_1 - \mu_2$ and not independtly on $\mu_1$ or $\mu_2$.
+
+In section \ref{sect:addquant}, we introduce several other metrics. These are all based on the distribution 
+\begin{equation}
+r_1=\ln[1-\delta+\delta \exp(b_1-b_2)] \label{r1distribution}
+\end{equation}
+
+\noindent But, again making the substitutions $\beta_1 = b_1 + \eta$ and $\beta_2 = b_2 + \eta$, we can see via the same resoning as above that the whole distribution $r_1$ depends only on $\mu_1 - \mu_2$, and not on $\mu_1$ and $\mu_2$ independently, for all three of our tail association cases.
+
+To see that our various metrics are the same for out left- and right-tail associated noise ?parameters?, let $(b_1,b_2)$ denote our left-tail associated random variable with parameters $\mu_1,\mu_2,\sigma$.
+Then define $\beta_1 = -b_2 + \mu_1 +\mu_2$, $\beta_2=-b_1+\mu_1+\mu_2$.
+It is easy to see that $\EX(\beta_1)=\mu_1$, $\EX(\beta_2)=\mu_2$, $\sd(\beta_1)=\sd(\beta_2)=\sigma$, so $\beta_1$ and $\beta_2$ are normally distributed.
+From there it is easy to see that $\beta_1$ and $beta_2$ are distributed in the same wat as our right-tail associated noise with parameter $\mu_1,\mu_2$, and $\sigma$.
+Making the substitutions $b_2 = -\beta_1 + \mu_1 + \mu_2$ and $b_1 = -\beta_2 + \mu_1+\mu_2$ in ref{r1distribution}, we get
+
+\begin{align}
+r_1 &= \ln[1-\delta+\delta \exp(-\beta_2+\mu_1+\mu_2-(\beta_1+\mu_1+\mu_2))]\\
+&= \ln[1-\delta+\delta \exp(\beta_1-\beta_2)].
+\end{align}
+
+\noindent And since $(\beta_1, \beta_2)$ is distributed in the same way as our right-tail associated noise, this proves all metrics based on the distribution of $r_1$ \ref{r1distribution} are the same for left- and right-tail associated noise.
 

### 2021-06-22 paper/Paper.tex

--- 2021-04-18 Paper.Rnw
@@
 
+\usepackage{Sweave}
 \begin{document}
-\SweaveOpts{concordance=TRUE}
+\input{Paper-concordance}
 
 %The following is where you load in the numeric results that will be embedded in the text
-<<eval=T,echo=F,message=F,warning=F>>=
-#system(paste("biber", sub("\\.Rnw$", "", current_input())))
-
-#place R code here for loading in necessary variables from Results
-#ordinarily this would be a readRDS command, but I'm just defining a variable here for now
-allregres<-matrix(rnorm(10),2,5)
-@
 
@@
 You can include the value of previously loaded R variables in the text, e.g., the value of allregres[1,5] is
-\Sexpr{allregres[1,5]}, which can also be given rounded: \Sexpr{round(allregres[1,5],3)}. You can do simple R
-on the fly, e.g., allregres[1,5] plus 1 is \Sexpr{allregres[1,5]+1}. 
+-1.16948109243002, which can also be given rounded: -1.169. You can do simple R
+on the fly, e.g., allregres[1,5] plus 1 is -0.169481092430019. 
 You should never manually type any results, they should always be autolinks like this.
@@
 even & the & values & of \\
-loaded & R & variables & \Sexpr{allregres[1,5]} \\
-you & can & round & \Sexpr{round(allregres[1,5],3)} \\ 
+loaded & R & variables & -1.16948109243002 \\
+you & can & round & -1.169 \\ 
 \hline

### 2021-06-22 paper/Paper.Rnw
new fig
added a couple paragraphs to theory and removed from git some files which should not be under version control

--- 2021-04-18 Paper.Rnw
@@
 
+\DeclareMathOperator{\EX}{E}% expected value
+\DeclareMathOperator{\VarX}{var}
+\DeclareMathOperator{\CovX}{cov}
+\DeclareMathOperator{\mean}{mean}
+\DeclareMathOperator{\se}{se}
+\DeclareMathOperator{\sd}{sd}
+\DeclareMathOperator{\Prob}{P}
+
 %external documents
@@
 %header material for paper
-\title{Place title here}
+\title{Relationships in the extremes and their influence on competition and coexistence}
 \date{}
+
+%Put other possible titles below, commented, and we will select the best one at the end
+%Relationships between extreme events and their influence on competition and coexistence
 
@@
 <<eval=T,echo=F,message=F,warning=F>>=
-#system(paste("biber", sub("\\.Rnw$", "", current_input())))
+system(paste("biber", sub("\\.Rnw$", "", current_input())))
 
@@
 
+\section{Theory}\label{section:theory}
+
+For pedgogical clarity, we develop each step of the theory both for a simple classical model, 
+the lottery model, and in general. The model is as follows. 
+%***DAN: Note the inclusion of the model here means we will have to edit the "Model" section of sup mat, since we don't need to intro the model twice. Do later.
+Letting $N_i(t)$ denote the population density of species $i=1,2$ at time $t$, and defining $N = N_1(t)+N_2(t)$,
+model equations are
+\begin{equation}
+N_i(t+1)=(1-\delta)N_i(t)+\delta N \frac{B_i(t)N_i(t)}{B_1(t)N_1(t)+B_2(t)N_2(t)}  \label{model_eq}
+\end{equation}
+for $i=1,2$. Here, $\delta$ is a mortality rate, and $B_i(t)$ is the fecundiity of species $i$ at time $t$. 
+The model postulates that individuals die at rate $\delta$ at each time step, and are replaced by juveniles 
+in proportion to the reproductive outputs of the two species that year.
+For each $i$, we assume the random variables $B_i(t)$ are independent and identically distributed (iid) through
+time. We let $B_i = \exp(b_i)$, where $\left(b_1, b_2\right)$ is some bivariate random variable with the properties 
+that $b_i$ is normally distributed with mean $\mu_i$ and variance $\sigma^2$. 
+We denote $\CovX(b_1, b_2)$ by $\rho$. However, we do \emph{not} assume that $\left(b_1, b_2\right)$ is a 
+bivariate normal distribution. We will consider various distributions of $\left(b_1, b_2\right)$ with the above 
+properties, corresponding to symmetric and asymmetric tail association cases.
+We assume without loss of generality that $\mu_1 \leq \mu_2$, so that species 1 is the weaker competitor.
+Note that $N$ is constant through time.
+
+Modern coexistence theory \citep{Chesson_2000} and its recent computational extensions \citep{Ellner_2016,Ellner_2019} 
+%***DAN: Search for REF and replace all with appropriate citations
+quantify the contributions of 
+different mechanisms to species coexistence. But we show in Sup Mat section X
+%***DAN: Make this reference to the sup mat an autolink, and write the section
+that, for purely temporal variation, only one mechanism relates to tail associations: storage effects. 
+We here define storage effects abstractly and then for the lottery model, using an approach and notation
+that follows elements of the presentations in \cite{Ellner_2016} and \cite{Ellner_2019} closely (and see
+those papers for additional details and explanation). We 
+return briefly to other mechanisms of coexistence in the Discussion.
+%***DAN: Don't forget to do this. Basically want to say intuitively why tail associations don't influence those,
+%though might be worth mentioning that events are getting more extreme. We have studied how *relationships*
+%in the extremes influence coexistence, but probably worth studying (or cite papers if it's been done) how changes
+%in the commonness and extremeness of extreme events themselves influence coexistence, and for that you'd have
+%to also look at the other mechanisms. 
+It is assumed that the growth rate $r_i(t)$ of species $i$ can be written as an increasing function of
+an environment-dependent factor $E_i(t)$ and as a decreasing function of a competitive pressure, $C_i(t)$.
+For the lottery model, we take $E_i(t)=B_i(t)$, henceforth assuming that fecundity depends strictly on
+the environment. Competition $C_i(t)$ is taken to be
+
 \section{Methods}\label{section:methods}
@@
 \section{Discussion}\label{section:discussion}
-
-\section{Reference code for Jasmin to learn Latex/Sweave}
-
-To get normal text you just type.
-
-To get a paragraph break you leave a blank line, indentation is automatic.
-
-\noindent If you want to remove the indent you use this.
-
-In-line math is always surrounded by dollar signs, $x+1=4$.
-Even when you just have single variables, like if you say ``the notation we use for synchrony is $x$'', you put the $x$ in dollar signs.
-The reason is x and $x$ look different, in a way that a mathematician will notice and wonder if they are meant to be different variables.
-
-For quotes, do it ``like this'' not "like this" or else one of the quotes will be backwards.
-
-Some basic math, inline, by example: $x_{11}$, $y^{22}$, $\frac{1+y}{x+3}$, $\sum_{i=1}^N x_x$,
-$\alpha$, $\beta$.
-
-There is also display math:
-\begin{equation}
-x=y+12-z \label{eq_about_x_and_y}
-\end{equation}
-The ``label'' command allows a textual label for refering to equations later.
-
-The math gets pretty complex. Keep in mind, whatever you want to do, there is a way to do it. We'll save most of that for later.
-
-Now we can refer to equation/expression (\ref{eq_about_x_and_y}). Latex will sort
-out equation numbering and insert the correct number labels for you.
-
-You can also refer to any section in the same way: section \ref{section:introduction}, section \ref{section:methods}.
-
-You can include the value of previously loaded R variables in the text, e.g., the value of allregres[1,5] is
-\Sexpr{allregres[1,5]}, which can also be given rounded: \Sexpr{round(allregres[1,5],3)}. You can do simple R
-on the fly, e.g., allregres[1,5] plus 1 is \Sexpr{allregres[1,5]+1}. 
-You should never manually type any results, they should always be autolinks like this.
-
-%Comments are lines starting with %
-
-You make a percent with \%
-
-You can refer to Fig. \ref{fig:random_fig} by its label and latex will number figures and use the correct numbers.
-The label is defined below where the figure is placed.
-
-You can refer to Table \ref{tab:random_table} in the same way.
-
-%You can also refer to figures and tables and equations from the sup mat, just prefix the labels
-%by SM- in your calls to \ref{}
-
-You can cite like this \cite{Abanetal2006} or this \citep{Abanetal2006} or this \citealp{Abanetal2006}.
-You cal also cite books \citep{BurnhamAnderson2002}, of course.
 
@@
 
-\clearpage
-\newpage
-
-\section{Tables}
-
-%You make tables like this:
-\begin{table}[H]
-\caption[Summary caption, for table of contents, can be skipped.]{Full caption here.}
-\begin{tabular}{llll} %The four ls mean four columns, each left justified. Can use r or c, too.
-\hline 
-Col 1 header & Col 2 header & Col 3 header & Col 4 header \\
-\hline
-text & or & numbers & 3.2 \\
-or & math & $x+1$ & or \\
-even & the & values & of \\
-loaded & R & variables & \Sexpr{allregres[1,5]} \\
-you & can & round & \Sexpr{round(allregres[1,5],3)} \\ 
-\hline
-\end{tabular}
-\label{tab:random_table}
-\end{table}
-
-\clearpage
-\newpage
-
-\section{Figures}
-
-\begin{figure}[!h]
-\includegraphics[width=.5\textwidth]{figures/summaryDocFigures/figure8.png}
-\caption{Type the caption here.} \label{fig:random_fig}
-\end{figure}
-
 \end{document}

### 2021-06-22 paper/Paper.tex
added a couple paragraphs to theory and removed from git some files which should not be under version control

@@
-\documentclass[letterpaper,11pt]{article}
+\documentclass[letterpaper,11pt]{article}\usepackage[]{graphicx}\usepackage[]{color}
+% maxwidth is the original width if it is less than linewidth
+% otherwise use linewidth (to make sure the graphics do not exceed the margin)
+\makeatletter
+\def\maxwidth{ %
+  \ifdim\Gin@nat@width>\linewidth
+    \linewidth
+  \else
+    \Gin@nat@width
+  \fi
+}
+\makeatother
+
+\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
+\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
+\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
+\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
+\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
+\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
+\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
+\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
+\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
+\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%
+\let\hlipl\hlkwb
+
+\usepackage{framed}
+\makeatletter
+\newenvironment{kframe}{%
+ \def\at@end@of@kframe{}%
+ \ifinner\ifhmode%
+  \def\at@end@of@kframe{\end{minipage}}%
+  \begin{minipage}{\columnwidth}%
+ \fi\fi%
+ \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
+ \colorbox{shadecolor}{##1}\hskip-\fboxsep
+     % There is no \\@totalrightmargin, so:
+     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
+ \MakeFramed {\advance\hsize-\width
+   \@totalleftmargin\z@ \linewidth\hsize
+   \@setminipage}}%
+ {\par\unskip\endMakeFramed%
+ \at@end@of@kframe}
+\makeatother
+
+\definecolor{shadecolor}{rgb}{.97, .97, .97}
+\definecolor{messagecolor}{rgb}{0, 0, 0}
+\definecolor{warningcolor}{rgb}{1, 0, 1}
+\definecolor{errorcolor}{rgb}{1, 0, 0}
+\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX
+
+\usepackage{alltt}
 
@@
 %***Need to indicate corresponding authors
-
-\usepackage{Sweave}
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
 \begin{document}
-\input{Paper-concordance}
+%\SweaveOpts{concordance=TRUE}
 
@@
 
+\section{Theory}\label{section:theory}
+
+For pedgogical clarity, we develop each step of the theory both for a simple classical model, 
+the lottery model, and in general. The model is as follows. 
+%***DAN: Note the inclusion of the model here means we will have to edit the "Model" section of sup mat, since we don't need to intro the model twice. Do later.
+Letting $N_i(t)$ denote the population density of species $i=1,2$ at time $t$, and defining $N = N_1(t)+N_2(t)$,
+model equations are
+\begin{equation}
+N_i(t+1)=(1-\delta)N_i(t)+\delta N \frac{B_i(t)N_i(t)}{B_1(t)N_1(t)+B_2(t)N_2(t)}  \label{model_eq}
+\end{equation}
+for $i=1,2$. Here, $\delta$ is a mortality rate, and $B_i(t)$ is the fecundiity of species $i$ at time $t$. 
+The model postulates that individuals die at rate $\delta$ at each time step, and are replaced by juveniles 
+in proportion to the reproductive outputs of the two species that year.
+For each $i$, we assume the random variables $B_i(t)$ are independent and identically distributed (iid) through
+time. We let $B_i = \exp(b_i)$, where $\left(b_1, b_2\right)$ is some bivariate random variable with the properties 
+that $b_i$ is normally distributed with mean $\mu_i$ and variance $\sigma^2$. 
+We denote $\CovX(b_1, b_2)$ by $\rho$. However, we do \emph{not} assume that $\left(b_1, b_2\right)$ is a 
+bivariate normal distribution. We will consider various distributions of $\left(b_1, b_2\right)$ with the above 
+properties, corresponding to symmetric and asymmetric tail association cases.
+We assume without loss of generality that $\mu_1 \leq \mu_2$, so that species 1 is the weaker competitor.
+Note that $N$ is constant through time.
+
+Modern coexistence theory (REFS) and its recent computational extensions \citep{Ellner_2016,Ellner_2019} 
+%***DAN: Search for REF and replace all with appropriate citations
+quantify the contributions of 
+different mechanisms to species coexistence. But we show in Sup Mat section X
+%***DAN: Make this reference to the sup mat an autolink, and write the section
+that, for purely temporal variation, only one mechanism relates to tail associations: storage effects. 
+We here define storage effects abstractly and then for the lottery model, using an approach and notation
+similar to that of \cite{Ellner_2016} and \cite{Ellner_2019}.
+
 \section{Methods}\label{section:methods}
@@
 \section{Discussion}\label{section:discussion}
-
-\section{Reference code for Jasmin to learn Latex/Sweave}
-
-To get normal text you just type.
-
-To get a paragraph break you leave a blank line, indentation is automatic.
-
-\noindent If you want to remove the indent you use this.
-
-In-line math is always surrounded by dollar signs, $x+1=4$.
-Even when you just have single variables, like if you say ``the notation we use for synchrony is $x$'', you put the $x$ in dollar signs.
-The reason is x and $x$ look different, in a way that a mathematician will notice and wonder if they are meant to be different variables.
-
-For quotes, do it ``like this'' not "like this" or else one of the quotes will be backwards.
-
-Some basic math, inline, by example: $x_{11}$, $y^{22}$, $\frac{1+y}{x+3}$, $\sum_{i=1}^N x_x$,
-$\alpha$, $\beta$.
-
-There is also display math:
-\begin{equation}
-x=y+12-z \label{eq_about_x_and_y}
-\end{equation}
-The ``label'' command allows a textual label for refering to equations later.
-
-The math gets pretty complex. Keep in mind, whatever you want to do, there is a way to do it. We'll save most of that for later.
-
-Now we can refer to equation/expression (\ref{eq_about_x_and_y}). Latex will sort
-out equation numbering and insert the correct number labels for you.
-
-You can also refer to any section in the same way: section \ref{section:introduction}, section \ref{section:methods}.
-
-You can include the value of previously loaded R variables in the text, e.g., the value of allregres[1,5] is
--1.16948109243002, which can also be given rounded: -1.169. You can do simple R
-on the fly, e.g., allregres[1,5] plus 1 is -0.169481092430019. 
-You should never manually type any results, they should always be autolinks like this.
-
-%Comments are lines starting with %
-
-You make a percent with \%
-
-You can refer to Fig. \ref{fig:random_fig} by its label and latex will number figures and use the correct numbers.
-The label is defined below where the figure is placed.
-
-You can refer to Table \ref{tab:random_table} in the same way.
-
-%You can also refer to figures and tables and equations from the sup mat, just prefix the labels
-%by SM- in your calls to \ref{}
-
-You can cite like this \cite{Abanetal2006} or this \citep{Abanetal2006} or this \citealp{Abanetal2006}.
-You cal also cite books \citep{BurnhamAnderson2002}, of course.
 
@@
 
-\clearpage
-\newpage
-
-\section{Tables}
-
-%You make tables like this:
-\begin{table}[H]
-\caption[Summary caption, for table of contents, can be skipped.]{Full caption here.}
-\begin{tabular}{llll} %The four ls mean four columns, each left justified. Can use r or c, too.
-\hline 
-Col 1 header & Col 2 header & Col 3 header & Col 4 header \\
-\hline
-text & or & numbers & 3.2 \\
-or & math & $x+1$ & or \\
-even & the & values & of \\
-loaded & R & variables & -1.16948109243002 \\
-you & can & round & -1.169 \\ 
-\hline
-\end{tabular}
-\label{tab:random_table}
-\end{table}
-
-\clearpage
-\newpage
-
-\section{Figures}
-
-\begin{figure}[!h]
-\includegraphics[width=.5\textwidth]{figures/summaryDocFigures/figure8.png}
-\caption{Type the caption here.} \label{fig:random_fig}
-\end{figure}
-
 \end{document}

### 2021-06-22 paper/Paper.tex
failed previously to removed some files, committing changes so I can then try again

@@
 
-Modern coexistence theory (REFS) and its recent computational extensions \citep{Ellner_2016,Ellner_2019} 
+Modern coexistence theory \citep{Chesson_2000} and its recent computational extensions \citep{Ellner_2016,Ellner_2019} 
 %***DAN: Search for REF and replace all with appropriate citations
@@
 We here define storage effects abstractly and then for the lottery model, using an approach and notation
-similar to that of \cite{Ellner_2016} and \cite{Ellner_2019}.
+that follows elements of the presentations in \cite{Ellner_2016} and \cite{Ellner_2019} closely (and see
+those papers for additional details and explanation). We 
+return briefly to other mechanisms of coexistence in the Discussion.
+%***DAN: Don't forget to do this. Basically want to say intuitively why tail associations don't influence those,
+%though might be worth mentioning that events are getting more extreme. We have studied how *relationships*
+%in the extremes influence coexistence, but probably worth studying (or cite papers if it's been done) how changes
+%in the commonness and extremeness of extreme events themselves influence coexistence, and for that you'd have
+%to also look at the other mechanisms. 
+It is assumed that the growth rate $r_i(t)$ of species $i$ can be written as an increasing function of
+an environment-dependent factor $E_i(t)$ and as a decreasing function of a competitive pressure, $C_i(t)$.
+For the lottery model, we take $E_i(t)=B_i(t)$, henceforth assuming that fecundity depends strictly on
+the environment. Competition $C_i(t)$ is taken to be
 

### 2021-06-23 paper/Paper.Rnw
making progress on the theory section

@@
 \usepackage{amsmath}
+\usepackage{authblk}
 \usepackage{xr}
-\usepackage{authblk}
+\usepackage{relsize}
+\usepackage{tikz}
+
+%external documents
+\externaldocument[SI-]{SupMat}
 
@@
 
-\DeclareMathOperator{\EX}{E}% expected value
-\DeclareMathOperator{\VarX}{var}
-\DeclareMathOperator{\CovX}{cov}
+\DeclareMathOperator{\E}{E}% expected value
+\DeclareMathOperator{\var}{var}
+\DeclareMathOperator{\cov}{cov}
+\DeclareMathOperator{\cor}{cor}
 \DeclareMathOperator{\mean}{mean}
@@
 \DeclareMathOperator{\sd}{sd}
-\DeclareMathOperator{\Prob}{P}
+\DeclareMathOperator{\prob}{P}
 
-%external documents
-\externaldocument[SM-]{SupMat}
+%attempt 1 at nat and sharp
+%\newcommand{\nat}{\mathlarger{\natural}}
+%\newcommand{\shp}{\mathlarger{\sharp}}
+
+%attempt 2 at nat and sharp
+%\newcommand{\nat}{\raisebox{1pt}{\mathsmaller{\mathsmaller{/\hspace{-2pt}/}}}}
+%\newcommand{\shp}{\#}
+
+%attempt 3 at nat and sharp
+\newcommand{\nat}{%
+\text{\hspace{-1.5pt}
+\begin{tikzpicture}[scale=1.8]%
+\draw (.333ex,.333ex) -- (.666ex,.333ex);%
+\draw (.333ex,.666ex) -- (.666ex,.666ex);%
+\draw (.333ex,.333ex) -- (.333ex,1ex);%
+\draw (.66ex,0) -- (.66ex,.66ex);
+\end{tikzpicture}%
+}}
+\newcommand{\shp}{%
+\text{\hspace{-1.5pt}
+\begin{tikzpicture}[scale=1.8]%
+\draw (0,.333ex) -- (1ex,.333ex);%
+\draw (0,.666ex) -- (1ex,.666ex);%
+\draw (.333ex,0) -- (.333ex,1ex);%
+\draw (.666ex,0) -- (.666ex,1ex);
+\end{tikzpicture}%
+}}
+\newcommand{\test}{%
+\text{
+\begin{tikzpicture}[scale=1.8]%
+\draw (0,0) -- (1ex,0ex);
+\draw (0,0) -- (0ex,1ex);
+\draw (0,1ex) -- (1ex,1ex);
+\draw (1ex,0) -- (1ex,1ex);
+\draw (0,.333ex) -- (2ex,.333ex);%
+\draw (0,.666ex) -- (1ex,.666ex);%
+\draw (.333ex,0) -- (.333ex,1ex);%
+\draw (.666ex,0) -- (.666ex,1ex);
+\end{tikzpicture}%
+}}
+
+\newcommand{\olr}{\overline{r}}
+\newcommand{\olrs}{\overline{r}^{\shp}}
+\newcommand{\olrn}{\overline{r}^{\nat}}
+\newcommand{\bs}{\backslash}
 
@@
 \author[a,b]{Daniel C. Reuman}
+%***DAN: Jasmin, you should choose one way your name will appear, exactly, on all your publications,
+%so everyone knows its you. For instance, I always appear as Daniel C. Reuman, not Daniel Reuman, not
+%Dan Reuman, etc. This makes it possible to find all my papers by searching. So please decide what
+%name to use, Jasmin Albert or Pimsupa Albert or P. Jasmin Albert or whatever. Also keep in mind
+%that because "Albert" is a common name you might want something less common for the rest of it 
+%so that your papers are not confused with anyone elses, though of course I emphasize that how you
+%are going to be known it totally up to you, I'm just recommending you think it through now so you
+%can be consistent. I recognize that this all may be moot if you one day change your name and want to
+%be known by the new name on papers after that point, but anyway in case you don't ever change your 
+%name you might as well be consistent now, at least. So please change your name above to whatever
+%you want to be officially known as on scientific papers. Same change in the sup mat, please.
 
@@
 
-For pedgogical clarity, we develop each step of the theory both for a simple classical model, 
-the lottery model, and in general. The model is as follows. 
+For pedgogical clarity, we develop each step of the theory both for a simple classical model 
+(the lottery model), and in general. The lottery model is as follows. 
 %***DAN: Note the inclusion of the model here means we will have to edit the "Model" section of sup mat, since we don't need to intro the model twice. Do later.
@@
 For each $i$, we assume the random variables $B_i(t)$ are independent and identically distributed (iid) through
-time. We let $B_i = \exp(b_i)$, where $\left(b_1, b_2\right)$ is some bivariate random variable with the properties 
+time. We let $B_i = \exp(b_i)$, where $\left(b_1, b_2\right)$ is some bivariate random variable with the property 
 that $b_i$ is normally distributed with mean $\mu_i$ and variance $\sigma^2$. 
-We denote $\CovX(b_1, b_2)$ by $\rho$. However, we do \emph{not} assume that $\left(b_1, b_2\right)$ is a 
+We denote $\cor(b_1, b_2)$ by $\rho$. However, we do \emph{not} assume that $\left(b_1, b_2\right)$ is a 
 bivariate normal distribution. We will consider various distributions of $\left(b_1, b_2\right)$ with the above 
-properties, corresponding to symmetric and asymmetric tail association cases.
-We assume without loss of generality that $\mu_1 \leq \mu_2$, so that species 1 is the weaker competitor.
-Note that $N$ is constant through time.
+properties, corresponding to various tail association cases. Note that $N$ is constant through time.
 
@@
 quantify the contributions of 
-different mechanisms to species coexistence. But we show in Sup Mat section X
-%***DAN: Make this reference to the sup mat an autolink, and write the section
+different mechanisms to species coexistence. But we show below
 that, for purely temporal variation, only one mechanism relates to tail associations: storage effects. 
-We here define storage effects abstractly and then for the lottery model, using an approach and notation
-that follows elements of the presentations in \cite{Ellner_2016} and \cite{Ellner_2019} closely (and see
-those papers for additional details and explanation). We 
-return briefly to other mechanisms of coexistence in the Discussion.
-%***DAN: Don't forget to do this. Basically want to say intuitively why tail associations don't influence those,
-%though might be worth mentioning that events are getting more extreme. We have studied how *relationships*
+We now define storage effects for a general two species model, and specifically 
+for the lottery model, using an approach and notation
+similar to those of \cite{Ellner_2016} and \cite{Ellner_2019}; see those papers for additional details. 
+%***DAN: Don't forget to return to the other mechanisms of coexistence below. Basically want to say intuitively 
+%why tail associations don't influence those,
+%though might be worth mentioning in the Discussion that events are getting more extreme. We have studied how *relationships*
 %in the extremes influence coexistence, but probably worth studying (or cite papers if it's been done) how changes
 %in the commonness and extremeness of extreme events themselves influence coexistence, and for that you'd have
-%to also look at the other mechanisms. 
-It is assumed that the growth rate $r_i(t)$ of species $i$ can be written as an increasing function of
-an environment-dependent factor $E_i(t)$ and as a decreasing function of a competitive pressure, $C_i(t)$.
+%to also look at the other mechanisms. May still need a sup mat laying out all the math.
+
+For a general model, it is assumed that the growth rate $r_i(t)=\ln(N_i(t+1)/N_i(t))$ 
+of species $i$ can be written as an increasing function of
+an environment-dependent factor $E_i(t)$ and as a decreasing function of a quantity representing
+competitive pressure on species $i$, $C_i(t)$. For simplicity, we here assume 
+the $E_i(t)$ are iid through time, and likewise for the
+$C_i(t)$. Though this assumption can be circumvented, we make the assumption because it simplifies the
+presentation without obscuring main ideas. 
 For the lottery model, we take $E_i(t)=B_i(t)$, henceforth assuming that fecundity depends strictly on
-the environment. Competition $C_i(t)$ is taken to be
+the environment. Competition, $C_i(t)$, in the lottery model is the number of new offspring
+divided by the number of open sites, $C_i(t)=( B_1(t)N_1(t)+B_2(t)N_2(t) )/\delta N$, which
+does not depend on $i$. Competition in the lottery model occurs at the stage of juveniles occupying open sites. 
+Species growth rates for the lottery model can then be written 
+$r_i(t)=\ln[1-\delta+E_i(t)/C_i(t)]=\ln[1-\delta+\delta N B_i(t)/(B_1(t)N_1(t)+B_2(t)N_2(t))]$.
+
+In general, the storage effect for species $i$ is the contribution of covariation between $E_i(t)$ and 
+$C_i(t)$ to a difference between invader and resident mean growth rates, when $i$ is the invader.
+Specifically, storage effects for species $i$ are 
+\begin{equation}
+\Delta I_i = (\olr_{i\bs i} - q_{ij} \olr_{j \bs i})-(\olrs_{i \bs i}-q_{ij} \olrs_{j \bs i}),\label{eq:storage_effects_gen}
+\end{equation}
+where: $\olr_{i \bs i}$ refers to the mean ``invasion growth rate'' of species $i$, that is, the growth rate of 
+species $i$ when it is at negligible abundance and species $j$ (the other species) is at stochastic steady state;
+$\olr_{j \bs i}$ refers to the mean steady-state 
+growth rate of species $j$ when species $i$ is absent (this is $0$, \emph{a priori});
+and $\olrs_{i \bs i}$ and $\olrs_{j \bs i}$ are analogous quantities, but for an alternative model scenario, identical except that
+the bivariate random variable $(E_i,C_i)$ has been replaced by a random variable $(E_i^\shp,C_i^\shp)$ 
+with the same marginal distributions but independent components. The so-called \emph{scaling factor}
+$q_{ij}$ measures the relative sensitivities to competition of the two species, see \cite{Chesson_1994},
+\cite{Ellner_2016} and SI section \ref{SI-SIsect:scaling_factors} for the definition of $q_{ij}$, here 
+omitted for brevity.
+
+For the lottery model, it is straightforward to show that $\olr_{i \bs i}=\E[\ln(1-\delta+\delta B_i(t)/B_j(t))]$,
+$\olrs_{i \bs i} = \E[\ln(1-\delta+\delta B_i^\shp(t)/B_j(t))]$, and $\olrs_{j \bs i} = \E[\ln(1-\delta+\delta B_j^\shp(t)/B_j(t))]$, where $\E(\cdot)$ is expected value.
+Here, $B_i^\shp$ is a random variable distributed in the same way as $B_i$, but independent of it; and
+$B_j^\shp$ is distributed the same as $B_j$, but is independent of it. We show in SI section \ref{SI-SIsect:scaling_factors}
+that $q_{ij}=\overline{B}_i/((1-\delta)\overline{B}_j+\delta \overline{B}_i)$, where, as above, the overbar 
+denotes expected value. 
+%***DAN: Note this might be wrong. Some of the language in Ellner et al (2016) indicates q_{ij} here should be 1,
+%even if \mu_1 and \mu_2 are distinct, as long as we use the same \delta for both species (which we do).
+%But when I try to follow the definitions of the same paper to compute q_{ij} I get the quantity listed above,
+%which is not 1. I'll need to get to the bottom of this, possibly by speaking to Ellner, and then put the right 
+%value here, and also write the corresponding sup mat section accordingly (though you may not need that section, 
+%it may be enough to just cite Ellner and Chesson if you go with q_ij=1.
+Since we use log-normally distributed $B_i$, 
+$\overline{B}_i=\exp(\mu_i+\sigma^2/2)$, so $q_{ij}=\exp(\mu_i)/[(1-\delta)\exp(\mu_j)+\delta\exp(\mu_i)]$.
+%***DAN: This statement also needs adjusting if you go with q_ij=1
+So $\Delta I_i$ for the lottery model can be computed by using simulations \citep{Ellner_2016}
+to compute arbitrarily good approximations for $\olr_{i \bs i}$, $\olrs_{i \bs i}$,
+and $\olrs_{j \bs i}$, and then plugging them and the value for $q_{ij}$ into (\ref{eq:storage_effects_gen}).
+
+Having just reviewed storage effects, we now specify how to decompose storage effects into
+contrbutions due to asymmetric tail associations and contributions due
+to $EC$ covariance \emph{per se}. The storage effect for species $i$ was computed by comparing
+a difference between invader and resident mean growth rates for the original model to the same quantity calculated for a 
+modified model for which $E_i$ and $C_i$ were rendered independent (see (\ref{eq:storage_effects_gen})). 
+Because $E_i$ and $C_i$ were rendered 
+completely independent in the modified model, storage effects quantify the contribution of the 
+whole relationship between $E_i$ and $C_i$, in all its aspects, to the invader/resident growth rate
+difference. However, we can instead consider a distribution $(E_i^\nat,C_i^\nat)$ with the same
+marginal distributions as $(E_i,C_i)$ and with some but not all aspects of the relationship 
+between the components modified relative to $(E_i,C_i)$; specifically, tail associations are rendered symmetric
+(see below for how this is done). We then define 
+$\olrn_{i \bs i}$ and $\olrn_{j \bs i}$ based on a modified model using  $(E_i^\nat,C_i^\nat)$,
+and we consider the decomposition 
+\begin{align}
+\Delta I_i &= [(\olr_{i\bs i} - q_{ij} \olr_{j \bs i})-(\olrn_{i \bs i}-q_{ij} \olrn_{j \bs i})]+
+[(\olrn_{i \bs i}-q_{ij} \olrn_{j \bs i})-(\olrs_{i \bs i}-q_{ij} \olrs_{j \bs i})] \\
+&= I_i^{(\text{ata})} + I_i^{(\text{cps})}.\label{eq:storage_effects_further}
+\end{align}
+The superscript ``ata" stands for ``asymmetric tail associations'' because this component of storage effects is due to 
+asymmetric tail associations between $E_i$ and $C_i$. The superscript ``cps" stands for ``covariance
+\emph{per se}'' because this component of storage effects is due to $EC$ covariance \emph{per se}.
 

### 2021-06-23 paper/SupMat.Rnw
making progress on the theory section

@@
 \usepackage{xr}
-%\usepackage{authblk}
+\usepackage{authblk}
+\usepackage{mathrsfs}
 %\usepackage{accents}
 
+%new commands and so on
 \DeclareRobustCommand{\bbone}{\text{\usefont{U}{bbold}{m}{n}1}}
@@
 
-%new commands and so on
+\newcommand{\sC}{\mathscr{C}} %for a script C like Ellner et al (2016), Ecol Lett
+\newcommand{\bs}{\backslash}
 
@@
 %header material for paper
-\title{Drivers of spatial synchrony in giant kelp populations}
+\title{Relationships in the extremes and their influence on competition and coexistence: Supporting information}
 \date{}
 
-%***Tentative author order
-\author{Lawrence W. Sheppard}
-\author{Daniel C. Reuman}
+\author[a]{Jasmin Albert}
+\author[a,b]{Daniel C. Reuman}
+%***DAN: Jasmin, you should choose one way your name will appear, exactly, on all your publications,
+%so everyone knows its you. For instance, I always appear as Daniel C. Reuman, not Daniel Reuman, not
+%Dan Reuman, etc. This makes it possible to find all my papers by searching. So please decide what
+%name to use, Jasmin Albert or Pimsupa Albert or P. Jasmin Albert or whatever. Also keep in mind
+%that because "Albert" is a common name you might want something less common for the rest of it 
+%so that your papers are not confused with anyone elses, though of course I emphasize that how you
+%are going to be known it totally up to you, I'm just recommending you think it through now so you
+%can be consistent. I recognize that this all may be moot if you one day change your name and want to
+%be known by the new name on papers after that point, but anyway in case you don't ever change your 
+%name you might as well be consistent now, at least. So please change your name above to whatever
+%you want to be officially known as on scientific papers. Same change in main text, please.
+
+\affil[a]{Department of Ecology and Evolutionary Biology and Kansas Biological Survey, University of Kansas}
+\affil[b]{Laboratory of Populations, Rockefeller University}
+
+%***Need to indicate corresponding authors
 
 \begin{document}
-\SweaveOpts{concordance=TRUE}
+%\SweaveOpts{concordance=TRUE}
 
@@
 <<eval=T,echo=F,message=F,warning=F>>=
+system(paste("biber", sub("\\.Rnw$", "", current_input())))
+
 #place R code here for loading in necessary variables from Results
+#ordinarily this would be a readRDS command, but I'm just defining a variable here for now
+allregres<-matrix(rnorm(10),2,5)
 @
@@
 %probably want table of contents, figure and table lists
+
+\section{Scaling factors}\label{SIsect:scaling_factors}
+%***DAN: I am not completely sure about anything in this section and need to get Ellner to look at it. I 
+%did give it a reasonable shot though.
+
+Following \cite{Ellner_2016}, we define
+\begin{equation}
+\sC_j = -r_j(\overline{E}_j,C_j),
+\end{equation}
+and 
+\begin{equation}
+\sC_{j \bs i} = -r_j(\overline{E}_j,C_{j \bs i}),
+\end{equation}
+i.e., $\sC_{j \bs i}$ is the negative of the growth rate of species $j$ when $E_j$ is at its mean value
+and when $i$ is at negligible abundance. Then $q_{ij}$ is defined by computing
+\begin{equation}
+\frac{\partial \sC_{i \bs i}}{\partial \sC_{j \bs i}},
+\end{equation}
+and evaluating it at $\sC_{j \bs i}=0$ [see \cite{Chesson_1994} and \cite{Ellner_2016} for background
+on this definition]. To make sense of this definition, it must be possible to write $\sC_{i \bs i}$, explicitly 
+or implicitly, uniquely as a function of $\sC_{j \bs i}$ \citep{Ellner_2016}.
+
+For the lottery model, for $i \neq j$, we have
+\begin{align}
+\sC_j &= -\ln \left[ 1-\delta+\delta N \frac{\overline{B}_j}{B_1(t) N_1 (t) + B_2(t) N_2(t)}  \right]  \\
+\sC_{j \bs i} &= -\ln \left[ 1-\delta+\delta  \frac{\overline{B}_j}{B_j(t)} \right] \\
+\sC_{i \bs i} &= -\ln \left[ 1-\delta+\delta  \frac{\overline{B}_i}{B_j(t)} \right].
+\end{align}
+Both $\sC_{i \bs i}$ and $\sC_{j \bs i}$ are strictly monotonic functions of $B_j$, so the function
+$B_j \mapsto \sC_{j \bs i}$ is invertible, and using its inverse provides an expression of $\sC_{i \bs i}$ 
+in terms of $\sC_{j \bs i}$. One could write this expression explicitly and then differentiate it calculate $q_{ij}$,
+but we instead compute $q_{ij}$ through implicit differentiation:
+\begin{align}
+\frac{\partial \sC_{i \bs i}}{\partial \sC_{j \bs i}} &= \left. \frac{\partial \sC_{i \bs i}}{\partial B_j} \middle/ \frac{\partial \sC_{j \bs i}}{\partial B_j} \right. \\
+&= \left( \frac{1-\delta +\delta \frac{\overline{B}_j}{B_j}}{1-\delta+\delta \frac{\overline{B}_i}{B_j}} \right) \left( \frac{\overline{B}_i}{\overline{B}_j}  \right).
+\end{align}
+Solving $\sC_{j \bs i}=0$ gives $B_j=\overline{B}_j$, so 
+\begin{align}
+q_{ij} &= \left( \frac{1-\delta +\delta \frac{\overline{B}_j}{\overline{B}_j}}{1-\delta+\delta \frac{\overline{B}_i}{\overline{B}_j}} \right) \left( \frac{\overline{B}_i}{\overline{B}_j}  \right) \\
+&= \frac{\overline{B}_i}{\overline{B}_j (1-\delta) +\overline{B}_i \delta},
+\end{align}
+as stated in the main text.
+
+\cite{Chesson_1994} and \cite{Ellner_2016} use $q_{ij}=1$ for this same model. Why the discrepancy? 
+DAN: You need to explain this discrepancy here after talking to Ellner.
+%***DAN: Add some text explaining the discrepancy
+
+\section{Tail association is not relevant to other mechanisms of coexistence for purely temporal variation}\label{SIsect:other_mechanisms}
+
+%***DAN: Sections below this point are from the earlier reasoning and may have to be edited and placed in other
+%locations. This comment written 2021 06 22
 
@@
 
+\clearpage
+\newpage
+
+%you store your bibliographic info in refs.bib, see that file
+\bibliographystyle{ecology_letters2}
+\bibliography{refs}
+
 \end{document}

### 2021-06-23 paper/Paper.Rnw
further progress on theory section

@@
 \begin{tikzpicture}[scale=1.8]%
-\draw (.333ex,.333ex) -- (.666ex,.333ex);%
-\draw (.333ex,.666ex) -- (.666ex,.666ex);%
-\draw (.333ex,.333ex) -- (.333ex,1ex);%
-\draw (.66ex,0) -- (.66ex,.66ex);
+\draw (.333ex,0) -- (.333ex,1ex);%
+\draw (.666ex,0) -- (.666ex,1ex);
 \end{tikzpicture}%
@@
 that $b_i$ is normally distributed with mean $\mu_i$ and variance $\sigma^2$. 
-We denote $\cor(b_1, b_2)$ by $\rho$. However, we do \emph{not} assume that $\left(b_1, b_2\right)$ is a 
+However, we do \emph{not} assume that $\left(b_1, b_2\right)$ is a 
 bivariate normal distribution. We will consider various distributions of $\left(b_1, b_2\right)$ with the above 
@@
 contrbutions due to asymmetric tail associations and contributions due
-to $EC$ covariance \emph{per se}. The storage effect for species $i$ was computed by comparing
+to $EC$ correlation \emph{per se}. The storage effect for species $i$ was computed by comparing
 a difference between invader and resident mean growth rates for the original model to the same quantity calculated for a 
@@
 marginal distributions as $(E_i,C_i)$ and with some but not all aspects of the relationship 
-between the components modified relative to $(E_i,C_i)$; specifically, tail associations are rendered symmetric
+between the components modified relative to $(E_i,C_i)$; specifically, asymmetries of tail association are removed
 (see below for how this is done). We then define 
@@
 The superscript ``ata" stands for ``asymmetric tail associations'' because this component of storage effects is due to 
-asymmetric tail associations between $E_i$ and $C_i$. The superscript ``cps" stands for ``covariance
-\emph{per se}'' because this component of storage effects is due to $EC$ covariance \emph{per se}.
+asymmetries of tail associations between $E_i$ and $C_i$. The superscript ``cps" stands for ``correlation
+\emph{per se}'' because this component of storage effects is due to $EC$ correlation \emph{per se}.
+The notation $\nat$ was chosen because the new symbol comprises part of the symbol $\shp$ (i.e., two of its four lines), 
+corresponding to the fact that part of the relationship between $E$ and $C$ components is removed.
+
+How are asymmetries of tail association removed while leaving $E$ and $C$ marginal distributions and overall 
+correlation unchanged? We assume that the cummulative distribution functions (cdfs) of $E_i$ and 
+$C_i$, which we denote $F_{E_i}$ and $F_{C_i}$, respectively, are strictly monotonic and continuous. 
+Letting $\varphi$ be the cdf of a standard normal distribution, we consider the random variables
+$\varphi^{-1} \circ F_{E_i} (E_i)$ and $\varphi^{-1} \circ F_{C_i}(C_i)$, which are standard normally
+distributed, although they are not necessarily jointly bivariate normal.  Letting $\rho_i$ be the covariance of
+these random variables, we then define a bivariate normal random variable $(e_i,c_i)$ such that
+$\var(e_i)=\var(c_i)=1$ and such that $\cov(e_i,c_i)=\tilde{\rho}_i$. We define $E_i^\nat = F_{E_i}^{-1} \circ \varphi(e_i)$
+and $C_i^\nat=F_{C_i}^{-1} \circ \varphi(c_i)$. The value of $\tilde{\rho}_i$ is selected in one of 
+four alternative ways: $\tilde{\rho}_i = \rho_i$; or so that the Pearson, Spearman or Kendall correlation of 
+$E_i^\nat$ and $C_i^\nat$ matches that of $E_i$ and $C_i$. The choice of bivariate normal 
+$(e_i,c_i)$ ensures that $E_i^\nat$ and $C_i^\nat$ are symmetrically tail associated in the sense
+described in the Introduction.
+%***DAN: This means you have to introduce this well, and explain why the way we define tail assoc is
+%the *right* way
+For those familar with the language of copula statistics (see, e.g., \citealp{Ghosh_2020_AER} and \citealp{Genest_2007}), 
+we have replaced the copula of $(E_i,C_i)$ with a normal copula to form $(E_i^\nat,C_i^\nat)$.
+The choice of $\tilde{\rho}_i$ ensures that $E_i^\nat$ and $C_i^\nat$ are correlated to the same extent as 
+$E_i$ and $C_i$, according to whichever way of defining correlation is preferable to the researcher. Thus $E_i^\nat$ and 
+$C_i^\nat$ have had their tail association rendered symmetric while retaining their correlation \emph{per se},
+as desired.
+
+%***DAN: next step - work it all out for the lottery model now
 

### 2021-06-24 paper/Paper.Rnw
made some presentational errors, getting started fixing them. they have to do with the description of the general case.

@@
 $C_i(t)$. Though this assumption can be circumvented, we make the assumption because it simplifies the
-presentation without obscuring main ideas. 
+presentation without obscuring ideas. 
 For the lottery model, we take $E_i(t)=B_i(t)$, henceforth assuming that fecundity depends strictly on
@@
 Species growth rates for the lottery model can then be written 
-$r_i(t)=\ln[1-\delta+E_i(t)/C_i(t)]=\ln[1-\delta+\delta N B_i(t)/(B_1(t)N_1(t)+B_2(t)N_2(t))]$.
+$r_i(t)=\ln[1-\delta+\delta N B_i(t)/(B_1(t)N_1(t)+B_2(t)N_2(t))]=\ln[1-\delta+E_i(t)/C_i(t)]$.
 
@@
 and $\olrs_{i \bs i}$ and $\olrs_{j \bs i}$ are analogous quantities, but for an alternative model scenario, identical except that
-the bivariate random variable $(E_i,C_i)$ has been replaced by a random variable $(E_i^\shp,C_i^\shp)$ 
-with the same marginal distributions but independent components. The so-called \emph{scaling factor}
+the influences of environment and competition in the invader and resident scenarios have been rendered independent.  
+The so-called \emph{scaling factor}
 $q_{ij}$ measures the relative sensitivities to competition of the two species, see \cite{Chesson_1994},

### 2021-06-24 paper/SupMat.Rnw
typo

--- 2021-06-05 paper/SupMat.Rnw
@@
 \begin{equation}
-\bar \epsilon^{(\dot {EC})} = \bar \epsilon^{(EC)} - \bar \epsilon^{(\ddot {EC})}
+\bar \epsilon^{(\dot {EC})} = \bar \epsilon^{(EC)} - \bar \epsilon^{( {EC_0})}
 \end{equation}
@@
 
+\begin{equation}
+\bar \epsilon^{( {EC_0})} = \bar r_1 - \bar r_1^{\#} + \bar r_2^{\#},
+\end{equation}
+
+\noindent where $ \bar r_1, \bar r_1^{\#}$ and $\bar r_2^{\#}$ are in the symmetric case. In our model,
+
+\begin{equation}
+\bar \epsilon^{( {EC_0})} = \EX(\ln[1-\delta+\delta \exp(u_1)]) - \EX(\ln[1-\delta+\delta \exp(u_1^{\#})]) + \EX(\ln[1-\delta+\delta \exp(u_2^{\#})]),
+\end{equation}
+
+\noindent where $u_1 \sim N(\mu_1-\mu_2, 2\sigma^2-2\rho)$, $u_1^{\#} \sim N(\mu_1-\mu_2, 2\sigma^2)$, and $u_2^{\#} \sim N(0, 2\sigma^2)$
+
+\noindent *Note that the contribution of asymmetry to storage effects is not the same as the storage effect that we computed above for asymmetriv cases.
+
+In generalized terms, this further decomposition could be applied to test for contributions of asymmetry between any two variables.
+Here we tested the asymmetry between $b_1$ and $b_2$.
+Why did we not look at E and C?
+Do they have to be normally distrubuted?
+
+This break down would only work if the variables are normally distributed. Thus one can convert their distribution to normal to impliment this decomposition.
+
 \end{document}

### 2021-06-24 paper/Paper.Rnw

--- 2021-06-23 paper/Paper.Rnw
@@
 Species growth rates for the lottery model can then be written 
-$r_i(t)=\ln[1-\delta+E_i(t)/C_i(t)]=\ln[1-\delta+\delta N B_i(t)/(B_1(t)N_1(t)+B_2(t)N_2(t))]$.
+$r_i(t)=\ln[1-\delta+\delta N B_i(t)/(B_1(t)N_1(t)+B_2(t)N_2(t))]=\ln[1-\delta+E_i(t)/C_i(t)]$.
 
-In general, the storage effect for species $i$ is the contribution of covariation between $E_i(t)$ and 
-$C_i(t)$ to a difference between invader and resident mean growth rates, when $i$ is the invader.
+In general, the storage effect for species $i$ is the contribution of covariation between $E$ and 
+$C$ to a difference between invader and resident mean growth rates, when $i$ is the invader.
 Specifically, storage effects for species $i$ are 
@@
 \end{equation}
-where: $\olr_{i \bs i}$ refers to the mean ``invasion growth rate'' of species $i$, that is, the growth rate of 
+where: $\olr_{i \bs i}$ refers to the mean ``invasion growth rate'' of species $i$, that is, the mean growth rate of 
 species $i$ when it is at negligible abundance and species $j$ (the other species) is at stochastic steady state;
 $\olr_{j \bs i}$ refers to the mean steady-state 
-growth rate of species $j$ when species $i$ is absent (this is $0$, \emph{a priori});
+growth rate of species $j$ when species $i$ is absent or at negligible abundance (this is $0$, \emph{a priori});
 and $\olrs_{i \bs i}$ and $\olrs_{j \bs i}$ are analogous quantities, but for an alternative model scenario, identical except that
-the bivariate random variable $(E_i,C_i)$ has been replaced by a random variable $(E_i^\shp,C_i^\shp)$ 
-with the same marginal distributions but independent components. The so-called \emph{scaling factor}
+influences of environment and competition on the invader or resident have been rendered independent.  
+A subscript 
+$i \bs i$ always indicates a quantity computed for species $i$ when that species is at negligible abundance, and 
+a subscript $j \bs i$ indicates a quantity computed for $j$ when $i$ is absent.
+The \emph{scaling factor}
 $q_{ij}$ measures the relative sensitivities to competition of the two species, see \cite{Chesson_1994},
-\cite{Ellner_2016} and SI section \ref{SI-SIsect:scaling_factors} for the definition of $q_{ij}$, here 
-omitted for brevity.
+\cite{Ellner_2016} and SI section \ref{SI-SIsect:scaling_factors} for the definition of $q_{ij}$.
 
-For the lottery model, it is straightforward to show that $\olr_{i \bs i}=\E[\ln(1-\delta+\delta B_i(t)/B_j(t))]$,
-$\olrs_{i \bs i} = \E[\ln(1-\delta+\delta B_i^\shp(t)/B_j(t))]$, and $\olrs_{j \bs i} = \E[\ln(1-\delta+\delta B_j^\shp(t)/B_j(t))]$, where $\E(\cdot)$ is expected value.
-Here, $B_i^\shp$ is a random variable distributed in the same way as $B_i$, but independent of it; and
-$B_j^\shp$ is distributed the same as $B_j$, but is independent of it. We show in SI section \ref{SI-SIsect:scaling_factors}
-that $q_{ij}=\overline{B}_i/((1-\delta)\overline{B}_j+\delta \overline{B}_i)$, where, as above, the overbar 
-denotes expected value. 
+For the lottery model, when $i$ is an invader, the competition experienced by either 
+species is obtained from the expression for $C_i=C_j$ by setting $N_i=0$ and $N_j=N$; and likewise to get
+$\olr_{i \bs i}$ from the expression for $r_i$ for the model. Thus 
+$C_{i \bs i}(t)=C_{j \bs i}(t)=B_j(t)/\delta=E_j(t)/\delta$, and 
+$\olr_{i \bs i} = \E[\ln(1-\delta+\delta B_i/B_j)]=\E[\ln(1-\delta+E_i /C_{i \bs i})]$. 
+Here $\E(\cdot)$ denotes expected value just as does an overbar, but we use $\E$ for longer expressions for which
+an overbar would be cumbersome. Likewise, 
+$\olr_{j \bs i} = \E[\ln(1-\delta+\delta B_j/B_j)]=\E[\ln(1-\delta+E_j/C_{j \bs i})]$,
+though we already knew this equals $0$. An expression for $\olrs_{i \bs i}$ is then obtained from the 
+above expression for $\olr_{i \bs i}$ by replacing the bivariate random variable $(E_i,C_{i \bs i})=(B_i,B_j/\delta)$
+by a random variable with the same marginals, but independent components, $(B_i^\shp,B_j^\shp /\delta)$.
+Thus $\olrs_{i \bs i} = \E[\ln(1-\delta+\delta B_i^\shp/B_j^\shp)]$. Likewise, a formula for 
+$\olrs_{j \bs i}$ is available by replacing $(E_j,C_{j \bs i})$ in an analogous way. We use 
+$(E_j^\shp,C_{j \bs i}^\shp)=(B_j^\shp, B_j /\delta)$, so 
+$\olrs_{j \bs i}=\E[\ln(1-\delta+\delta B_j^\shp / B_j)]$.
+We show in SI section \ref{SI-SIsect:scaling_factors}
+that $q_{ij}=\overline{B}_i/((1-\delta)\overline{B}_j+\delta \overline{B}_i)$. 
 %***DAN: Note this might be wrong. Some of the language in Ellner et al (2016) indicates q_{ij} here should be 1,
@@
 So $\Delta I_i$ for the lottery model can be computed by using simulations \citep{Ellner_2016}
-to compute arbitrarily good approximations for $\olr_{i \bs i}$, $\olrs_{i \bs i}$,
-and $\olrs_{j \bs i}$, and then plugging them and the value for $q_{ij}$ into (\ref{eq:storage_effects_gen}).
+to very accurately approximate $\olr_{i \bs i}$, $\olrs_{i \bs i}$,
+and $\olrs_{j \bs i}$, and then plugging them and $q_{ij}$ into (\ref{eq:storage_effects_gen}).
 
 Having just reviewed storage effects, we now specify how to decompose storage effects into
-contrbutions due to asymmetric tail associations and contributions due
-to $EC$ covariance \emph{per se}. The storage effect for species $i$ was computed by comparing
-a difference between invader and resident mean growth rates for the original model to the same quantity calculated for a 
-modified model for which $E_i$ and $C_i$ were rendered independent (see (\ref{eq:storage_effects_gen})). 
-Because $E_i$ and $C_i$ were rendered 
+contributions due to asymmetric tail associations and due
+to $EC$ correlation \emph{per se}. 
+Storage effects for species $i$ were computed by comparing
+differences between invader and resident mean growth rates for the original model and for a 
+modified model for which the components of $(E_i,C_{i \bs i})$ and those of $(E_j,C_{j \bs i})$
+were rendered independent of each other [see (\ref{eq:storage_effects_gen})]. 
+Because these components were rendered 
 completely independent in the modified model, storage effects quantify the contribution of the 
-whole relationship between $E_i$ and $C_i$, in all its aspects, to the invader/resident growth rate
-difference. However, we can instead consider a distribution $(E_i^\nat,C_i^\nat)$ with the same
-marginal distributions as $(E_i,C_i)$ and with some but not all aspects of the relationship 
-between the components modified relative to $(E_i,C_i)$; specifically, tail associations are rendered symmetric
-(see below for how this is done). We then define 
-$\olrn_{i \bs i}$ and $\olrn_{j \bs i}$ based on a modified model using  $(E_i^\nat,C_i^\nat)$,
-and we consider the decomposition 
+whole $EC$ relationship, in all its aspects, to the invader/resident growth rate
+difference. But we can instead consider distributions $(E_i^\nat,C_{i \bs i}^\nat)$ and 
+$(E_j^\nat,C_{j \bs i}^\nat)$ with the same
+marginal distributions, respectively, as $(E_i,C_{i \bs i})$ and $(E_j,C_{j \bs i})$,
+and with some but not all aspects of the relationship 
+between the components altered relative to the original distributions. Specifically, asymmetries of tail association are removed -- see below for how this is done. We then define 
+$\olrn_{i \bs i}$ and $\olrn_{j \bs i}$ based on a modified model using  $(E_i^\nat,C_{i \bs i}^\nat)$
+and $(E_j^\nat,C_{j \bs i}^\nat)$, and we consider the decomposition 
 \begin{align}
@@
 The superscript ``ata" stands for ``asymmetric tail associations'' because this component of storage effects is due to 
-asymmetric tail associations between $E_i$ and $C_i$. The superscript ``cps" stands for ``covariance
-\emph{per se}'' because this component of storage effects is due to $EC$ covariance \emph{per se}.
+asymmetries of tail association between $E_i$ and $C_{i \bs i}$, and between $E_j$ and $C_{j \bs i}$. 
+The superscript ``cps" stands for ``correlation
+\emph{per se}'' because this component is due to $EC$ correlation \emph{per se}, after 
+removing the effects of asymmetric tail associations.
+The notation $\nat$ was chosen because this new symbol is part of the symbol $\shp$ (i.e., two of its four lines), 
+corresponding to the removal of part of the relationship between $E$ and $C$ components.
+
+How are asymmetries of tail association removed while leaving $E$ and $C$ marginal distributions and overall 
+correlation unchanged? We assume that the cummulative distribution functions (cdfs) of $E_i$ and 
+$C_{i \bs i}$, which we denote $F_{E_i}$ and $F_{C_{i \bs i}}$, respectively, are strictly monotonic and continuous. 
+Letting $\varphi$ be the cdf of a standard normal distribution, we consider the random variables
+$\varphi^{-1} \circ F_{E_i} (E_i)$ and $\varphi^{-1} \circ F_{C_{i \bs i}}(C_{i \bs i})$, which are standard normally
+distributed, although they are not necessarily jointly bivariate normal.  Let $\rho_i$ be the covariance of
+these random variables. We then define a bivariate normal random variable $(e_i,c_{i \bs i})$ such that
+$\var(e_i)=\var(c_{i \bs i})=1$, and denote $\cov(e_i,c_{i \bs i})$ by $\tilde{\rho}_i$. We define 
+$E_i^\nat = F_{E_i}^{-1} \circ \varphi(e_i)$
+and $C_{i \bs i}^\nat=F_{C_{i \bs i}}^{-1} \circ \varphi(c_{i \bs i})$. The value of $\tilde{\rho}_i$ is then determined 
+in one of 
+four alternative ways: $\tilde{\rho}_i = \rho_i$; or so that the Pearson, Spearman or Kendall correlation of 
+$E_i^\nat$ and $C_{i \bs i}^\nat$ matches that of $E_i$ and $C_{i \bs i}$. The choice of bivariate normal 
+$(e_i,c_{i \bs i})$ ensures that $E_i^\nat$ and $C_{i \bs i}^\nat$ are symmetrically tail associated in the sense
+described in the Introduction.
+%***DAN: This means you have to introduce this well, and explain why the way we define tail assoc is
+%the *right* way
+For those familar with the language of copula statistics (see, e.g., \citealp{Ghosh_2020_AER} and \citealp{Genest_2007}), 
+we replaced the copula of $(E_i,C_{i \bs i})$ with a normal copula to form $(E_i^\nat,C_{i \bs i}^\nat)$.
+The choice of $\tilde{\rho}_i$ ensures that $E_i^\nat$ and $C_{i \bs i}^\nat$ are correlated to the same extent as 
+$E_i$ and $C_{i \bs i}$, according to whichever definition of correlation is considered preferable. Thus $E_i^\nat$ and 
+$C_{i \bs i}^\nat$ have had their tail association rendered symmetric while retaining their correlation \emph{per se},
+as desired. The random variable $(E_j^\nat, C_{j \bs i}^\nat)$ is created in an analogous way.
+
+For the lottery model, recall that $E_i=B_i$ and $C_{i \bs i} = B_j/\delta$. So 
+$\varphi^{-1} \circ F_{E_i}(E_i) = \frac{b_i-\mu_i}{\sigma}$ and 
+$\varphi^{-1} \circ F_{C_{i \bs i}}(C_{i \bs i})=\frac{b_j - \mu_j}{\sigma}$.  
+Also, $\rho_i = \cov(\frac{b_i-\mu_i}{\sigma},\frac{b_j-\mu_j}{\sigma})$, which equals the 
+Pearson correlation of $b_i$ and $b_j$. For the lottery model, it seems natural to make the choice
+$\tilde{\rho}_i=\rho_i$, which straightforwardly implies that $(E_i^\nat,C_{i \bs i}^\nat)$
+can be written as $(\exp(\tilde{b}_i),\exp(\tilde{b}_j))$, where $(\tilde{b}_i,\tilde{b}_j)$
+is bivariate normally distributed with mean $(\mu_i,\mu_j)$, $\var(\tilde{b}_i)=\var(\tilde{b}_j)=\sigma$,
+and $\cov(\tilde{b}_i,\tilde{b}_j)=\cov(b_i,b_j)$. So $(\tilde{b}_i,\tilde{b}_j)$ has the same
+marginal distributions and covariance matrix as $(b_i,b_j)$, but it is bivariate normal, whereas
+$(b_i,b_j)$ itself need not be. Given these definitions, it is straightforward to compute the 
+components of (\ref{eq:storage_effects_further}) for the lottery model using the simulation 
+approach of \cite{Ellner_2016,Ellner_2019}, though in the next section we describe some tricks
+that make these computations efficient.
+
+The other mechanism in modern coexistence theory that apply in our modelling context (i.e., purely temporal variation)
+make no reference to covariation between quantities, whereas asymmetries of tail associations are 
+defined in a covariation context. Therefore, storage effects are the only mechanism of coexistence
+for which asymmetries of tail association are relevant. Details are given in SI section 
+\ref{SIsect:other_mechanisms}, and we return briefly to related ideas in the Discussion.
+%***DAN: In the Discussion, I intend to talk about how possible changes in tail association 
+%patterns of env vars in nature may influence coexistence of species. It won't influence the 
+%other mechanisms, though changes in the extreme values of environmental variables, known
+%to be occurring, might.
 
 \section{Methods}\label{section:methods}
+
+%***DAN: Next steps:
+%Write to Ellner asking question about scaling factors
+%Decide what we ought to be computing 
+%Write Methods correspondingly
+%Talk to Jasmin about it - did I miss anything that we already computed?
+%Then work with Jasmin to get the code edited to compute that stuff
+%Write results
+%Once results have solidified, make a code suite version that is clean
 

### 2021-06-28 paper/Paper.Rnw
name
titles
further work on editing theory, editing and writing sup mat sections to get ready to write methods

@@
 
+Jasmin, I think the below makes a pretty good start at tryinbg to nail down a good way to express our goals
+in the last para of the Intro, so please incorporate these into your last para. If you have a better way
+to phrase these, please put both so we can compare at some point when we meet.
+G1) To augment the modern theory of coexistence (MTC) to show how asymmetric tail association (ATA)
+plays a role. The MTC decomposes a measure of coexistence into components due to each of several mechanisms. 
+We address how ATA contributes to these components. We here examine mechanisms of coexistence that apply in 
+models exhibiting solely temporal variation; spatial variation can be considered in future work.
+G2) To explore using a simple model the relative magnitude of the contributions of coexistence that come
+from ATA compared to other mechanisms.
+
 \section{Theory}\label{section:theory}
 
-For pedgogical clarity, we develop each step of the theory both for a simple classical model 
-(the lottery model), and in general. The lottery model is as follows. 
+In this section we pursue goal G1 from the Introduction: to augment the MTC to show how ATA
+plays a role in species coexistence.
+%***DAN: make sure you define MTC and ATA once (probably in the Intro) and then use the abbreviations onle after that.
+For pedgogical clarity, we develop each step of the theory in general and for a simple classical model 
+(the lottery model), as an example.
+
+The lottery model is as follows. 
 %***DAN: Note the inclusion of the model here means we will have to edit the "Model" section of sup mat, since we don't need to intro the model twice. Do later.
@@
 similar to those of \cite{Ellner_2016} and \cite{Ellner_2019}; see those papers for additional details.
-%***DAN: Don't forget to return to the other mechanisms of coexistence below. Basically want to say intuitively 
-%why tail associations don't influence those,
-%though might be worth mentioning in the Discussion that events are getting more extreme. We have studied how *relationships*
-%in the extremes influence coexistence, but probably worth studying (or cite papers if it's been done) how changes
-%in the commonness and extremeness of extreme events themselves influence coexistence, and for that you'd have
-%to also look at the other mechanisms. May still need a sup mat laying out all the math.
 
@@
 an environment-dependent factor $E_i(t)$ and as a decreasing function of a quantity representing
-competitive pressure on species $i$, $C_i(t)$. For simplicity, we here assume 
-the $E_i(t)$ are iid through time, and likewise for the
+competitive pressure on species $i$, $C_i(t)$. As for the lottery model example, we here assume, for simpicity, 
+that the $E_i(t)$ are iid through time, and likewise for the
 $C_i(t)$. Though this assumption can be circumvented, we make the assumption because it simplifies the
-presentation without obscuring ideas. 
+presentation of new ideas.
+
 For the lottery model, we take $E_i(t)=B_i(t)$, henceforth assuming that fecundity depends strictly on
@@
 \begin{equation}
-\Delta I_i = (\olr_{i\bs i} - q_{ij} \olr_{j \bs i})-(\olrs_{i \bs i}-q_{ij} \olrs_{j \bs i}),\label{eq:storage_effects_gen}
+\Delta_i^{(EC)} = (\olr_{i\bs i} - q_{ij} \olr_{j \bs i})-(\olrs_{i \bs i}-q_{ij} \olrs_{j \bs i}),\label{eq:storage_effects_gen}
 \end{equation}
@@
 growth rate of species $j$ when species $i$ is absent or at negligible abundance (this is $0$, \emph{a priori});
-and $\olrs_{i \bs i}$ and $\olrs_{j \bs i}$ are analogous quantities, but for an alternative model scenario, identical except that
+and $\olrs_{i \bs i}$ and $\olrs_{j \bs i}$ are analogous quantities, but for an alternative modelling scenario, identical except that
 influences of environment and competition on the invader or resident have been rendered independent.  
-A subscript 
+Following \cite{Ellner_2016,Ellner_2019}, a subscript 
 $i \bs i$ always indicates a quantity computed for species $i$ when that species is at negligible abundance, and 
 a subscript $j \bs i$ indicates a quantity computed for $j$ when $i$ is absent.
+The notation $\Delta_i^{(EC)}$ for storage effects also follows \cite{Ellner_2019}.
 The \emph{scaling factor}
@@
 \cite{Ellner_2016} and SI section \ref{SI-SIsect:scaling_factors} for the definition of $q_{ij}$. 
+Since it was assumed that $r_i=r_i(E_i,C_i)$, we can write $\olr_{i \bs i}=\E[r_i(E_i,C_{i \bs i})]$, where
+$\E(\cdot)$ denotes expected value just as does an overbar, but we use $\E$ for longer expressions. Also, $\olr_{j \bs i}=\E[r_j(E_j,C_{j \bs i})]$.
+An expression for $\olrs_{i \bs i}$ is obtained from $\olr_{i \bs i}$ by replacing the bivariate random 
+variable $(E_i,C_{i \bs i})$ by a random variable with the same marginal distributions, but independent components, 
+$(E_i^\shp,C_{i \bs i}^\shp)$, so $\olrs_{i \bs i}=\E[r_i(E_i^\shp,C_{i \bs i}^\shp)]$.
+Likewise, $\olrs_{j \bs i}=\E[r_j(E_j^\shp,C_{j \bs i}^\shp)]$.
 
@@
 $\olr_{i \bs i} = \E[\ln(1-\delta+\delta B_i/B_j)]=\E[\ln(1-\delta+E_i /C_{i \bs i})]$. 
-Here $\E(\cdot)$ denotes expected value just as does an overbar, but we use $\E$ for longer expressions for which
-an overbar would be cumbersome. Likewise, 
+Likewise, 
 $\olr_{j \bs i} = \E[\ln(1-\delta+\delta B_j/B_j)]=\E[\ln(1-\delta+E_j/C_{j \bs i})]$,
-though we already knew this equals $0$. An expression for $\olrs_{i \bs i}$ is then obtained from the 
-above expression for $\olr_{i \bs i}$ by replacing the bivariate random variable $(E_i,C_{i \bs i})=(B_i,B_j/\delta)$
+though we already knew this equals $0$. 
+An expression for $\olrs_{i \bs i}$ is obtained by replacing $(E_i,C_{i \bs i})=(B_i,B_j/\delta)$
 by a random variable with the same marginals, but independent components, $(B_i^\shp,B_j^\shp /\delta)$.
-Thus $\olrs_{i \bs i} = \E[\ln(1-\delta+\delta B_i^\shp/B_j^\shp)]$. Likewise, a formula for 
-$\olrs_{j \bs i}$ is available by replacing $(E_j,C_{j \bs i})$ in an analogous way. We use 
+Thus $\olrs_{i \bs i} = \E[\ln(1-\delta+\delta B_i^\shp/B_j^\shp)]$. Likewise,
+$\olrs_{j \bs i}$ is computed by replacing $(E_j,C_{j \bs i})$ in an analogous way; we use 
 $(E_j^\shp,C_{j \bs i}^\shp)=(B_j^\shp, B_j /\delta)$, so 
@@
 %it may be enough to just cite Ellner and Chesson if you go with q_ij=1.
-Since we use log-normally distributed $B_i$, 
+Since we assumed log-normally distributed $B_i$, 
 $\overline{B}_i=\exp(\mu_i+\sigma^2/2)$, so $q_{ij}=\exp(\mu_i)/[(1-\delta)\exp(\mu_j)+\delta\exp(\mu_i)]$.
 %***DAN: This statement also needs adjusting if you go with q_ij=1
-So $\Delta I_i$ for the lottery model can be computed by using simulations \citep{Ellner_2016}
+So $\Delta_i^{(EC)}$ for the lottery model can be computed by using simulations \citep{Ellner_2016}
 to very accurately approximate $\olr_{i \bs i}$, $\olrs_{i \bs i}$,
@@
 
-Having just reviewed storage effects, we now specify how to decompose storage effects into
+We now specify how to decompose storage effects into
 contributions due to asymmetric tail associations and due
@@
 \begin{align}
-\Delta I_i &= [(\olr_{i\bs i} - q_{ij} \olr_{j \bs i})-(\olrn_{i \bs i}-q_{ij} \olrn_{j \bs i})]+
-[(\olrn_{i \bs i}-q_{ij} \olrn_{j \bs i})-(\olrs_{i \bs i}-q_{ij} \olrs_{j \bs i})] \\
-&= I_i^{(\text{ata})} + I_i^{(\text{cps})}.\label{eq:storage_effects_further}
+\Delta_i^{(EC)} &= [(\olr_{i\bs i} - q_{ij} \olr_{j \bs i})-(\olrn_{i \bs i}-q_{ij} \olrn_{j \bs i})]+
+[(\olrn_{i \bs i}-q_{ij} \olrn_{j \bs i})-(\olrs_{i \bs i}-q_{ij} \olrs_{j \bs i})] \label{eq:storage_effects_further1}\\
+&= \Delta_i^{[E \nat C]} + \Delta_i^{[EC]}.\label{eq:storage_effects_further}
 \end{align}
-The superscript ``ata" stands for ``asymmetric tail associations'' because this component of storage effects is due to 
+The term $\Delta_i^{[E \nat C]}$ is the component of storage effects due to 
 asymmetries of tail association between $E_i$ and $C_{i \bs i}$, and between $E_j$ and $C_{j \bs i}$.
-The superscript ``cps" stands for ``correlation
-\emph{per se}'' because this component is due to $EC$ correlation \emph{per se}, after 
-removing the effects of asymmetric tail associations.
 The notation $\nat$ was chosen because this new symbol is part of the symbol $\shp$ (i.e., two of its four lines), 
 corresponding to the removal of part of the relationship between $E$ and $C$ components.
+The term $\Delta_i^{[EC]}$ is the component due to $EC$ correlation \emph{per se}, after 
+removing the effects of asymmetric tail associations. 
+The brackets in the superscripts of $\Delta_i^{[E \nat C]}$ and $\Delta_i^{[EC]}$ are intended to indicate that 
+this is a decompostion of storage effects, which were denoted by \cite{Ellner_2019} using a superscript $(EC)$,
+i.e., $\Delta_i^{(EC)}$, to indicate that storage effects are, in turn, part of a decomposition of a difference between invader and
+resident growth rates (see below).
 
@@
 these random variables. We then define a bivariate normal random variable $(e_i,c_{i \bs i})$ such that
-$\var(e_i)=\var(c_{i \bs i})=1$, and denote $\cov(e_i,c_{i \bs i})$ by $\tilde{\rho}_i$. We define 
+$\var(e_i)=\var(c_{i \bs i})=1$, and we denote $\cov(e_i,c_{i \bs i})$ by $\tilde{\rho}_i$. We define 
 $E_i^\nat = F_{E_i}^{-1} \circ \varphi(e_i)$
@@
 can be written as $(\exp(\tilde{b}_i),\exp(\tilde{b}_j))$, where $(\tilde{b}_i,\tilde{b}_j)$
-is bivariate normally distributed with mean $(\mu_i,\mu_j)$, $\var(\tilde{b}_i)=\var(\tilde{b}_j)=\sigma$,
+is bivariate-normally distributed with mean $(\mu_i,\mu_j)$, $\var(\tilde{b}_i)=\var(\tilde{b}_j)=\sigma$,
 and $\cov(\tilde{b}_i,\tilde{b}_j)=\cov(b_i,b_j)$. So $(\tilde{b}_i,\tilde{b}_j)$ has the same
@@
 $(b_i,b_j)$ itself need not be. Given these definitions, it is straightforward to compute the 
-components of (\ref{eq:storage_effects_further}) for the lottery model using the simulation 
-approach of \cite{Ellner_2016,Ellner_2019}, though in the next section we describe some tricks
-that make these computations efficient.
+component $\olrn_{i \bs i}$ of (\ref{eq:storage_effects_further}) for the lottery model using the simulation 
+approach of \cite{Ellner_2016,Ellner_2019}. Note that $E_j$ and $C_{j \bs i}$ are linearly related for the lottery model,
+so $E_j^\nat=E_j$ and $C_{j \bs i}^\nat = C_{j \bs i}$. Thus $\olr_{j \bs i}=\olrn_{j \bs i}$.
 
-The other mechanism in modern coexistence theory that apply in our modelling context (i.e., purely temporal variation)
+The other mechanisms in modern coexistence theory that apply in our modelling context (of purely temporal variation)
 make no reference to covariation between quantities, whereas asymmetries of tail associations are 
 defined in a covariation context. Therefore, storage effects are the only mechanism of coexistence
-for which asymmetries of tail association are relevant. Details are given in SI section 
-\ref{SIsect:other_mechanisms}, and we return briefly to related ideas in the Discussion.
-%***DAN: In the Discussion, I intend to talk about how possible changes in tail association 
-%patterns of env vars in nature may influence coexistence of species. It won't influence the 
-%other mechanisms, though changes in the extreme values of environmental variables, known
-%to be occurring, might.
+for which asymmetries of tail association are relevant. Mathematical details are given in SI section 
+\ref{SI-SIsect:other_mechanisms}, but we here provide a conceptual understanding. \cite{Ellner_2016} and
+\cite{Ellner_2019} provide a decomposition of $\olr_{i \bs i} -q_{ij}\olr_{i \bs i}$ which is similar to and motivated by the 
+original decomposition of \cite{Chesson_1994}, and which we use; the decomposition is how the mechanisms of 
+coexistence opperating in a given case are illuminated. The decomposition is 
+\begin{equation}
+\olr_{i \bs i} -q_{ij}\olr_{i \bs i}=\Delta_i^0+\Delta_i^E+\Delta_i^C+\Delta_i^{(E\shp C)}+\Delta_i^{(EC)},\label{eq:previous_decomp}
+\end{equation}
+where: $\Delta_i^0$ is a baseline;
+$\Delta_i^E$ is a contribution of environmental variation;
+$\Delta_i^C$ is a contribution of variation in competition;
+$\Delta_i^{(E\shp C)}$ is a contribution of having simultaneous variation in both environment and competition, but not including the effects 
+of covariation in these quantities; and
+$\Delta_i^{(EC)}$ is a contribution of $EC$ covariation -- storage effects. 
+Only storage effects involve $EC$ covariation,
+so the only mechanism of coexistence influenced by ATA is storage effects. Pairing our decomposition 
+(\ref{eq:storage_effects_further1})-(\ref{eq:storage_effects_further}) with the 
+previous MTC decomposition (\ref{eq:previous_decomp}) gives the extension of the MTC which was goal G1 of the Introduction.
 
@@
 
-%***DAN: Next steps:
-%Write to Ellner asking question about scaling factors
-%Decide what we ought to be computing 
-%Write Methods correspondingly
-%Talk to Jasmin about it - did I miss anything that we already computed?
-%Then work with Jasmin to get the code edited to compute that stuff
-%Write results
-%Once results have solidified, make a code suite version that is clean
+%Goals, for reference:
+%G1) To augment the modern theory of coexistence (MTC) to show how asymmetric tail association (ATA)
+%plays a role. The MTC decomposes a measure of coexistence into components due to each of several mechanisms. 
+%We address how ATA contributes to these components. We here examine mechanisms of coexistence that apply in 
+%models exhibiting solely temporal variation; spatial variation can be considered in future work.
+%G2) To explore using a simple model the relative magnitude of the contributions of coexistence that come
+%from ATA compared to other mechanisms.
+
+%Figs, for reference:
+%1) pedagog fig showing what tail assoc is, also later used for introing the noise we use, in methods
+%2) pedgog fig showing a simple set of sims, somehow, the only difference being ATA, showing good things 
+%happening in one panel and bad in another - cited in Intro, only serves to prove to people that ATA matters for 
+%coexistence, so they keep reading
+%3) Fig with a grid of panels, each panel plots against sigma, and the grid controls a few values of 
+%delta and a few of mu1-mu1. Plotted is all the components of coexistence, for comparison.
+%4) Theoretical fig exploring why ATA matters to r_1\1 for the lottery model and hopefully more gernerally - try to get
+%at the intuition here. See if you can edge to more generality here, or not?
 
@@
 \section{Discussion}\label{section:discussion}
+
+Need a para about multi-species models - how would things change? One difference is there are more different
+things that might have asymmetric tail association patterns. See the section "General Theory" of Ellner et al 2019 - might
+just want write a short para saying it seems like it should be possible to generalize what we've done along those lines.
+
+Might be interesting to study how tail associations of environmental variables are changing and how this might 
+alter competition, coexistence, and the maintenance of biodiversity generally. This would be a whole potential 
+research program which presumably would start with a more detailed study of how tail associations of environmental
+variables may be changing under global change.
+
+Search for comments in the rest of the text indicating things that need to be done in the Discussion
+
+Ellner et al (2016) points out that coexistence is also mediated by other factors that competition: shared enemies, 
+mutualist, facilitation etc. (these are the exact factors Ellner lists), so future work could explore 
+the importance of ATA for coexistence mediated by these factors.
+
+Something about spatial mechanisms of coexistence? We considered only mechanisms that applied in a model with
+solely temporal variation, but there are also spatial mechanisms. Need to learn about those and think about
+how to say something about how future work may be able to explore the importance of ATA in that context?
+
+Also need to talk about how this stuff will apply to data.
+
+In the Discussion, I intend to talk about how possible changes in tail association 
+patterns of env vars in nature may influence coexistence of species. It won't influence the 
+other mechanisms, though changes in the extreme values of environmental variables, known
+to be occurring, might. Should also mention this second point in the Discussion as something that merits further research.
+%though might be worth mentioning in the Discussion that events are getting more extreme. We have studied how *relationships*
+%in the extremes influence coexistence, but probably worth studying (or cite papers if it's been done) how changes
+%in the commonness and extremeness of extreme events themselves influence coexistence, and for that you'd have
+%to also look at the other mechanisms. May still need a sup mat laying out all the math.
 

### 2021-06-28 paper/SupMat.Rnw
finished a draft of the Theory section
further work on editing theory, editing and writing sup mat sections to get ready to write methods

--- 2021-06-23 paper/SupMat.Rnw
@@
 \usepackage{mathrsfs}
-%\usepackage{accents}
+\usepackage{relsize}
+\usepackage{tikz}
 
@@
 \DeclareMathOperator{\EX}{E}% expected value
+\DeclareMathOperator{\E}{E}
 \DeclareMathOperator{\VarX}{var}
+\DeclareMathOperator{\var}{var}
 \DeclareMathOperator{\CovX}{cov}
+\DeclareMathOperator{\cov}{cov}
 \DeclareMathOperator{\mean}{mean}
@@
 \newcommand{\bs}{\backslash}
+
+%attempt 3 at nat and sharp (see Paper.Rnw for attempts 1 and 2, which I decide not to use)
+\newcommand{\nat}{%
+\text{\hspace{-1.5pt}
+\begin{tikzpicture}[scale=1.8]%
+\draw (.333ex,0) -- (.333ex,1ex);%
+\draw (.666ex,0) -- (.666ex,1ex);
+\end{tikzpicture}%
+}}
+\newcommand{\shp}{%
+\text{\hspace{-1.5pt}
+\begin{tikzpicture}[scale=1.8]%
+\draw (0,.333ex) -- (1ex,.333ex);%
+\draw (0,.666ex) -- (1ex,.666ex);%
+\draw (.333ex,0) -- (.333ex,1ex);%
+\draw (.666ex,0) -- (.666ex,1ex);
+\end{tikzpicture}%
+}}
+\newcommand{\test}{%
+\text{
+\begin{tikzpicture}[scale=1.8]%
+\draw (0,0) -- (1ex,0ex);
+\draw (0,0) -- (0ex,1ex);
+\draw (0,1ex) -- (1ex,1ex);
+\draw (1ex,0) -- (1ex,1ex);
+\draw (0,.333ex) -- (2ex,.333ex);%
+\draw (0,.666ex) -- (1ex,.666ex);%
+\draw (.333ex,0) -- (.333ex,1ex);%
+\draw (.666ex,0) -- (.666ex,1ex);
+\end{tikzpicture}%
+}}
+
+\newcommand{\olr}{\overline{r}}
+\newcommand{\olrs}{\overline{r}^{\shp}}
+\newcommand{\olrn}{\overline{r}^{\nat}}
+\newcommand{\olep}{\overline{\epsilon}}
 
@@
 
-\section{Tail association is not relevant to other mechanisms of coexistence for purely temporal variation}\label{SIsect:other_mechanisms}
+\section{Mechanisms other than storage effects}\label{SIsect:other_mechanisms}
 
-%***DAN: Sections below this point are from the earlier reasoning and may have to be edited and placed in other
+We review the decompositions of \cite{Ellner_2016} and \cite{Ellner_2019}, which are similar to and motivated by the 
+original decompositions of \cite{Chesson_1994}, and then we include our extension.
+
+We begin by decomposing 
+the invasion growth rate,
+\begin{equation}
+\olr_{i \bs i}=\E[r_i(E_i,C_{\i \bs i})].
+\end{equation}
+The invasion growth rate with both $E_i$ and $C_{i \bs i}$ constant at their means is 
+\begin{align}
+\epsilon_i^0 = r_i(\overline{E}_i,\overline{C}_{i \bs i}).
+\end{align}
+The additional contribution of variation in $E_i$ to the invasion growth rate is then
+\begin{equation}
+\olep_i^E = \E[r_i(E_i,\overline{C}_{i \bs i})]-\epsilon_i^0,
+\end{equation}
+whereas the additional contribution of variation in $C_{i \bs i}$ is 
+\begin{equation}
+\olep_i^C=\E[r_i(\overline{E}_i,C_{i \bs i})]-\epsilon_i^0.
+\end{equation}
+We then define
+\begin{align}
+\olep_i^{EC} &= \E[r_i(E_i,C_{i \bs i})]-\olep_i^E-\olep_i^C-\epsilon_i^0 \\
+&= \olr_{i \bs i} - \olep_i^E - \olep_i^C - \epsilon_i^0.
+\end{align}
+Thus we have the primary decomposition 
+\begin{equation}
+\olr_{i \bs i} = \epsilon_i^0 + \olep_i^E + \olep_i^C + \olep_i^{EC}.
+\end{equation}
+But some of $\olep_i^{EC}$ comes from the fact that $E_i$ and $C_{i \bs i}$ are
+both varying, and some comes from covariation in these two quantities. So we next separate these
+contributions in a secondary decomposition of $\olep_i^{EC}$.
+
+Let
+\begin{equation}
+\olep_i^{(E\shp C)}=\E[r_i(E_i^\shp,C^\shp_{i \bs i})]-\olep_i^E-\olep_i^C-\epsilon_i^0
+\end{equation}
+and let
+\begin{equation}
+\olep_i^{(EC)} = \olep_i^{EC}-\olep_i^{(E\shp C)},
+\end{equation}
+so
+\begin{equation}
+\olep_i^{(EC)}=\E[r_i(E_i,C_{i \bs i})]-\E[r_i(E_i^\shp,C_{i \bs i}^\shp)].
+\end{equation}
+The parentheses in the superscripts of $\olep_i^{(EC)}$ and $\olep_i^{(E\shp C)}$
+indicate that we are talking about a term in the secondary decomposition, i.e., in the decomposition of 
+$\olep_i^{EC}$. We now have
+\begin{align}
+\olr_{i \bs i} &= \epsilon_i^0 + \olep_i^E + \olep_i^C + \olep_i^{EC} \\
+\olep_i^{EC} &= \olep_i^{(E\shp C)}+\olep_i^{(EC)},
+\end{align}
+i.e.,
+\begin{equation}
+\olr_{i \bs i} = \epsilon_i^0 + \olep_i^E + \olep_i^C +\olep_i^{(E\shp C)}+\olep_i^{(EC)}. 
+\end{equation}
+The term $\olep_i^{(EC)}$ is due to covariation between $E_i$ and $C_{i \bs i}$, whereas the term 
+$\olep_i^{(E\shp C)}$ is due to variation \emph{per se} in both $E_i$ and $C_{i \bs i}$, after the effects of 
+covariation itself are removed. 
+The term $\olep_i^{(EC)}$ relates to storage effects, as we will see.
+
+As in the main text, we can now define the tertiary decomposition
+\begin{align}
+\olep_i^{(EC)} &= \E[r_i(E_i,C_{i \bs i})]-\E[r_i(E_i^\shp,C_{i \bs i}^\shp)] \\
+&= \left[ \E[r_i(E_i,C_{i \bs i})]-\E[r_i(E_i^\nat,C_{i \bs i}^\nat)] \right]+\left[ \E[r_i(E_i^\nat,C_{i \bs i}^\nat)]-\E[r_i(E_i^\shp,C_{i \bs i}^\shp)] \right] \\
+&= \olep_i^{[EC]}+\olep_i^{[E \nat C]}.
+\end{align}
+The brackets in the superscripts of $\olep_i^{[EC]}$ and $\olep_i^{[E \nat C]}$ indicate that we are talking about a term in the
+tertiary decomposition, i.e., in the decomposition of $\olep_i^{(EC)}$. 
+The term $\olep_i^{[E \nat C]}$ quantifies the influence of asymmetric tail associations of $E_i$ and $C_{i \bs i}$ on the invasion growth rate,
+and the the term $\olep_i^{[EC]}$ corresponds to $EC$ correlations \emph{per se}, after the removal of asymmetries of tail association.
+The symbol $\nat$ denotes the removal of asymmetries of tail association. We therefore have 
+\begin{equation}
+\olr_{i \bs i} = \epsilon_i^0 + \olep_i^E + \olep_i^C + \olep_i^{(E\shp C)}+\olep_i^{[EC]}+\olep_i^{[E \nat C]},
+\end{equation}
+our overall decompsition of $r_{i \bs i}$.
+The derivation also applies to provide 
+\begin{equation}
+\olr_{j \bs i} = \epsilon_j^0 + \olep_j^E + \olep_j^C + \olep_j^{(E\shp C)}+\olep_j^{[EC]}+\olep_j^{[E \nat C]},
+\end{equation}
+an analogous decomposition of $\olr_{j \bs i}$.
+
+Combining the decompositions of $\olr_{i \bs i}$ and $\olr_{j \bs i}$ and assuming without loss of generality that $\mu_i \leq \mu_j$ so that
+species i is the weaker competitor, we have
+\begin{equation}
+\olr_{i \bs i}-q_{ij}\olr_{j \bs i}=\Delta_i^0+\Delta_i^E+\Delta_i^C+\Delta_i^{(E\shp C)}+\Delta_i^{[EC]}+\Delta_i^{[E\nat C]},
+\end{equation}
+where $\Delta_i^0=\epsilon_i^0- q_{ij} \epsilon_j^0$;
+$\Delta_i^E=\olep_i^E -q_{ij}\olep_j^E$ is the contribution to coexistence of environmental variation;
+$\Delta_i^C=\olep_i^C -q_{ij}\olep_j^C$ is the contribution to coexistence of variation in competitive pressure;
+$\Delta_i^{(E\shp C)}=\olep_i^{(E\shp C)} - q_{ij}\olep_j^{(E\shp C)}$ is the contribution to coexistence of 
+variation \emph{per se} in both environment and competition, not including the effects of covariation of these quantities;
+$\Delta_i^{[EC]}=\olep_i^{[EC]} - q_{ij} \olep_j^{[EC]}$ is the contribution to coexistence of $EC$ covariation \emph{per se},
+not including the effects of asymmetric tail associations between environment and competition;
+and $\Delta_i^{[E\nat C]}=\olep_i^{[E \nat C]}-q_{ij} \olep_j^{[E \nat C]}$ is the contribution to coexistence of 
+asymmetric tail assocations. The subscript $i$ on all these newly defined quantities $\Delta_i$ refers to species $i$.
+The notation $\Delta I_i$ from the main text equals $\Delta_i^{(EC)}=\Delta_i^{[EC]}+\Delta_i^{[E\nat C]}$,
+and $\Delta I_i^{(\text{ata})}$ from the main text is the same as $\Delta_i^{[E\nat C]}$ and $\Delta I_i^{(\text{cps})}$ 
+from the main text is $\Delta_i^{[EC]}$ (see (\ref{MT-eq:storage_effects_further1}) and (\ref{MT-eq:storage_effects_further}) from the main text).
+
+%***DAN: Next go through everything for the lottery model
+%***DAN: Then argue that everything depends only on mu1-mu2
+
+\vspace{2in}
+
+\noindent By substituting the definitions,
+
+\begin{align}
+\bar r_i &=\EX(\ln(1-\delta + \frac{B_i}{B_2/ \delta})) \\
+&= \EX(\ln(1-\delta +\delta \frac{B_i}{B_2})) 
+\end{align}
+
+The null growth rate of species $i$ is computed by setting both E and C at their means.
+
+\begin{align}
+\epsilon_i^0 &=\ln(1-\delta + \frac{\bar E_i}{\bar C_i}) \\
+&=\ln(1-\delta + \delta \frac{\bar B_i}{\bar B_2}) 
+\end{align}
+
+The contribution of variance E in to the IGR is computed by setting C to its mean and letting E vary subtracted by the null.
+
+\begin{align}
+\epsilon_i^E &=\ln(1-\delta + \frac{E_i}{\bar C_i}) -  \epsilon_i^0 \\
+&=\ln(1-\delta + \delta \frac{B_i}{\bar B_2}) -  \epsilon_i^0 
+\end{align}
+
+\begin{equation}
+\bar \epsilon_i^E=\EX(\epsilon_i^E) \label{Evar_mean}
+\end{equation}
+
+Similarly, the contribution of variance C in to the IGR is computed by setting E to its mean and letting C vary subtracted by the null.
+
+\begin{align}
+\epsilon_i^C &=\ln(1-\delta + \frac{\bar E_i}{C_i}) -  \epsilon_i^0 \\
+ &=\ln(1-\delta + \delta \frac{\bar B_i}{ B_2}) -  \epsilon_i^0 
+\end{align}
+
+\begin{equation}
+\bar \epsilon_i^C=\EX(\epsilon_i^C) \label{Cvar_mean}
+\end{equation}
+
+The contribution of both E and C varying to the IGR is computed by letting both variables vary subtracted by the null and the contributions of only one variable varying.
+
+\begin{equation}
+\epsilon_i^{EC}=\ln(1-\delta + \delta \frac{B_i}{B_j}) -  [\epsilon_i^0 + \epsilon_i^E + \epsilon_i^C] \label{ECvar}
+\end{equation}
+
+\begin{equation}
+\bar \epsilon_i^{EC}=\EX(\ln(1-\delta + \delta \frac{B_i}{B_j})) -[\epsilon_i^0 +\bar \epsilon_i^E + \bar \epsilon_i^C] \label{ECvar_mean}
+\end{equation}
+
+So,
+
+\begin{equation}
+\bar r_i=\epsilon_i^0 +\bar \epsilon_i^E + \bar \epsilon_i^C + \bar \epsilon_i^{EC}. \label{rbari_decomp}
+\end{equation}
+
+But some of $\bar \epsilon_i^{EC}$ comes from covariance between E and C and some comes from both varying independently.
+
+\begin{align}
+\bar \epsilon_i^{(E\#C)} &=\EX(\ln(1-\delta + \frac{E_i}{C_i^{\#}})) -[\epsilon_i^0 +\bar \epsilon_i^E + \bar \epsilon_i^C] \\
+ &=\EX(\ln(1-\delta + \delta \frac{B_i}{B_2^{\#}})) -[\epsilon_i^0 +\bar \epsilon_i^E + \bar \epsilon_i^C] 
+\end{align}
+
+\begin{align}
+\bar \epsilon_i^{(EC)} &=\bar \epsilon_i^{EC}-\bar \epsilon_i^{(E\#C)}\\
+&= \EX(\ln(1-\delta + \frac{E_i}{C_i}))-[\epsilon_i^0 +\bar \epsilon_i^E + \bar \epsilon_i^C]\\
+&- [\EX(\ln(1-\delta + \frac{E_i}{C_i^{\#}})) -[\epsilon_i^0 +\bar \epsilon_i^E + \bar \epsilon_i^C]] \\
+&= \EX(\ln(1-\delta + \delta \frac{B_i}{B_2}))-\epsilon_i^0 -\bar \epsilon_i^E - \bar \epsilon_i^C\\
+&- \EX(\ln(1-\delta + \delta \frac{B_i}{B_2^{\#}})) +\epsilon_i^0 +\bar \epsilon_i^E + \bar \epsilon_i^C \\
+&= \EX(\ln(1-\delta + \delta \frac{B_i}{B_2}))-\EX(\ln(1-\delta + \delta \frac{B_i}{B_2^{\#}}))\\
+\end{align}
+
+%***DAN: Sections below this point are from the earlier version/reasoning and may have to be edited and placed in other
 %locations. This comment written 2021 06 22
@@
 \section{The Model}
+
+%***DAN: Probably delete this section, since the model is now elaborated in the main text
 
@@
 \section{Invasion growth rate of species 1, and storage effects} \label{sect:IGR}
+
+%***DAN: This and probably all or most of the following sections will have to be edited, not least to make the
+%notation consistent with the notation now used in the Theory section of the main text
+%***DAN: At the end of Theory, I mention that in methods the tricks for computational efficiency are
+%elaborated, but actually most of those will be in a sup mat section, the material for which will be 
+%an edited version of a lot of what is in this section.
 
@@
 
-\section{E and C definitions}
-
-The lottery model is given by
-
-\begin{equation}
-N_i(t+1)=(1-\delta)N(t)+\delta N \frac{B_i(t)N_i(t)}{B_1(t)N_1(t)+B_2(t)N_2(t)}.  
-\end{equation}
-
-REF also defines $E_i = B_i$ and $C_1=C_2=\frac{B_1 N_1 + B_2 N_2}{\delta N}$.
-Then in the invader-resident scenario which defines the invader to be essentially zero and $N_2 = N$, $C_1 = C_2 = \frac{B_2}{\delta}$.
-The growth rate equation relates the two variables of $E$ and $C$ by
-
-\begin{equation}
-\bar r_i=\EX(\ln(1-\delta + \frac{E_i}{C_i})).\label{rbarEC}
-\end{equation}
-
-\noindent By substituting the definitions,
-
-\begin{align}
-\bar r_i &=\EX(\ln(1-\delta + \frac{B_i}{B_2/ \delta})) \\
-&= \EX(\ln(1-\delta +\delta \frac{B_i}{B_2})) 
-\end{align}
-
-The null growth rate of species $i$ is computed by setting both E and C at their means.
-
-\begin{align}
-\epsilon_i^0 &=\ln(1-\delta + \frac{\bar E_i}{\bar C_i}) \\
-&=\ln(1-\delta + \delta \frac{\bar B_i}{\bar B_2}) 
-\end{align}
-
-The contribution of variance E in to the IGR is computed by setting C to its mean and letting E vary subtracted by the null.
-
-\begin{align}
-\epsilon_i^E &=\ln(1-\delta + \frac{E_i}{\bar C_i}) -  \epsilon_i^0 \\
-&=\ln(1-\delta + \delta \frac{B_i}{\bar B_2}) -  \epsilon_i^0 
-\end{align}
-
-\begin{equation}
-\bar \epsilon_i^E=\EX(\epsilon_i^E) \label{Evar_mean}
-\end{equation}
-
-Similarly, the contribution of variance C in to the IGR is computed by setting E to its mean and letting C vary subtracted by the null.
-
-\begin{align}
-\epsilon_i^C &=\ln(1-\delta + \frac{\bar E_i}{C_i}) -  \epsilon_i^0 \\
- &=\ln(1-\delta + \delta \frac{\bar B_i}{ B_2}) -  \epsilon_i^0 
-\end{align}
-
-\begin{equation}
-\bar \epsilon_i^C=\EX(\epsilon_i^C) \label{Cvar_mean}
-\end{equation}
-
-The contribution of both E and C varying to the IGR is computed by letting both variables vary subtracted by the null and the contributions of only one variable varying.
-
-\begin{equation}
-\epsilon_i^{EC}=\ln(1-\delta + \delta \frac{B_i}{B_j}) -  [\epsilon_i^0 + \epsilon_i^E + \epsilon_i^C] \label{ECvar}
-\end{equation}
-
-\begin{equation}
-\bar \epsilon_i^{EC}=\EX(\ln(1-\delta + \delta \frac{B_i}{B_j})) -[\epsilon_i^0 +\bar \epsilon_i^E + \bar \epsilon_i^C] \label{ECvar_mean}
-\end{equation}
-
-So,
-
-\begin{equation}
-\bar r_i=\epsilon_i^0 +\bar \epsilon_i^E + \bar \epsilon_i^C + \bar \epsilon_i^{EC}. \label{rbari_decomp}
-\end{equation}
-
-But some of $\bar \epsilon_i^{EC}$ comes from covariance between E and C and some comes from both varying independently.
-
-\begin{align}
-\bar \epsilon_i^{(E\#C)} &=\EX(\ln(1-\delta + \frac{E_i}{C_i^{\#}})) -[\epsilon_i^0 +\bar \epsilon_i^E + \bar \epsilon_i^C] \\
- &=\EX(\ln(1-\delta + \delta \frac{B_i}{B_2^{\#}})) -[\epsilon_i^0 +\bar \epsilon_i^E + \bar \epsilon_i^C] 
-\end{align}
-
-\begin{align}
-\bar \epsilon_i^{(EC)} &=\bar \epsilon_i^{EC}-\bar \epsilon_i^{(E\#C)}\\
-&= \EX(\ln(1-\delta + \frac{E_i}{C_i}))-[\epsilon_i^0 +\bar \epsilon_i^E + \bar \epsilon_i^C]\\
-&- [\EX(\ln(1-\delta + \frac{E_i}{C_i^{\#}})) -[\epsilon_i^0 +\bar \epsilon_i^E + \bar \epsilon_i^C]] \\
-&= \EX(\ln(1-\delta + \delta \frac{B_i}{B_2}))-\epsilon_i^0 -\bar \epsilon_i^E - \bar \epsilon_i^C\\
-&- \EX(\ln(1-\delta + \delta \frac{B_i}{B_2^{\#}})) +\epsilon_i^0 +\bar \epsilon_i^E + \bar \epsilon_i^C \\
-&= \EX(\ln(1-\delta + \delta \frac{B_i}{B_2}))-\EX(\ln(1-\delta + \delta \frac{B_i}{B_2^{\#}}))\\
-\end{align}
-
-\section{Decomposition}
+\section{Decomposition}\label{Decomposition}
 

### 2021-06-28 paper/SupMat.Rnw

@@
 \usepackage{tikz}
+\usepackage{accents}
 
@@
 
-\author[a]{Jasmin Albert}
+\author[a]{Pimsupa Jasmin Albert}
 \author[a,b]{Daniel C. Reuman}
@@
 \begin{document}
+\SweaveOpts{concordance=TRUE}
 %\SweaveOpts{concordance=TRUE}
@@
 The descrition of $\bar \epsilon^{(\ddot {EC})}$ is given in section 3 and equations 9 and 21-23.
+
+\begin{equation}
+\bar \epsilon^{( {EC_0})} = \bar r_1 - \bar r_1^{\#} + \bar r_2^{\#},
+\end{equation}
+
+\noindent where $ \bar r_1, \bar r_1^{\#}$ and $\bar r_2^{\#}$ are in the symmetric case. In our model,
+
+\begin{equation}
+\bar \epsilon^{( {EC_0})} = \EX(\ln[1-\delta+\delta \exp(u_1)]) - \EX(\ln[1-\delta+\delta \exp(u_1^{\#})]) + \EX(\ln[1-\delta+\delta \exp(u_2^{\#})]),
+\end{equation}
+
+\noindent where $u_1 \sim N(\mu_1-\mu_2, 2\sigma^2-2\rho)$, $u_1^{\#} \sim N(\mu_1-\mu_2, 2\sigma^2)$, and $u_2^{\#} \sim N(0, 2\sigma^2)$
+
+\noindent *Note that the contribution of asymmetry to storage effects is not the same as the storage effect that we computed above for asymmetriv cases.
+
+In generalized terms, this further decomposition could be applied to test for contributions of asymmetry between any two variables.
+Here we tested the asymmetry between $b_1$ and $b_2$.
+Why did we not look at E and C?
+Do they have to be normally distrubuted?
+
+This break down would only work if the variables are normally distributed. Thus one can convert their distribution to normal to impliment this decomposition.
 

### 2021-06-29 paper/SupMat.Rnw
fixed the use of \hat in several places to make the latexc compile

@@
 \begin{document}
-\SweaveOpts{concordance=TRUE}
 %\SweaveOpts{concordance=TRUE}
@@
 \begin{equation}
-\olr_{i \bs i}=\E[r_i(E_i,C_{\i \bs i})].
+\olr_{i \bs i}=\E[r_i(E_i,C_{i \bs i})].
 \end{equation}
@@
 
-\item Estimate $\bar r_1$ for the left-tail associated case as $\hat \bar r_1 = \mean_i (\ln [1-\delta+\delta\exp(b_{l1}^{(i)}-b_{l2}^{(i)})])$ where $b_{l1}^{(i)} = \sigma \tilde b_{l1}^{(i)} + \mu_1$ and $b_{l2}^{(i)} = \sigma \tilde b_{l2}^{(i)} + \mu_2$.
-The standard error of this estimate is $\se (\hat \bar r_1) =\frac{\sd _i (\ln[1-\delta +\delta \exp(b_{l1}^{(i)}-b_{l2}^{(i)})])}{\sqrt M}$, where $\sd _i(\cdot)$ is standard deviation.
+\item Estimate $\bar r_1$ for the left-tail associated case as $\hat{\bar r_1} = \mean_i (\ln [1-\delta+\delta\exp(b_{l1}^{(i)}-b_{l2}^{(i)})])$ where $b_{l1}^{(i)} = \sigma \tilde b_{l1}^{(i)} + \mu_1$ and $b_{l2}^{(i)} = \sigma \tilde b_{l2}^{(i)} + \mu_2$.
+The standard error of this estimate is $\se (\hat{\bar r_1}) =\frac{\sd _i (\ln[1-\delta +\delta \exp(b_{l1}^{(i)}-b_{l2}^{(i)})])}{\sqrt M}$, where $\sd _i(\cdot)$ is standard deviation.
 
-\item Estiate $\bar r_1$ for the left-tail associated case as $\hat \bar r_1 = \mean_i (\ln [1-\delta+\delta\exp(b_{r1}^{(i)}-b_{r2}^{(i)})])$ where $b_{r1}^{(i)} = \sigma \tilde b_{r1}^{(i)} + \mu_1$ and $b_{r2}^{(i)} = \sigma \tilde b_{r2}^{(i)} + \mu_2$.
-The standard error of this estimate is $\se (\hat \bar r_1) =\frac{\sd _i (\ln[1-\delta +\delta \exp(b_{r1}^{(i)}-b_{r2}^{(i)})])}{\sqrt M}$.
+\item Estiate $\bar r_1$ for the left-tail associated case as $\hat{\bar r_1} = \mean_i (\ln [1-\delta+\delta\exp(b_{r1}^{(i)}-b_{r2}^{(i)})])$ where $b_{r1}^{(i)} = \sigma \tilde b_{r1}^{(i)} + \mu_1$ and $b_{r2}^{(i)} = \sigma \tilde b_{r2}^{(i)} + \mu_2$.
+The standard error of this estimate is $\se (\hat{\bar r_1}) =\frac{\sd _i (\ln[1-\delta +\delta \exp(b_{r1}^{(i)}-b_{r2}^{(i)})])}{\sqrt M}$.
 
-\item Estimate $\bar r_1$ in the symmetric tail association case as $\hat \bar r_1=\mean_i(\ln[1-\delta+\delta \exp(u_i)])$, where $u_i = \sqrt{(2\sigma^2-2\rho)}\tilde u_i + \mu_1 - \mu_2$.
-The standard error of this estimation is $\se(\hat \bar r_1) = \frac {\sd_i(\ln[1-\delta+\delta \exp(u_i)])}{\sqrt M}$.
+\item Estimate $\bar r_1$ in the symmetric tail association case as $\hat{\bar r_1}=\mean_i(\ln[1-\delta+\delta \exp(u_i)])$, where $u_i = \sqrt{(2\sigma^2-2\rho)}\tilde u_i + \mu_1 - \mu_2$.
+The standard error of this estimation is $\se(\hat{\bar r_1}) = \frac {\sd_i(\ln[1-\delta+\delta \exp(u_i)])}{\sqrt M}$.
 
-\item Estimate $\bar r_1^{\#}$ (for all three tail-assocaition cases) as $\hat \bar r_1^{\#} = \mean_i(\ln[1-\delta+\delta \exp(u_i)])$, where $u_i = \sqrt{2\sigma^2}\tilde u_i + \mu_1 - \mu_2$.
-The standard error is $\se(\hat \bar r_1^{\#}) = \sd_i \frac{ln[1-\delta+\delta \exp(u_i)]}{\sqrt M}$.
+\item Estimate $\bar r_1^{\#}$ (for all three tail-assocaition cases) as $\hat{\bar r_1^{\#}} = \mean_i(\ln[1-\delta+\delta \exp(u_i)])$, where $u_i = \sqrt{2\sigma^2}\tilde u_i + \mu_1 - \mu_2$.
+The standard error is $\se(\hat{\bar r_1^{\#}}) = \sd_i \frac{ln[1-\delta+\delta \exp(u_i)]}{\sqrt M}$.
 
-\item Estimate $\bar r_2^{\#}$ (for all three tail-assocaition cases) as $\hat \bar r_2^{\#} = \mean_i(\ln[1-\delta+\delta \exp(u_i)])$, where $u_i = \sqrt{2\sigma^2}\tilde u_i$.
-The standard error is $\se(\hat \bar r_2^{\#}) = \sd_i \frac{ln[1-\delta+\delta \exp(u_i)]}{\sqrt M}$.
+\item Estimate $\bar r_2^{\#}$ (for all three tail-assocaition cases) as $\hat{\bar r_2^{\#}} = \mean_i(\ln[1-\delta+\delta \exp(u_i)])$, where $u_i = \sqrt{2\sigma^2}\tilde u_i$.
+The standard error is $\se(\hat{\bar r_2^{\#}}) = \sd_i \frac{ln[1-\delta+\delta \exp(u_i)]}{\sqrt M}$.
 
-\item $\hat {\Delta I} = \hat \bar r_1 - \hat \bar r_1^{\#} + \hat \bar r_2^{\#}$, estimated separately for the left-tail associated, right-tail associated and symmetric noise cases.
-We obtained conservative standard error estimates by adding the standard error of the estimates of $\hat \bar r_1, \hat \bar r_1^{\#}$, and $\hat \bar r_2^{\#}$.
+\item $\hat {\Delta I} = \hat{\bar r_1} - \hat{\bar r_1^{\#}} + \hat{\bar r_2^{\#}}$, estimated separately for the left-tail associated, right-tail associated and symmetric noise cases.
+We obtained conservative standard error estimates by adding the standard error of the estimates of $\hat{\bar r_1}, \hat{\bar r_1^{\#}}$, and $\hat{\bar r_2^{\#}}$.
 

### 2021-06-29 paper/Paper.Rnw
made additions to sup mat to support Methods - still working on it though

@@
 %Relationships between extreme events and their influence on competition and coexistence
+%Coexistence and consequences of extreme environmental relationships
+%Coexistence and consequences of environmental relationships in their extremes
+%Coexistence and implications of asymmetric environmental relationships
+%Extreme [climatic] events and their implications for biodiversity
+%Deductions to coexistence, extreme events, and asymmetric relationships
+%Extreme relationships (relationships in their extremes) and their deductions ([negative] contributions) to coexistence
 
-\author[a]{Jasmin Albert}
+\author[a]{Pimsupa Jasmin Albert}
 \author[a,b]{Daniel C. Reuman}
-%***DAN: Jasmin, you should choose one way your name will appear, exactly, on all your publications,
-%so everyone knows its you. For instance, I always appear as Daniel C. Reuman, not Daniel Reuman, not
-%Dan Reuman, etc. This makes it possible to find all my papers by searching. So please decide what
-%name to use, Jasmin Albert or Pimsupa Albert or P. Jasmin Albert or whatever. Also keep in mind
-%that because "Albert" is a common name you might want something less common for the rest of it 
-%so that your papers are not confused with anyone elses, though of course I emphasize that how you
-%are going to be known it totally up to you, I'm just recommending you think it through now so you
-%can be consistent. I recognize that this all may be moot if you one day change your name and want to
-%be known by the new name on papers after that point, but anyway in case you don't ever change your 
-%name you might as well be consistent now, at least. So please change your name above to whatever
-%you want to be officially known as on scientific papers. Same change in the sup mat, please.
 
@@
 
-Jasmin, I think the below makes a pretty good start at tryinbg to nail down a good way to express our goals
-in the last para of the Intro, so please incorporate these into your last para. If you have a better way
-to phrase these, please put both so we can compare at some point when we meet.
+%Jasmin, I think the below makes a pretty good start at tryinbg to nail down a good way to express our goals
+%in the last para of the Intro, so please incorporate these into your last para. If you have a better way
+%to phrase these, please put both so we can compare at some point when we meet.
 G1) To augment the modern theory of coexistence (MTC) to show how asymmetric tail association (ATA)
@@
 models exhibiting solely temporal variation; spatial variation can be considered in future work.
-G2) To explore using a simple model the relative magnitude of the contributions of coexistence that come
+G2) To explore using a simple model the relative magnitudes of the contributions to coexistence that come
 from ATA compared to other mechanisms.
@@
 
-In this section we pursue goal G1 from the Introduction: to augment the MTC to show how ATA
-plays a role in species coexistence.
+We here pursue goal G1 from the Introduction, to augment the MTC to show how ATA
+plays a role.
 %***DAN: make sure you define MTC and ATA once (probably in the Intro) and then use the abbreviations onle after that.
-For pedgogical clarity, we develop each step of the theory in general and for a simple classical model 
-(the lottery model), as an example.
+For pedgogical clarity, we develop each step of the theory in general and for a simple example model, the lottery model.
 
@@
 \end{equation}
-for $i=1,2$. Here, $\delta$ is a mortality rate, and $B_i(t)$ is the fecundiity of species $i$ at time $t$. 
+for $i=1,2$. Here, $\delta$ is a mortality rate, and $B_i(t)$ is the fecundity of species $i$ at time $t$. 
 The model postulates that individuals die at rate $\delta$ at each time step, and are replaced by juveniles 
@@
 the environment. Competition, $C_i(t)$, in the lottery model is the number of new offspring
-divided by the number of open sites, $C_i(t)=( B_1(t)N_1(t)+B_2(t)N_2(t) )/\delta N$, which
+divided by the number of open sites, $C_i(t)=( B_1(t)N_1(t)+B_2(t)N_2(t) )/(\delta N$), which
 does not depend on $i$. Competition in the lottery model occurs at the stage of juveniles occupying open sites. 
@@
 %***DAN: This statement also needs adjusting if you go with q_ij=1
-So $\Delta_i^{(EC)}$ for the lottery model can be computed by using simulations \citep{Ellner_2016}
+Thus $\Delta_i^{(EC)}$ for the lottery model can be computed by using simulations \citep{Ellner_2016}
 to very accurately approximate $\olr_{i \bs i}$, $\olrs_{i \bs i}$,
@@
 The notation $\nat$ was chosen because this new symbol is part of the symbol $\shp$ (i.e., two of its four lines), 
-corresponding to the removal of part of the relationship between $E$ and $C$ components.
-The term $\Delta_i^{[EC]}$ is the component due to $EC$ correlation \emph{per se}, after 
-removing the effects of asymmetric tail associations. 
+corresponding to the removal of part of the relationship between $E$ and $C$ components (namely, ATA).
+The term $\Delta_i^{[EC]}$ is the component of storage effects due to $EC$ correlation \emph{per se}, remaining after 
+having removed the effects of asymmetric tail associations.
+%The brackets in the superscripts of $\Delta_i^{[E \nat C]}$ and $\Delta_i^{[EC]}$ are intended to indicate that 
+%this is a decompostion of storage effects, which were denoted by \cite{Ellner_2019} using a superscript $(EC)$,
+%i.e., $\Delta_i^{(EC)}$, to indicate that storage effects are, in turn, part of a decomposition of a difference between invader and
+%resident growth rates (see below).  
 The brackets in the superscripts of $\Delta_i^{[E \nat C]}$ and $\Delta_i^{[EC]}$ are intended to indicate that 
-this is a decompostion of storage effects, which were denoted by \cite{Ellner_2019} using a superscript $(EC)$,
-i.e., $\Delta_i^{(EC)}$, to indicate that storage effects are, in turn, part of a decomposition of a difference between invader and
-resident growth rates (see below).
+(\ref{eq:storage_effects_further}) is a decompostion within a decomposition, i.e., (\ref{eq:storage_effects_further}) is a decomposition
+of storage effects, which are themselves a term within the decomposition provided by the MTC \citep{Ellner_2019} -- see below for details.
 
 How are asymmetries of tail association removed while leaving $E$ and $C$ marginal distributions and overall 
-correlation unchanged? We assume that the cummulative distribution functions (cdfs) of $E_i$ and 
+correlation unchanged? We assume, for simplicity, that the cummulative distribution functions (CDFs) of $E_i$ and 
 $C_{i \bs i}$, which we denote $F_{E_i}$ and $F_{C_{i \bs i}}$, respectively, are strictly monotonic and continuous. 
-Letting $\varphi$ be the cdf of a standard normal distribution, we consider the random variables
+Letting $\varphi$ be the CDF of a standard normal distribution, we consider the random variables
 $\varphi^{-1} \circ F_{E_i} (E_i)$ and $\varphi^{-1} \circ F_{C_{i \bs i}}(C_{i \bs i})$, which are standard normally
@@
 these random variables. We then define a bivariate normal random variable $(e_i,c_{i \bs i})$ such that
-$\var(e_i)=\var(c_{i \bs i})=1$, and we denote $\cov(e_i,c_{i \bs i})$ by $\tilde{\rho}_i$. We define 
+$\var(e_i)=\var(c_{i \bs i})=1$, and we denote $\cov(e_i,c_{i \bs i})$ by $\ddot{\rho}_i$. We define 
 $E_i^\nat = F_{E_i}^{-1} \circ \varphi(e_i)$
-and $C_{i \bs i}^\nat=F_{C_{i \bs i}}^{-1} \circ \varphi(c_{i \bs i})$. The value of $\tilde{\rho}_i$ is then determined 
+and $C_{i \bs i}^\nat=F_{C_{i \bs i}}^{-1} \circ \varphi(c_{i \bs i})$. The value of $\ddot{\rho}_i$ is then determined 
 in one of 
-four alternative ways: $\tilde{\rho}_i = \rho_i$; or so that the Pearson, Spearman or Kendall correlation of 
+four alternative ways: $\ddot{\rho}_i = \rho_i$; or so that the Pearson, Spearman or Kendall correlation of 
 $E_i^\nat$ and $C_{i \bs i}^\nat$ matches that of $E_i$ and $C_{i \bs i}$. The choice of bivariate normal 
@@
 we replaced the copula of $(E_i,C_{i \bs i})$ with a normal copula to form $(E_i^\nat,C_{i \bs i}^\nat)$.
-The choice of $\tilde{\rho}_i$ ensures that $E_i^\nat$ and $C_{i \bs i}^\nat$ are correlated to the same extent as 
+The choice of $\ddot{\rho}_i$ ensures that $E_i^\nat$ and $C_{i \bs i}^\nat$ are correlated to the same extent as 
 $E_i$ and $C_{i \bs i}$, according to whichever definition of correlation is considered preferable. Thus $E_i^\nat$ and 
@@
 Pearson correlation of $b_i$ and $b_j$. For the lottery model, it seems natural to make the choice
-$\tilde{\rho}_i=\rho_i$, which straightforwardly implies that $(E_i^\nat,C_{i \bs i}^\nat)$
-can be written as $(\exp(\tilde{b}_i),\exp(\tilde{b}_j))$, where $(\tilde{b}_i,\tilde{b}_j)$
-is bivariate-normally distributed with mean $(\mu_i,\mu_j)$, $\var(\tilde{b}_i)=\var(\tilde{b}_j)=\sigma$,
-and $\cov(\tilde{b}_i,\tilde{b}_j)=\cov(b_i,b_j)$. So $(\tilde{b}_i,\tilde{b}_j)$ has the same
+$\ddot{\rho}_i=\rho_i$, which straightforwardly implies that $(E_i^\nat,C_{i \bs i}^\nat)$
+can be written as $(\exp(\ddot{b}_i),\exp(\ddot{b}_j))$, where $(\ddot{b}_i,\ddot{b}_j)$
+is bivariate-normally distributed with mean $(\mu_i,\mu_j)$, $\var(\ddot{b}_i)=\var(\ddot{b}_j)=\sigma$,
+and $\cov(\ddot{b}_i,\ddot{b}_j)=\cov(b_i,b_j)$. So $(\ddot{b}_i,\ddot{b}_j)$ has the same
 marginal distributions and covariance matrix as $(b_i,b_j)$, but it is bivariate normal, whereas
 $(b_i,b_j)$ itself need not be. Given these definitions, it is straightforward to compute the 
-component $\olrn_{i \bs i}$ of (\ref{eq:storage_effects_further}) for the lottery model using the simulation 
+component $\olrn_{i \bs i}$ of (\ref{eq:storage_effects_further1}) for the lottery model using the simulation 
 approach of \cite{Ellner_2016,Ellner_2019}. Note that $E_j$ and $C_{j \bs i}$ are linearly related for the lottery model,
-so $E_j^\nat=E_j$ and $C_{j \bs i}^\nat = C_{j \bs i}$. Thus $\olr_{j \bs i}=\olrn_{j \bs i}$.
+so $E_j^\nat=E_j$, $C_{j \bs i}^\nat = C_{j \bs i}$, and $\olrn_{j \bs i}=\olr_{j \bs i}$.
 
-The other mechanisms in modern coexistence theory that apply in our modelling context (of purely temporal variation)
+The other mechanisms in modern coexistence theory that apply in our modelling context of purely temporal variation
 make no reference to covariation between quantities, whereas asymmetries of tail associations are 
 defined in a covariation context. Therefore, storage effects are the only mechanism of coexistence
-for which asymmetries of tail association are relevant. Mathematical details are given in SI section 
+for which asymmetries of tail association are relevant. Mathematical details supporting this claim are in SI section 
 \ref{SI-SIsect:other_mechanisms}, but we here provide a conceptual understanding. \cite{Ellner_2016} and
-\cite{Ellner_2019} provide a decomposition of $\olr_{i \bs i} -q_{ij}\olr_{i \bs i}$ which is similar to and motivated by the 
+\cite{Ellner_2019} provide a decomposition of the coexistence metric $\olr_{i \bs i} -q_{ij}\olr_{i \bs i}$ which is similar to and motivated by the 
 original decomposition of \cite{Chesson_1994}, and which we use; the decomposition is how the mechanisms of 
@@
 where: $\Delta_i^0$ is a baseline;
-$\Delta_i^E$ is a contribution of environmental variation;
-$\Delta_i^C$ is a contribution of variation in competition;
-$\Delta_i^{(E\shp C)}$ is a contribution of having simultaneous variation in both environment and competition, but not including the effects 
+$\Delta_i^E$ is a contribution to coexistence of environmental variation;
+$\Delta_i^C$ is a contribution to coexistence of variation in competition;
+$\Delta_i^{(E\shp C)}$ is a contribution to coexistence of having simultaneous variation in both environment and competition, but not including the effects 
 of covariation in these quantities; and
-$\Delta_i^{(EC)}$ is a contribution of $EC$ covariation -- storage effects. 
+$\Delta_i^{(EC)}$ is a contribution to coexistence of $EC$ covariation itself -- storage effects. 
 Only storage effects involve $EC$ covariation,
-so the only mechanism of coexistence influenced by ATA is storage effects. Pairing our decomposition 
-(\ref{eq:storage_effects_further1})-(\ref{eq:storage_effects_further}) with the 
-previous MTC decomposition (\ref{eq:previous_decomp}) gives the extension of the MTC which was goal G1 of the Introduction.
+so the only mechanism of coexistence influenced by ATA is storage effects. Combining our decomposition 
+(\ref{eq:storage_effects_further}) with the decomposition (\ref{eq:previous_decomp}) of \cite{Ellner_2019} gives 
+\begin{equation}
+\olr_{i \bs i} -q_{ij}\olr_{i \bs i}=\Delta_i^0+\Delta_i^E+\Delta_i^C+\Delta_i^{(E\shp C)}+\Delta_i^{[E \nat C]} + \Delta_i^{[EC]},\label{eq:full_decomp}
+\end{equation}
+which is the extension of the MTC which was goal G1 of the Introduction.
 
@@
 %models exhibiting solely temporal variation; spatial variation can be considered in future work.
-%G2) To explore using a simple model the relative magnitude of the contributions of coexistence that come
+%G2) To explore using a simple model the relative magnitudes of the contributions to coexistence that come
 %from ATA compared to other mechanisms.
@@
 %at the intuition here. See if you can edge to more generality here, or not?
+
+Left- or right-tail associated noise $(b_1,b_2)$, with $b_i$ normally distributed with mean $\mu_i$ and standard deviation 
+$\sigma$, was generated as pictured in Fig. \ref{fig:noise}, with details in SI section \ref{SI-sect:noise}.
+To achieve goal G2 from the Introduction, i.e., to explore using a simple model the relative magnitudes of the contributions 
+to coexistence that come from ATA versus other mechanisms, we plotted the terms $\Delta_i^0$, $\Delta_i^E$, $\Delta_i^C$, 
+$\Delta_i^{(E\shp C)}$, $\Delta_i^{[E \nat C]}$ and $\Delta_i^{[EC]}$ [see (\ref{eq:full_decomp})] for the lottery model
+against the independent parameters of the model, $\mu_1$, $\mu_2$, $\sigma$ and $\delta$; and then we compared their
+values on the plots. Some mathematical shortcuts, summarized here, greatly simplified this plotting. First, we showed that, for the 
+lottery model, none of the components in (\ref{eq:full_decomp}) depends independently on $\mu_1$ and $\mu_2$,
+instead they all depend only on the difference $\mu_1-\mu_2$ (SI section \ref{SI-sect:math_simplifications}); so we only plotted 
+against $\mu_1 - \mu_2$. Second, we showed that the values of all components in (\ref{eq:full_decomp}) were the same 
+for both our left- and right-tail associated noises (SI section \ref{SI-sect:math_simplifications}); so we only generated plots 
+for left-tail associated noise. For each combination of $\delta=W, X, Y, Z$
+%***DAN: Jasmin, please select 4 sensible values, including \delta=1, and put the ones you select here
+and $\mu_1-\mu_2=A,B,C,D$,
+%***DAN: Jasmin, please select 3-4 sensible values and put the ones you select here
+we plotted all the components of (\ref{eq:full_decomp}) against $\sigma$. Relative importance of contributions was judged from the 
+plots to assess whether and to what extent ATA effects on coexistence were meaningfully large for the lottery model.
+%***DAN: Jasmin, if you select 4 values for mu1-mu2 and 4 for delta, then Fig. 3 is going to have 16 panels arranged in a 
+%4 by 4 grid, each of which will plot all the components against sigma. You'll need to use different colors and/or line
+%types for the components and have a legend.
+The simulation methods of \cite{Ellner_2019} were used to compute the components of (\ref{eq:full_decomp}).
+Computationally efficient means of carrying out these simulations, for the lottery model, are elaborated in 
+SI section X.
+%***DAN: Dan to add the reference to the right sup mat section once things get sufficiently organized in sup mat to know what section.
 
@@
 
+\clearpage
+\newpage
+
+\begin{figure}
+\caption{Pedagogical figure somehow displaying the different outcomes of competition in ATA and symmetric cases for an example.
+This will be cited from the Intro, so needs to be understandable to someone who does not know any of the ideas we introduce yet.
+So in particular you cannot use IGR as your metric of coexistence, you need to use something more intuitive. You will pick
+model parameter values and do two sims, one with ATA noise and one with symmtric noise, and show different outcomes. We should
+discuss and possibly take several iterations to get this figure right since it's a pedagogically important figure - if the reader
+is not convinced in the Intro that we are studying something that makes a difference, they are not going to slog through
+the Theory section, which is inevitably going to be tough going for many readers.\label{fig:pedagog}}
+\end{figure}
+
+\begin{figure}
+\caption{Figure displaying the noise. Three panels, left-, symmetric, and right-tail assoc noise. Use standard normal marginals.
+Explain in caption that alternatives for $(b_1,b_2)$ are obtained by multipling by $\sigma$ and then adding $(\mu_1,\mu_2)$.
+Jasmin to make and insert the figure here using the includegraphics command, and then draft a caption along the lines described here.\label{fig:noise}}
+\end{figure}
+
 \end{document}

### 2021-06-29 paper/SupMat.Rnw
made additions to sup mat to support Methods - still working on it though

@@
 \author[a,b]{Daniel C. Reuman}
-%***DAN: Jasmin, you should choose one way your name will appear, exactly, on all your publications,
-%so everyone knows its you. For instance, I always appear as Daniel C. Reuman, not Daniel Reuman, not
-%Dan Reuman, etc. This makes it possible to find all my papers by searching. So please decide what
-%name to use, Jasmin Albert or Pimsupa Albert or P. Jasmin Albert or whatever. Also keep in mind
-%that because "Albert" is a common name you might want something less common for the rest of it 
-%so that your papers are not confused with anyone elses, though of course I emphasize that how you
-%are going to be known it totally up to you, I'm just recommending you think it through now so you
-%can be consistent. I recognize that this all may be moot if you one day change your name and want to
-%be known by the new name on papers after that point, but anyway in case you don't ever change your 
-%name you might as well be consistent now, at least. So please change your name above to whatever
-%you want to be officially known as on scientific papers. Same change in main text, please.
 
@@
 
-We review the decompositions of \cite{Ellner_2016} and \cite{Ellner_2019}, which are similar to and motivated by the 
-original decompositions of \cite{Chesson_1994}, and then we include our extension.
+We first review the decompositions of \cite{Ellner_2016} and \cite{Ellner_2019}, which are similar to and motivated by the 
+original decompositions of \cite{Chesson_1994}; and then we include our extensions of the decomposition.
 
@@
 &= \left[ \E[r_i(E_i,C_{i \bs i})]-\E[r_i(E_i^\nat,C_{i \bs i}^\nat)] \right]+\left[ \E[r_i(E_i^\nat,C_{i \bs i}^\nat)]-\E[r_i(E_i^\shp,C_{i \bs i}^\shp)] \right] \\
-&= \olep_i^{[EC]}+\olep_i^{[E \nat C]}.
+&= \olep_i^{[E \nat C]}+\olep_i^{[EC]}.
 \end{align}
@@
 \begin{equation}
-\olr_{i \bs i} = \epsilon_i^0 + \olep_i^E + \olep_i^C + \olep_i^{(E\shp C)}+\olep_i^{[EC]}+\olep_i^{[E \nat C]},
+\olr_{i \bs i} = \epsilon_i^0 + \olep_i^E + \olep_i^C + \olep_i^{(E\shp C)}+\olep_i^{[EC]}+\olep_i^{[E \nat C]},\label{eq:decomp_overall_IGR}
 \end{equation}
 our overall decompsition of $r_{i \bs i}$.
-The derivation also applies to provide 
+The same derivation also applies to provide 
 \begin{equation}
@@
 Combining the decompositions of $\olr_{i \bs i}$ and $\olr_{j \bs i}$ and assuming without loss of generality that $\mu_i \leq \mu_j$ so that
-species i is the weaker competitor, we have
+species $i$ is the weaker competitor and $\olr_{i \bs i}-q_{ij}\olr_{j \bs i}$ can be regarded as a coexistence metric, we have
 \begin{equation}
@@
 asymmetric tail assocations. The subscript $i$ on all these newly defined quantities $\Delta_i$ refers to species $i$.
-The notation $\Delta I_i$ from the main text equals $\Delta_i^{(EC)}=\Delta_i^{[EC]}+\Delta_i^{[E\nat C]}$,
-and $\Delta I_i^{(\text{ata})}$ from the main text is the same as $\Delta_i^{[E\nat C]}$ and $\Delta I_i^{(\text{cps})}$ 
-from the main text is $\Delta_i^{[EC]}$ (see (\ref{MT-eq:storage_effects_further1}) and (\ref{MT-eq:storage_effects_further}) from the main text).
-
-%***DAN: Next go through everything for the lottery model
-%***DAN: Then argue that everything depends only on mu1-mu2
-
-\vspace{2in}
-
-\noindent By substituting the definitions,
-
-\begin{align}
-\bar r_i &=\EX(\ln(1-\delta + \frac{B_i}{B_2/ \delta})) \\
-&= \EX(\ln(1-\delta +\delta \frac{B_i}{B_2})) 
-\end{align}
-
-The null growth rate of species $i$ is computed by setting both E and C at their means.
-
-\begin{align}
-\epsilon_i^0 &=\ln(1-\delta + \frac{\bar E_i}{\bar C_i}) \\
-&=\ln(1-\delta + \delta \frac{\bar B_i}{\bar B_2}) 
-\end{align}
-
-The contribution of variance E in to the IGR is computed by setting C to its mean and letting E vary subtracted by the null.
-
-\begin{align}
-\epsilon_i^E &=\ln(1-\delta + \frac{E_i}{\bar C_i}) -  \epsilon_i^0 \\
-&=\ln(1-\delta + \delta \frac{B_i}{\bar B_2}) -  \epsilon_i^0 
-\end{align}
-
-\begin{equation}
-\bar \epsilon_i^E=\EX(\epsilon_i^E) \label{Evar_mean}
-\end{equation}
-
-Similarly, the contribution of variance C in to the IGR is computed by setting E to its mean and letting C vary subtracted by the null.
-
-\begin{align}
-\epsilon_i^C &=\ln(1-\delta + \frac{\bar E_i}{C_i}) -  \epsilon_i^0 \\
- &=\ln(1-\delta + \delta \frac{\bar B_i}{ B_2}) -  \epsilon_i^0 
-\end{align}
-
-\begin{equation}
-\bar \epsilon_i^C=\EX(\epsilon_i^C) \label{Cvar_mean}
-\end{equation}
-
-The contribution of both E and C varying to the IGR is computed by letting both variables vary subtracted by the null and the contributions of only one variable varying.
-
-\begin{equation}
-\epsilon_i^{EC}=\ln(1-\delta + \delta \frac{B_i}{B_j}) -  [\epsilon_i^0 + \epsilon_i^E + \epsilon_i^C] \label{ECvar}
-\end{equation}
-
-\begin{equation}
-\bar \epsilon_i^{EC}=\EX(\ln(1-\delta + \delta \frac{B_i}{B_j})) -[\epsilon_i^0 +\bar \epsilon_i^E + \bar \epsilon_i^C] \label{ECvar_mean}
-\end{equation}
-
-So,
-
-\begin{equation}
-\bar r_i=\epsilon_i^0 +\bar \epsilon_i^E + \bar \epsilon_i^C + \bar \epsilon_i^{EC}. \label{rbari_decomp}
-\end{equation}
-
-But some of $\bar \epsilon_i^{EC}$ comes from covariance between E and C and some comes from both varying independently.
-
-\begin{align}
-\bar \epsilon_i^{(E\#C)} &=\EX(\ln(1-\delta + \frac{E_i}{C_i^{\#}})) -[\epsilon_i^0 +\bar \epsilon_i^E + \bar \epsilon_i^C] \\
- &=\EX(\ln(1-\delta + \delta \frac{B_i}{B_2^{\#}})) -[\epsilon_i^0 +\bar \epsilon_i^E + \bar \epsilon_i^C] 
-\end{align}
-
-\begin{align}
-\bar \epsilon_i^{(EC)} &=\bar \epsilon_i^{EC}-\bar \epsilon_i^{(E\#C)}\\
-&= \EX(\ln(1-\delta + \frac{E_i}{C_i}))-[\epsilon_i^0 +\bar \epsilon_i^E + \bar \epsilon_i^C]\\
-&- [\EX(\ln(1-\delta + \frac{E_i}{C_i^{\#}})) -[\epsilon_i^0 +\bar \epsilon_i^E + \bar \epsilon_i^C]] \\
-&= \EX(\ln(1-\delta + \delta \frac{B_i}{B_2}))-\epsilon_i^0 -\bar \epsilon_i^E - \bar \epsilon_i^C\\
-&- \EX(\ln(1-\delta + \delta \frac{B_i}{B_2^{\#}})) +\epsilon_i^0 +\bar \epsilon_i^E + \bar \epsilon_i^C \\
-&= \EX(\ln(1-\delta + \delta \frac{B_i}{B_2}))-\EX(\ln(1-\delta + \delta \frac{B_i}{B_2^{\#}}))\\
-\end{align}
-
-%***DAN: Sections below this point are from the earlier version/reasoning and may have to be edited and placed in other
-%locations. This comment written 2021 06 22
-
-\section{The Model}
-
-%***DAN: Probably delete this section, since the model is now elaborated in the main text
-
-We used the lottery model, previously studied, for instance, by REFS. Model equations are
-
-\begin{equation}
-N_i(t+1)=(1-\delta)N(t)+\delta N \frac{B_i(t)N_i(t)}{B_1(t)N_1(t)+B_2(t)N_2(t)}  \label{model_eq}
-\end{equation}
-
-\noindent for species $i=1,2$. 
-Here, $N_i(t)$ is the population density of species $i$ at time $t$, and $N = N_1(t)+N_2(t)$ is fixed through time. 
-The parameter, $\delta$ is a mortality rate, and $B_i(t)$ is the fecundiity of specise $i$ at time $t$. 
-The model postulates that individuals die at rate $\delta$ at each time step, and are replaced by juveniles in population to the reproductive output of the two species that year.
-We take, for each $i$, $B_i(t)$ to be $iid$ through time. 
-We let $B_i = \exp(b_i)$, where $\left(b_1, b_2\right)$ is some distribution such that $b_i \sim N\left(\mu_i, \sigma^2\right)$ and such that $\CovX(b_1, b_2) = \rho$, a fixed quantity we will specify later.
-However, $\left(b_1, b_2\right)$ is not necessarily a bivariate normal distribution.
-We will consider various distributions of $\left(b_1, b_2\right)$ with these properties, corresponding to our symmetric and asymmetric tail associations cases defined in the main text and elaborated below.
-We assume without loss of generality that $\mu_1 \leq \mu_2$, so that species 1 is the weaker competitor.
 
@@
 
-We created three cases for $(b_1,b_2)$, the case of symmetric tail association and the cases of asymmetrical left- and right-tail association. 
+We created two cases for $(b_1,b_2)$, left- and right-tail association. 
 To generate $M$ points $\left(b_1^{(i)}, b_2^{(i)}\right)$, $i=1,..,M$, for the left tail association case, we first generated M points $(a_1^{(i)},a_2^{(i)})$, $i=1,...,M$, from a bivariate normal distribution $N\left( \vec{0}, \Sigma \right)$, where here $\Sigma = \begin{pmatrix} 1&0\\0&1\end{pmatrix}$.
-Then, for each index $i$, we randomly, with 50\% probability, either (A) let $\tilde b_1^{(i)} = -|a_1^{(i)}|$ and $\tilde b_2^{(i)} = -a|_1^{(i)}|$, or (B) let $\tilde b_1^{(i)} = |a_1^{(i)}|$ and $\tilde b_2^{(i)} = |a_2^{(i)}|$.
+Then, for each index $i$, we randomly, with 50\% probability, either (A) let $\tilde b_1^{(i)} = -|a_1^{(i)}|$ and $\tilde b_2^{(i)} = -|a_1^{(i)}|$, or (B) let $\tilde b_1^{(i)} = |a_1^{(i)}|$ and $\tilde b_2^{(i)} = |a_2^{(i)}|$.
 We then let $b_1^{(i)} = \sigma \tilde b_1^{(i)} + \mu_1$ and $b_2^{(i)} = \sigma \tilde b_2^{(i)} + \mu_2$. 
 Perfect association in the left tails of the resulting distribution $\left(b_1,b_2\right)$ results from the fact that, in case (A) above, both $\tilde b_1^{(i)}$ and $\tilde b_2^{(i)}$ were assigned to the same value.
-For speed of computation, when samples $(\tilde b_1^{(i)},\tilde b_2^{(i)})$, $i=1,...,M$ were needed for every distinct set of values of the parameters $\mu_1$, $\mu_2$, and $\sigma^2$, we used the same $\left(\tilde b_1^{(i)},\tilde b_2^{(i)}\right)$, $i=1,...,M$, simply transforming these values differently (using the different valus of $\mu_1$, $\mu_2$, $\sigma$) for each set of parameter values.
+For speed of computation, when samples $(b_1^{(i)},b_2^{(i)})$, $i=1,...,M$ were needed for distinct sets of values of the parameters $\mu_1$, $\mu_2$, and $\sigma$, we used the same $\left(\tilde b_1^{(i)},\tilde b_2^{(i)}\right)$, $i=1,...,M$, simply transforming these values differently (using the different valus of $\mu_1$, $\mu_2$, $\sigma$) for each set of parameter values.
 Right-tail associated noise was generated in an analogous way.
+Overall, our version of the lottery model was completely specified by specifying $\delta$, $\mu_1$, $\mu_2$, $\sigma$ and whether 
+noise was generated from the left- or right-tail association cases.
 
-Noise with symmetric tail association was generated by taking a very large sample (100000) of left-tail associated noise, as described above, and then calculating the Pearson correaltion coefficient, rho.
-The value obtained was ... . 
-The symmetric noise was then generated once from $N\left(\vec{0}, \Sigma \right)$ where $\Sigma = \begin{pmatrix} 1&\rho \\ \rho&1\end{pmatrix}$, denoting the points so generated by $\left(\tilde b_1^{(i)},\tilde b_2^{(i)}\right)$ for $i=1,...,M$.
-We then set $b_1^{(i)} = \sigma \tilde b_1^{(i)} + \mu_1$ and $b_2^{(i)} = \sigma \tilde b_2^{(i)}+ \mu_2$.
+When needed, noise with symmetric tail association was generated by taking a very large sample (100000) of left-tail associated noise generated as 
+described above, and then calculating the Pearson correaltion coefficient, $\rho$, of the sample.
+Symmetric noise $\left(\tilde b_1^{(i)},\tilde b_2^{(i)}\right)$ for $i=1,...,M$ was then generated 
+from the bivariate normal distribution $N\left(\vec{0}, \Sigma \right)$, where $\Sigma = \begin{pmatrix} 1&\rho \\ \rho&1\end{pmatrix}$, 
+and then we set $b_1^{(i)} = \sigma \tilde b_1^{(i)} + \mu_1$ and $b_2^{(i)} = \sigma \tilde b_2^{(i)}+ \mu_2$.
 
-Thus, overall, our model was completely specified by specifying $\delta$, $\mu_1$, $\mu_2$, $\sigma$ and whether noise was generated from the left-tailed association, right-tailed association or symmetric tail association cases specified above.
+\section{Mathematical simplifications for plotting}\label{sect:math_simplifications}
+
+For the lottery model, we can straightforwardly show from the formulas of SI section \ref{SIsect:other_mechanisms} that:
+\begin{equation}
+E_i=B_i=\exp(b_i)
+\end{equation}
+\begin{equation}
+C_i=\frac{B_1 N_1 + B_2 N_2}{\delta N}
+\end{equation}
+\begin{equation}
+C_{i \bs i} = B_j/\delta=\exp(b_j)/\delta
+\end{equation}
+\begin{equation}
+r_i=\ln(1-\delta+E_i/C_i)
+\end{equation}
+%IGR
+\begin{align}
+\olr_{i \bs i} &= \E[\ln(1-\delta+E_i/C_{i \bs i})] \\
+&= \E[\ln(1-\delta+\delta B_i/B_j)] \\
+&= \E[\ln(1-\delta+\delta\exp(b_i-b_j))]
+\end{align}
+\begin{equation}
+\overline{E}_i = \exp(\mu_i + \sigma^2/2)
+\end{equation}
+\begin{equation}
+\overline{C}_{i \bs i}=\exp(\mu_j+\sigma^2/2)/\delta
+\end{equation}
+%baseline
+\begin{equation}
+\epsilon_i^0 = \ln(1-\delta+\delta \exp(\mu_i - \mu_j))\label{eq:decomp:baseline}
+\end{equation}
+%contrib of variation in E
+\begin{align}
+\olep_i^E &= \E[\ln(1-\delta+\delta\exp(b_i)/\exp(\mu_j + \sigma^2/2))]-\epsilon_i^0 \\
+&= \E[\ln(1-\delta+\delta\exp(\sigma \tilde{b}_i+\mu_i)/\exp(\mu_j+\sigma^2/2))]-\epsilon_i^0 \\
+&= \E[\ln(1-\delta+\delta\exp(\sigma \tilde{b}_i + \mu_i - \mu_j - \sigma^2/2))]-\epsilon_i^0, \label{eq:decomp:contrib_of_var_E}
+\end{align}
+where $\tilde{b}_i$ is standard normally distributed (see SI section \ref{sect:noise}). Continuing,
+%contrib of variation in C
+\begin{align}
+\olep_i^C &= \E[\ln(1-\delta + \delta \exp(\mu_i+\sigma^2/2)/\exp(b_j))]-\epsilon_i^0 \\
+&= \E[\ln(1-\delta + \delta \exp(\mu_i+\sigma^2/2)/\exp(\sigma \tilde{b}_j +\mu_j))]-\epsilon_i^0 \\
+&= \E[\ln(1-\delta + \delta \exp(-\sigma \tilde{b}_j +\mu_i -\mu_j+\sigma^2/2))]-\epsilon_i^0, \label{eq:decomp:contrib_of_var_C} 
+\end{align}
+where $\tilde{b}_j$ is standard normally distributed (see SI section \ref{sect:noise}). Continuing,
+%last term of primary decomp, later splits into storage effects and effect of variation in E and C, per se  
+\begin{align}
+\olep_i^{EC} &= \E[\ln(1-\delta+\delta\exp(b_i)/\exp(b_j))]-\olep_i^E-\olep_i^C-\epsilon_i^0 \\
+&= \E[\ln(1-\delta+\delta\exp(\sigma \tilde{b}_i +\mu_i)/\exp(\sigma \tilde{b}_j +\mu_j))]-\olep_i^E-\olep_i^C-\epsilon_i^0 \\
+&= \E[\ln(1-\delta+\delta\exp(\sigma (\tilde{b}_i - \tilde{b}_j)+\mu_i-\mu_j))]-\olep_i^E-\olep_i^C-\epsilon_i^0, \label{eq:decomp:last_term_primary}
+\end{align}
+where $(\tilde{b}_i, \tilde{b}_j)$ is distributed as in SI section \ref{sect:noise}, i.e., 
+$(\tilde{b}_i,\tilde{b}_j)=\left(\frac{b_i-\mu_i}{\sigma},\frac{b_j-\mu_j}{\sigma} \right)$. Continuing,
+%contrib of variation in E and C, per se  
+\begin{align}
+\olep_i^{(E\shp C)} &= \E[\ln(1-\delta+\delta\exp(b_i^\shp)/\exp(b_j^\shp))]-\olep_i^E-\olep_i^C-\epsilon_i^0 \\
+&= \E[\ln(1-\delta+\delta\exp(\sigma \tilde{b}_i^\shp +\mu_i)/\exp(\sigma \tilde{b}_j^\shp + \mu_j))]-\olep_i^E-\olep_i^C-\epsilon_i^0 \\
+&= \E[\ln(1-\delta+\delta\exp(\sigma(\tilde{b}_i^\shp - \tilde{b}_j^\shp)+\mu_i-\mu_j))]-\olep_i^E-\olep_i^C-\epsilon_i^0, \label{eq:decomp:contrib_of_EC_var_per_se}
+\end{align}
+where $\tilde{b}_i^\shp$ and $\tilde{b}_j^\shp$ are standard normally distributed and independent. Continuing,
+%storage effects
+\begin{equation}
+\olep_i^{(EC)} = \E[\ln(1-\delta+\delta\exp(\sigma (\tilde{b}_i - \tilde{b}_j)+\mu_i-\mu_j))]-
+\E[\ln(1-\delta+\delta\exp(\sigma (\tilde{b}_i^\shp - \tilde{b}_j^\shp)+\mu_i-\mu_j))],\label{eq:decomp:storage_effects}
+\end{equation}
+where $(\tilde{b}_i,\tilde{b}_j)$ is distributed as in SI section \ref{sect:noise}, i.e., 
+$(\tilde{b}_i,\tilde{b}_j)=\left(\frac{b_i-\mu_i}{\sigma},\frac{b_j-\mu_j}{\sigma} \right)$, and 
+$\tilde{b}_i^\shp$ and $\tilde{b}_j^\shp$ are standard normal and independent. Continuing,
+%contrib of ATA
+\begin{equation}
+\olep_i^{[E \nat C]} = \E[\ln(1-\delta+\delta\exp(b_i-b_j))]-\E[\ln(1-\delta+\delta\exp(\ddot{b}_i-\ddot{b}_j))],\label{eq:decomp:contrib_ATA_1}
+\end{equation}
+where $(\ddot{b}_i,\ddot{b}_j)$ have the same (normal) marginals and covariance as $(b_i,b_j)$ but are 
+bivariate normal. Thus
+\begin{equation}
+\olep_i^{[E \nat C]} = \E[\ln(1-\delta+\delta\exp(\sigma(\tilde{b}_i-\tilde{b}_j)+\mu_i-\mu_j))]-
+\E[\ln(1-\delta+\delta\exp(\sigma(\tilde{\ddot{b}}_i-\tilde{\ddot{b}}_j)+\mu_i-\mu_j))], \label{eq:decomp:contrib_ATA_2}
+\end{equation}
+where $(\tilde{b}_i,\tilde{b}_j)=\left( \frac{b_i-\mu_i}{\sigma},\frac{b_j-\mu_j}{\sigma} \right)$ and 
+$(\tilde{\ddot{b}}_i,\tilde{\ddot{b}}_j)=\left( \frac{\ddot{b}_i-\mu_i}{\sigma},\frac{\ddot{b}_j-\mu_j}{\sigma} \right)$.
+Continuing, $\olep_i^{[EC]}=\olep_i^{(EC)}-\olep_i^{[E \nat C]}$ is available by subtracting expression (\ref{eq:decomp:contrib_ATA_2})
+from expression (\ref{eq:decomp:storage_effects}), i.e., 
+\begin{equation}
+\olep_i^{[EC]} = \E[\ln(1-\delta+\delta\exp(\sigma(\tilde{\ddot{b}}_i-\tilde{\ddot{b}}_j)+\mu_i-\mu_j))] - 
+\E[\ln(1-\delta+\delta\exp(\sigma (\tilde{b}_i^\shp - \tilde{b}_j^\shp)+\mu_i-\mu_j))].
+\end{equation}
+This completes the calculation of all the terms of the decomposition (\ref{eq:decomp_overall_IGR})  of $\olr_{i \bs i}$ for the lottery model.
+
+%***DAN: Next think it all through for j\i and determine what needs to be said for that case before combining the two cases 
+%(and also scaling factors) to get the delta terms in extended MTC decomp
+%Once you do those things you can observe that all these terms depend only on mui-muj and not independently on mui or muj.
 

### 2021-07-06 paper/Paper.Rnw
further progress in writing sup mat sections associated with Methods

@@
 models exhibiting solely temporal variation; spatial variation can be considered in future work.
-G2) To explore using a simple model the relative magnitudes of the contributions to coexistence that come
+G2) To explore, using a simple model, the relative magnitudes of the contributions to coexistence that come
 from ATA compared to other mechanisms.
@@
 that $q_{ij}=\overline{B}_i/((1-\delta)\overline{B}_j+\delta \overline{B}_i)$. 
-%***DAN: Note this might be wrong. Some of the language in Ellner et al (2016) indicates q_{ij} here should be 1,
-%even if \mu_1 and \mu_2 are distinct, as long as we use the same \delta for both species (which we do).
-%But when I try to follow the definitions of the same paper to compute q_{ij} I get the quantity listed above,
-%which is not 1. I'll need to get to the bottom of this, possibly by speaking to Ellner, and then put the right 
-%value here, and also write the corresponding sup mat section accordingly (though you may not need that section, 
-%it may be enough to just cite Ellner and Chesson if you go with q_ij=1.
 Since we assumed log-normally distributed $B_i$, 
 $\overline{B}_i=\exp(\mu_i+\sigma^2/2)$, so $q_{ij}=\exp(\mu_i)/[(1-\delta)\exp(\mu_j)+\delta\exp(\mu_i)]$.
-%***DAN: This statement also needs adjusting if you go with q_ij=1
+We alternatively also consider $q_{ij}=1$, for reasons explained in SI section \ref{SI-SIsect:scaling_factors}.
+%Say this here if it turns out to be true, or say something like it:
+%These alternatives turn out to make very little difference to our results. 
 Thus $\Delta_i^{(EC)}$ for the lottery model can be computed by using simulations \citep{Ellner_2016}
@@
 corresponding to the removal of part of the relationship between $E$ and $C$ components (namely, ATA).
+The new symbol should correspondingly be pronounced ``partial sharp''.
 The term $\Delta_i^{[EC]}$ is the component of storage effects due to $EC$ correlation \emph{per se}, remaining after 
@@
 %models exhibiting solely temporal variation; spatial variation can be considered in future work.
-%G2) To explore using a simple model the relative magnitudes of the contributions to coexistence that come
+%G2) To explore, using a simple model, the relative magnitudes of the contributions to coexistence that come
 %from ATA compared to other mechanisms.
@@
 values on the plots. Some mathematical shortcuts, summarized here, greatly simplified this plotting. First, we showed that, for the 
-lottery model, none of the components in (\ref{eq:full_decomp}) depends independently on $\mu_1$ and $\mu_2$,
-instead they all depend only on the difference $\mu_1-\mu_2$ (SI section \ref{SI-sect:math_simplifications}); so we only plotted 
+lottery model, none of the components in (\ref{eq:full_decomp}) depends independently on $\mu_1$ and $\mu_2$;
+instead they all depend only on the difference $\mu_1-\mu_2$ (SI section \ref{SI-sect:math_simplifications}). So we only plotted 
 against $\mu_1 - \mu_2$. Second, we showed that the values of all components in (\ref{eq:full_decomp}) were the same 
-for both our left- and right-tail associated noises (SI section \ref{SI-sect:math_simplifications}); so we only generated plots 
+for both our left- and right-tail associated noises (SI section \ref{SI-sect:math_simplifications}). So we only generated plots 
 for left-tail associated noise. For each combination of $\delta=W, X, Y, Z$
@@
 we plotted all the components of (\ref{eq:full_decomp}) against $\sigma$. Relative importance of contributions was judged from the 
-plots to assess whether and to what extent ATA effects on coexistence were meaningfully large for the lottery model.
+plots to assess whether and to what extent ATA effects on coexistence were meaningfully large for the lottery model, compared to 
+the contributions of other mechanisms.
 %***DAN: Jasmin, if you select 4 values for mu1-mu2 and 4 for delta, then Fig. 3 is going to have 16 panels arranged in a 
@@
 Computationally efficient means of carrying out these simulations, for the lottery model, are elaborated in 
-SI section X.
-%***DAN: Dan to add the reference to the right sup mat section once things get sufficiently organized in sup mat to know what section.
+SI section \ref{SI-sect:efficient_computation}.
+%***DAN: Dan, you still need to write that sup mat section
 

### 2021-07-06 paper/SupMat.Rnw
further progress in writing sup mat sections associated with Methods

@@
 \section{Scaling factors}\label{SIsect:scaling_factors}
-%***DAN: I am not completely sure about anything in this section and need to get Ellner to look at it. I 
-%did give it a reasonable shot though.
 
@@
 
-\cite{Chesson_1994} and \cite{Ellner_2016} use $q_{ij}=1$ for this same model. Why the discrepancy? 
-DAN: You need to explain this discrepancy here after talking to Ellner.
-%***DAN: Add some text explaining the discrepancy
+%***DAN: Change the below paragraph if needed. Some reasons it might be needed: Steve joins as a coauthor;
+%Steve reads this and thinks it makes sense to change it because something is factually or presentationally 
+%sub-optimal; or Dan or Jasmin actually develops a more rigorous understanding of scaling factors and based
+%on that thinks something here is sub-optimal. Anyway, the "do both" approach seems like it can't go wrong,
+%especially if Steve's prediction is correct that it won't make much difference.
+%***DAN: You probably need to spend a bit more time with the cited papers to better understand this. 
+We also use $q_{ij}=1$ as an alternative value, based on communication with Steve Ellner, for the following reasons.
+First, as discussed in the supporting information of \cite{Ellner_2016}, and as further elaborated by 
+\cite{Barabas_2018}, it takes very special circumstances for the scaling factors to exist and be unique, 
+especially with size-structured models. Second, in empirically parameterized models, scaling factor values are 
+often not robust to arbitrary modelling choices. 
+Although the lottery model is not size structured and we do not empirically parameterize it, the above shortcomings of 
+scaling factors, generally, suggest we should find alternatives to their use. The alternative 
+suggested by Ellner for empirical applications is to do a series of pairwise invader-resident comparisons, each with an 
+equal weighting of invader and resident $r$ values; for our context this amounts to setting scaling factors to $1$. 
+We consider both approaches because the above-mentioned shortcomings do not apply to our case, but it seems 
+possible that future applications will often set scaling factors to $1$, and we want our ideas to also be considered 
+by researchers making that choice.
 
@@
 \begin{equation}
-\olr_{j \bs i} = \epsilon_j^0 + \olep_j^E + \olep_j^C + \olep_j^{(E\shp C)}+\olep_j^{[EC]}+\olep_j^{[E \nat C]},
+\olr_{j \bs i} = \epsilon_j^0 + \olep_j^E + \olep_j^C + \olep_j^{(E\shp C)}+\olep_j^{[EC]}+\olep_j^{[E \nat C]},\label{eq:decomp_overall_rjbsi}
 \end{equation}
@@
 \begin{equation}
-\olr_{i \bs i}-q_{ij}\olr_{j \bs i}=\Delta_i^0+\Delta_i^E+\Delta_i^C+\Delta_i^{(E\shp C)}+\Delta_i^{[EC]}+\Delta_i^{[E\nat C]},
+\olr_{i \bs i}-q_{ij}\olr_{j \bs i}=\Delta_i^0+\Delta_i^E+\Delta_i^C+\Delta_i^{(E\shp C)}+\Delta_i^{[EC]}+\Delta_i^{[E\nat C]},\label{eq:decomp_overall_delta}
 \end{equation}
@@
 
-When needed, noise with symmetric tail association was generated by taking a very large sample (100000) of left-tail associated noise generated as 
+When needed, noise with symmetric tail association was generated by taking a very large sample (100000) 
+%***DAN: Dan, later, check whether 100000 is what was actually used. I'm thinking we could probably do a million.
+of left-tail associated noise generated as 
 described above, and then calculating the Pearson correaltion coefficient, $\rho$, of the sample.
@@
 
-For the lottery model, we can straightforwardly show from the formulas of SI section \ref{SIsect:other_mechanisms} that:
+We first compute all terms of the decomposition (\ref{eq:decomp_overall_IGR}) of $\olr_{i \bs i}$ for the lottery model. 
+One can straightforwardly show from the formulas of SI section \ref{SIsect:other_mechanisms} that
 \begin{equation}
@@
 &= \E[\ln(1-\delta+\delta B_i/B_j)] \\
-&= \E[\ln(1-\delta+\delta\exp(b_i-b_j))]
+&= \E[\ln(1-\delta+\delta\exp(b_i-b_j))]\label{eq:base_IGR}
 \end{align}
@@
 %contrib of variation in E
+and
 \begin{align}
@@
 \olep_i^{[EC]} = \E[\ln(1-\delta+\delta\exp(\sigma(\tilde{\ddot{b}}_i-\tilde{\ddot{b}}_j)+\mu_i-\mu_j))] - 
-\E[\ln(1-\delta+\delta\exp(\sigma (\tilde{b}_i^\shp - \tilde{b}_j^\shp)+\mu_i-\mu_j))].
+\E[\ln(1-\delta+\delta\exp(\sigma (\tilde{b}_i^\shp - \tilde{b}_j^\shp)+\mu_i-\mu_j))].\label{eq:decomp:contrib_EC}
 \end{equation}
@@
 
-%***DAN: Next think it all through for j\i and determine what needs to be said for that case before combining the two cases 
-%(and also scaling factors) to get the delta terms in extended MTC decomp
-%Once you do those things you can observe that all these terms depend only on mui-muj and not independently on mui or muj.
+We next decompose $\olr_{j \bs i}$ for the lottery model. We have 
+\begin{equation}
+E_j = B_j = \exp(b_j)
+\end{equation}
+\begin{equation}
+C_j=\frac{B_1 N_1 + B_2 N_2}{\delta N}
+\end{equation}
+\begin{equation}
+C_{j \bs i} = B_j/\delta = \exp(b_j)/\delta
+\end{equation}
+\begin{equation}
+r_j=\ln(1-\delta+E_j/C_j)
+\end{equation}
+\begin{align}
+\olr_{j \bs i} &= \E[\ln(1-\delta+E_j/C_{j \bs i})] \\
+&= \E[\ln(1-\delta+\delta B_j/B_j)] \\
+&= 0
+\end{align}
+\begin{equation}
+\overline{E}_j = \exp(\mu_j+\sigma^2/2)
+\end{equation}
+\begin{equation}
+\overline{C}_{j \bs i} = \exp(\mu_j + \sigma^2/2)/\delta
+\end{equation}
+%baseline
+\begin{align}
+\epsilon_j^0 &= r_j(\overline{E}_j,\overline{C}_{j \bs i}) \\
+&= \ln \left( 1-\delta+\delta \frac{\exp(\mu_j + \sigma^2/2)}{\exp(\mu_j + \sigma^2/2)} \right) \\
+&= 0 \label{eq:decomp:baseline_j}
+\end{align}
+and
+%contrib of variation in E
+\begin{align}
+\olep_j^E &= \E[r_j(E_j,\overline{C}_{j \bs i})]-\epsilon_j^0 \\
+&= \E[\ln(1-\delta+\delta B_j/\overline{B}_j)]-\epsilon_j^0 \\
+&= \E \left[ \ln \left(1-\delta+\delta \frac{\exp(b_j)}{\exp(\mu_j+\sigma^2/2)} \right) \right]-\epsilon_j^0 \\
+&= \E \left[ \ln \left(1-\delta+\delta\frac{\exp(\sigma \tilde{b}_j + \mu_j)}{\exp(\mu_j+\sigma^2/2)} \right) \right]-\epsilon_j^0 \\
+&= \E[\ln(1-\delta+\delta \exp(\sigma \tilde{b}_j - \sigma^2/2))]-\epsilon_j^0, \label{eq:decomp:contrib_of_var_E_j}
+\end{align}
+where $\tilde{b}_j$ is standard normally distributed (see SI section \ref{sect:noise}). Continuing,
+%contrib of variation in C
+\begin{align}
+\olep_j^C &= \E[r_j(\overline{E}_j,C_{j \bs i})]-\epsilon_j^0 \\
+&= \E \left[ \ln \left(1-\delta+\delta \frac{\overline{B}_j}{B_j} \right) \right]-\epsilon_j^0 \\
+&= \E \left[ \ln \left(1-\delta+\delta \frac{\exp(\mu_j+\sigma^2/2)}{\exp(b_j)} \right) \right]-\epsilon_j^0 \\
+&= \E \left[ \ln \left(1-\delta+\delta \frac{\exp(\mu_j+\sigma^2/2)}{\exp(\sigma \tilde{b}_j +\mu_j)} \right) \right]-\epsilon_j^0 \\
+&= \E[\ln(1-\delta+\delta \exp(-\sigma \tilde{b}_j+\sigma^2/2))]-\epsilon_j^0,\label{eq:decomp:contrib_of_var_C_j}
+\end{align}
+where again $\tilde{b}_j$ is standard normally distributed (see SI section \ref{sect:noise}). Continuing,
+%last term of primary decomp, later splits into storage effects and effect of variation in E and C, per se  
+\begin{align}
+\olep_j^{EC} &= \E[r_j(E_j,C_{j \bs i})]-\olep_j^E-\olep_j^C-\epsilon_j^0 \\
+&= \E[\ln(1-\delta+\delta B_j/B_j)]-\olep_j^E-\olep_j^C-\epsilon_j^0 \\
+&= -\olep_j^E-\olep_j^C-\epsilon_j^0.\label{eq:decomp:j:EC}
+\end{align}
+Continuing,
+%contrib of variation in E and C, per se  
+\begin{align}
+\olep_j^{(E\shp C)} &= \E[r_j(E_j^\shp , C_{j \bs i}^\shp )]-\olep_j^E-\olep_j^C-\epsilon_j^0 \\
+&= \E \left[ \ln \left(1-\delta+\delta \frac{B_j^\shp}{B_j} \right) \right]-\olep_j^E-\olep_j^C-\epsilon_j^0 \\
+&= \E \left[ \ln \left(1-\delta+\delta \frac{\exp(b_j^\shp)}{\exp(b_j)} \right) \right]-\olep_j^E-\olep_j^C-\epsilon_j^0 \\
+&= \E \left[ \ln \left(1-\delta+\delta \frac{\exp(\sigma \tilde{b}_j^\shp + \mu_j)}{\exp(\sigma \tilde{b}_j + \mu_j)} \right) \right]-\olep_j^E-\olep_j^C-\epsilon_j^0 \\
+&= \E[\ln(1-\delta+\delta \exp(\sigma(\tilde{b}_j^\shp - \tilde{b}_j)))]-\olep_j^E-\olep_j^C-\epsilon_j^0, \label{eq:decomp:j:EshpC}
+\end{align}
+where $(E_j^\shp , C_{j \bs i}^\shp)$ has the same marginals as $(E_j,C_{j \bs i})$ but has independent components;
+$B_j^\shp$ is distributed in the same way as $B_j$ but is independent of it; $b_j^\shp$ is distributed in the same
+way as $b_j$ but is independent of it; and $\tilde{b}_j$ and $\tilde{b}_j^\shp$ are standard normally distributed and 
+independent of each other. Continuing,
+%storage effects
+\begin{align}
+\olep_j^{(EC)} &= \E[r_j(E_j,C_{j \bs i})]-\E[r_j(E_j^\shp , C_{j \bs i}^\shp)] \\
+&= -\E[r_j(E_j^\shp , C_{j \bs i}^\shp)] \\
+&= -\E[\ln(1-\delta+\delta \exp(\sigma (\tilde{b}_j^\shp - \tilde{b}_j)))],
+\end{align}
+these results obtained from (\ref{eq:decomp:j:EC}) and (\ref{eq:decomp:j:EshpC}). Again, $\tilde{b}_j$ and 
+$\tilde{b}_j^\shp$ are standard normally distributed and independent. Continuing,
+\begin{align}
+\olep_j^{[E \nat C]} &= \E[r_j(E_j,C_{j \bs i})] - \E[r_j(E_j^\nat , C_{j \bs i}^\nat)] \\
+&= \olr_{j \bs i} - \olr_{j \bs i}^\nat \\
+&= 0
+\end{align}
+because we showed in the Theory section of the main text that $\olr_{j \bs i}^\nat = \olr_{j \bs i}$ due to the fact that
+$E_j$ and $C_{j \bs i}$ are perfectly linearly related to each other. Continuing,
+\begin{align}
+\olep_j^{[EC]} &= \olep_j^{(EC)} - \olep_j^{[E \nat C]} \\
+&= - \E[\ln(1-\delta+\delta \exp(\sigma(\tilde{b}_j^\shp - \tilde{b}_j)))],
+\end{align}
+where again $\tilde{b}_j$ and $\tilde{b}_j^\shp$ are standard normally distributed and independent of each other.
+This completes the calculation of all the terms of the decomposition of $\olr_{j \bs i}$ for the lottery model.
+
+Using the above formulas, we observe that all the terms of the decomposition of $\olr_{i \bs i}$ depend on 
+$\delta$, $\sigma$, and $\mu_i - \mu_j$, but not separately on $\mu_i$ or $\mu_j$. The terms of the decomposition
+of $\olr_{j \bs i}$ depend on $\delta$ and $\sigma$, but do not depend on $\mu_i$ or $\mu_j$ at all.
+As discussed in the main text and SI section \ref{SIsect:scaling_factors}, we consider two alternative
+values for the scaling factors, $q_{ij}$, namely $q_{ij}=1$ and 
+\begin{align}
+q_{ij} &= \frac{\exp(\mu_i)}{(1-\delta) \exp(\mu_j)+\delta \exp(\mu_i)} \\
+&= \frac{\exp(-\mu_i)\exp(\mu_i)}{\exp(-\mu_i)[(1-\delta) \exp(\mu_j)+\delta \exp(\mu_i)]} \\
+&= \frac{1}{(1-\delta)\exp(-(\mu_i-\mu_j))+\delta}.
+\end{align}
+These also either do not depend on $\mu_i$ or $\mu_j$ at all or they depend on the difference 
+$\mu_i-\mu_j$. Thus each of the terms of the decomposition (\ref{MT-eq:full_decomp}) from the main text
+depends only on $\mu_i-\mu_j$ and not separately on $\mu_i$ or $\mu_j$, as stated in the Methods section of the main text.
+
+We now want to show that, for the lottery model, 
+all elements of our decompositions (\ref{eq:decomp_overall_IGR}) and (\ref{eq:decomp_overall_rjbsi}), as well
+as $q_{ij}$, and therefore all elements of our decomposition (\ref{eq:decomp_overall_delta}), are the same 
+for our left- and right-tail associated cases, as was stated in Methods in the main text. Using the notation 
+$(E_i^{(l)},C_{i \bs i}^{(l)})=(B_i^{(l)},B_j^{(l)}/\delta)=(\exp(b_i^{(l)}),\exp(b_j^{(l)})/\delta)$ and 
+$(E_i^{(r)},C_{i \bs i}^{(r)})=(B_i^{(r)},B_j^{(r)}/\delta)=(\exp(b_i^{(r)}),\exp(b_j^{(r)})/\delta)$
+for the left- and right-tail associated cases, respectively, we note that the marginal distributions of
+$(b_i^{(l)},b_j^{(l)})$ are the same as those of $(b_i^{(r)},b_j^{(r)})$, and likewise the marginals of 
+$(B_i^{(l)},B_j^{(l)}/\delta)$ are the same as those of $(B_i^{(r)},B_j^{(r)}/\delta)$.
+So any component of (\ref{eq:decomp_overall_IGR}), or $q_{ij}$, will not depend on the choice
+of our left- or right-tail associated noise cases as long as it depends only on marginal distributions.
+But it is easy to see that $q_{ij}$ depends only on marginals, since $q_{ij}=\exp(\mu_i)/[(1-\delta)\exp(\mu_j)+\delta\exp(\mu_i)]$  
+depends only on $\mu_i$, $\mu_j$ and $\delta$, and these are parameters of the marginal distributions of 
+$(b_i,b_j)$. The alternative value $q_{ij}=1$ is constant, so also takes the same value for both our
+left- and right-tail association noise cases. It is also easy to see from the definitions of $\epsilon_i^0$, 
+$\olep_i^E$, $\olep_i^C$, and $\olep_i^{(E \shp C)}$ that these quantities only depend on marginals.
+
+Whereas $\olep_i^{[EC]} = \E[r_i(E_i^\nat , C_{i \bs i}^\nat )]-\E[r_i(E_i^\shp ,C_{i \bs i}^\shp)]$ depends on
+more than just marginals, this quantity is still independent of the choice between our left- and right-tail-association 
+cases because the bivariate random variables $(E_i^\nat , C_{i \bs i}^\nat )$ and 
+$(E_i^\shp ,C_{i \bs i}^\shp)$ have had their tail association removed, and the only differences 
+between $(E_i^{(l)},C_{i \bs i}^{(l)})$ and $(E_i^{(r)},C_{i \bs i}^{(r)})$ are in their tail association
+properties (i.e., these two bivariate distributions have the same marginals and covariances).
+In other words, $(E_i^\nat , C_{i \bs i}^\nat )$ is distributed in the same way for the left- and right-tail-association cases,
+as is $(E_i^\shp ,C_{i \bs i}^\shp)$.
+
+We now examine $\olep_i^{[E \nat C ]}$, which, by (\ref{eq:decomp:contrib_ATA_1}), equals 
+\begin{equation}
+\olep_i^{[E \nat C]} = \E[\ln(1-\delta+\delta\exp(b_i-b_j))]-\E[\ln(1-\delta+\delta\exp(\ddot{b}_i-\ddot{b}_j))].\label{eq:rep1}
+\end{equation}
+But $(\ddot{b}_i,\ddot{b}_j)$ is a bivariate normal random variable constructed to have the same marginals and 
+covariance as $(b_i,b_j)$, and the marginals and covariance of $(b_i^{(l)},b_j^{(l)})$ are the same as those of
+$(b_i^{(r)},b_j^{(r)})$, so $(\ddot{b}_i^{(l)},\ddot{b}_j^{(l)})$ is distributed in the same way as 
+$(\ddot{b}_i^{(r)},\ddot{b}_j^{(r)})$. So the second term on the right of (\ref{eq:rep1}) is independent
+of the choice between our left- and right-tail-associated noise cases. Now examining 
+$\E[\ln(1-\delta+\delta\exp(b_i-b_j))]$, we define $\beta_1=-b_2+\mu_1+\mu_2$ and $\beta_2=-b_1+\mu_1+\mu_2$,
+and we note that $(\beta_1^{(l)},\beta_2^{(l)})$ is distributed in the same way as 
+$(b_1^{(r)},b_2^{(r)})$; likewise, $(\beta_1^{(r)},\beta_2^{(r)})$ is distributed in the same way as 
+$(b_1^{(l)},b_2^{(l)})$. But then we have 
+\begin{align}
+\E[\ln(1-\delta+\delta\exp(b_i^{(l)}-b_j^{(l)}))] &= \E[\ln(1-\delta+\delta\exp(-\beta_j^{(l)}+\mu_1+\mu_2-(-\beta_i^{(l)}+\mu_1+\mu_2)))] \\
+&= \E[\ln(1-\delta+\delta\exp(\beta_i^{(l)}-\beta_j^{(l)}))] \\
+&= \E[\ln(1-\delta+\delta\exp(b_i^{(r)}-b_j^{(r)}))], 
+\end{align}
+which completes the proof that $\olep_i^{[E \nat C]}$ is independent of left- versus right-tail associated noise.
+For the terms of the decomposition (\ref{eq:decomp_overall_rjbsi}), it suffices to look at the equations for those terms, 
+above, in this section, and note that all those terms depend only on marginals.
+
+\section{Efficient computation}\label{sect:efficient_computation}
+
+We here elaborate how the terms in the decompositions (\ref{eq:decomp_overall_IGR}), (\ref{eq:decomp_overall_rjbsi})
+and (\ref{eq:decomp_overall_delta}) are computed rapidly, for the lottery model.
+
+First, for a large integer $M$ (we used $M=$),
+%***DAN: fill in the value, 1000000 if you can swing it
+carry out the following steps once.
+\begin{enumerate}
+\item Generate $M$ independent points $u^{(k)}$, $k=1,\ldots,M$, from a standard normal distribution.
+\item Generate $M$ independent points $(\tilde{b}_i^{(k)},\tilde{b}_j^{(k)})$, as in SI section \ref{sect:noise}, which are left-tail associated
+and which have standard-normal marginals. 
+\item Again as in SI section \ref{sect:noise}, generate $M$ independent points $(\tilde{\ddot{b}}_i^{(k)},\tilde{\ddot{b}}_j^{(k)})$, 
+$k=1,\ldots,M$, from a bivariate normal distribution with component variances $1$ and Pearson correlation the same as 
+$(\tilde{b}_i^{(k)},\tilde{b}_j^{(k)})$.
+\end{enumerate}
+These data are stored and used repeatedly below. This provides computational efficiency because 
+random number generation is computationally expensive.
+
+Next, given values of the parameters $\delta$, $\mu_i - \mu_j$, and $\sigma$, proceed as follows to get estimates of the 
+terms of the decomposition (\ref{eq:decomp_overall_IGR}).
+\begin{enumerate}
+\item Compute $\epsilon_i^0$ following (\ref{eq:decomp:baseline}), which is a simple formula of $\delta$ and 
+$\mu_i - \mu_j$, not requiring use of the random samples described above. This gives an exact value of
+$\epsilon_i^0$, not an estimate.
+\item Estimate $\olep_i^E$, following (\ref{eq:decomp:contrib_of_var_E}), as 
+$\widehat{\olep_i^E}=\mean_k [\ln(1-\delta+\delta \exp(\sigma u^{(k)}+\mu_i-\mu_j-\sigma^2/2))]-\epsilon_i^0$, an estimate with 
+standard error $\se \left( \widehat{\olep_i^E} \right) = \sd_k [\ln(1-\delta+\delta \exp(\sigma u^{(k)}+\mu_i-\mu_j-\sigma^2/2))]/\sqrt{M}$,
+where $\sd(\cdot)$ denotes sample standard deviation.
+\item Estimate $\olep_i^C$, following (\ref{eq:decomp:contrib_of_var_C}), as 
+$\widehat{\olep_i^C}=\mean_k [\ln(1-\delta+\delta \exp(-\sigma u^{(k)}+\mu_i-\mu_j+\sigma^2/2))]-\epsilon_i^0$, which has 
+standard error $\se \left( \widehat{\olep_i^C} \right) = \sd_k[\ln(1-\delta+\delta \exp(-\sigma u^{(k)}+\mu_i-\mu_j+\sigma^2/2))]/\sqrt{M}$.
+\item Note that $\tilde{b}_i^\shp-\tilde{b}_j^\shp$ in (\ref{eq:decomp:contrib_of_EC_var_per_se}) is normally distributed
+with mean $0$ and variance $2$, so we can estimate $\olep_i^{(E \shp C)}$, following (\ref{eq:decomp:contrib_of_EC_var_per_se}),
+as $\widehat{\olep_i^{(E \shp C)}}=\mean_k[\ln(1-\delta+\delta\exp(\sigma\sqrt{2}u^{(k)}+\mu_i-\mu_j))]-\widehat{\olep_i^E}-\widehat{\olep_i^C}-\epsilon_i^0.$ The standard error of this expression is 
+$\se \left( \widehat{\olep_i^{(E \shp C)}} \right) = \sd_k [\ln(1-\delta+\delta\exp(\sigma \sqrt{2} u^{(k)}+\mu_i-\mu_j))-
+\ln(1-\delta+\delta\exp(\sigma u^{(k)}+\mu_i-\mu_j-\sigma^2/2))-
+\ln(1-\delta+\delta\exp(-\sigma u^{(k)}+\mu_i-\mu_j+\sigma^2/2))]/\sqrt{M}$. 
+\item Estimate $\olep_i^{[EC]}$, following (\ref{eq:decomp:contrib_EC}), as
+$\widehat{\olep_i^{[EC]}}=\mean_k[\ln(1-\delta+\delta\exp(\sigma(\tilde{\ddot{b}}_i^{(k)}-\tilde{\ddot{b}}_j^{(k)})+\mu_i-\mu_j))]-\mean_k[\ln(1-\delta+\delta\exp(\sigma\sqrt{2}u^{(k)}+\mu_i-\mu_j))]$.
+The standard error of this expression is
+$\se \left( \widehat{\olep_i^{[EC]}} \right) = \sqrt{s_1^2+s_2^2}$, where $s_1=\sd_k [\ln(1-\delta+\delta\exp(\sigma(\tilde{\ddot{b}}_i^{(k)}-\tilde{\ddot{b}}_j^{(k)})+\mu_i-\mu_j))]/\sqrt{M}$
+and $s_2=\sd_k [\ln(1-\delta+\delta\exp(\sigma\sqrt{2}u^{(k)}+\mu_i-\mu_j))]/\sqrt{M}$.
+\item Estimate $\olep_i^{[E \nat C]}$, following (\ref{eq:decomp:contrib_ATA_2}), as 
+$\widehat{\olep_i^{[E \nat C]}} = \mean_k[\ln(1-\delta+\delta\exp(\sigma(\tilde{b}_i^{(k)}-\tilde{b}_j^{(k)})+\mu_i-\mu_j))]-
+\mean_k[\ln(1-\delta+\delta\exp(\sigma(\tilde{\ddot{b}}_i^{(k)}-\tilde{\ddot{b}}_j^{(k)})+\mu_i-\mu_j))].$ The standard
+error of this expression is $\se \left( \widehat{\olep_i^{[E \nat C]}} \right)=\sqrt{s_3^2+s_1^2}$, where
+$s_1$ is as above, and $s_3=\sd_k[\ln(1-\delta+\delta\exp(\tilde{b}_i^{(k)}-\tilde{b}_j^{(k)}))]/\sqrt{M}.$
+\item Finally, estimate $\olr_{i \bs i}$, following (\ref{eq:base_IGR}), as 
+$\widehat{\olr_{i \bs i}}=\mean_k[\ln(1-\delta+\delta\exp(\sigma(\tilde{b}_i^{(k)}-\tilde{b}_j^{(k)})+\mu_i-\mu_j))],$
+which has standard error $s_3$ from above.
+\end{enumerate}
+It is easy to see that $\widehat{\olr_{i \bs i}}=\epsilon_i^0+\widehat{\olep_i^E}+\widehat{\olep_i^C}+\widehat{\olep_i^{(E \shp C)}}+
+\widehat{\olep_i^{[EC]}}+\widehat{\olep_i^{[E \nat C]}},$ i.e., the decomposition (\ref{eq:decomp_overall_IGR}) applies 
+at the level of the estimators, as well
+
+Next, proceed as follows to get estimates of the terms in the decomposition (\ref{eq:decomp_overall_rjbsi}). 
+\begin{enumerate}
+\item Use $\epsilon_i^0=0$, as in (\ref{eq:decomp:baseline_j}), and this is the exact value.
+\item Estimate $\olep_j^E$, following (\ref{eq:decomp:contrib_of_var_E_j}), as
+$\widehat{\olep_j^E}=\mean_k[\ln(1-\delta+\delta\exp(\sigma u^{(k)}-\sigma^2/2))]-\epsilon_j^0$, an estimate
+with standard error $\se \left( \widehat{\olep_j^E} \right) = \sd_k [\ln(1-\delta+\delta\exp(\sigma u^{(k)}-\sigma^2/2))]/\sqrt{M}$.
+\item Estimate $\olep_j^C$, following (\ref{eq:decomp:contrib_of_var_C_j}), as
+$\widehat{\olep_j^C}=\mean_k[\ln(1-\delta+\delta\exp(-\sigma u^{(k)}+\sigma^2/2))]-\epsilon_j^0$, which has 
+standard error $\se \left( \widehat{\olep_j^C} \right) = \sd_k [\ln(1-\delta+\delta\exp(-\sigma u^{(k)}+\sigma^2/2))]/\sqrt{M}$.
+\item Note that $\tilde{b}_j$ and $\tilde{b}_j^\shp$ in (\ref{eq:decomp:j:EshpC}) are standard normally distributed and
+independent, so $\tilde{b}_j^\shp - \tilde{b}_j$ is normally distributed with mean $0$ and variance $2$. So we can 
+estimate $\olep_j^{(E \shp C)}$, following (\ref{eq:decomp:j:EshpC}), as 
+$\widehat{\olep_j^{(E \shp C)}} = \mean_k [\ln(1-\delta+\delta\exp(\sigma\sqrt{2}u^{(k)}))]-\widehat{\olep_j^E}-\widehat{\olep_j^C}-\epsilon_j^0$.
+The standard error of this expression is $\se \left(  \widehat{\olep_j^{(E \shp C)}} \right) = 
+\sd_k[\ln(1-\delta+\delta\exp(\sigma\sqrt{2}u^{(k)}))-\ln(1-\delta+\delta\exp(u^{(k)}-\sigma^2/2))-
+\ln(1-\delta+\delta\exp(-\sigma u^{(k)}+\sigma^2/2))]/\sqrt{M}$.
+\item Estimate
+%***DAN: DAN, proceed from here
+\end{enumerate}
 

### 2021-07-07 paper/Paper.Rnw
small changes to compile

--- 2021-06-28 paper/Paper.Rnw
@@
 %Relationships between extreme events and their influence on competition and coexistence
+%Coexistence and consequences of extreme environmental relationships
+%Coexistence and consequences of environmental relationships in their extremes
+%Coexistence and implications of asymmetric environmental relationships
+%Extreme [climatic] events and their implications for biodiversity
+%Deductions to coexistence, extreme events, and asymmetric relationships
+%Extreme relationships (relationships in their extremes) and their deductions ([negative] contributions) to coexistence
 
-\author[a]{Jasmin Albert}
+\author[a]{Pimsupa Jasmin Albert}
 \author[a,b]{Daniel C. Reuman}
@@
 \begin{document}
-%\SweaveOpts{concordance=TRUE}
+\SweaveOpts{concordance=TRUE}
+\SweaveOpts{concordance=TRUE}
 
 %The following is where you load in the numeric results that will be embedded in the text
-<<eval=T,echo=F,message=F,warning=F>>=
-system(paste("biber", sub("\\.Rnw$", "", current_input())))
+%<<eval=T,echo=F,message=F,warning=F>>=
+%system(paste("biber", sub("\\.Rnw$", "", current_input())))
 
-#place R code here for loading in necessary variables from Results
-#ordinarily this would be a readRDS command, but I'm just defining a variable here for now
-allregres<-matrix(rnorm(10),2,5)
+%#place R code here for loading in necessary variables from Results
+%#ordinarily this would be a readRDS command, but I'm just defining a variable here for now
+%allregres<-matrix(rnorm(10),2,5)
 @
@@
 
-Jasmin, I think the below makes a pretty good start at tryinbg to nail down a good way to express our goals
-in the last para of the Intro, so please incorporate these into your last para. If you have a better way
-to phrase these, please put both so we can compare at some point when we meet.
-G1) To augment the modern theory of coexistence (MTC) to show how asymmetric tail association (ATA)
-plays a role. The MTC decomposes a measure of coexistence into components due to each of several mechanisms. 
-We address how ATA contributes to these components. We here examine mechanisms of coexistence that apply in 
-models exhibiting solely temporal variation; spatial variation can be considered in future work.
-G2) To explore using a simple model the relative magnitude of the contributions of coexistence that come
-from ATA compared to other mechanisms.
-
 \section{Theory}\label{section:theory}
 
-In this section we pursue goal G1 from the Introduction: to augment the MTC to show how ATA
-plays a role in species coexistence.
-%***DAN: make sure you define MTC and ATA once (probably in the Intro) and then use the abbreviations onle after that.
-For pedgogical clarity, we develop each step of the theory in general and for a simple classical model 
-(the lottery model), as an example.
-
-The lottery model is as follows. 
+For pedgogical clarity, we develop each step of the theory both for a simple classical model 
+(the lottery model), and in general. The lottery model is as follows. 
 %***DAN: Note the inclusion of the model here means we will have to edit the "Model" section of sup mat, since we don't need to intro the model twice. Do later.
@@
 similar to those of \cite{Ellner_2016} and \cite{Ellner_2019}; see those papers for additional details. 
+%***DAN: Don't forget to return to the other mechanisms of coexistence below. Basically want to say intuitively 
+%why tail associations don't influence those,
+%though might be worth mentioning in the Discussion that events are getting more extreme. We have studied how *relationships*
+%in the extremes influence coexistence, but probably worth studying (or cite papers if it's been done) how changes
+%in the commonness and extremeness of extreme events themselves influence coexistence, and for that you'd have
+%to also look at the other mechanisms. May still need a sup mat laying out all the math.
 
@@
 an environment-dependent factor $E_i(t)$ and as a decreasing function of a quantity representing
-competitive pressure on species $i$, $C_i(t)$. As for the lottery model example, we here assume, for simpicity, 
-that the $E_i(t)$ are iid through time, and likewise for the
+competitive pressure on species $i$, $C_i(t)$. For simplicity, we here assume 
+the $E_i(t)$ are iid through time, and likewise for the
 $C_i(t)$. Though this assumption can be circumvented, we make the assumption because it simplifies the
-presentation of new ideas.
-
+presentation without obscuring ideas. 
 For the lottery model, we take $E_i(t)=B_i(t)$, henceforth assuming that fecundity depends strictly on
@@
 \begin{equation}
-\Delta_i^{(EC)} = (\olr_{i\bs i} - q_{ij} \olr_{j \bs i})-(\olrs_{i \bs i}-q_{ij} \olrs_{j \bs i}),\label{eq:storage_effects_gen}
+\Delta I_i = (\olr_{i\bs i} - q_{ij} \olr_{j \bs i})-(\olrs_{i \bs i}-q_{ij} \olrs_{j \bs i}),\label{eq:storage_effects_gen}
 \end{equation}
@@
 growth rate of species $j$ when species $i$ is absent or at negligible abundance (this is $0$, \emph{a priori});
-and $\olrs_{i \bs i}$ and $\olrs_{j \bs i}$ are analogous quantities, but for an alternative modelling scenario, identical except that
+and $\olrs_{i \bs i}$ and $\olrs_{j \bs i}$ are analogous quantities, but for an alternative model scenario, identical except that
 influences of environment and competition on the invader or resident have been rendered independent.  
-Following \cite{Ellner_2016,Ellner_2019}, a subscript 
+A subscript 
 $i \bs i$ always indicates a quantity computed for species $i$ when that species is at negligible abundance, and 
 a subscript $j \bs i$ indicates a quantity computed for $j$ when $i$ is absent.
-The notation $\Delta_i^{(EC)}$ for storage effects also follows \cite{Ellner_2019}.
 The \emph{scaling factor}
@@
 \cite{Ellner_2016} and SI section \ref{SI-SIsect:scaling_factors} for the definition of $q_{ij}$.
-Since it was assumed that $r_i=r_i(E_i,C_i)$, we can write $\olr_{i \bs i}=\E[r_i(E_i,C_{i \bs i})]$, where
-$\E(\cdot)$ denotes expected value just as does an overbar, but we use $\E$ for longer expressions. Also, $\olr_{j \bs i}=\E[r_j(E_j,C_{j \bs i})]$.
-An expression for $\olrs_{i \bs i}$ is obtained from $\olr_{i \bs i}$ by replacing the bivariate random 
-variable $(E_i,C_{i \bs i})$ by a random variable with the same marginal distributions, but independent components, 
-$(E_i^\shp,C_{i \bs i}^\shp)$, so $\olrs_{i \bs i}=\E[r_i(E_i^\shp,C_{i \bs i}^\shp)]$.
-Likewise, $\olrs_{j \bs i}=\E[r_j(E_j^\shp,C_{j \bs i}^\shp)]$.
 
@@
 $\olr_{i \bs i} = \E[\ln(1-\delta+\delta B_i/B_j)]=\E[\ln(1-\delta+E_i /C_{i \bs i})]$. 
-Likewise, 
+Here $\E(\cdot)$ denotes expected value just as does an overbar, but we use $\E$ for longer expressions for which
+an overbar would be cumbersome. Likewise, 
 $\olr_{j \bs i} = \E[\ln(1-\delta+\delta B_j/B_j)]=\E[\ln(1-\delta+E_j/C_{j \bs i})]$,
-though we already knew this equals $0$. 
-An expression for $\olrs_{i \bs i}$ is obtained by replacing $(E_i,C_{i \bs i})=(B_i,B_j/\delta)$
+though we already knew this equals $0$. An expression for $\olrs_{i \bs i}$ is then obtained from the 
+above expression for $\olr_{i \bs i}$ by replacing the bivariate random variable $(E_i,C_{i \bs i})=(B_i,B_j/\delta)$
 by a random variable with the same marginals, but independent components, $(B_i^\shp,B_j^\shp /\delta)$.
-Thus $\olrs_{i \bs i} = \E[\ln(1-\delta+\delta B_i^\shp/B_j^\shp)]$. Likewise,
-$\olrs_{j \bs i}$ is computed by replacing $(E_j,C_{j \bs i})$ in an analogous way; we use 
+Thus $\olrs_{i \bs i} = \E[\ln(1-\delta+\delta B_i^\shp/B_j^\shp)]$. Likewise, a formula for 
+$\olrs_{j \bs i}$ is available by replacing $(E_j,C_{j \bs i})$ in an analogous way. We use 
 $(E_j^\shp,C_{j \bs i}^\shp)=(B_j^\shp, B_j /\delta)$, so 
@@
 %it may be enough to just cite Ellner and Chesson if you go with q_ij=1.
-Since we assumed log-normally distributed $B_i$, 
+Since we use log-normally distributed $B_i$, 
 $\overline{B}_i=\exp(\mu_i+\sigma^2/2)$, so $q_{ij}=\exp(\mu_i)/[(1-\delta)\exp(\mu_j)+\delta\exp(\mu_i)]$.
 %***DAN: This statement also needs adjusting if you go with q_ij=1
-So $\Delta_i^{(EC)}$ for the lottery model can be computed by using simulations \citep{Ellner_2016}
+So $\Delta I_i$ for the lottery model can be computed by using simulations \citep{Ellner_2016}
 to very accurately approximate $\olr_{i \bs i}$, $\olrs_{i \bs i}$,
@@
 
-We now specify how to decompose storage effects into
+Having just reviewed storage effects, we now specify how to decompose storage effects into
 contributions due to asymmetric tail associations and due
@@
 \begin{align}
-\Delta_i^{(EC)} &= [(\olr_{i\bs i} - q_{ij} \olr_{j \bs i})-(\olrn_{i \bs i}-q_{ij} \olrn_{j \bs i})]+
-[(\olrn_{i \bs i}-q_{ij} \olrn_{j \bs i})-(\olrs_{i \bs i}-q_{ij} \olrs_{j \bs i})] \label{eq:storage_effects_further1}\\
-&= \Delta_i^{[E \nat C]} + \Delta_i^{[EC]}.\label{eq:storage_effects_further}
+\Delta I_i &= [(\olr_{i\bs i} - q_{ij} \olr_{j \bs i})-(\olrn_{i \bs i}-q_{ij} \olrn_{j \bs i})]+
+[(\olrn_{i \bs i}-q_{ij} \olrn_{j \bs i})-(\olrs_{i \bs i}-q_{ij} \olrs_{j \bs i})] \\
+&= I_i^{(\text{ata})} + I_i^{(\text{cps})}.\label{eq:storage_effects_further}
 \end{align}
-The term $\Delta_i^{[E \nat C]}$ is the component of storage effects due to 
+The superscript ``ata" stands for ``asymmetric tail associations'' because this component of storage effects is due to 
 asymmetries of tail association between $E_i$ and $C_{i \bs i}$, and between $E_j$ and $C_{j \bs i}$. 
+The superscript ``cps" stands for ``correlation
+\emph{per se}'' because this component is due to $EC$ correlation \emph{per se}, after 
+removing the effects of asymmetric tail associations.
 The notation $\nat$ was chosen because this new symbol is part of the symbol $\shp$ (i.e., two of its four lines), 
 corresponding to the removal of part of the relationship between $E$ and $C$ components.
-The term $\Delta_i^{[EC]}$ is the component due to $EC$ correlation \emph{per se}, after 
-removing the effects of asymmetric tail associations. 
-The brackets in the superscripts of $\Delta_i^{[E \nat C]}$ and $\Delta_i^{[EC]}$ are intended to indicate that 
-this is a decompostion of storage effects, which were denoted by \cite{Ellner_2019} using a superscript $(EC)$,
-i.e., $\Delta_i^{(EC)}$, to indicate that storage effects are, in turn, part of a decomposition of a difference between invader and
-resident growth rates (see below).
 
@@
 these random variables. We then define a bivariate normal random variable $(e_i,c_{i \bs i})$ such that
-$\var(e_i)=\var(c_{i \bs i})=1$, and we denote $\cov(e_i,c_{i \bs i})$ by $\tilde{\rho}_i$. We define 
+$\var(e_i)=\var(c_{i \bs i})=1$, and denote $\cov(e_i,c_{i \bs i})$ by $\tilde{\rho}_i$. We define 
 $E_i^\nat = F_{E_i}^{-1} \circ \varphi(e_i)$
@@
 can be written as $(\exp(\tilde{b}_i),\exp(\tilde{b}_j))$, where $(\tilde{b}_i,\tilde{b}_j)$
-is bivariate-normally distributed with mean $(\mu_i,\mu_j)$, $\var(\tilde{b}_i)=\var(\tilde{b}_j)=\sigma$,
+is bivariate normally distributed with mean $(\mu_i,\mu_j)$, $\var(\tilde{b}_i)=\var(\tilde{b}_j)=\sigma$,
 and $\cov(\tilde{b}_i,\tilde{b}_j)=\cov(b_i,b_j)$. So $(\tilde{b}_i,\tilde{b}_j)$ has the same
@@
 $(b_i,b_j)$ itself need not be. Given these definitions, it is straightforward to compute the 
-component $\olrn_{i \bs i}$ of (\ref{eq:storage_effects_further}) for the lottery model using the simulation 
-approach of \cite{Ellner_2016,Ellner_2019}. Note that $E_j$ and $C_{j \bs i}$ are linearly related for the lottery model,
-so $E_j^\nat=E_j$ and $C_{j \bs i}^\nat = C_{j \bs i}$. Thus $\olr_{j \bs i}=\olrn_{j \bs i}$.
+components of (\ref{eq:storage_effects_further}) for the lottery model using the simulation 
+approach of \cite{Ellner_2016,Ellner_2019}, though in the next section we describe some tricks
+that make these computations efficient.
 
-The other mechanisms in modern coexistence theory that apply in our modelling context (of purely temporal variation)
+The other mechanism in modern coexistence theory that apply in our modelling context (i.e., purely temporal variation)
 make no reference to covariation between quantities, whereas asymmetries of tail associations are 
 defined in a covariation context. Therefore, storage effects are the only mechanism of coexistence
-for which asymmetries of tail association are relevant. Mathematical details are given in SI section 
-\ref{SI-SIsect:other_mechanisms}, but we here provide a conceptual understanding. \cite{Ellner_2016} and
-\cite{Ellner_2019} provide a decomposition of $\olr_{i \bs i} -q_{ij}\olr_{i \bs i}$ which is similar to and motivated by the 
-original decomposition of \cite{Chesson_1994}, and which we use; the decomposition is how the mechanisms of 
-coexistence opperating in a given case are illuminated. The decomposition is 
-\begin{equation}
-\olr_{i \bs i} -q_{ij}\olr_{i \bs i}=\Delta_i^0+\Delta_i^E+\Delta_i^C+\Delta_i^{(E\shp C)}+\Delta_i^{(EC)},\label{eq:previous_decomp}
-\end{equation}
-where: $\Delta_i^0$ is a baseline;
-$\Delta_i^E$ is a contribution of environmental variation;
-$\Delta_i^C$ is a contribution of variation in competition;
-$\Delta_i^{(E\shp C)}$ is a contribution of having simultaneous variation in both environment and competition, but not including the effects 
-of covariation in these quantities; and
-$\Delta_i^{(EC)}$ is a contribution of $EC$ covariation -- storage effects. 
-Only storage effects involve $EC$ covariation,
-so the only mechanism of coexistence influenced by ATA is storage effects. Pairing our decomposition 
-(\ref{eq:storage_effects_further1})-(\ref{eq:storage_effects_further}) with the 
-previous MTC decomposition (\ref{eq:previous_decomp}) gives the extension of the MTC which was goal G1 of the Introduction.
+for which asymmetries of tail association are relevant. Details are given in SI section 
+\ref{SIsect:other_mechanisms}, and we return briefly to related ideas in the Discussion.
+%***DAN: In the Discussion, I intend to talk about how possible changes in tail association 
+%patterns of env vars in nature may influence coexistence of species. It won't influence the 
+%other mechanisms, though changes in the extreme values of environmental variables, known
+%to be occurring, might.
 
@@
 
-%Goals, for reference:
-%G1) To augment the modern theory of coexistence (MTC) to show how asymmetric tail association (ATA)
-%plays a role. The MTC decomposes a measure of coexistence into components due to each of several mechanisms. 
-%We address how ATA contributes to these components. We here examine mechanisms of coexistence that apply in 
-%models exhibiting solely temporal variation; spatial variation can be considered in future work.
-%G2) To explore using a simple model the relative magnitude of the contributions of coexistence that come
-%from ATA compared to other mechanisms.
-
-%Figs, for reference:
-%1) pedagog fig showing what tail assoc is, also later used for introing the noise we use, in methods
-%2) pedgog fig showing a simple set of sims, somehow, the only difference being ATA, showing good things 
-%happening in one panel and bad in another - cited in Intro, only serves to prove to people that ATA matters for 
-%coexistence, so they keep reading
-%3) Fig with a grid of panels, each panel plots against sigma, and the grid controls a few values of 
-%delta and a few of mu1-mu1. Plotted is all the components of coexistence, for comparison.
-%4) Theoretical fig exploring why ATA matters to r_1\1 for the lottery model and hopefully more gernerally - try to get
-%at the intuition here. See if you can edge to more generality here, or not?
+%***DAN: Next steps:
+%Write to Ellner asking question about scaling factors
+%Decide what we ought to be computing 
+%Write Methods correspondingly
+%Talk to Jasmin about it - did I miss anything that we already computed?
+%Then work with Jasmin to get the code edited to compute that stuff
+%Write results
+%Once results have solidified, make a code suite version that is clean
 
@@
 \section{Discussion}\label{section:discussion}
-
-Need a para about multi-species models - how would things change? One difference is there are more different
-things that might have asymmetric tail association patterns. See the section "General Theory" of Ellner et al 2019 - might
-just want write a short para saying it seems like it should be possible to generalize what we've done along those lines.
-
-Might be interesting to study how tail associations of environmental variables are changing and how this might 
-alter competition, coexistence, and the maintenance of biodiversity generally. This would be a whole potential 
-research program which presumably would start with a more detailed study of how tail associations of environmental
-variables may be changing under global change.
-
-Search for comments in the rest of the text indicating things that need to be done in the Discussion
-
-Ellner et al (2016) points out that coexistence is also mediated by other factors that competition: shared enemies, 
-mutualist, facilitation etc. (these are the exact factors Ellner lists), so future work could explore 
-the importance of ATA for coexistence mediated by these factors.
-
-Something about spatial mechanisms of coexistence? We considered only mechanisms that applied in a model with
-solely temporal variation, but there are also spatial mechanisms. Need to learn about those and think about
-how to say something about how future work may be able to explore the importance of ATA in that context?
-
-Also need to talk about how this stuff will apply to data.
-
-In the Discussion, I intend to talk about how possible changes in tail association 
-patterns of env vars in nature may influence coexistence of species. It won't influence the 
-other mechanisms, though changes in the extreme values of environmental variables, known
-to be occurring, might. Should also mention this second point in the Discussion as something that merits further research.
-%though might be worth mentioning in the Discussion that events are getting more extreme. We have studied how *relationships*
-%in the extremes influence coexistence, but probably worth studying (or cite papers if it's been done) how changes
-%in the commonness and extremeness of extreme events themselves influence coexistence, and for that you'd have
-%to also look at the other mechanisms. May still need a sup mat laying out all the math.
 

### 2021-07-07 parameter_loop_w_q12.R
new looping with q12 computations

@@ -1,0 +1,113 @@
+# loop through parameters
+# computations with q12 expression and not q12=1
+
+load("./params.RData")
+
+load("./noise_etc.RData")
+
+source("./transform.R")
+
+source("./growth_rate_metrics.R")
+
+rbar <- rep(NA, 5)
+se_rbar <- rbar
+DeltaI <- rep(NA,3)
+se_DeltaI <- DeltaI
+sd_r <- DeltaI
+epECdot <- DeltaI
+invPrb <- DeltaI
+invE <- DeltaI
+q12 <- DeltaI
+params <- rep(NA, 4)
+
+for (g in 1:length(sigma)){
+  for (m in 1:nrow(mu)){
+    
+    rv <- transform(b_tilde, u_tilde, rho, sigma[g], mu[m,]) #makes random vars defined by parameters
+    list2env(rv, globalenv())
+    
+    #probability r1>0
+    invPrb_l <- length(b_l1[b_l1>b_l2])/M
+    invPrb_r <- length(b_r1[b_r1>b_r2])/M
+    invPrb_s <- length(u_s[u_s>0])/M
+    
+    for (d in 1:length(delta)){
+      #rbar1
+      #left tail (1)
+      l <- r_mets(b_l1-b_l2, delta[d], M)
+      
+      invE_l <- mean(l$r[l$r>0])
+      
+      q12_l <- mean(exp(b_l1))/((1-delta[d])*mean(exp(b_l2))+delta[d]*mean(exp(b_l1)))
+      
+      #righ tail (2)
+      r <- r_mets(b_r1-b_r2, delta[d], M)
+      
+      invE_r <- mean(r$r[r$r>0])
+      
+      q12_r <- mean(exp(b_r1))/((1-delta[d])*mean(exp(b_r2))+delta[d]*mean(exp(b_r1)))
+      
+      #symmetric (3)
+      s <- r_mets(u_s, delta[d], M)
+      
+      invE_s <- mean(s$r[s$r>0])
+      
+      q12_s <- mean(exp(b_s1))/((1-delta[d])*mean(exp(b_s2))+delta[d]*mean(exp(b_s1)))
+      
+      #rbar1sharp (4)
+      sharp1 <- r_mets(u_1, delta[d], M)
+      
+      #rbar2sharp (5)
+      sharp2 <- r_mets(u_2, delta[d], M)
+      
[…]

### 2021-07-07 parameter_loop.R
naming resutl columns

@@ -8,15 +8,20 @@ source("./transform.R")
 
 source("./growth_rate_metrics.R")
 
+noisename <- c("left","right","sym")
+
 rbar <- rep(NA, 5)
+names(rbar) <- c(noisename, "sharp1","sharp2")
 se_rbar <- rbar
 DeltaI <- rep(NA,3)
+names(DeltaI) <- noisename
 se_DeltaI <- DeltaI
 sd_r <- DeltaI
 epECdot <- DeltaI
 invPrb <- DeltaI
 invE <- DeltaI
 params <- rep(NA, 4)
+names(params) <- c("sigma","mu1","mu2","delta")
 
 for (g in 1:length(sigma)){
   for (m in 1:nrow(mu)){

### 2021-07-07 parameter_loop_w_q12.R

@@ -9,9 +9,13 @@ source("./transform.R")
 
 source("./growth_rate_metrics.R")
 
+noisename <- c("left","right","sym")
+
 rbar <- rep(NA, 5)
+names(rbar) <- c(noisename, "sharp1","sharp2")
 se_rbar <- rbar
 DeltaI <- rep(NA,3)
+names(DeltaI) <- noisename
 se_DeltaI <- DeltaI
 sd_r <- DeltaI
 epECdot <- DeltaI
@@ -19,6 +23,7 @@ invPrb <- DeltaI
 invE <- DeltaI
 q12 <- DeltaI
 params <- rep(NA, 4)
+names(params) <- c("sigma","mu1","mu2","delta")
 
 for (g in 1:length(sigma)){
   for (m in 1:nrow(mu)){

### 2021-07-07 decomposition_fxn.R
decomposition fxn to use for plot generation

@@ -1,0 +1,112 @@
+#decomposition function
+decomp <- function(sigma, mu, delta, M){
+  
+  #exponentiate noise to get B
+  load("./noise_etc.RData")
+  rv <- transform(b_tilde, u_tilde, rho, sigma, mu, b_s=TRUE) #makes random vars defined by parameters
+  Bl <- cbind(exp(rv$b_l1), exp(rv$b_l2))
+  Br <- cbind(exp(rv$b_r1), exp(rv$b_r2))
+  Bs <- cbind(exp(rv$b_s1), exp(rv$b_s2))
+  
+  #make B sharp
+  bsharp_tilde <- makenoise(M)
+  rho <- cor(bsharp_tilde$l)[1,2]
+  rvsharp <- transform(bsharp_tilde, u_tilde, rho, sigma, mu, b_s=TRUE)
+  Bl2sharp <- exp(rvsharp$b_l2)
+  Br2sharp <- exp(rvsharp$b_r2)
+  Bs2sharp <- exp(rvsharp$b_s2)
+  
+  #epsilon_0
+  null_l <- log(1-delta+delta*(mean(Bl[,1])/mean(Bl[,2])))
+  null_r <- log(1-delta+delta*(mean(Bl[,1])/mean(Bl[,2])))
+  null_s <- log(1-delta+delta*(mean(Bl[,1])/mean(Bl[,2])))
+  
+  #epsilon_C
+  epC1l <- mean(log(1-delta+delta*(mean(Bl[,1])/Bl[,2])))- null_l
+  epC1r <- mean(log(1-delta+delta*(mean(Br[,1])/Br[,2])))- null_r
+  epC1s <- mean(log(1-delta+delta*(mean(Bs[,1])/Bs[,2])))- null_s
+  
+  epC2l <- mean(log(1-delta+delta*(mean(Bl[,2])/Bl[,2])))
+  epC2r <- mean(log(1-delta+delta*(mean(Br[,2])/Br[,2])))
+  epC2s <- mean(log(1-delta+delta*(mean(Bs[,2])/Bs[,2])))
+  
+  #epsilon_E
+  epE1l <- mean(log(1-delta+delta*(Bl[,1]/mean(Bl[,2]))))- null_l
[…]

### 2021-07-07 decomp_plotting.R
starting decomposition plot

@@ -1,0 +1,17 @@
+x0 <- decomp(0.8, c(0.3,0.3), 0.5, M)
+x1 <- decomp(1.6, c(0.3,0.3), 0.5, M)
+x2 <- decomp(3.2, c(0.3,0.3), 0.5, M)
+
+plot(0, xlab="", ylab="", ylim=c(-0.08,0.35), xlim=c(0.8,3.2), col="white")
+polygon(c(0.8,0.8,1.6,3.2,3.2),c(0,x0$DEC_[1],x1$DEC_[1], x2$DEC_[1],0)
+        ,col="darkred", border=NA)
+polygon(c(0.8,0.8,1.6,3.2,3.2),c(0,x0$Dnull[1],x1$Dnull[1], x2$Dnull[1],0)
+        ,col="darkblue", border=NA)
+polygon(c(0.8,0.8,1.6,3.2,3.2),c(0,x0$DE[1],x1$DE[1], x2$DE[1],0)
+        ,col="darkorange1", border=NA)
+polygon(c(0.8,0.8,1.6,3.2,3.2),c(0,x0$DC[1],x1$DC[1], x2$DC[1],0)
+        ,col="gold", border=NA)
+polygon(c(0.8,0.8,1.6,3.2,3.2),c(0,x0$DECsharp[1],x1$DECsharp[1], x2$DECsharp[1],0)
+        ,col="darkgreen", border=NA)
+lines(c(0.8,1.6,3.2), c(sum(x0[1,]), sum(x1[1,]), sum(x2[1,])), lwd=3)
+abline(h=0, lty=2)

### 2021-07-07 parameter_loop_w_q12.R
typo

@@ -28,7 +28,7 @@ names(params) <- c("sigma","mu1","mu2","delta")
 for (g in 1:length(sigma)){
   for (m in 1:nrow(mu)){
     
-    rv <- transform(b_tilde, u_tilde, rho, sigma[g], mu[m,]) #makes random vars defined by parameters
+    rv <- transform(b_tilde, u_tilde, rho, sigma[g], mu[m,], b_s=TRUE) #makes random vars defined by parameters
     list2env(rv, globalenv())
     
     #probability r1>0

### 2021-07-07 paper/SupMat.Rnw
further progress in methods and associated sup mat

@@
 &= \olr_{j \bs i} - \olr_{j \bs i}^\nat \\
-&= 0
+&= 0 \label{eq:decomp:j:EnatC}
 \end{align}
@@
 \olep_j^{[EC]} &= \olep_j^{(EC)} - \olep_j^{[E \nat C]} \\
-&= - \E[\ln(1-\delta+\delta \exp(\sigma(\tilde{b}_j^\shp - \tilde{b}_j)))],
+&= - \E[\ln(1-\delta+\delta \exp(\sigma(\tilde{b}_j^\shp - \tilde{b}_j)))],\label{eq:decomp:j:EChard}
 \end{align}
@@
 &= \frac{\exp(-\mu_i)\exp(\mu_i)}{\exp(-\mu_i)[(1-\delta) \exp(\mu_j)+\delta \exp(\mu_i)]} \\
-&= \frac{1}{(1-\delta)\exp(-(\mu_i-\mu_j))+\delta}.
+&= \frac{1}{(1-\delta)\exp(-(\mu_i-\mu_j))+\delta}.\label{eq:decomp:qnot1}
 \end{align}
@@
 \ln(1-\delta+\delta\exp(-\sigma u^{(k)}+\sigma^2/2))]/\sqrt{M}$.
-\item Estimate
-%***DAN: DAN, proceed from here
+\item Estimate $\olep_j^{[EC]}$, following (\ref{eq:decomp:j:EChard}), as $\widehat{\olep_j^{[EC]}}=\mean_k [-\ln(1-\delta+\delta \exp(\sigma \sqrt{2} u^{(k)}))],$ which has $\se \left( \widehat{\olep_j^{[EC]}} \right) = \sd_k[-\ln(1-\delta+\delta \exp(\sigma \sqrt{2} u^{(k)}))]/\sqrt{M}$.
+\item Use $\olep_j^{[E \nat C]}=0$, following (\ref{eq:decomp:j:EnatC}). This is exact.
+\item Finally, use $\olr_{j \bs i} = 0$, again exact. 
 \end{enumerate}
+As for (\ref{eq:decomp_overall_IGR}), it is again easy to see that (\ref{eq:decomp_overall_rjbsi}) applies for the 
+estimates as well as for the quantities being estimated.
+
+Next, $q_{ij}$ can be taken (see SI section \ref{SIsect:scaling_factors}) to be, alternatively, $1$ or 
+the quantity displayed in (\ref{eq:decomp:qnot1}). Hence the terms of the decomposition 
+(\ref{eq:decomp_overall_delta}) can straightforwardly be etimated using our estimates elaborated 
+above, as follows. There are two alternative versions of the estiamtors for this decomposition, corresponding 
+to the alternative values for $q_{ij}$.
+\begin{enumerate}
+\item Compute $\Delta_i^0 = \epsilon_i^0 - q_{ij} \epsilon_j^0$, which is computed exactly, since 
+the quantities described above for the components $\epsilon_i^0$, $\epsilon_j^0$ and $q_{ij}$ are exact, as opposed to being
+estimates. Simplifying, since $\epsilon_j^0=0$, we have $\Delta_i^0 = \epsilon_i^0$.
+\item Estimate $\Delta_i^E$ as $\widehat{\Delta_i^E}=\widehat{\olep_i^E}-q_{ij}\widehat{\olep_j^E}=
+\mean_k[\ln(1-\delta+\delta\exp(\sigma u^{(k)}+\mu_i-\mu_j-\sigma^2/2))]-\epsilon_i^0-
+q_{ij} \mean_k[\ln(1-\delta+\delta\exp(\sigma u^{(k)}-\sigma^2/2))],$ which has standard error 
+$\se\left( \widehat{\Delta_i^E} \right)=\sd_k[\ln(1-\delta+\delta\exp(\sigma u^{(k)}+\mu_i-\mu_j-\sigma^2/2))-
+q_{ij} \ln(1-\delta+\delta\exp(\sigma u^{(k)}-\sigma^2/2))]/\sqrt{M}.$
+\item Estimate $\Delta_i^C$ as $\widehat{\Delta_i^C}=\widehat{\olep_i^C}-q_{ij}\widehat{\olep_j^C}=
+\mean_k[\ln(1-\delta+\delta\exp(-\sigma u^{(k)}+\mu_i-\mu_j+\sigma^2/2))]-
+\epsilon_i^0-q_{ij}\mean_k[\ln(1-\delta+\delta\exp(-\sigma u^{(k)}+\sigma^2/2))],$ which has standard error
+$\se \left( \widehat{\Delta_i^C} \right)=\sd_k[\ln(1-\delta+\delta\exp(-\sigma u^{(k)}+\mu_i-\mu_j+\sigma^2/2))-
+q_{ij}\ln(1-\delta+\delta\exp(-\sigma u^{(k)}+\sigma^2/2))]/\sqrt{M}$.
+\item Estimate $\Delta_i^{(E \shp C)}$ as $\widehat{\Delta_i^{(E \shp C)}} = \widehat{\olep_i^{(E \shp C)}}-q_{ij} \widehat{\olep_j^{(E \shp C)}}=
+\mean_k[\ln(1-\delta+\delta\exp(\sigma\sqrt{2}u^{(k)}+\mu_i-\mu_j))]-\widehat{\olep_i^E}-\widehat{\olep_i^C}-\epsilon_i^0-
+q_{ij}[\mean_k [\ln(1-\delta+\delta\exp(\sigma\sqrt{2}u^{(k)}))]-\widehat{\olep_j^E}-\widehat{\olep_j^C}-\epsilon_j^0].$
+This simplifies to $\mean_k[\ln(1-\delta+\delta\exp(\sigma\sqrt{2}u^{(k)}+\mu_i-\mu_j))]-
+q_{ij}\mean_k[\ln(1-\delta+\delta\exp(\sigma\sqrt{2}u^{(k)}))]-\widehat{\Delta_i^E}-\widehat{\Delta_i^C}-\Delta_i^0$, 
+which can be straightforwardly shown has standard error
+$\se \left( \widehat{\Delta_i^{(E \shp C)}} \right) = 
+\sd_k[\ln(1-\delta+\delta\exp(\sigma\sqrt{2}u^{(k)}+\mu_i-\mu_j))-
+\ln(1-\delta+\delta\exp(\sigma u^{(k)}+\mu_i-\mu_j-\sigma^2/2))-
+\ln(1-\delta+\delta\exp(-\sigma u^{(k)}+\mu_i-\mu_j+\sigma^2/2))-
+q_{ij}\ln(1-\delta+\delta\exp(\sigma\sqrt{2}u^{(k)}))+
+q_{ij}\ln(1-\delta+\delta\exp(\sigma u^{(k)}-\sigma^2/2))+
+q_{ij}\ln(1-\delta+\delta\exp(-\sigma u^{(k)}+\sigma^2/2))]/\sqrt{M}$.
+\item Estimate $\Delta_i^{[EC]}$ as $\widehat{\Delta_i^{[EC]}}=\widehat{\olep_i^{[EC]}}-q_{ij}\widehat{\olep_j^{[EC]}}=
+\mean_k[\ln(1-\delta+\delta\exp(\sigma(\tilde{\ddot{b}}_i^{(k)}-\tilde{\ddot{b}}_j^{(k)})+\mu_i-\mu_j))]-
+\mean_k[\ln(1-\delta+\delta\exp(\sigma\sqrt{2}u^{(k)}+\mu_i-\mu_j))]+
+q_{ij}\mean_k[\ln(1-\delta+\delta\exp(\sigma\sqrt{2}u^{(k)}))],$ which one can straightforwardly show has standard error
+$\sqrt{s_1^2+s_2^2}$, where $s_1=\sd_k[\ln(1-\delta+\delta\exp(\sigma(\tilde{\ddot{b}}_i^{(k)}-\tilde{\ddot{b}}_j^{(k)})+\mu_i-\mu_j))]/\sqrt{M}$ 
+and $s_2=\sd_k[\ln(1-\delta+\delta\exp(\sigma\sqrt{2}u^{(k)}+\mu_i-\mu_j))-q_{ij}\ln(1-\delta+\delta\exp(\sigma\sqrt{2}u^{(k)}))]/\sqrt{M}$.
+\item For $\Delta_i^{[E \nat C]}$, we recall that $\olep_j^{[E \nat C]}=0$, so $\Delta_i^{[E \nat C]}=\olep_i^{[E \nat C]}$. So we use
+$\widehat{\olep_i^{[E \nat C]}}$ for $\widehat{\Delta_i^{[E \nat C]}}$.
+\item Finally, we can also estimate $\olr_{i \bs i}-q_{ij}\olr_{j \bs i}$ simply as $\widehat{\olr_{i \bs i}}$
+since $\olr_{j \bs i}=0$.
+\end{enumerate}
+As for (\ref{eq:decomp_overall_IGR}) and (\ref{eq:decomp_overall_rjbsi}), it is again the case that (\ref{eq:decomp_overall_delta})
+applies for the estimators as well as for the quantities being estimated.
+
+%***DAN: Most but not all of the below is obsolete now, replaced and upgraded by stuff above this point
 
 \section{Invasion growth rate of species 1, and storage effects} \label{sect:IGR}
-
-%***DAN: This and probably all or most of the following sections will have to be edited, not least to make the
-%notation consistent with the notation now used in the Theory section of the main text
-%***DAN: At the end of Theory, I mention that in methods the tricks for computational efficiency are
-%elaborated, but actually most of those will be in a sup mat section, the material for which will be 
-%an edited version of a lot of what is in this section.
 

### 2021-07-07 paper/Paper.Rnw
chose delta and mudif vals
further progress in methods and associated sup mat

--- 2021-07-06 paper/Paper.Rnw
@@
 %The following is where you load in the numeric results that will be embedded in the text
-<<eval=T,echo=F,message=F,warning=F>>=
-system(paste("biber", sub("\\.Rnw$", "", current_input())))
+%<<eval=T,echo=F,message=F,warning=F>>=
+%system(paste("biber", sub("\\.Rnw$", "", current_input())))
 
-#place R code here for loading in necessary variables from Results
-#ordinarily this would be a readRDS command, but I'm just defining a variable here for now
-allregres<-matrix(rnorm(10),2,5)
-@
+%#place R code here for loading in necessary variables from Results
+%#ordinarily this would be a readRDS command, but I'm just defining a variable here for now
+%allregres<-matrix(rnorm(10),2,5)
+%@
 
@@
 for both our left- and right-tail associated noises (SI section \ref{SI-sect:math_simplifications}). So we only generated plots 
-for left-tail associated noise. For each combination of $\delta=W, X, Y, Z$
-%***DAN: Jasmin, please select 4 sensible values, including \delta=1, and put the ones you select here
-and $\mu_1-\mu_2=A,B,C,D$,
-%***DAN: Jasmin, please select 3-4 sensible values and put the ones you select here
+for left-tail associated noise. For each combination of $\delta=0.2, 0.5, 0.8, 1.0$
+and $\mu_1-\mu_2=0.0,-0.3,-0.6,-0.8$,
 we plotted all the components of (\ref{eq:full_decomp}) against $\sigma$. Relative importance of contributions was judged from the 
@@
 the contributions of other mechanisms.
-%***DAN: Jasmin, if you select 4 values for mu1-mu2 and 4 for delta, then Fig. 3 is going to have 16 panels arranged in a 
-%4 by 4 grid, each of which will plot all the components against sigma. You'll need to use different colors and/or line
-%types for the components and have a legend.
+%***DAN: If we select 4 values for mu1-mu2 and 4 for delta, then Fig. 3 is going to have 16 panels arranged in a 
+%4 by 4 grid, each of which will plot all the components against sigma. We'll need to use different colors and/or line
+%types for the components and have a legend. Sort it all out as one big multi-panel pdf fig, after running all the sims.
 The simulation methods of \cite{Ellner_2019} were used to compute the components of (\ref{eq:full_decomp}).
@@
 SI section \ref{SI-sect:efficient_computation}.
-%***DAN: Dan, you still need to write that sup mat section
+
+%What's next
+%1) Write code implementing the above and making a figure - that'll be Fig. 3
+%2) Get Jasmin to work on Fig. 1 and Fig. 2 (Fig. 2 is largely done). Fig 1 should be facilitated by the work to make Fig. 3 in the previous item.
+%3) Decide what other results we will put in the main text and write a paragraph on methods for the other results (for both the main text
+%and sup mat results), put it above.
+%4) Think about this: Theoretical fig exploring why ATA matters to r_1\1 for the lottery model and hopefully more gernerally - try to get
+%at the intuition here. See if you can edge to more generality here, or not?
+
+%Ideas while driving to pick up Max:
+%1) I think in every case for the lottery model, making tail associations asymmetric reduced the potential for coexistence - I think that's what Jasmin told me, but if we present it we should make sure I'm not remembering wrong, of course.
+Is that generally the case, for other models?
+What are the features of a model, or of the growth rate of the invader, that make it so that more asymmetric tail associations means reduced coexistence?
+Does this have implications for climate change?
+%2) In this paper with Jasmin you need to cite the two papers you just got from the professor at University of Massachusetts, Lowell.
+I mean the papers about changes in the special extent of heat waves: https://agupubs.onlinelibrary.wiley.com/doi/full/10.1029/2020GL087097 and https://iopscience.iop.org/article/10.1088/1748-9326/ab4b41
+%3) Discussion content: We should (in future) study tail associations between environmental variables, and how these might be changing.
+However, more than that, the variables E in the lottery model are not actual environmental variables but rather are the effects of environmental variables on the species of the model.
+Probably real environmental variables are actually related to vital rates in a non-linear and probably often threshold-like way.
+Threshold relationships can produce tail association, as we have showed elsewhere.
+Changes in the means of true environmental variables may change where those distributions of environmental variables are compared to thresholds, which may change the nature of tail associations between the effects of those environmental variables.
 
@@
 \end{figure}
+%possible examples:
+%population simulations - shows symmetric with faster fluctatuations to extremes of coexistence while asymmetric has longer periods of noticeable coexistence and dominace/exclusion; may be unclear
 

### 2021-07-08 paper/Paper.Rnw
added some  intro

@@
 \section{Introduction}\label{section:introduction}
+
+The maintenance of biodiversity in competitive ecosystems is a central idea in ecology.
+Among others, the modern theory of coexistence seeks to assess and quantify the mechanisms that result in coexisting species and thus the maintenance of biodiversity.
+Many are familiar with the principle of competitive exclusion - an idea which postulates if species within the same trophic level and similar niches exist, one will eventually outcompete the other (Hardin 1960).
+Modern coexistence theory provides a quantitative breakdown of contributing mechanisms as criteria for coexistence and competitive exclusion (Chesson 1994).
+
+Equally central to ecology is the study of relationships between ecological variables; however, it is almost exclusively blanketed by simple measures of associativity.
+Recent work has shown that tail associations, or associations between the tail ends of data, and specifically the asymmetry between tail associations have important considerations for ecological phenomena; although, to our knowledge, the consequences/implications of asymmetric tail associations for competition and coexistence have not yet been investigated.
+Correlation measures such as Pearson or Spearman correlation disguise the asymmetry of tail associations, and datasets with different tail associations can have the same correlation coefficient (see Figure).
+Asymmetry in tail associations describes tighter correlations in regions of a dataset compared to other regions of the dataset.
+A simple example of asymmetric tail associations is a dataset which is more correlated in the lower (upper) half than the upper (lower) half, described as left (right) tail associated.
+Current research has shown that ATAs have important considerations for extinction risks, “skewness of spatial averages”, Taylor’s Law, Moran Effect/ spatial synchrony (Gosh et al 2020 Ecosphere, Gosh et al 2020 And.
+EcoRes,).
+Despite the existence of tail associations within ecological data, to our knowledge, the implications of tail associations for competition and coexistence has not yet been examined; doing so is the main goal of this paper.
+
+Asymmetries in tail associations have the potential to influence competition and coexistence via one particular mechanism, namely, storage effects.
+Covariation between environmental fluctuations and competitive pressures allows for storage effect to be a contributing stabalizing mechanism to coexistence.
+Because the contributive value of storage effects relies on the association of competition and environment (Chesson 1994, Barbaras ..), it is reasonable to suspect the contribution of storage effects to coexistence can be altered by an association between competition and environment that is asymmetric.
+While we aimed to investigate the importance of ATAs on coexistence in general and all the mechanisms which compose the modern theory of coesustence, it was quickly revealed that ATAs are specifically important for storage effects and not for other mechanisms on coexistence (due to the fact that storage effect is the only mechanism that depends of the variance between two drivers).
+
+While there are multiple studies on the influence of extremes (due to climate change), the influence of relationships in their extremes (also due to climate change), has not yet been explored.
+The lottery model within the framework of coexistence theory provides a suitable opportunity to explore extreme relationships and the consequences to coexistence.
 

### 2021-07-08 paper/Paper.Rnw
more intro work

@@
 
-While there are multiple studies on the influence of extremes (due to climate change), the influence of relationships in their extremes (also due to climate change), has not yet been explored.
+[As we know], anthropogenic climate change has large consequences for almost all aspects of life.
+Extreme climatic events have only become more extreme - increasing in magnitude and frequency - as climate change has progressed and have contributed to an increase in global means as well an an increase in environmental variability.
+While there are multiple studies on the influence of environmental extremes and varaibility (Parepa et al 2013, Diez et al 2012), the influence of relationships and interactions between drivers in their extremes, has not yet been explored.
+Encapsulated by ATAs Tail associations are relationships in the extremes .
+Environmental variability has been classically shown to promote coexistence via storage mechanisms (Chesson & Warner 1981, Warner & Chesson 1985), thus we continue/move forward with this simple model to examine extreme relationships between environmental variables on the contribution of storage effects to coexistence.
+
 The lottery model within the framework of coexistence theory provides a suitable opportunity to explore extreme relationships and the consequences to coexistence.

### 2021-07-08 paper/Paper.Rnw
workingon intro

--- 2021-07-08 paper/Paper.Rnw
@@
 
-While there are multiple studies on the influence of extremes (due to climate change), the influence of relationships in their extremes (also due to climate change), has not yet been explored.
+paragraph to inuitively explain storage effects
+We emphasize that although this paper focuses on storage effects, it is still a paper generally about the importance of asymmetrical tail associations for coexistence because we do consider whether they are important for other mechanisms of modern coexistence theory, though we quickly realize their importance is solely for storage effects
+
+[As we know], anthropogenic climate change has large consequences for almost all aspects of life including via extreme climatic events, which have only become more extreme - increasing in magnitude and frequency - and have contributed to the increase in global means as well as an increase in environmental variability.
+While there are multiple studies on the influence of environmental extremes and varaibility on competition and coexistence (Parepa et al 2013, Diez et al 2012), the influence of relationships and interactions between drivers in their extremes, has not yet been explored.
+By definition, interacting environmental drivers that are  moreso related in the extreme values are tail associated, thus we have implemented our tools for the study of tail associations to explore the importance of environmental relationships in their exrtemes on competition and coexistence.
+Environmental variability has been classically shown to promote coexistence via storage mechanisms (Chesson and Warner 1981, Warner and Chesson 1985), hence we continue/move forward with this simple model to examine the effect of extreme relationships between environmental variables on the contribution of storage effects to coexistence.
+
 The lottery model within the framework of coexistence theory provides a suitable opportunity to explore extreme relationships and the consequences to coexistence.
 
+Results main ideas + example
+
+introduce new decomposition
+
+final parahraph:goals
 %Jasmin, I think the below makes a pretty good start at tryinbg to nail down a good way to express our goals

### 2021-07-08 parameters.R
shortened param choices

@@ -15,4 +15,9 @@ for (i in 1:length(mu1)){
 mu <- mu[-1,]
 #print(mu) #45x2
 
-save(delta, mu, sigma, file = "./params.RData")
+delta <- seq(0.2, 0.5, 0.8, 1.0)
+sigma <- seq(1.6, 3.2, 6.4)
+mudif <- seq(0.0, -0.3, -0.6, -0.8)
+
+
+save(delta, mu, mudif, sigma, file = "./params.RData")

### 2021-07-08 parameters.R
typo

@@ -15,9 +15,9 @@ for (i in 1:length(mu1)){
 mu <- mu[-1,]
 #print(mu) #45x2
 
-delta <- seq(0.2, 0.5, 0.8, 1.0)
-sigma <- seq(1.6, 3.2, 6.4)
-mudif <- seq(0.0, -0.3, -0.6, -0.8)
+delta <- c(0.2, 0.5, 0.8, 1.0)
+sigma <- c(1.6, 3.2, 6.4)
+mudif <- c(0.0, -0.3, -0.6, -0.8)
 
 
 save(delta, mu, mudif, sigma, file = "./params.RData")

### 2021-07-08 param_loop_short_w_q12.R
new loop for short params

@@ -1,0 +1,118 @@
+# loop through parameters
+# computations with q12 expression and not q12=1
+
+load("./params.RData")
+
+load("./noise_etc.RData")
+
+source("./transform2.R")
+
+source("./growth_rate_metrics.R")
+
+noisename <- c("left","right","sym")
+
+rbar <- rep(NA, 5)
+names(rbar) <- c(noisename, "sharp1","sharp2")
+se_rbar <- rbar
+DeltaI <- rep(NA,3)
+names(DeltaI) <- noisename
+se_DeltaI <- DeltaI
+sd_r <- DeltaI
+epECdot <- DeltaI
+invPrb <- DeltaI
+invE <- DeltaI
+q12 <- DeltaI
+params <- rep(NA, 4)
+names(params) <- c("sigma","mu1","mu2","delta")
+
+for (g in 1:length(sigma)){
+  for (m in 1:length(mudif)){
+    
+    rv <- transform(b_tilde, u_tilde, rho, sigma[g], mudif[m], b_s=TRUE) #makes random vars defined by parameters
+    list2env(rv, globalenv())
+    
+    #probability r1>0
+    invPrb_l <- length(b_l1[b_l1>b_l2])/M
+    invPrb_r <- length(b_r1[b_r1>b_r2])/M
+    invPrb_s <- length(u_s[u_s>0])/M
+    
+    for (d in 1:length(delta)){
+      #rbar1
+      #left tail (1)
+      l <- r_mets(b_l1-b_l2, delta[d], M)
+      
+      invE_l <- mean(l$r[l$r>0])
+      
+      q12_l <- mean(exp(b_l1))/((1-delta[d])*mean(exp(b_l2))+delta[d]*mean(exp(b_l1)))
+      
+      #righ tail (2)
+      r <- r_mets(b_r1-b_r2, delta[d], M)
+      
+      invE_r <- mean(r$r[r$r>0])
+      
+      q12_r <- mean(exp(b_r1))/((1-delta[d])*mean(exp(b_r2))+delta[d]*mean(exp(b_r1)))
+      
+      #symmetric (3)
+      s <- r_mets(u_s, delta[d], M)
+      
+      invE_s <- mean(s$r[s$r>0])
+      
[…]

### 2021-07-09 decomp_plotting.R
added source function

@@ -1,3 +1,5 @@
+source("./decomposition_fxn.R")
+
 x0 <- decomp(0.8, c(0.3,0.3), 0.5, M)
 x1 <- decomp(1.6, c(0.3,0.3), 0.5, M)
 x2 <- decomp(3.2, c(0.3,0.3), 0.5, M)

### 2021-07-09 parameters.R
modifies sigma vals

@@ -16,7 +16,7 @@ mu <- mu[-1,]
 #print(mu) #45x2
 
 delta <- c(0.2, 0.5, 0.8, 1.0)
-sigma <- c(1.6, 3.2, 6.4)
+sigma <- c(0.8, 1.6, 3.2, 6.4)
 mudif <- c(0.0, -0.3, -0.6, -0.8)
 
 

### 2021-07-12 exploring_q12.R
checking different formulas for q12

@@ -1,0 +1,37 @@
+source("./makenoise.R") 
+load("./noise_etc.RData")
+m1 <- 0.3; m2 <- 0.9; sig <- 6.4
+
+rv <- transform(b_tilde, u_tilde, rho, sig, c(m1,m2), b_s=TRUE)
+list2env(rv, globalenv())
+d <- 0.4
+#bar B
+q12_l <- mean(exp(b_l1))/((1-d)*mean(exp(b_l2))+d*mean(exp(b_l1)))
+
+q12_r <- mean(exp(b_r1))/((1-d)*mean(exp(b_r2))+d*mean(exp(b_r1)))
+
+q12_s <- mean(exp(b_s1))/((1-d)*mean(exp(b_s2))+d*mean(exp(b_s1)))
+B<-c(q12_l, q12_r, q12_s)
+
+#exp(bar b + sd(b)^2/2)
+q12_l <- exp(mean(b_l1)+var(b_l1)/2)/((1-d)*exp(mean(b_l2)+var(b_l2)/2)+d*exp(mean(b_l1)+var(b_l1)/2))
+
+q12_r <- exp(mean(b_r1)+var(b_r1)/2)/((1-d)*exp(mean(b_r2)+var(b_r2)/2)+d*exp(mean(b_r1)+var(b_r1)/2))
+
+q12_s <- exp(mean(b_s1)+var(b_s1)/2)/((1-d)*exp(mean(b_s2)+var(b_s2)/2)+d*exp(mean(b_s1)+var(b_s1)/2))
+b<-c(q12_l, q12_r, q12_s)
+
+#exp(mu + sig^2/2)
+q12_l <- exp(m1+(sig^2)/2)/((1-d)*exp(m2+(sig^2)/2)+d*exp(m1+(sig^2)/2))
+
+q12_r <- exp(m1+(sig^2)/2)/((1-d)*exp(m2+(sig^2)/2)+d*exp(m1+(sig^2)/2))
+
+q12_s <- exp(m1+(sig^2)/2)/((1-d)*exp(m2+(sig^2)/2)+d*exp(m1+(sig^2)/2))
+mu<-c(q12_l, q12_r, q12_s)
+
+rbind(B,b,mu)
+
+
+#str(res)
+#round(head(cbind(res$params,res$DeltaI, resq$DeltaI)),3)
+#round(head(cbind(res$params,res$epECdot[,1:2], resq$epECdot[,1:2]),100),3)

### 2021-07-12 decomposition_fxn.R
working on decomp function

@@ -1,4 +1,4 @@
-#decomposition function
+#decomposition function #need to check
 decomp <- function(sigma, mu, delta, M){
   
   #exponentiate noise to get B
@@ -109,4 +109,143 @@ decomp <- function(sigma, mu, delta, M){
 }
 
 #decomp(1.6, c(0.3,0.4), 0.4, M)
+
+decompose <- function(mudif,sigma,delta,b_tilde,u) {
+  
+  bi_til<-b_tilde$l[,1]
+  bj_til<-b_tilde$l[,2]
+  
+  bi_um<-b_tilde$s[,1]
+  bj_um<-b_tilde$s[,2]
+  
+  #epsilon^C_i bar (45)
+  #hat
+  e_Ci_hat <- mean(log(1-delta+delta*exp(-sigma*u + mudif + (sigma^2)/2))) - e_0i
+  #SE hat
+  e_Ci_se <- sd(log(1-delta+delta*exp(-sigma*u + mudif + (sigma^2)/2)))/sqrt(M)
+  
+  #epsilon^(E#C) bar (51)
+  #hat
+  e_ECsharpi_hat <- mean(log(1-delta+delta*exp(sigma*sqrt(2)*u + mudif))) - e_Ei_hat - e_Ci_hat - e_0i
+  #SE hat
+  e_ECsharpi_se <- sd( log(1-delta+delta*exp(sigma*sqrt(2)*u + mudif)) 
+                    - log(1-delta+delta*exp(sigma*u + mudif - (sigma^2)/2)) 
+                    - log(1-delta+delta*exp(-sigma*u + mudif + (sigma^2)/2)) )/sqrt(M)  
+  
+  #epsilon^[EC] bar (55)
+  #hat
+  e_ECi_hat <- mean(log(1-delta+delta*exp(sigma*(bi_um - bj_um)+mudif)))-mean(log(1-delta+delta*exp(sigma*sqrt(2)*u+mudif)))
+  #SE hat
+  s1 <- sd(log(1-delta+delta*exp(sigma*(bi_um - bj_um)+mudif)))/sqrt(M)
+  s2 <- sd(log(1-delta+delta*exp(sigma*sqrt(2)*u+mudif)))/sqrt(M)
+  e_ECi_se <- sqrt(s1^2 + s2^2)
+  
+  #epsilon^[E||C] bar (54)
+  #hat
[…]

### 2021-07-12 decomp_plotting.R
writing decomp plot function

@@ -1,19 +1,64 @@
+#THIS IS A FUNCTION
+
 source("./decomposition_fxn.R")
+dePlot <- function(mudif, delta){
+  load("./noise_etc.RData")
+  store <- vector(mode='list', length=7)
+  sigma <- seq(0,7,1)
+  
+  for (i in 1:length(sigma)){
+    store[[i]] <- decompose(mudif,sigma[i],delta,b_tilde,u_tilde)
+  }
+  range <- range(unlist(lapply(store, function(X){(X$D)})))
 
-x0 <- decomp(0.8, c(0.3,0.3), 0.5, M)
-x1 <- decomp(1.6, c(0.3,0.3), 0.5, M)
-x2 <- decomp(3.2, c(0.3,0.3), 0.5, M)
+  plot(0, xlab="", ylab="", ylim=range*1.1, xlim=c(0,7), col="white")
+  
+  for (i in 1:length(sigma)){
+    if(i>=2){
+      dat <- cbind(store[[i-1]]$D, store[[i]]$D)
+      lines(c(sigma[i-1], sigma[i]), dat[1,1:2])#null
+      lines(c(sigma[i-1], sigma[i]), dat[2,1:2], lty=2)#E
+      lines(c(sigma[i-1], sigma[i]), dat[3,1:2], lty=4)#C
+      lines(c(sigma[i-1], sigma[i]), dat[4,1:2], lty=3)#(E#C)
+      lines(c(sigma[i-1], sigma[i]), dat[5,1:2], col="blue")#[EC]
+      lines(c(sigma[i-1], sigma[i]), dat[6,1:2], col="red")#[E||C]
+      lines(c(sigma[i-1], sigma[i]), dat[7,1:2], col="orange",lwd=2)#r
+    }
+  }
+  abline(h=0, col="gray",lwd=0.7)
+  title(main=paste("mu1-mu2=",mudif,"delta=",delta), cex.main=0.7)
+  legend("topleft", 
+         legend=c(expression(Delta[0]), expression(Delta[E]),
+                  expression(Delta[C]), expression(Delta[("E#C")]),
+                  expression(Delta["[EC]"]), expression(Delta["[E||C]"])
[…]

### 2021-07-12 decomposition_fxn.R

@@ -114,9 +114,20 @@ decompose <- function(mudif,sigma,delta,b_tilde,u) {
   
   bi_til<-b_tilde$l[,1]
   bj_til<-b_tilde$l[,2]
+  bi_um <-b_tilde$s[,1]
+  bj_um <-b_tilde$s[,2]
+  M <- nrow(b_tilde$s)
   
-  bi_um<-b_tilde$s[,1]
-  bj_um<-b_tilde$s[,2]
+  #invader i
+  #epsilon^0_i
+  e_0i <- log(1-delta+delta*exp(mudif))
+  
+  #epsilon^E_i bar (42)
+  #hat
+  e_Ei_hat <- mean(log(1-delta+delta*exp(sigma*u + mudif - (sigma^2)/2))) - e_0i
+  #SE hat
+  e_Ei_se <- sd(log(1-delta+delta*exp(sigma*u + mudif - (sigma^2)/2)))/sqrt(M)
+  
   
   #epsilon^C_i bar (45)
   #hat
@@ -129,8 +140,8 @@ decompose <- function(mudif,sigma,delta,b_tilde,u) {
   e_ECsharpi_hat <- mean(log(1-delta+delta*exp(sigma*sqrt(2)*u + mudif))) - e_Ei_hat - e_Ci_hat - e_0i
   #SE hat
   e_ECsharpi_se <- sd( log(1-delta+delta*exp(sigma*sqrt(2)*u + mudif)) 
-                    - log(1-delta+delta*exp(sigma*u + mudif - (sigma^2)/2)) 
-                    - log(1-delta+delta*exp(-sigma*u + mudif + (sigma^2)/2)) )/sqrt(M)  
+                       - log(1-delta+delta*exp(sigma*u + mudif - (sigma^2)/2)) 
+                       - log(1-delta+delta*exp(-sigma*u + mudif + (sigma^2)/2)) )/sqrt(M)  
   
   #epsilon^[EC] bar (55)
   #hat
@@ -198,6 +209,7 @@ decompose <- function(mudif,sigma,delta,b_tilde,u) {
   DECsharp <- e_ECsharpi_hat - e_ECsharpj_hat #Deltai_(E#C)
   DEC <- e_ECi_hat - e_ECj_hat #Deltai_[EC]
   DECpip <- e_ECpipi_hat  #Deltai_[E||C]
+  Dr <- r_i_hat 
   
[…]

### 2021-07-13 Figure3.R
configuring figure 3

@@ -1,0 +1,17 @@
+load("./params.RData")
+source("./decomp_plotting.R")
+base <- par()
+
+pdf("fig3.pdf")
+
+par(mfrow=c(4,4))
+
+res <- vector(mode='list',length=1)
+for (m in 1:length(mudif)){
+  for (d in 1:length(delta)){
+    res <- append(res,dePlot(mudif[m], delta[d]))
+   }
+}
+
+dev.off()
+res <- res[-1]

### 2021-07-13 decomp_plotting.R

@@ -1,6 +1,7 @@
 #THIS IS A FUNCTION
 
 source("./decomposition_fxn.R")
+
 dePlot <- function(mudif, delta){
   load("./noise_etc.RData")
   store <- vector(mode='list', length=7)
@@ -34,10 +35,12 @@ dePlot <- function(mudif, delta){
                   ,expression(r)), 
          col = c("black","black","black","black","blue","red","orange"),
          lty = c(1,2,4,3,1,1,1), bty="n", cex=0.8)
+  
+  return(store)
 }
 
 
-dePlot(-0.3,0.8)
+#dePlot(-0.3,0.8)
 
 
 

### 2021-07-13 decomp_plotting.R
adding legend option

@@ -2,7 +2,7 @@
 
 source("./decomposition_fxn.R")
 
-dePlot <- function(mudif, delta){
+dePlot <- function(mudif, delta, legend=FALSE){
   load("./noise_etc.RData")
   store <- vector(mode='list', length=7)
   sigma <- seq(0,7,1)
@@ -28,20 +28,23 @@ dePlot <- function(mudif, delta){
   }
   abline(h=0, col="gray",lwd=0.7)
   title(main=paste("mu1-mu2=",mudif,"delta=",delta), cex.main=0.7)
-  legend("topleft", 
-         legend=c(expression(Delta[0]), expression(Delta[E]),
-                  expression(Delta[C]), expression(Delta[("E#C")]),
-                  expression(Delta["[EC]"]), expression(Delta["[E||C]"])
-                  ,expression(r)), 
-         col = c("black","black","black","black","blue","red","orange"),
-         lty = c(1,2,4,3,1,1,1), bty="n", cex=0.8)
+  
+  if(legend==TRUE){
+    legend("topleft", 
+           legend=c(expression(Delta[0]), expression(Delta[E]),
+                    expression(Delta[C]), expression(Delta[("E#C")]),
+                    expression(Delta["[EC]"]), expression(Delta["[E||C]"])
+                    ,expression(r)), 
+           col = c("black","black","black","black","blue","red","orange"),
+           lty = c(1,2,4,3,1,1,1), bty="n", cex=0.8)
+  }
+  
   
   return(store)
 }
 
 
 #dePlot(-0.3,0.8)
-
 
 
 

### 2021-07-14 paper/Paper.Rnw
working on sentence structure in intro

@@
 
-The maintenance of biodiversity in competitive ecosystems is a central idea in ecology.
-Among others, the modern theory of coexistence seeks to assess and quantify the mechanisms that result in coexisting species and thus the maintenance of biodiversity.
-Many are familiar with the principle of competitive exclusion - an idea which postulates if species within the same trophic level and similar niches exist, one will eventually outcompete the other (Hardin 1960).
-Modern coexistence theory provides a quantitative breakdown of contributing mechanisms as criteria for coexistence and competitive exclusion (Chesson 1994).
+Central to ecology is the maintenance of biodiversity in competitive ecosystems. 
+Populations subject to strong competitive pressure may succumb to competitive exclusion - an idea which postulates if species within the same trophic level and similar niches exist, one will eventually outcompete the other (Hardin 1960).
+Mediation of competitive pressure and exclusion allow for species to stably coexist. 
+The criteria for coexistence and competitive exclusion can be broken down into contributing mechanisms provided by modern coexistence theory, a framework which seeks to assess and quantify mechanisms that result in coexisting species and thus the maintenance of biodiversity (Chesson 1994).
 
-Equally central to ecology is the study of relationships between ecological variables; however, it is almost exclusively blanketed by simple measures of associativity.
-Recent work has shown that tail associations, or associations between the tail ends of data, and specifically the asymmetry between tail associations have important considerations for ecological phenomena; although, to our knowledge, the consequences/implications of asymmetric tail associations for competition and coexistence have not yet been investigated.
-Correlation measures such as Pearson or Spearman correlation disguise the asymmetry of tail associations, and datasets with different tail associations can have the same correlation coefficient (see Figure).
-Asymmetry in tail associations describes tighter correlations in regions of a dataset compared to other regions of the dataset.
-A simple example of asymmetric tail associations is a dataset which is more correlated in the lower (upper) half than the upper (lower) half, described as left (right) tail associated.
-Current research has shown that ATAs have important considerations for extinction risks, “skewness of spatial averages”, Taylor’s Law, Moran Effect/ spatial synchrony (Gosh et al 2020 Ecosphere, Gosh et al 2020 And.
+Equally central to ecology is the study of relationships between ecological variables; however, it is almost exclusively blanketed by simple measures of associativity, i.e., correlation coefficients.
+Although two sets of variables may have equal correlation coefficients, they can show different association structures (see Figure).
+More specifically, the association within the tail ends of data and the asymmetry between the tail associations, termed asymmetric tail associations (ATAs), tell a more interesting story about the relationship between variables...and?...
+Recent work has shown that ATAs have important considerations for ecological phenomena; although, the implications of ATAs for competition and coexistence have not yet been investigated. 
+Asymmetry of tail associations within a dataset of variables describes uneven correlations between the tails of a dataset.
+For a simple example, a dataset with more association in the lower (upper) half and lesser association in the upper (lower) half is termed left (right) tail associated. 
+The application of ATAs in ecology is relatively new, and in exploration of ATAs in ecology, Gosh et al (AER) found ATAs in all eight ecological datasets they considered.
+Thus, the existence and functioning role of ATAs in nature is reasonable to assume and question. 
+In fact, ATAs can have influence on ecosystem service stability, Taylor’s law, and extinction risks via extreme climatic events.
+Conversely, ATA’s can be influenced by Moran Effect and Leibeig’s Law of Minimum (Gosh et al 2020 Ecosphere, Gosh et al 2020 And.
 EcoRes,).
-Despite the existence of tail associations within ecological data, to our knowledge, the implications of tail associations for competition and coexistence has not yet been examined; doing so is the main goal of this paper.
+Despite the frequent occurrence of ATAs in ecological data, the implications of tail associations for competition and coexistence, to our knowledge, has not yet been examined; doing so is the main goal of this paper.
 

### 2021-07-14 paper/Paper.Rnw
more work on intro and sentence strucutre

@@
 
-Asymmetries in tail associations have the potential to influence competition and coexistence via one particular mechanism, namely, storage effects.
-Covariation between environmental fluctuations and competitive pressures allows for storage effect to be a contributing stabalizing mechanism to coexistence.
-Because the contributive value of storage effects relies on the association of competition and environment (Chesson 1994, Barbaras ..), it is reasonable to suspect the contribution of storage effects to coexistence can be altered by an association between competition and environment that is asymmetric.
-While we aimed to investigate the importance of ATAs on coexistence in general and all the mechanisms which compose the modern theory of coesustence, it was quickly revealed that ATAs are specifically important for storage effects and not for other mechanisms on coexistence (due to the fact that storage effect is the only mechanism that depends of the variance between two drivers).
+Asymmetries in tail associations can potentially influence competition and coexistence via one particular mechanism, namely, storage effects.
+Storage effect contributes as a stabilizing mechanism to coexistence when there is covariation between competitive pressures and environmental fluctuations.
+Because the contributive value of storage effects relies on the relationship between environment and competition (Chesson 1994, Barbaras ..), it reasonably follows to suspect the contribution of storage effectsto be altered by an association between environment and competition that is asymmetric, i.e., ATAs between environment and competition.
+Uniquely, storage effect is the only mechanism of modern coexistence theory which depends on the variance between two drivers rather than one.
+While we aimed to investigate the importance of ATAs on coexistence in general and all the mechanisms which compose the modern theory of coexistence, it was quickly revealed that ATAs are specifically important for storage effects and not for other mechanisms of coexistence.
 
@@
 
-[As we know], anthropogenic climate change has large consequences for almost all aspects of life including via extreme climatic events, which have only become more extreme - increasing in magnitude and frequency - and have contributed to the increase in global means as well as an increase in environmental variability.
-While there are multiple studies on the influence of environmental extremes and varaibility on competition and coexistence (Parepa et al 2013, Diez et al 2012), the influence of relationships and interactions between drivers in their extremes, has not yet been explored.
-By definition, interacting environmental drivers that are  moreso related in the extreme values are tail associated, thus we have implemented our tools for the study of tail associations to explore the importance of environmental relationships in their exrtemes on competition and coexistence.
-Environmental variability has been classically shown to promote coexistence via storage mechanisms (Chesson and Warner 1981, Warner and Chesson 1985), hence we continue/move forward with this simple model to examine the effect of extreme relationships between environmental variables on the contribution of storage effects to coexistence.
+[As we know], anthropogenic climate change has large consequences for almost all aspects of life including via extreme climatic events, which have only become more extreme - increasing in magnitude and frequency - and have contributed to both the increase in global means as well as an increase in environmental variability.
+There are many studies on the influence of environmental extremes and variability on competition and coexistence (Parepa et al 2013, Diez et al 2012); however, the study of the influence of relationships and interactions between drivers in their extremes on competition and coexistence, is lacking in the scientific literature.
+Interacting environmental drivers can be asymmetrically related in the extreme values (as shown in Gosh et al 2020 AER).
+Thus we have implemented our tools for the study ATAs to explore the importance of environmental relationships in their extremes on competition and coexistence.
+Environmental variability has been classically shown to promote coexistence via storage mechanisms in a lottery model of competition (Chesson and Warner 1981, Warner and Chesson 1985), hence we continue/move forward with this simple model to examine the effect of extreme relationships between environmental variables on the contribution of storage effects to coexistence.
 

### 2021-07-14 paper/Paper.Rnw
worked on storage effect section

@@
 
-paragraph to inuitively explain storage effects
-We emphasize that although this paper focuses on storage effects, it is still a paper generally about the importance of asymmetrical tail associations for coexistence because we do consider whether they are important for other mechanisms of modern coexistence theory, though we quickly realize their importance is solely for storage effects
+For clarification, we explain the how storage effects operate in maintaining biodiversity and the conditions that are required for it to do so.
+We emphasize that although this paper focuses on storage effects, it is still a paper generally about the importance of ATAs for coexistence because we do consider whether they are important for other mechanisms of modern coexistence theory, though we quickly realize their importance is solely for storage effects.
+Storage effects allow for species to fluctuate over time while escaping extinction.
+In order for storage effect to be a functioning mechanism in coexistence populations must fluctuate differently in resonse to environmental fluctuations, there must be a relationship between envrionmental responses and competition, and there must variation in responses at different life history stages.
+In a simple two species system, differences in environmental response allows for one species to grow to high densities while the other is very low.
+Competition is positively related to environment for the high density species (species A), stabalizing its growth.
+Conversely, competition is negatively/less positively?/more negative?
+related to environment for the low density species (species B), leading to large population growth rate fluctuations in both directions.
+At first glance, this may seem like a scenario for extiction of species B, but it can, in fact, be at an advantage due to the differences in life history stages.
+Different responses at different stages of life buffer the species against extinction by limiting negative growth rates, but not positive ones.
+Buffering against negative growth rates can be seen in insensitive adults or dormant seeds that persist even when environmental response amoung others in the population is low (Chesson et al 2012).
+Thus, higher fluctuation in growth rate can lead to a higher long term mean growth rate.
+In this paper, we investigate the consequences of having ATAs in the relationship between environmental fluctuations and competitive pressure.
+More specifically, we answer: Does having higher association in a tail end of the EC relationship, while maintaining overall association, change the contribution of storage effect?
 

### 2021-07-15 Figure3.R
working on figure3

@@ -2,10 +2,35 @@ load("./params.RData")
 source("./decomp_plotting.R")
 base <- par()
 
-pdf("fig3.pdf")
+#pdf("fig3.pdf")
 
-par(mfrow=c(4,4))
+par(mfrow=c(1,1), mar = c(2,2,2,1))
 
+				
+layout(matrix(c(1,2,3,4,5,
+				6,10,11,12,13,	
+				7,14,15,16,17,
+				8,18,19,20,21,
+				9,22,23,24,25), ncol=5, byrow=TRUE), 
+				heights=c(1,2,2,2,2), widths=c(1,2.5,2.5,2.5,2.5))
+#layout.show(n=25)
+				
+plot(0, bty="n")
+plot.new() #1
+
+#2-5
+for (d in 1:length(delta)){
+	plot.new()
+	text(4,-0.5,labels=paste(delta[d]))
+}
+
+#6-9
+for (m in 1:length(mudif)){
+	plot.new()
+	text(0.5,0,labels=paste(mudif[m]))
+}
+
+#10-25
 res <- vector(mode='list',length=1)
 for (m in 1:length(mudif)){
   for (d in 1:length(delta)){
@@ -13,5 +38,7 @@ for (m in 1:length(mudif)){
    }
 }
 
-dev.off()
+#dev.off()
 res <- res[-1]
+
+

### 2021-07-15 decomp_plotting.R

@@ -27,7 +27,7 @@ dePlot <- function(mudif, delta, legend=FALSE){
     }
   }
   abline(h=0, col="gray",lwd=0.7)
-  title(main=paste("mu1-mu2=",mudif,"delta=",delta), cex.main=0.7)
+  #title(main=paste("mu1-mu2=",mudif,"delta=",delta), cex.main=0.7)
   
   if(legend==TRUE){
     legend("topleft", 

### 2021-07-16 decomp_plotting.R
small changes to plotting function

@@ -11,7 +11,7 @@ dePlot <- function(mudif, delta, legend=FALSE){
     store[[i]] <- decompose(mudif,sigma[i],delta,b_tilde,u_tilde)
   }
   range <- range(unlist(lapply(store, function(X){(X$D)})))
-
+  
   plot(0, xlab="", ylab="", ylim=range*1.1, xlim=c(0,7), col="white")
   
   for (i in 1:length(sigma)){
@@ -27,7 +27,7 @@ dePlot <- function(mudif, delta, legend=FALSE){
     }
   }
   abline(h=0, col="gray",lwd=0.7)
-  title(main=paste("mu1-mu2=",mudif,"delta=",delta), cex.main=0.7)
+  #title(main=paste("mu1-mu2=",mudif,"delta=",delta), cex.main=0.7)
   
   if(legend==TRUE){
     legend("topleft", 
@@ -57,7 +57,7 @@ dePlot <- function(mudif, delta, legend=FALSE){
 
 #plot(0, xlab="", ylab="", ylim=c(-0.08,0.35), xlim=c(0.8,3.2), col="white")
 #polygon(c(0.8,0.8,1.6,3.2,3.2),c(0,x0$DEC_[1],x1$DEC_[1], x2$DEC_[1],0)
-        #,col="darkred", border=NA)
+#,col="darkred", border=NA)
 #polygon(c(0.8,0.8,1.6,3.2,3.2),c(0,x0$Dnull[1],x1$Dnull[1], x2$Dnull[1],0)
         #,col="darkblue", border=NA)
 #polygon(c(0.8,0.8,1.6,3.2,3.2),c(0,x0$DE[1],x1$DE[1], x2$DE[1],0)

### 2021-07-16 paper/Paper.Rnw
intro work

--- 2021-07-14 paper/Paper.Rnw
@@
 
-paragraph to inuitively explain storage effects
-We emphasize that although this paper focuses on storage effects, it is still a paper generally about the importance of asymmetrical tail associations for coexistence because we do consider whether they are important for other mechanisms of modern coexistence theory, though we quickly realize their importance is solely for storage effects
+For clarification, we explain the how storage effects operate in maintaining biodiversity and the conditions that are required for it to do so.
+We emphasize that although this paper focuses on storage effects, we do consider other mechanisms of modern coexistence theory, though we quickly realize the importance of ATAs is solely for storage effects; hence, it is still a paper generally about the importance of ATAs for coexistence because.
+Storage effects allow for species to fluctuate over time while escaping extinction. 
+In order for storage effect to be a functioning mechanism in coexistence, populations must fluctuate differently in response to environmental fluctuations, there must be a relationship between envrionmental responses and competition, and there must variation in responses at different life history stages.
+In a simple two species system, differences in environmental response allow for one species to grow to high densities while the other is very low. 
+Competition is positively related to environmental response for the high density species (species A), stabalizing its growth. 
+Conversely, competition is more negatively related to environmental response for the low density species (species B), leading to large fluctuations in population growth rate in both directions. 
+Seemingly, this may appear like a scenario for extiction of species B, but it can, in fact, be at an advantage due to differences in life history stages. 
+Different responses at different life stages buffer the species against extinction by limiting negative growth rates, but not positive ones. 
+Examples of life stages that can buffering against negative growth rates are insensitive adults or dormant seeds that persist even when environmental response amoung others in the population is low (Chesson et al 2012).
+Thus, higher fluctuation in growth rate can lead to a higher long term mean growth rate. 
+In this paper, we investigate the consequences of having ATAs in the relationship between fluctuations environmental response and competitive pressure.
+More specifically, we answer: Does having higher association in a tail end of the EC relationship, while maintaining overall association, change the contribution of storage effect?
 
-[As we know], anthropogenic climate change has large consequences for almost all aspects of life including via extreme climatic events, which have only become more extreme - increasing in magnitude and frequency - and have contributed to both the increase in global means as well as an increase in environmental variability.
+[As we know], anthropogenic climate change threatens almost all aspects of life including via extreme climatic events, which have only become more extreme - increasing in magnitude and frequency - and have contributed to both the increase in global means as well as an increase in environmental variability.
 There are many studies on the influence of environmental extremes and variability on competition and coexistence (Parepa et al 2013, Diez et al 2012); however, the study of the influence of relationships and interactions between drivers in their extremes on competition and coexistence, is lacking in the scientific literature.
@@
 
+(establish criteria for storage effects and establish connection to extreme events to introduce example)
 Results main ideas + example
 
-introduce new decomposition
+Previously, storage effects were described as the contribution of covariance between the environment and competition to coexistence (Chesson et al 2012, Ellner et al 2016, Ellner at al 2019). 
+However, the description of storage effects as the covariance between drivers is a (potentially dangerous/irresponsible?) misnomer in terminology, and we argue that storage effects, instead, should be described as the contribution of the relationship between drivers.
+All aspects and contributing values of the relationship between environmental response and competitive pressure are not fully captured by only the covariance between them because a measure of covariance does not capture the strucutre of associativity throuhgout the distribution of the relationship and any asymmetries there may be.
+Thus, following the decomposition presented in Ellner et al (2019), we propose storage effect can be broken down further into the contribution of ATAs [to storage effect/coexistence] and the contribution of symmetric association [to storage effect/coexistence].
 

### 2021-07-16 paper/Paper.Rnw
fleshed out example

@@
 Different responses at different life stages buffer the species against extinction by limiting negative growth rates, but not positive ones. 
-Examples of life stages that can buffering against negative growth rates are insensitive adults or dormant seeds that persist even when environmental response amoung others in the population is low (Chesson et al 2012).
+Examples of life stages that can buffering against negative growth rates are insensitive adults or dormant seeds that persist even when environmental response amoung others in the population is low (Chesson et al 2012); environment and competiton do not effect the buffered life stages.
 Thus, higher fluctuation in growth rate can lead to a higher long term mean growth rate. 
@@
 
-[As we know], anthropogenic climate change threatens almost all aspects of life including via extreme climatic events, which have only become more extreme - increasing in magnitude and frequency - and have contributed to both the increase in global means as well as an increase in environmental variability.
+[As we know], anthropogenic climate change threatens almost all aspects of life including via extreme climatic events, which have only become more extreme - increasing in magnitude and frequency - and have contributed to both the increase in global means as well as an increase in environmental variability (citation).
 There are many studies on the influence of environmental extremes and variability on competition and coexistence (Parepa et al 2013, Diez et al 2012); however, the study of the influence of relationships and interactions between drivers in their extremes on competition and coexistence, is lacking in the scientific literature.
 Interacting environmental drivers can be asymmetrically related in the extreme values (as shown in Gosh et al 2020 AER). 
-Thus we have implemented our tools for the study ATAs to explore the importance of environmental relationships in their extremes on competition and coexistence.
-Environmental variability has been classically shown to promote coexistence via storage mechanisms in a lottery model of competition (Chesson and Warner 1981, Warner and Chesson 1985), hence we continue/move forward with this simple model to examine the effect of extreme relationships between environmental variables on the contribution of storage effects to coexistence.
+Environmental drivers directly influence competitive pressure experienced by populations through environmental response; competition is always a function of environmental response (Chesson et al 2012, Ellner et al 2016).
+Thus we have implemented our tools for the study of ATAs to explore the importance of environmental relationships in their extremes on competition and coexistence. 
+Environmental variability has been classically shown to promote coexistence via storage mechanisms in a temporal lottery model of competition (Chesson and Warner 1981, Warner and Chesson 1985), hence we continue/move forward with this simple model to examine the effect of extreme relationships between environmental responses on the contribution of storage effects to coexistence.
+Moreover, the lottery model within the framework of coexistence theory provides a suitable opportunity to explore relationships in their extremes and the consequences to coexistence.
 
-The lottery model within the framework of coexistence theory provides a suitable opportunity to explore extreme relationships and the consequences to coexistence.
-
-(establish criteria for storage effects and establish connection to extreme events to introduce example)
-Results main ideas + example
+To flesh out the idea of how ATAs in environmental response might manifest and modfiy the outcome of coexistence, we introduce an example of a two species grass-forb plant community (inspired by Chesson et al 2012).
+The community follows the criteria required for temporal storage effects to operate.
+The forb dominates and has high fecundity when a sufficient threhold of rain is reached in the early season.
+The grass dominates and has high fecundity when there is sufficeint rain in the late season.
+Competition is influenced by the resource comsumption of growing juvenile plants.
+Dormant seeds are unaffected by competiton and unfavorable environment and can persist in the community.
+When rain conditions fluctuate from year to year, relative abundance and dominance of the grass-forb community also fluctuate year to year.
+This produces a relationship between the grass's and forb's environmental responses (fecundity) that is symmetrically associated.
+ATAs in environmental response can be caused by extreme climatic events.
+If the community is affected by frequent and long drought with some years of normalcy, the grass-forb environmental response relationship may become more correlated in the lower values or left tail associated.
+Conversely, if the community is effected by frequent and long floods with some years of normalcy, the grass-forb environmental relationship may become more correlated in the higher values or right-tail associated.
+In Figure 2 we see a possible outcome of ATAs between environmental responses.
+In simulating their populations,  ATAs can cause species to have longer periods of extreme dominance and lower growth rates overall.
 
@@
 G2) To explore, using a simple model, the relative magnitudes of the contributions to coexistence that come
-from ATA compared to other mechanisms.
+from ATAs compared to other mechanisms.
 

### 2021-07-16 Figure3.R
fine tunign fig3

@@ -2,9 +2,9 @@ load("./params.RData")
 source("./decomp_plotting.R")
 base <- par()
 
-#pdf("fig3.pdf")
+pdf("fig3.pdf")
 
-par(mfrow=c(1,1), mar = c(2,2,2,1))
+par(mfrow=c(1,1), mar = c(1,2,1,1), oma=c(1.5,3,3,1))
 
 				
 layout(matrix(c(1,2,3,4,5,
@@ -12,33 +12,48 @@ layout(matrix(c(1,2,3,4,5,
 				7,14,15,16,17,
 				8,18,19,20,21,
 				9,22,23,24,25), ncol=5, byrow=TRUE), 
-				heights=c(1,2,2,2,2), widths=c(1,2.5,2.5,2.5,2.5))
+				heights=c(1,3,3,3,3), widths=c(1,2.5,2.5,2.5,2.5))
 #layout.show(n=25)
 				
-plot(0, bty="n")
+#plot(0, bty="n")
 plot.new() #1
 
 #2-5
 for (d in 1:length(delta)){
 	plot.new()
-	text(4,-0.5,labels=paste(delta[d]))
+	text(0.5,0.5,labels=paste(delta[d]),font=2)
 }
 
 #6-9
 for (m in 1:length(mudif)){
 	plot.new()
-	text(0.5,0,labels=paste(mudif[m]))
+	text(0.3,0.5,labels=paste(mudif[m]), font=2)
 }
 
 #10-25
 res <- vector(mode='list',length=1)
 for (m in 1:length(mudif)){
   for (d in 1:length(delta)){
-    res <- append(res,dePlot(mudif[m], delta[d]))
+    res <- append(res,dePlot(mudif[m], delta[d], xaxt="n"))
+    axis(1, labels=FALSE, tick=TRUE)
    }
 }
 
-#dev.off()
+title(ylab=expression(mu[1]-mu[2]), outer=TRUE, line=0, font.lab=2, cex.lab=1.5)
+title(xlab=expression(delta), outer=TRUE, line=-50, font.lab=2, cex.lab=1.5)
+title(xlab=expression(sigma), outer=TRUE, line=0, cex.lab=1.3)
+
+layout(matrix(c(1)))
+plot.new() #1
+legend("center", 
+       legend=c(expression(Delta[0]), expression(Delta[E]),
+                expression(Delta[C]), expression(Delta[("E#C")]),
+                expression(Delta["[EC]"]), expression(Delta["[E||C]"])
[…]

### 2021-07-16 decomp_plotting.R

@@ -2,7 +2,7 @@
 
 source("./decomposition_fxn.R")
 
-dePlot <- function(mudif, delta, legend=FALSE){
+dePlot <- function(mudif, delta, legend=FALSE,...){
   load("./noise_etc.RData")
   store <- vector(mode='list', length=7)
   sigma <- seq(0,7,1)
@@ -12,7 +12,7 @@ dePlot <- function(mudif, delta, legend=FALSE){
   }
   range <- range(unlist(lapply(store, function(X){(X$D)})))
   
-  plot(0, xlab="", ylab="", ylim=range*1.1, xlim=c(0,7), col="white")
+  plot(0, xlab="", ylab="", ylim=range*1.1, xlim=c(0,7), col="white",...)
   
   for (i in 1:length(sigma)){
     if(i>=2){

### 2021-07-16 paper/Paper.Rnw
left last paragraph

@@
 All aspects and contributing values of the relationship between environmental response and competitive pressure are not fully captured by only the covariance between them because a measure of covariance does not capture the strucutre of associativity throuhgout the distribution of the relationship and any asymmetries there may be.
-Thus, following the decomposition presented in Ellner et al (2019), we propose storage effect can be broken down further into the contribution of ATAs [to storage effect/coexistence] and the contribution of symmetric association [to storage effect/coexistence].
+Thus, aligning with the decomposition presented in Ellner et al (2019), we propose storage effect can be broken down further into the contribution of ATA [to storage effect/coexistence] and the contribution of symmetric association [to storage effect/coexistence].
 
-final parahraph:goals
-%Jasmin, I think the below makes a pretty good start at tryinbg to nail down a good way to express our goals
-%in the last para of the Intro, so please incorporate these into your last para. If you have a better way
-%to phrase these, please put both so we can compare at some point when we meet.
-G1) To augment the modern theory of coexistence (MTC) to show how asymmetric tail association (ATA)
-plays a role. The MTC decomposes a measure of coexistence into components due to each of several mechanisms. 
-We address how ATA contributes to these components. We here examine mechanisms of coexistence that apply in 
-models exhibiting solely temporal variation; spatial variation can be considered in future work.
+This paper has the following goals. 
+G1) To augment the modern theory of coexistence (MTC) to show how ATA plays a role. 
+The MTC decomposes a measure of coexistence into components due to each of several mechanisms. 
+We address how ATA contributes to these components. 
+We here examine mechanisms of coexistence that apply in models exhibiting solely temporal variation; spatial variation can be considered in future work.
 G2) To explore, using a simple model, the relative magnitudes of the contributions to coexistence that come
 from ATAs compared to other mechanisms.
-
+Our examination and separation of the role that ATA play in contributing to coexistence is a novel approach to the study of coexistence... 
 \section{Theory}\label{section:theory}

### 2021-07-19 transform2.R
transfrom fuction with mudif instead of mu1 mu2

@@ -1,0 +1,34 @@
+# transform noise by sigma and mu to make all the noise needed to compute DeltaI and related quantities
+
+#b_tilde - 2 sets bivariate noise 
+#u_tilde - univariate noise
+#rho - common correlation of bivariate noise sets
+#sigma - common sd of bivariate noise sets
+#mudif - mu1-mu2
+
+transform <- function(b_tilde, u_tilde, rho, sigma, mudif, b_s=FALSE){
+  b_l1 <- sigma*b_tilde$l[,1] + mudif
+  b_l2 <- sigma*b_tilde$l[,2]
+  
+  b_r1 <- sigma*b_tilde$r[,1] + mudif
+  b_r2 <- sigma*b_tilde$r[,2]
+  
+  u_s <- sqrt(2*sigma^2 - 2*rho*sigma^2)*u_tilde + mudif 
+  u_1 <- sqrt(2*sigma^2)*u_tilde + mudif
+  u_2 <- sqrt(2*sigma^2)*u_tilde  
+  
+  if (b_s==FALSE){
+    return(list(b_l1=b_l1, b_l2=b_l2, b_r1=b_r1, b_r2=b_r2, 
+                u_s=u_s, u_1=u_1, u_2=u_2))
+  } else{
+    b_s1 <- sigma*b_tilde$s[,1] + mudif
+    b_s2 <- sigma*b_tilde$s[,2] 
+    
+    return(list(b_l1=b_l1, b_l2=b_l2, b_r1=b_r1, b_r2=b_r2, 
+                b_s1=b_s1, b_s2=b_s2, u_s=u_s, u_1=u_1, u_2=u_2))
+  }
+  
+  
+}
+
+

### 2021-07-19 transform2.R
fixinf mudif

@@ -7,11 +7,11 @@
 #mudif - mu1-mu2
 
 transform <- function(b_tilde, u_tilde, rho, sigma, mudif, b_s=FALSE){
-  b_l1 <- sigma*b_tilde$l[,1] + mudif
-  b_l2 <- sigma*b_tilde$l[,2]
+  b_l1 <- sigma*b_tilde$l[,1] 
+  b_l2 <- sigma*b_tilde$l[,2] - mudif
   
-  b_r1 <- sigma*b_tilde$r[,1] + mudif
-  b_r2 <- sigma*b_tilde$r[,2]
+  b_r1 <- sigma*b_tilde$r[,1] 
+  b_r2 <- sigma*b_tilde$r[,2] - mudif
   
   u_s <- sqrt(2*sigma^2 - 2*rho*sigma^2)*u_tilde + mudif 
   u_1 <- sqrt(2*sigma^2)*u_tilde + mudif
@@ -21,8 +21,8 @@ transform <- function(b_tilde, u_tilde, rho, sigma, mudif, b_s=FALSE){
     return(list(b_l1=b_l1, b_l2=b_l2, b_r1=b_r1, b_r2=b_r2, 
                 u_s=u_s, u_1=u_1, u_2=u_2))
   } else{
-    b_s1 <- sigma*b_tilde$s[,1] + mudif
-    b_s2 <- sigma*b_tilde$s[,2] 
+    b_s1 <- sigma*b_tilde$s[,1] 
+    b_s2 <- sigma*b_tilde$s[,2] - mudif
     
     return(list(b_l1=b_l1, b_l2=b_l2, b_r1=b_r1, b_r2=b_r2, 
                 b_s1=b_s1, b_s2=b_s2, u_s=u_s, u_1=u_1, u_2=u_2))

### 2021-07-19 simsPlot_fxn.R
fucntion fore plotting figure 2 (pedagogical fig)

@@ -1,0 +1,18 @@
+#function to plot simulations side by side to compare effect of ATA contributions
+
+source("./transform2.R")
+
+simsPlot <- function(mudif, delta, sigma, start=1, end=500){
+  load("./noise_etc.RData")
+  
+  #noise
+  n <- transform(b_tilde, u_tilde, rho, sigma=sigma, mudif=mudif, b_s=T)
+  
+  #asymmetric pop
+  popA <- popsim(n$b_l1,n$b_l2,N=50,N1=25,delta=delta,M)
+  #symmetric pop
+  popS <- popsim(n$b_s1,n$b_s2,N=50,N1=25,delta=delta,M)
+  plot(popS$N1[start:end], type='l')
+  plot(popA$N1[start:end], type='l')
+  
+}

### 2021-07-19 simsPlot_fxn.R
add source fxns

@@ -1,5 +1,5 @@
 #function to plot simulations side by side to compare effect of ATA contributions
-
+source("./pop_sim.R")
 source("./transform2.R")
 
 simsPlot <- function(mudif, delta, sigma, start=1, end=500){

### 2021-07-19 Figure2.R
figuring out figure 2

@@ -1,0 +1,25 @@
+source("./simsPlot_fxn.R")
+par(mfrow=c(2,1), mar=c(3,3,2,1))
+#impeding
+simsPlot(mudif=-0.8, delta=0.8, sigma=5)
+simsPlot(mudif=-0.6, delta=0.8, sigma=4)
+
+#facilitating?
+simsPlot(mudif=0, delta=1, sigma=7)
+simsPlot(mudif=0, delta=1, sigma=6)
+simsPlot(mudif=0, delta=1, sigma=5)
+simsPlot(mudif=0, delta=1, sigma=3)
+simsPlot(mudif=0, delta=1, sigma=2)
+simsPlot(mudif=0, delta=1, sigma=1)
+
+
+#max r
+simsPlot(mudif=0, delta=0.5, sigma=7)
+
+#min r (negative)
+simsPlot(mudif=-0.8, delta=1, sigma=1)
+
+
+
+
+

### 2021-07-20 paper/Paper.Rnw
adding refs

@@
 Central to ecology is the maintenance of biodiversity in competitive ecosystems. 
-Populations subject to strong competitive pressure may succumb to competitive exclusion - an idea which postulates if species within the same trophic level and similar niches exist, one will eventually outcompete the other (Hardin 1960).
+Populations subject to strong competitive pressure may succumb to competitive exclusion - an idea which postulates if species within the same trophic level and similar niches exist, one will eventually outcompete the other \citep{Hardin_1960}.
 Mediation of competitive pressure and exclusion allow for species to stably coexist. 
-The criteria for coexistence and competitive exclusion can be broken down into contributing mechanisms provided by modern coexistence theory, a framework which seeks to assess and quantify mechanisms that result in coexisting species and thus the maintenance of biodiversity (Chesson 1994).
+The criteria for coexistence and competitive exclusion can be broken down into contributing mechanisms provided by modern coexistence theory, a framework which seeks to assess and quantify mechanisms that result in coexisting species and thus the maintenance of biodiversity \citep{Chesson_1994}.
 
@@
 In fact, ATAs can have influence on ecosystem service stability, Taylor’s law, and extinction risks via extreme climatic events.
-Conversely, ATA’s can be influenced by Moran Effect and Leibeig’s Law of Minimum (Gosh et al 2020 Ecosphere, Gosh et al 2020 And.
-EcoRes,).
+Conversely, ATA’s can be influenced by Moran Effect and Leibeig’s Law of Minimum \citep{Gosh_2020_Eco, Gosh_2020_AER}.
 Despite the frequent occurrence of ATAs in ecological data, the implications of tail associations for competition and coexistence, to our knowledge, has not yet been examined; doing so is the main goal of this paper.
@@
 Storage effect contributes as a stabilizing mechanism to coexistence when there is covariation between competitive pressures and environmental fluctuations.
-Because the contributive value of storage effects relies on the relationship between environment and competition (Chesson 1994, Barbaras ..), it reasonably follows to suspect the contribution of storage effectsto be altered by an association between environment and competition that is asymmetric, i.e., ATAs between environment and competition.
+Because the contributive value of storage effects relies on the relationship between environment and competition \citep{Chesson_1994, Barabas_2018}, it reasonably follows to suspect the contribution of storage effectsto be altered by an association between environment and competition that is asymmetric, i.e., ATAs between environment and competition.
 Uniquely, storage effect is the only mechanism of modern coexistence theory which depends on the variance between two drivers rather than one.
@@
 Different responses at different life stages buffer the species against extinction by limiting negative growth rates, but not positive ones. 
-Examples of life stages that can buffering against negative growth rates are insensitive adults or dormant seeds that persist even when environmental response amoung others in the population is low (Chesson et al 2012); environment and competiton do not effect the buffered life stages.
+Examples of life stages that can buffering against negative growth rates are insensitive adults or dormant seeds that persist even when environmental response amoung others in the population is low \citep{Chesson_2012}; environment and competiton do not effect the buffered life stages.
 Thus, higher fluctuation in growth rate can lead to a higher long term mean growth rate. 
@@
 [As we know], anthropogenic climate change threatens almost all aspects of life including via extreme climatic events, which have only become more extreme - increasing in magnitude and frequency - and have contributed to both the increase in global means as well as an increase in environmental variability (citation).
-There are many studies on the influence of environmental extremes and variability on competition and coexistence (Parepa et al 2013, Diez et al 2012); however, the study of the influence of relationships and interactions between drivers in their extremes on competition and coexistence, is lacking in the scientific literature.
-Interacting environmental drivers can be asymmetrically related in the extreme values (as shown in Gosh et al 2020 AER). 
-Environmental drivers directly influence competitive pressure experienced by populations through environmental response; competition is always a function of environmental response (Chesson et al 2012, Ellner et al 2016).
+There are many studies on the influence of environmental extremes and variability on competition and coexistence \citep{Parepa_2013, Diez_2012}; however, the study of the influence of relationships and interactions between drivers in their extremes on competition and coexistence, is lacking in the scientific literature.
+Interacting environmental drivers can be asymmetrically related in the extreme values (as shown in \cite{Gosh_2020_AER}). 
+Environmental drivers directly influence competitive pressure experienced by populations through environmental response; competition is always a function of environmental response \citep{Chesson_2012, Ellner_2016}.
 Thus we have implemented our tools for the study of ATAs to explore the importance of environmental relationships in their extremes on competition and coexistence. 
-Environmental variability has been classically shown to promote coexistence via storage mechanisms in a temporal lottery model of competition (Chesson and Warner 1981, Warner and Chesson 1985), hence we continue/move forward with this simple model to examine the effect of extreme relationships between environmental responses on the contribution of storage effects to coexistence.
+Environmental variability has been classically shown to promote coexistence via storage mechanisms in a temporal lottery model of competition \citep{Chesson_1981, Warner_1985}, hence we continue/move forward with this simple model to examine the effect of extreme relationships between environmental responses on the contribution of storage effects to coexistence.
 Moreover, the lottery model within the framework of coexistence theory provides a suitable opportunity to explore relationships in their extremes and the consequences to coexistence.
 
-To flesh out the idea of how ATAs in environmental response might manifest and modfiy the outcome of coexistence, we introduce an example of a two species grass-forb plant community (inspired by Chesson et al 2012).
+To flesh out the idea of how ATAs in environmental response might manifest and modfiy the outcome of coexistence, we introduce an example of a two species grass-forb plant community (inspired by \cite{Chesson_2012}.
 The community follows the criteria required for temporal storage effects to operate.
@@
 
-Previously, storage effects were described as the contribution of covariance between the environment and competition to coexistence (Chesson et al 2012, Ellner et al 2016, Ellner at al 2019). 
+Previously, storage effects were described as the contribution of covariance between the environment and competition to coexistence \citep{Chesson_2012, Ellner_2016, Ellner_2019}. 
 However, the description of storage effects as the covariance between drivers is a (potentially dangerous/irresponsible?) misnomer in terminology, and we argue that storage effects, instead, should be described as the contribution of the relationship between drivers.
 All aspects and contributing values of the relationship between environmental response and competitive pressure are not fully captured by only the covariance between them because a measure of covariance does not capture the strucutre of associativity throuhgout the distribution of the relationship and any asymmetries there may be.
-Thus, aligning with the decomposition presented in Ellner et al (2019), we propose storage effect can be broken down further into the contribution of ATA [to storage effect/coexistence] and the contribution of symmetric association [to storage effect/coexistence].
+Thus, aligning with the decomposition presented in \cite{Ellner_2019}, we propose storage effect can be broken down further into the contribution of ATA [to storage effect/coexistence] and the contribution of symmetric association [to storage effect/coexistence].
 

### 2021-07-20 paper/Paper.Rnw
added paragraph level 'rules' of riting as a comment, for reference

--- 2021-07-16 paper/Paper.Rnw
@@
 
-Central to ecology is the maintenance of biodiversity in competitive ecosystems. 
-Populations subject to strong competitive pressure may succumb to competitive exclusion - an idea which postulates if species within the same trophic level and similar niches exist, one will eventually outcompete the other (Hardin 1960).
-Mediation of competitive pressure and exclusion allow for species to stably coexist. 
-The criteria for coexistence and competitive exclusion can be broken down into contributing mechanisms provided by modern coexistence theory, a framework which seeks to assess and quantify mechanisms that result in coexisting species and thus the maintenance of biodiversity (Chesson 1994).
+%Here are the paragraph-level "rules" of writing: 
+%Every paragraph needs to have a clear topic sentence.
+%It needs to be one of the first three sentences.
+%If it is not the first sentence, then the sentences preceding it are for background, in order to make it possible to state the topic sentence.
+%The topic sentence is the main *result* or *claim* of the paragraph, not just a statement of what the topic is.
+%All sentences in the paragraph support and validate the claim made in the topic sentence.
 
-Equally central to ecology is the study of relationships between ecological variables; however, it is almost exclusively blanketed by simple measures of associativity, i.e., correlation coefficients.
-Although two sets of variables may have equal correlation coefficients, they can show different association structures (see Figure).
-More specifically, the association within the tail ends of data and the asymmetry between the tail associations, termed asymmetric tail associations (ATAs), tell a more interesting story about the relationship between variables...and?...
-Recent work has shown that ATAs have important considerations for ecological phenomena; although, the implications of ATAs for competition and coexistence have not yet been investigated. 
-Asymmetry of tail associations within a dataset of variables describes uneven correlations between the tails of a dataset.
-For a simple example, a dataset with more association in the lower (upper) half and lesser association in the upper (lower) half is termed left (right) tail associated. 
-The application of ATAs in ecology is relatively new, and in exploration of ATAs in ecology, Gosh et al (AER) found ATAs in all eight ecological datasets they considered.
-Thus, the existence and functioning role of ATAs in nature is reasonable to assume and question. 
-In fact, ATAs can have influence on ecosystem service stability, Taylor’s law, and extinction risks via extreme climatic events.
-Conversely, ATA’s can be influenced by Moran Effect and Leibeig’s Law of Minimum (Gosh et al 2020 Ecosphere, Gosh et al 2020 And.
+The maintenance of biodiversity in competitive ecosystems is a central idea in ecology.
+Among others, the modern theory of coexistence seeks to assess and quantify the mechanisms that result in coexisting species and thus the maintenance of biodiversity.
+Many are familiar with the principle of competitive exclusion - an idea which postulates if species within the same trophic level and similar niches exist, one will eventually outcompete the other (Hardin 1960).
+Modern coexistence theory provides a quantitative breakdown of contributing mechanisms as criteria for coexistence and competitive exclusion (Chesson 1994).
+
+Equally central to ecology is the study of relationships between ecological variables; however, it is almost exclusively blanketed by simple measures of associativity.
+Recent work has shown that tail associations, or associations between the tail ends of data, and specifically the asymmetry between tail associations have important considerations for ecological phenomena; although, to our knowledge, the consequences/implications of asymmetric tail associations for competition and coexistence have not yet been investigated.
+Correlation measures such as Pearson or Spearman correlation disguise the asymmetry of tail associations, and datasets with different tail associations can have the same correlation coefficient (see Figure).
+Asymmetry in tail associations describes tighter correlations in regions of a dataset compared to other regions of the dataset.
+A simple example of asymmetric tail associations is a dataset which is more correlated in the lower (upper) half than the upper (lower) half, described as left (right) tail associated.
+Current research has shown that ATAs have important considerations for extinction risks, “skewness of spatial averages”, Taylor’s Law, Moran Effect/ spatial synchrony (Gosh et al 2020 Ecosphere, Gosh et al 2020 And.
 EcoRes,).
-Despite the frequent occurrence of ATAs in ecological data, the implications of tail associations for competition and coexistence, to our knowledge, has not yet been examined; doing so is the main goal of this paper.
+Despite the existence of tail associations within ecological data, to our knowledge, the implications of tail associations for competition and coexistence has not yet been examined; doing so is the main goal of this paper.
 
-Asymmetries in tail associations can potentially influence competition and coexistence via one particular mechanism, namely, storage effects.
-Storage effect contributes as a stabilizing mechanism to coexistence when there is covariation between competitive pressures and environmental fluctuations.
-Because the contributive value of storage effects relies on the relationship between environment and competition (Chesson 1994, Barbaras ..), it reasonably follows to suspect the contribution of storage effectsto be altered by an association between environment and competition that is asymmetric, i.e., ATAs between environment and competition.
-Uniquely, storage effect is the only mechanism of modern coexistence theory which depends on the variance between two drivers rather than one.
-While we aimed to investigate the importance of ATAs on coexistence in general and all the mechanisms which compose the modern theory of coexistence, it was quickly revealed that ATAs are specifically important for storage effects and not for other mechanisms of coexistence.
+Asymmetries in tail associations have the potential to influence competition and coexistence via one particular mechanism, namely, storage effects.
+Covariation between environmental fluctuations and competitive pressures allows for storage effect to be a contributing stabalizing mechanism to coexistence.
+Because the contributive value of storage effects relies on the association of competition and environment (Chesson 1994, Barbaras ..), it is reasonable to suspect the contribution of storage effects to coexistence can be altered by an association between competition and environment that is asymmetric.
+While we aimed to investigate the importance of ATAs on coexistence in general and all the mechanisms which compose the modern theory of coesustence, it was quickly revealed that ATAs are specifically important for storage effects and not for other mechanisms on coexistence (due to the fact that storage effect is the only mechanism that depends of the variance between two drivers).
 
-For clarification, we explain the how storage effects operate in maintaining biodiversity and the conditions that are required for it to do so.
-We emphasize that although this paper focuses on storage effects, we do consider other mechanisms of modern coexistence theory, though we quickly realize the importance of ATAs is solely for storage effects; hence, it is still a paper generally about the importance of ATAs for coexistence because.
-Storage effects allow for species to fluctuate over time while escaping extinction. 
-In order for storage effect to be a functioning mechanism in coexistence, populations must fluctuate differently in response to environmental fluctuations, there must be a relationship between envrionmental responses and competition, and there must variation in responses at different life history stages.
-In a simple two species system, differences in environmental response allow for one species to grow to high densities while the other is very low. 
-Competition is positively related to environmental response for the high density species (species A), stabalizing its growth. 
-Conversely, competition is more negatively related to environmental response for the low density species (species B), leading to large fluctuations in population growth rate in both directions. 
-Seemingly, this may appear like a scenario for extiction of species B, but it can, in fact, be at an advantage due to differences in life history stages. 
-Different responses at different life stages buffer the species against extinction by limiting negative growth rates, but not positive ones. 
-Examples of life stages that can buffering against negative growth rates are insensitive adults or dormant seeds that persist even when environmental response amoung others in the population is low (Chesson et al 2012); environment and competiton do not effect the buffered life stages.
-Thus, higher fluctuation in growth rate can lead to a higher long term mean growth rate. 
-In this paper, we investigate the consequences of having ATAs in the relationship between fluctuations environmental response and competitive pressure.
-More specifically, we answer: Does having higher association in a tail end of the EC relationship, while maintaining overall association, change the contribution of storage effect?
+paragraph to inuitively explain storage effects
+We emphasize that although this paper focuses on storage effects, it is still a paper generally about the importance of asymmetrical tail associations for coexistence because we do consider whether they are important for other mechanisms of modern coexistence theory, though we quickly realize their importance is solely for storage effects
 
-[As we know], anthropogenic climate change threatens almost all aspects of life including via extreme climatic events, which have only become more extreme - increasing in magnitude and frequency - and have contributed to both the increase in global means as well as an increase in environmental variability (citation).
-There are many studies on the influence of environmental extremes and variability on competition and coexistence (Parepa et al 2013, Diez et al 2012); however, the study of the influence of relationships and interactions between drivers in their extremes on competition and coexistence, is lacking in the scientific literature.
-Interacting environmental drivers can be asymmetrically related in the extreme values (as shown in Gosh et al 2020 AER). 
-Environmental drivers directly influence competitive pressure experienced by populations through environmental response; competition is always a function of environmental response (Chesson et al 2012, Ellner et al 2016).
-Thus we have implemented our tools for the study of ATAs to explore the importance of environmental relationships in their extremes on competition and coexistence. 
-Environmental variability has been classically shown to promote coexistence via storage mechanisms in a temporal lottery model of competition (Chesson and Warner 1981, Warner and Chesson 1985), hence we continue/move forward with this simple model to examine the effect of extreme relationships between environmental responses on the contribution of storage effects to coexistence.
-Moreover, the lottery model within the framework of coexistence theory provides a suitable opportunity to explore relationships in their extremes and the consequences to coexistence.
+[As we know], anthropogenic climate change has large consequences for almost all aspects of life including via extreme climatic events, which have only become more extreme - increasing in magnitude and frequency - and have contributed to the increase in global means as well as an increase in environmental variability.
+While there are multiple studies on the influence of environmental extremes and varaibility on competition and coexistence (Parepa et al 2013, Diez et al 2012), the influence of relationships and interactions between drivers in their extremes, has not yet been explored.
+By definition, interacting environmental drivers that are  moreso related in the extreme values are tail associated, thus we have implemented our tools for the study of tail associations to explore the importance of environmental relationships in their exrtemes on competition and coexistence.
+Environmental variability has been classically shown to promote coexistence via storage mechanisms (Chesson and Warner 1981, Warner and Chesson 1985), hence we continue/move forward with this simple model to examine the effect of extreme relationships between environmental variables on the contribution of storage effects to coexistence.
 
-To flesh out the idea of how ATAs in environmental response might manifest and modfiy the outcome of coexistence, we introduce an example of a two species grass-forb plant community (inspired by Chesson et al 2012).
-The community follows the criteria required for temporal storage effects to operate.
-The forb dominates and has high fecundity when a sufficient threhold of rain is reached in the early season.
-The grass dominates and has high fecundity when there is sufficeint rain in the late season.
-Competition is influenced by the resource comsumption of growing juvenile plants.
-Dormant seeds are unaffected by competiton and unfavorable environment and can persist in the community.
-When rain conditions fluctuate from year to year, relative abundance and dominance of the grass-forb community also fluctuate year to year.
-This produces a relationship between the grass's and forb's environmental responses (fecundity) that is symmetrically associated.
-ATAs in environmental response can be caused by extreme climatic events.
-If the community is affected by frequent and long drought with some years of normalcy, the grass-forb environmental response relationship may become more correlated in the lower values or left tail associated.
-Conversely, if the community is effected by frequent and long floods with some years of normalcy, the grass-forb environmental relationship may become more correlated in the higher values or right-tail associated.
-In Figure 2 we see a possible outcome of ATAs between environmental responses.
-In simulating their populations,  ATAs can cause species to have longer periods of extreme dominance and lower growth rates overall.
+The lottery model within the framework of coexistence theory provides a suitable opportunity to explore extreme relationships and the consequences to coexistence.
 
-Previously, storage effects were described as the contribution of covariance between the environment and competition to coexistence (Chesson et al 2012, Ellner et al 2016, Ellner at al 2019). 
-However, the description of storage effects as the covariance between drivers is a (potentially dangerous/irresponsible?) misnomer in terminology, and we argue that storage effects, instead, should be described as the contribution of the relationship between drivers.
-All aspects and contributing values of the relationship between environmental response and competitive pressure are not fully captured by only the covariance between them because a measure of covariance does not capture the strucutre of associativity throuhgout the distribution of the relationship and any asymmetries there may be.
-Thus, following the decomposition presented in Ellner et al (2019), we propose storage effect can be broken down further into the contribution of ATAs [to storage effect/coexistence] and the contribution of symmetric association [to storage effect/coexistence].
+Results main ideas + example
+
+introduce new decomposition
 
@@
 G2) To explore, using a simple model, the relative magnitudes of the contributions to coexistence that come
-from ATAs compared to other mechanisms.
+from ATA compared to other mechanisms.
 

### 2021-07-20 paper/Paper.Rnw
added ref

--- 2021-07-20 paper/Paper.Rnw
@@
 We emphasize that although this paper focuses on storage effects, we do consider other mechanisms of modern coexistence theory, though we quickly realize the importance of ATAs is solely for storage effects; hence, it is still a paper generally about the importance of ATAs for coexistence because.
-Storage effects allow for species to fluctuate over time while escaping extinction. 
+Storage effects allow for species to fluctuate over time while escaping extinction, continously recovering from low densities. 
 In order for storage effect to be a functioning mechanism in coexistence, populations must fluctuate differently in response to environmental fluctuations, there must be a relationship between envrionmental responses and competition, and there must variation in responses at different life history stages.
@@
 [As we know], anthropogenic climate change threatens almost all aspects of life including via extreme climatic events, which have only become more extreme - increasing in magnitude and frequency - and have contributed to both the increase in global means as well as an increase in environmental variability (citation).
-There are many studies on the influence of environmental extremes and variability on competition and coexistence \citep{Parepa_2013, Diez_2012}; however, the study of the influence of relationships and interactions between drivers in their extremes on competition and coexistence, is lacking in the scientific literature.
+There are many studies on the influence of environmental extremes and variability on competition and coexistence \citep{Adler_2006, Parepa_2013, Diez_2012}; however, the study of the influence of relationships and interactions between drivers in their extremes on competition and coexistence, is lacking in the scientific literature.
 Interacting environmental drivers can be asymmetrically related in the extreme values (as shown in \cite{Gosh_2020_AER}). 

### 2021-07-20 paper/Paper.Rnw
ref
worked on editing the Intro

--- 2021-07-20 paper/Paper.Rnw
@@
 %system(paste("biber", sub("\\.Rnw$", "", current_input())))
-
+%
 %#place R code here for loading in necessary variables from Results
@@
 
+%Here are the paragraph-level "rules" of writing, for reference: 
+%Every paragraph needs to have a clear topic sentence.
+%It needs to be one of the first three sentences.
+%If it is not the first sentence, then the sentences preceding it are for background, in order to make it possible to state the topic sentence.
+%The topic sentence is the main *result* or *claim* of the paragraph, not just a statement of what the topic is.
+%All sentences in the paragraph support and validate the claim made in the topic sentence.
+
 \section{Introduction}\label{section:introduction}
 
-Central to ecology is the maintenance of biodiversity in competitive ecosystems. 
-Populations subject to strong competitive pressure may succumb to competitive exclusion - an idea which postulates if species within the same trophic level and similar niches exist, one will eventually outcompete the other \citep{Hardin_1960}.
-Mediation of competitive pressure and exclusion allow for species to stably coexist. 
-The criteria for coexistence and competitive exclusion can be broken down into contributing mechanisms provided by modern coexistence theory, a framework which seeks to assess and quantify mechanisms that result in coexisting species and thus the maintenance of biodiversity \citep{Chesson_1994}.
+%Old version, prior to my edits: Central to ecology is the maintenance of biodiversity in competitive ecosystems. Populations subject to strong
+%competitive pressure may succumb to competitive exclusion - an idea which postulates if species within the same
+%trophic level and similar niches exist, one will eventually outcompete the other (Hardin, 1960). Mediation of
+%competitive pressure and exclusion allow for species to stably coexist. The criteria for coexistence and competitive
+%exclusion can be broken down into contributing mechanisms provided by modern coexistence theory, a framework
+%which seeks to assess and quantify mechanisms that result in coexisting species and thus the maintenance of
+%biodiversity (Chesson, 1994).
+%Now the new, edited version, with some additional changes requested, to be made by Jasmin, please:
+The maintenance of biodiversity in competitive ecosystems is a research topic at the absolute core of modern ecology.
+%***DAN: Jasmin, add a few classic references. See the first sentence of Ellner et al 2019, read the abstracts of those papers
+%(more than the abstracts if you want), and consider using them as references here.
+The competitive exclusion principle, covered, for instance, in most or all introductory ecology courses and texts, states 
+that two species competing for the same limiting resource cannot coexist at constant densities - one will outcompete 
+the other and drive it extinction \citep{Hardin_1960}.
+Modern coexistence theory [MCT; \cite{Chesson_1994,Chesson_2000,Barabas_2018}] is a 
+framework for understanding how stable species coexistence 
+nevertheless occurs, and biodiversity in maintained, in real ecosystems, through stabilizing mechanisms and niche 
+differentiation, in spite of the conclusions drawn for idealized scenarios by the competitive exclusion principle. 
+MCT breaks down coexistence into the contributions of various mechanisms, with name such as `relative nonlinearity
+in competition' and `storage effects' \citep{Chesson_1994} which will be familiar to many readers. The MCT has been
+applied to quite a few empirical systems, including annual plants \citep{Angert_2009}, Y and Z, 
+%***DAN: Jasmin, fill in Y and Y with names of systems, and accompanying reference(s), 
+%in a similar way to how I filled in the Angert example.
+%Please also read the abstract of the Angert reference to make sure it is appropriate to cite it in this way (and so that
+%you become aware of that reference). One place to look for more references to put in for Y and Z is the top of the 
+%right column on the first page of the Barabas review (Ecol Monogr, 2018).
+even though it is mathematically complex in the eyes
+of most ecologists (but see \cite{Ellner_2016,Ellner_2019} for efforts to make it more readily applied).
+%***DAN: Jasmin, note the differences carefully, please, between your version (commented, above) and the edits I've 
+%made. It's the same in main structure, making all the same points. It's just a bit more fleshed out. One of the purposes
+%of an Intro is to link the reader to the appropriate background lit. You need a fair number of references and a little
+%more detail than you previously gave to do that.
 
-Equally central to ecology is the study of relationships between ecological variables; however, it is almost exclusively blanketed by simple measures of associativity, i.e., correlation coefficients.
-Although two sets of variables may have equal correlation coefficients, they can show different association structures (see Figure).
-More specifically, the association within the tail ends of data and the asymmetry between the tail associations, termed asymmetric tail associations (ATAs), tell a more interesting story about the relationship between variables...and?...
-Recent work has shown that ATAs have important considerations for ecological phenomena; although, the implications of ATAs for competition and coexistence have not yet been investigated. 
-Asymmetry of tail associations within a dataset of variables describes uneven correlations between the tails of a dataset.
-For a simple example, a dataset with more association in the lower (upper) half and lesser association in the upper (lower) half is termed left (right) tail associated. 
-The application of ATAs in ecology is relatively new, and in exploration of ATAs in ecology, Gosh et al (AER) found ATAs in all eight ecological datasets they considered.
-Thus, the existence and functioning role of ATAs in nature is reasonable to assume and question. 
-In fact, ATAs can have influence on ecosystem service stability, Taylor’s law, and extinction risks via extreme climatic events.
-Conversely, ATA’s can be influenced by Moran Effect and Leibeig’s Law of Minimum \citep{Gosh_2020_Eco, Gosh_2020_AER}.
-Despite the frequent occurrence of ATAs in ecological data, the implications of tail associations for competition and coexistence, to our knowledge, has not yet been examined; doing so is the main goal of this paper.
+%Old version, some comments interspersed: 
+%Equally central to ecology is the study of relationships between ecological variables; however, it is almost exclusively blanketed by simple measures of associativity, i.e., correlation coefficients.
+%***DAN: "Blanketed" is a weird verb in this context. 
+%***DAN: Also, MCT is a conceptually deep theory, and ATA is a fairly simple observation that happens to have some consequences.
+I would not be confortable with phrasing that seems to put them on the same footing!
+%Although two sets of variables may have equal correlation coefficients, they can show different association structures (see Figure).
+%More specifically, the association within the tail ends of data and the asymmetry between the tail associations, termed asymmetric tail associations (ATAs), tell a more interesting story about the relationship between variables...and?...
+%Recent work has shown that ATAs have important considerations for ecological phenomena; although, the implications of ATAs for competition and coexistence have not yet been investigated. 
+%Asymmetry of tail associations within a dataset of variables describes uneven correlations between the tails of a dataset.
+%For a simple example, a dataset with more association in the lower (upper) half and lesser association in the upper (lower) half is termed left (right) tail associated. 
+%The application of ATAs in ecology is relatively new, and in exploration of ATAs in ecology, Gosh et al (AER) found ATAs in all eight ecological datasets they considered.
+%Thus, the existence and functioning role of ATAs in nature is reasonable to assume and question. 
+%In fact, ATAs can have influence on ecosystem service stability, Taylor’s law, and extinction risks via extreme climatic events.
+Conversely, ATA’s can be influenced by Moran Effect and Leibeig’s Law of Minimum \citep{Ghosh_2020_Eco, Ghosh_2020_AER}.
+%Despite the frequent occurrence of ATAs in ecological data, the implications of tail associations for competition and coexistence, to our knowledge, has not yet been examined; doing so is the main goal of this paper.
+%New version:
+Every ecological discipline involves the study of relationships among biological and environmental variables,
+but standard methods for examining the association between variables provide only a small portion of the overall information content contained in the
+relationship \citep{Ghosh_2020_AER,Ghosh_2020_Eco,Ghosh_2020_ecolevol,Ghosh_2021}. For instance, 
+the three pairs of variables in figure \ref{fig:noise} all have the same
+Pearson correlation; and likewise have the same Spearman correlation. 
+%***DAN: Jasmin, when you make the figure, please display the Pearson and Spearman correlations on each panel, as in 
+%Fig. 1 of Ghosh et al (2020, Ecology and Evolution).
+But each panel of figure \ref{fig:noise} nevertheless
+depicts a very different type of association. Other metrics of association are used, in addition to correlations
+(e.g., various common forms of regression), but correlations are extremely common, and other common
+methods also provide only a small portion of the avaiable information \citep{Nelsen2006}.
+The variables in figure \ref{fig:noise}A,C 
+%***DAN: Jasmin, when you make the figure, put the two asymmetric cases in A and C and put the symmetric case in B, in the middle.
+exhibit what we have previously referred to as ``asymmetric tail association'' [ATA; \cite{Ghosh_2020_AER}], i.e., 
+they are more strongly related to each
+other in the left (figure \ref{fig:noise}A, termed ``left-tail association'') or right (figure \ref{fig:noise}A,
+termed `right-tail association') portions of the variables' distributions. Although ATAs and related ideas had previously been very seldom 
+considered in ecology and related areas (but see REFS), 
+%***DAN: Jasmin, please insert bib entries for the references Anderson et al., 2018; de Valpine et al., 2014; Popovic et al., 2019
+%from the Ghosh et al (2020; Adv Ecol Res) paper, and cite them here.
+our recent work has shown that, in fact, ATAs are extremely common in ecology \citep{Ghosh_2020_AER}, and can be mechanistically important for 
+many ecological phenomena, including extinction risk \citep{Ghosh_2020_Eco}, Taylor's law \citep{Ghosh_2020_AER}, 
+synchrony, and the temporal stability of ecosystem functioning \citep{Ghosh_2020_ecolevol,Ghosh_2021}.
+%***DAN: Again note the similarity between the old and new versions, the new is just a bit more fleshed out, and maybe a
+%bit more precise in some of the wordings in places.
 
-Asymmetries in tail associations can potentially influence competition and coexistence via one particular mechanism, namely, storage effects.
+%***DAN: Please combine the following two paragraphs into one single paragraph with this topic sentence:
+Although ATAs are now known to be common and important, the implications of ATAs for competition and species coexistence 
+have, to our knowledge, not previously been considered; doing so is the overarching mission of this study. It is straightforward to see, intuitively, 
+why ATAs may influence one mechanism of coexistence, in particular: storage effects. We review, conceptually, what storage effects are and we explain 
+why ATAs may be relevant to them.
+%***DAN: I have inserted some comments below that might help.
+
 Storage effect contributes as a stabilizing mechanism to coexistence when there is covariation between competitive pressures and environmental fluctuations.
 Because the contributive value of storage effects relies on the relationship between environment and competition \citep{Chesson_1994, Barabas_2018}, it reasonably follows to suspect the contribution of storage effectsto be altered by an association between environment and competition that is asymmetric, i.e., ATAs between environment and competition.
+%***DAN: You don't need to explain why it is reasonable to suspect something, that belabors the point too much. Just say indicate that storage effects
+%involve environment-coexistence associations, and hence there is the potential for ATAs to influence storage effects.
 Uniquely, storage effect is the only mechanism of modern coexistence theory which depends on the variance between two drivers rather than one.
 While we aimed to investigate the importance of ATAs on coexistence in general and all the mechanisms which compose the modern theory of coexistence, it was quickly revealed that ATAs are specifically important for storage effects and not for other mechanisms of coexistence.
+%***DAN: Now that you have written this, I see it belabors the point too much. It sounds defensive to explain this early on that none of the 
+%other mechanisms are influenced. I think instead it's enough to just explain, as laid out in 
+%new, proposed topic sentence above, that storage effects are influenced by ATAs. That justifies the study, and then the reader can 
+%discover later that only storage effects are influenced. If a referee complains, I think we will be able to effectively respond after the fact.
 
 For clarification, we explain the how storage effects operate in maintaining biodiversity and the conditions that are required for it to do so.
+%***DAN: the above sentence will be superceded by the new topic sentence, above, of the combined paragraph, so you can cut this.
 We emphasize that although this paper focuses on storage effects, we do consider other mechanisms of modern coexistence theory, though we quickly realize the importance of ATAs is solely for storage effects; hence, it is still a paper generally about the importance of ATAs for coexistence because.
+%***DAN: Protesting too much, sounds defensive. Maybe we can return to this in the Discussion, after we've made our points, but it 
+%currently seems to me even that won't be necessary. We can just point out in the first para of the Discussion something like "We have learned
+%that ATAs do meaningfully influence species coexistence, and they do so through storage effects." Less defensive. 
 Storage effects allow for species to fluctuate over time while escaping extinction. 
@@
 More specifically, we answer: Does having higher association in a tail end of the EC relationship, while maintaining overall association, change the contribution of storage effect?
+%***DAN: This explanation of storage effects is too long. See if you can cut it to approximately half its current length.
 
+%***DAN: stopped here 2021 07 20
 [As we know], anthropogenic climate change threatens almost all aspects of life including via extreme climatic events, which have only become more extreme - increasing in magnitude and frequency - and have contributed to both the increase in global means as well as an increase in environmental variability (citation).
@@
 Our examination and separation of the role that ATA play in contributing to coexistence is a novel approach to the study of coexistence...
+
 \section{Theory}\label{section:theory}
@@
 \begin{figure}
+\caption{Figure displaying the noise. Three panels, left-, symmetric, and right-tail assoc noise. Use standard normal marginals.
+Explain in caption that alternatives for $(b_1,b_2)$ are obtained by multipling by $\sigma$ and then adding $(\mu_1,\mu_2)$.
+Jasmin to make and insert the figure here using the includegraphics command, and then draft a caption along the lines described here.\label{fig:noise}}
+\end{figure}
+
+\begin{figure}
 \caption{Pedagogical figure somehow displaying the different outcomes of competition in ATA and symmetric cases for an example.
@@
 
-\begin{figure}
-\caption{Figure displaying the noise. Three panels, left-, symmetric, and right-tail assoc noise. Use standard normal marginals.
-Explain in caption that alternatives for $(b_1,b_2)$ are obtained by multipling by $\sigma$ and then adding $(\mu_1,\mu_2)$.
-Jasmin to make and insert the figure here using the includegraphics command, and then draft a caption along the lines described here.\label{fig:noise}}
-\end{figure}
-
 \end{document}

### 2021-07-20 paper/Paper.Rnw

@@
 %that ATAs do meaningfully influence species coexistence, and they do so through storage effects." Less defensive. 
-Storage effects allow for species to fluctuate over time while escaping extinction. 
+Storage effects allow for species to fluctuate over time while escaping extinction, continously recovering from low densities. 
 In order for storage effect to be a functioning mechanism in coexistence, populations must fluctuate differently in response to environmental fluctuations, there must be a relationship between envrionmental responses and competition, and there must variation in responses at different life history stages.
@@
 [As we know], anthropogenic climate change threatens almost all aspects of life including via extreme climatic events, which have only become more extreme - increasing in magnitude and frequency - and have contributed to both the increase in global means as well as an increase in environmental variability (citation).
-There are many studies on the influence of environmental extremes and variability on competition and coexistence \citep{Parepa_2013, Diez_2012}; however, the study of the influence of relationships and interactions between drivers in their extremes on competition and coexistence, is lacking in the scientific literature.
+There are many studies on the influence of environmental extremes and variability on competition and coexistence \citep{Adler_2006,  Parepa_2013, Diez_2012}; however, the study of the influence of relationships and interactions between drivers in their extremes on competition and coexistence, is lacking in the scientific literature.
 Interacting environmental drivers can be asymmetrically related in the extreme values (as shown in \cite{Gosh_2020_AER}). 

### 2021-07-21 new_functions/co.periods.R
working on figure2

@@ -1,39 +1,0 @@
-#function: co.periods
-#computes lengths of periods of noticeable coexistence
-#ARGS:
-# pop - T x 6 df from popsim
-# dom - dominance threshold, 0<dom<1
-# N - total spaces
-#OUTPUT: a list of three, co.periods for each noise regime
-
-co.periods <- function(pop, dom, N){
-	
-	co.periods <- vector(mode = 'list', length = 3)
-	
-	for (i in 0:2){
-		N1 <- pop[, 1+i*2]
-		
-		coexist <- N1 < dom*N & N1 > (1-dom)*N
-		
-		RLE <- rle(coexist)
-		
-		co.periods[[i+1]] <- RLE$lengths[RLE$values == TRUE]		
-	}
-	names(co.periods) <- c('ELT', 'sym', 'ERT')
-	return(co.periods)
-}
-
-
-
-
-#test 
-source("./popsim.R")
-
-testnoise <- matrix(rnorm(1200), ncol=12)
-
-pop <- popsim(exp(testnoise), 50, 25, 0.5)
-
-dom <- 0.90; N <- 50
-
-co.periods(pop, dom, N)
-

### 2021-07-21 simsPlot_fxn.R

@@ -1,6 +1,37 @@
 #function to plot simulations side by side to compare effect of ATA contributions
 source("./pop_sim.R")
 source("./transform2.R")
+source("./co.periods.R")
+
+co.pPlot <- function(co.p, pop, start, end){
+  plot(0, xlab="", ylab="", ylim=c(0,50), xlim=c(start,end), col="white")
+  
+  
+  ybottom <- 0
+  ytop <- 50
+    
+  if (co.p$values[1] == FALSE){
+    xleft <- sum(co.p$lengths[1]+1)
+    c <- 1
+  }
+    
+  c <- 0
+  odds <- seq(1,201,2)
+  evens <- seq(2,202,2)
+  xleft <- 0
+    
+  for (p in 1:length(co.p$lengths)){
+      
+    xright <- sum(co.p$lengths[1:(odds[p]+c)])
+      
+    rect(xleft, ybottom, xright, ytop, col="grey", border=NA)
+      
+    xleft <- sum(co.p$lengths[1:(evens[p]+c)]) + 1
+     
+  }
+  lines(pop$N1[start:end])
+
+}
 
 simsPlot <- function(mudif, delta, sigma, start=1, end=500){
   load("./noise_etc.RData")
@@ -10,9 +41,15 @@ simsPlot <- function(mudif, delta, sigma, start=1, end=500){
   
   #asymmetric pop
   popA <- popsim(n$b_l1,n$b_l2,N=50,N1=25,delta=delta,M)
+  copA <- co.periods(popA[start:end,], dom=0.95, N=50)
   #symmetric pop
   popS <- popsim(n$b_s1,n$b_s2,N=50,N1=25,delta=delta,M)
-  plot(popS$N1[start:end], type='l')
-  plot(popA$N1[start:end], type='l')
+  copS <- co.periods(popS[start:end,], dom=0.95, N=50)
+  
+  co.pPlot(copA, popA, start, end)
+  co.pPlot(copS, popS, start, end)
   
 }
+
+#simsPlot(0,0.5,6)
+

### 2021-07-21 paper/Paper.Rnw
adding refs per Dan request

@@
 %Now the new, edited version, with some additional changes requested, to be made by Jasmin, please:
-The maintenance of biodiversity in competitive ecosystems is a research topic at the absolute core of modern ecology.
+The maintenance of biodiversity in competitive ecosystems is a research topic at the absolute core of modern ecology (\citep{MacArthur_1958, Hutchinson_1961, Schmida_1984}.
 %***DAN: Jasmin, add a few classic references. See the first sentence of Ellner et al 2019, read the abstracts of those papers
@@
 in competition' and `storage effects' \citep{Chesson_1994} which will be familiar to many readers. The MCT has been
-applied to quite a few empirical systems, including annual plants \citep{Angert_2009}, Y and Z, 
+applied to quite a few empirical systems, including annual plants \citep{Angert_2009, Hallett_2019}, perennial plants \citep{Adler_2010}, tropical trees \citep{Usinowicz_2012}, plankton \citep{Narawani_2013, Caceres_1997}, and corals \citep{}.
 %***DAN: Jasmin, fill in Y and Y with names of systems, and accompanying reference(s), 
@@
 termed `right-tail association') portions of the variables' distributions. Although ATAs and related ideas had previously been very seldom 
-considered in ecology and related areas (but see REFS), 
+considered in ecology and related areas (but see \citep{Anderson_2018, deValpine_2014, Popovic_2019}), 
 %***DAN: Jasmin, please insert bib entries for the references Anderson et al., 2018; de Valpine et al., 2014; Popovic et al., 2019

### 2021-07-22 Figure2.R
making pngs for figure 2

@@ -1,24 +1,64 @@
 source("./simsPlot_fxn.R")
-par(mfrow=c(2,1), mar=c(3,3,2,1))
-#impeding
-simsPlot(mudif=-0.8, delta=0.8, sigma=5)
-simsPlot(mudif=-0.6, delta=0.8, sigma=4)
-
-#facilitating?
-simsPlot(mudif=0, delta=1, sigma=7)
-simsPlot(mudif=0, delta=1, sigma=6)
-simsPlot(mudif=0, delta=1, sigma=5)
-simsPlot(mudif=0, delta=1, sigma=3)
-simsPlot(mudif=0, delta=1, sigma=2)
-simsPlot(mudif=0, delta=1, sigma=1)
 
 
-#max r
-simsPlot(mudif=0, delta=0.5, sigma=7)
+
+
+#### ---- IMPEDING ---- ####
+#( r w/o ATA is positive and r with ATA is negative - ATA causes r to become negative )
+
+png("figure2.1.1.PNG", width=1000, height=600, pointsize=15)
+par(mfrow=c(2,1), mar=c(3,3,1,1), oma=c(2,1,1,1))
+simsPlot(mudif=-0.8, delta=0.8, sigma=5, end=1000) # r_woATA = 0.098, r = -0.0373, [E||C] = -0.1354
+dev.off()
+
+png("figure2.1.2.PNG", width=1000, height=600, pointsize=15)
+par(mfrow=c(2,1), mar=c(3,3,1,1), oma=c(2,1,1,1))
+simsPlot(mudif=-0.6, delta=0.8, sigma=4) # r_woATA = 0.0185, r = -0.0602, [E||C] = -0.0787
+dev.off()
+
+#### ---- FACILITATING ---- ####
+#( r w/o ATA is negative and r with ATA is positive - ATA causes r to become positive )
+pdf("figure2.2.pdf")
+#png("figure2.2.1.PNG", width=1000, height=600, pointsize=15)
+par(mfrow=c(2,1), mar=c(3,3,1,1), oma=c(2,1,1,1))
+simsPlot(mudif=0, delta=1, sigma=7) # r_woATA = -0.0071, r = 0.00036, [E||C] = 0.0074
+#dev.off()
+#png("figure2.2.2.PNG", width=1000, height=600, pointsize=15)
+#par(mfrow=c(2,1), mar=c(3,3,1,1), oma=c(2,1,1,1))
[…]

### 2021-07-22 finding_pedagogical_params.R

@@ -23,15 +23,16 @@ p3.df <- p2.df[order(p2.df$md, decreasing=T),]
 dat <- cbind(p3.df, df)
 
 
-cbind(p3.df[df$r <0 & df$woATA >0,],df[df$r <0 & df$woATA >0,]) #impeding
+
+dat[dat$r<0 & df$woATA >0,] #impeding
 
 #plot1<-dePlot(-0.8,0.8, legend=T)
 
-cbind(p3.df[df$r >0 & df$woATA <0,],df[df$r >0 & df$woATA <0,]) #facilitating
+dat[dat$r>0 & df$woATA <0,] #facilitating
 
-cbind(p3.df[df$r == max(df$r),],df[df$r == max(df$r),]) #max r
+dat[dat$r == max(dat$r),]  #max r
 
-#most negatice r
+S#most negatice r
 head(dat[order(dat$r),])
 
 #most negative [E||C]

### 2021-07-22 simsPlot_fxn.R

@@ -2,9 +2,16 @@
 source("./pop_sim.R")
 source("./transform2.R")
 source("./co.periods.R")
+#end must be an even number
 
-co.pPlot <- function(co.p, pop, start, end){
-  plot(0, xlab="", ylab="", ylim=c(0,50), xlim=c(start,end), col="white")
+co.pPlot <- function(co.p, pop, start, end,...){
+  
+  if(end %%2 != 0){
+    stop("end must be set to an even number")
+  }
+  
+  plot(0, xlab="", ylab="", ylim=c(0,50), xlim=c(start,end), 
+       col="white",bty="l",...)
   
   
   ybottom <- 0
@@ -16,15 +23,15 @@ co.pPlot <- function(co.p, pop, start, end){
   }
     
   c <- 0
-  odds <- seq(1,201,2)
-  evens <- seq(2,202,2)
+  odds <- seq(1,end-1,2)
+  evens <- seq(2,end,2)
   xleft <- 0
     
   for (p in 1:length(co.p$lengths)){
       
     xright <- sum(co.p$lengths[1:(odds[p]+c)])
       
-    rect(xleft, ybottom, xright, ytop, col="grey", border=NA)
+    rect(xleft, ybottom, xright, ytop, col="grey90", border=NA)
       
     xleft <- sum(co.p$lengths[1:(evens[p]+c)]) + 1
      
@@ -46,9 +53,27 @@ simsPlot <- function(mudif, delta, sigma, start=1, end=500){
   popS <- popsim(n$b_s1,n$b_s2,N=50,N1=25,delta=delta,M)
   copS <- co.periods(popS[start:end,], dom=0.95, N=50)
   
-  co.pPlot(copA, popA, start, end)
-  co.pPlot(copS, popS, start, end)
   
+  comeanA <- round(mean(copA$length[copA$values==TRUE]),3)
+  comeanS <- round(mean(copS$length[copS$values==TRUE]),3)
+  dommeanA <- round(mean(copA$length[copA$values==FALSE]),3)
+  dommeanS <- round(mean(copS$length[copS$values==FALSE]),3)
+    
+  co.pPlot(copS, popS, start, end)
+  title(main="without ATA", line=0)   
+  text(x=end, y=20, label=paste("mean time in coexistence =", comeanS), adj=1)
+  text(x=end, y=15, label=paste("mean time not in coexistence =", dommeanS), adj=1)
+  
[…]

### 2021-07-23 Figure2.R
changed y axis label

@@ -6,7 +6,8 @@ source("./simsPlot_fxn.R")
 #### ---- IMPEDING ---- ####
 #( r w/o ATA is positive and r with ATA is negative - ATA causes r to become negative )
 
-png("figure2.1.1.PNG", width=1000, height=600, pointsize=15)
+# *** use this one for paper *** #
+png("figure2.1.1.PNG", width=1000, height=600, pointsize=15) 
 par(mfrow=c(2,1), mar=c(3,3,1,1), oma=c(2,1,1,1))
 simsPlot(mudif=-0.8, delta=0.8, sigma=5, end=1000) # r_woATA = 0.098, r = -0.0373, [E||C] = -0.1354
 dev.off()

### 2021-07-28 finding_pedagogical_params.R
added loading of params

@@ -1,4 +1,5 @@
 source("./Figure3.R")
+load("./params.RData")
 D <- lapply(res, function(X){(X$D)})
 D1 <- unlist(D)
 D2 <- matrix(D1, nrow=7)

### 2021-07-28 finding_pedagogical_params.R
added makenoise source

@@ -1,5 +1,5 @@
+source("./makenoise.R")
 source("./Figure3.R")
-load("./params.RData")
 D <- lapply(res, function(X){(X$D)})
 D1 <- unlist(D)
 D2 <- matrix(D1, nrow=7)

### 2021-07-28 Figure3.R
editing sourcing

@@ -1,33 +1,34 @@
 load("./params.RData")
-source("./decomp_plotting.R")
+load("./noise_etc.RData")
+source("./decomp1_plotting.R")
 base <- par()
 
-pdf("fig3.pdf")
+#pdf("fig3.pdf")
 
-par(mfrow=c(1,1), mar = c(1,2,1,1), oma=c(1.5,3,3,1))
+#par(mfrow=c(1,1), mar = c(1,2,1,1), oma=c(1.5,3,3,1))
 
-				
-layout(matrix(c(1,2,3,4,5,
-				6,10,11,12,13,	
-				7,14,15,16,17,
-				8,18,19,20,21,
-				9,22,23,24,25), ncol=5, byrow=TRUE), 
-				heights=c(1,3,3,3,3), widths=c(1,2.5,2.5,2.5,2.5))
+
+#layout(matrix(c(1,2,3,4,5,
+#                6,10,11,12,13,	
+#                7,14,15,16,17,
+#                8,18,19,20,21,
+#                9,22,23,24,25), ncol=5, byrow=TRUE), 
+#       heights=c(1,3,3,3,3), widths=c(1,2.5,2.5,2.5,2.5))
 #layout.show(n=25)
-				
+
 #plot(0, bty="n")
 plot.new() #1
 
 #2-5
 for (d in 1:length(delta)){
-	plot.new()
-	text(0.5,0.5,labels=paste(delta[d]),font=2)
+  plot.new()
+  text(0.5,0.5,labels=paste(delta[d]),font=2)
 }
 
 #6-9
 for (m in 1:length(mudif)){
-	plot.new()
-	text(0.3,0.5,labels=paste(mudif[m]), font=2)
+  plot.new()
+  text(0.3,0.5,labels=paste(mudif[m]), font=2)
 }
 
 #10-25
@@ -36,14 +37,14 @@ for (m in 1:length(mudif)){
   for (d in 1:length(delta)){
     res <- append(res,dePlot(mudif[m], delta[d], xaxt="n"))
     axis(1, labels=FALSE, tick=TRUE)
-   }
+  }
 }
 
 title(ylab=expression(mu[1]-mu[2]), outer=TRUE, line=0, font.lab=2, cex.lab=1.5)
 title(xlab=expression(delta), outer=TRUE, line=-50, font.lab=2, cex.lab=1.5)
 title(xlab=expression(sigma), outer=TRUE, line=0, cex.lab=1.3)
 
-layout(matrix(c(1)))
[…]

### 2021-07-28 decomp2_plotting.R

@@ -1,0 +1,49 @@
+
+source("./decomposition_fxn.R")
+
+#choose deltas 1, 0.8, 0.5
+#choose sigmas 2, 4, 6
+
+sigma=5
+delta=0.8
+
+dePlot2 <- function(sigma, delta, legend=FALSE...){
+  load("./noise_etc.RData")
+  mudif <- seq(0,-0.8,-0.01)
+  store <- vector(mode='list', length=length(mudif))
+  
+  for (i in 1:length(mudif)){
+    store[[i]] <- decompose(mudif[i],sigma,delta,b_tilde,u=u_tilde)
+  }
+  
+  range <- range(unlist(lapply(store, function(X){(X$D[6:7])})))
+  
+  plot(0, xlab="", ylab="", ylim=range*1.2, xlim=c(0,min(mudif)), col="white")
+  abline(h=0, col="gray",lwd=0.7)
+  
+  for (i in 1:length(mudif)){
+    if(i>=2){
+      dat <- cbind(store[[i-1]]$D, store[[i]]$D)
+      rwoATA <- dat[7,1:2] - dat[6,1:2]
+      lines(c(mudif[i-1], mudif[i]), dat[6,1:2], col="red")#[E||C]
+      lines(c(mudif[i-1], mudif[i]), dat[7,1:2], col="orange",lwd=2)#r
+      lines(c(mudif[i-1], mudif[i]), rwoATA, col="forestgreen", lty=4)
+      
+      if(rwoATA[1] >0 & dat[7,1]<0){
+        lines(c(mudif[i-1], mudif[i]), rep(0,2), col="deeppink")
+      }
+      if(rwoATA[1] <0 & dat[7,1]>0){
+        lines(c(mudif[i-1], mudif[i]), rep(0,2), col="yellow")
+      }
+    }
+  }
+  
+}
+
+#dePlot2(5, 0.8)
+#dePlot2(7,1)
+
+
+
+#dePlot1(-0.8, 0.8, legend=FALSE)[[6]]$D
+

### 2021-07-28 Figure3.R
making figure 4

@@ -35,7 +35,7 @@ for (m in 1:length(mudif)){
 res <- vector(mode='list',length=1)
 for (m in 1:length(mudif)){
   for (d in 1:length(delta)){
-    res <- append(res,dePlot(mudif[m], delta[d], xaxt="n"))
+    res <- append(res,dePlot1(mudif[m], delta[d], xaxt="n"))
     axis(1, labels=FALSE, tick=TRUE)
   }
 }

### 2021-07-28 Figure4.R

@@ -1,0 +1,14 @@
+source("./decomp2_plotting.R")
+
+delta <- c(1, 0.8, 0.5)
+sigma <- c(2, 4 ,6)
+
+par(mfrow=c(3,3))
+
+res <- vector(mode='list',length=1)
+for (s in 1:length(sigma)){
+  for (d in 1:length(delta)){
+    res <- append(res,dePlot2(sigma[s], delta[d], xaxt="n"))
+    axis(1, labels=FALSE, tick=TRUE)
+  }
+}

### 2021-07-28 decomp2_plotting.R

@@ -7,7 +7,7 @@ source("./decomposition_fxn.R")
 sigma=5
 delta=0.8
 
-dePlot2 <- function(sigma, delta, legend=FALSE...){
+dePlot2 <- function(sigma, delta, legend=FALSE,...){
   load("./noise_etc.RData")
   mudif <- seq(0,-0.8,-0.01)
   store <- vector(mode='list', length=length(mudif))

### 2021-07-29 Figure4.R
working on figure 4

@@ -1,14 +1,54 @@
 source("./decomp2_plotting.R")
 
 delta <- c(1, 0.8, 0.5)
-sigma <- c(2, 4 ,6)
+sigma <- c(2, 4 ,5,6)
 
-par(mfrow=c(3,3))
+pdf("fig4.pdf")
 
+par(mfrow=c(1,1), mar = c(1,2,1,1), oma=c(3,2,3,1))
+
+layout(matrix(c(1,2,3,4,
+                5,9,10,11,	
+                6,12,13,14,
+                7,15,16,17,
+                8,18,19,20), ncol=4, byrow=TRUE), 
+       heights=c(1,3,3,3,3), widths=c(1,2.5,2.5,2.5))
+
+
+plot.new() #1
+legend("center", 
+       legend=c(expression(r), expression(Delta["[E||C]"])
+                ,expression(r-Delta["[E||C]"])), 
+       col = c("orange","red","navy"),
+       lty = c(1,1,5), bty="n", cex=0.6)
+
+#2-4
+for (d in 1:length(delta)){
+  plot.new()
+  text(0.5,0.3,labels=paste(delta[d]),font=2)
+}
+
+#5-8
+for (s in 1:length(sigma)){
+  plot.new()
+  text(0.8,0.5,labels=paste(sigma[s]), font=2)
+}
+
+#9-20
 res <- vector(mode='list',length=1)
+m <- 1
 for (s in 1:length(sigma)){
   for (d in 1:length(delta)){
-    res <- append(res,dePlot2(sigma[s], delta[d], xaxt="n"))
-    axis(1, labels=FALSE, tick=TRUE)
+    res <- append(res,dePlot2(mudif_list[[m]],sigma[s], delta[d]))
+    m <- m+1
   }
 }
+
+
+title(ylab=expression(sigma), outer=TRUE, line=-2.5, font.lab=2, cex.lab=1.5)
+title(xlab=expression(delta), outer=TRUE, line=-48, font.lab=2, cex.lab=1.5)
+title(xlab=expression(mu[1]-mu[2]), outer=TRUE, line=1.4, cex.lab=1.3)
+
+
+
+dev.off()

### 2021-07-29 decomp2_plotting.R

@@ -1,49 +1,94 @@
 
 source("./decomposition_fxn.R")
 
-#choose deltas 1, 0.8, 0.5
-#choose sigmas 2, 4, 6
+#function to plot decomposition against mudif to find "differential coexistence"
+  #the degree of difference in mu1-mu2 that can result in ATAs impeding or facilitating coexistence
 
-sigma=5
-delta=0.8
+#ARGS:
+#mudif  a descending vector of mu1-mu2 to plot against and compute decompostion for
+#sigma  an integer value of sigma
+#delta  an integer value of delta
 
-dePlot2 <- function(sigma, delta, legend=FALSE,...){
+dePlot2 <- function(mudif = seq(0, -0.8, -0.01), sigma, delta, legend=FALSE,...){
   load("./noise_etc.RData")
-  mudif <- seq(0,-0.8,-0.01)
+
   store <- vector(mode='list', length=length(mudif))
   
   for (i in 1:length(mudif)){
     store[[i]] <- decompose(mudif[i],sigma,delta,b_tilde,u=u_tilde)
   }
   
-  range <- range(unlist(lapply(store, function(X){(X$D[6:7])})))
+  range <- range(unlist(lapply(store, function(X){(X$D[6:8])})))
   
-  plot(0, xlab="", ylab="", ylim=range*1.2, xlim=c(0,min(mudif)), col="white")
+  plot(0, xlab="", ylab="", ylim=range*1.2, xlim=c(max(mudif),min(mudif)), col="white",bty="l")
   abline(h=0, col="gray",lwd=0.7)
   
   for (i in 1:length(mudif)){
     if(i>=2){
       dat <- cbind(store[[i-1]]$D, store[[i]]$D)
-      rwoATA <- dat[7,1:2] - dat[6,1:2]
+      
       lines(c(mudif[i-1], mudif[i]), dat[6,1:2], col="red")#[E||C]
       lines(c(mudif[i-1], mudif[i]), dat[7,1:2], col="orange",lwd=2)#r
-      lines(c(mudif[i-1], mudif[i]), rwoATA, col="forestgreen", lty=4)
+      lines(c(mudif[i-1], mudif[i]), dat[8,1:2], col="navy", lty=5)
       
-      if(rwoATA[1] >0 & dat[7,1]<0){
-        lines(c(mudif[i-1], mudif[i]), rep(0,2), col="deeppink")
[…]

### 2021-07-29 MASTER.R
sourcing figure scripts to Master

@@ -4,6 +4,15 @@ source("./makenoise.R") #make standard noise, change M here
 
 source("./parameters.R") #set parameters sigma, mu1, mu2, delta
 
-source("./parameter_loop.R") #loop through parameters to get results #see section 4
+#FIGURE 1
+
+#FIGURE 2
+source("./Figure2.R")
+#FIGURE 3
+source("./Figure3.R")
+#FIGURE 4
+source("/Figure4.R")
+
+#source("./parameter_loop.R") #loop through parameters to get results #see section 4
 
 

### 2021-07-29 paper/Paper.Rnw
comments
added captions

@@
 There are many studies on the influence of environmental extremes and variability on competition and coexistence \citep{Adler_2006,  Parepa_2013, Diez_2012}; however, the study of the influence of relationships and interactions between drivers in their extremes on competition and coexistence, is lacking in the scientific literature.
+%;we review these studies here 
 Interacting environmental drivers can be asymmetrically related in the extreme values (as shown in \cite{Gosh_2020_AER}). 
@@
 
+If extreme events are more temporally extensive/expansive than regular environmental fluctuations, extreme events may cause the environmental response of two species who depend on temporally separated indicators of growth to become more correlated in the extremes.
+For example, ...
+
+To demonstrate plausibility of ATA altering coexistence between species, we carried out long time scale simulations using a simple lottery model. 
 To flesh out the idea of how ATAs in environmental response might manifest and modfiy the outcome of coexistence, we introduce an example of a two species grass-forb plant community (inspired by \cite{Chesson_2012}.
@@
 \begin{figure}
-\caption{Figure displaying the noise. Three panels, left-, symmetric, and right-tail assoc noise. Use standard normal marginals.
-Explain in caption that alternatives for $(b_1,b_2)$ are obtained by multipling by $\sigma$ and then adding $(\mu_1,\mu_2)$.
-Jasmin to make and insert the figure here using the includegraphics command, and then draft a caption along the lines described here.\label{fig:noise}}
+\caption{
+%Figure displaying the noise.
+Three panels, left-, symmetric, and right-tail assoc noise.
+Use standard normal marginals.Explain in caption that alternatives for $(b_1,b_2)$ are obtained by multipling by $\sigma$ and then adding $(\mu_1,\mu_2)$.Jasmin to make and insert the figure here using the includegraphics command, and then draft a caption along the lines described here.
+Simple example of bivariate noise with standard normal marginals.
+A) is left-tail associated, B) is symmetric, and C) is right-tail associated.
+All three cases have the same Pearson correlation of $\rho=$, up to sampling.., yet show different symmetry of tail associaton.
+Asymmetrically tail associated cases are perfectly correlated in the lower (A)  or upper (C) halfs of the distribution.
+The variables $(b_1,b_2)$ are transformed by multipling by $\sigma$ and then adding $(\mu_1,\mu_2)$.
+\label{fig:noise}}
 \end{figure}
@@
 \begin{figure}
-\caption{Pedagogical figure somehow displaying the different outcomes of competition in ATA and symmetric cases for an example.
+\caption{%Pedagogical figure somehow displaying the different outcomes of competition in ATA and symmetric cases for an example.
 This will be cited from the Intro, so needs to be understandable to someone who does not know any of the ideas we introduce yet.
-So in particular you cannot use IGR as your metric of coexistence, you need to use something more intuitive. You will pick
-model parameter values and do two sims, one with ATA noise and one with symmtric noise, and show different outcomes. We should
-discuss and possibly take several iterations to get this figure right since it's a pedagogically important figure - if the reader
-is not convinced in the Intro that we are studying something that makes a difference, they are not going to slog through
-the Theory section, which is inevitably going to be tough going for many readers.\label{fig:pedagog}}
+So in particular you cannot use IGR as your metric of coexistence, you need to use something more intuitive.
+You will pick model parameter values and do two sims, one with ATA noise and one with symmtric noise, and show different outcomes.
+We should discuss and possibly take several iterations to get this figure right since it's a pedagogically important figure - if the reader is not convinced in the Intro that we are studying something that makes a difference, they are not going to slog through the Theory section, which is inevitably going to be tough going for many readers.
+Populations inlfuenced by asymmetric tail association...
+recover from low densities less often/ spend more time on average under one-species dominance.
+A) is a simulation over time of a population within a two species lottery model whose environmental response and competition is influenced by symmetric bivariate noise.
+B) is like A), but the population is influenced by bivariate noise with left-tailed asymmetric association.
+Both simulations were ran for total time of 100,000.
+Noticeable population density was defined at a threshold of 0.95 of total community space.
+Periods of noticable coexistence are marked in grey.
+While the population in B) had longer mean periods of coexistence, it had much longer mean periods of one-species dominace.
+The population in B) also recovered from low densities less often and thus has a lower ability to recover when rare and thus that community has lower coexistence.
+\label{fig:pedagog}}
 \end{figure}
-%possible examples:
-%population simulations - shows symmetric with faster fluctatuations to extremes of coexistence while asymmetric has longer periods of noticeable coexistence and dominace/exclusion; may be unclear
 

### 2021-07-30 paper/Paper.Rnw
connecting ECE to temporal ATA

@@
 
-If extreme events are more temporally extensive/expansive than regular environmental fluctuations, extreme events may cause the environmental response of two species who depend on temporally separated indicators of growth to become more correlated in the extremes.
-For example, ...
-
+%If extreme events are more temporally extensive/expansive than regular environmental fluctuations, extreme events may cause the environmental response of two species who depend on temporally separated indicators of growth to become more correlated in the extremes.
+Extreme events can have important ecological impacts for controphic competing communities where species growth and the maintenance of biodiversity depends on distinct temporal events/dynamics.
+If extreme events are more temporally extensive (last longer, have higher duration) than moderate disturbances, then it is more likely that two species will be influenced by the same events.
+Both species may become more alike (similar, correlated), an important consideration for competition and coexistence.
+For example, an extreme flood event that lasts across the growth season of two different species and increases recruitment in both species will cause recruitment of the two species to become more correlated in the upper tail end of the distribution.
+This is a change in ATA not just from the event alone, but from the relationship between species.
+Though previous literature considers impacts of extreme events, change in means, and increased variability on species coexistence and competition, they do not directly consider ATA between species response caused by environment.
 To demonstrate plausibility of ATA altering coexistence between species, we carried out long time scale simulations using a simple lottery model. 
@@
 In simulating their populations,  ATAs can cause species to have longer periods of extreme dominance and lower growth rates overall.
-
+%move this paragraph to discussion
 Previously, storage effects were described as the contribution of covariance between the environment and competition to coexistence \citep{Chesson_2012, Ellner_2016, Ellner_2019}. 
-However, the description of storage effects as the covariance between drivers is a (potentially dangerous/irresponsible?) misnomer in terminology, and we argue that storage effects, instead, should be described as the contribution of the relationship between drivers.
+However, the description of storage effects as the covariance between drivers is a (potentially dangerous) misnomer in terminology, and we argue that storage effects, instead, should be described as the contribution of the relationship between drivers.
 All aspects and contributing values of the relationship between environmental response and competitive pressure are not fully captured by only the covariance between them because a measure of covariance does not capture the strucutre of associativity throuhgout the distribution of the relationship and any asymmetries there may be.

### 2021-07-30 Figure4.R
same x axis

@@ -3,7 +3,7 @@ source("./decomp2_plotting.R")
 delta <- c(1, 0.8, 0.5)
 sigma <- c(2, 4 ,5,6)
 
-pdf("fig4.pdf")
+pdf("fig4_samexaxis50.pdf")
 
 par(mfrow=c(1,1), mar = c(1,2,1,1), oma=c(3,2,3,1))
 
@@ -39,7 +39,8 @@ res <- vector(mode='list',length=1)
 m <- 1
 for (s in 1:length(sigma)){
   for (d in 1:length(delta)){
-    res <- append(res,dePlot2(mudif_list[[m]],sigma[s], delta[d]))
+    res <- append(res,dePlot2(mudif_list[[m]],sigma[s], delta[d], xaxt="n"))
+    axis(1, labels=ifelse(m>1, yes=FALSE, no=TRUE), tick=TRUE)
     m <- m+1
   }
 }

### 2021-07-30 decomp2_plotting.R

@@ -20,7 +20,7 @@ dePlot2 <- function(mudif = seq(0, -0.8, -0.01), sigma, delta, legend=FALSE,...)
   
   range <- range(unlist(lapply(store, function(X){(X$D[6:8])})))
   
-  plot(0, xlab="", ylab="", ylim=range*1.2, xlim=c(max(mudif),min(mudif)), col="white",bty="l")
+  plot(0, xlab="", ylab="", ylim=range*1.2, xlim=c(max(mudif),min(mudif)), col="white",bty="l",... )
   abline(h=0, col="gray",lwd=0.7)
   
   for (i in 1:length(mudif)){
@@ -32,12 +32,12 @@ dePlot2 <- function(mudif = seq(0, -0.8, -0.01), sigma, delta, legend=FALSE,...)
       lines(c(mudif[i-1], mudif[i]), dat[8,1:2], col="navy", lty=5)
       
       if(all(dat[8,]>0) & all(dat[7,]<0) ){ #impeding
-        rect(mudif[i-1], range[1]*1.2, mudif[i], range[2]*1.2, col="lightpink", border=NA, 
-             density=30, lty=3)
+        rect(mudif[i-1], range[1]*1.2, mudif[i], range[2]*1.2, col="hotpink2", border=NA, 
+             density=40, lty=3)
       }
       if(all(dat[8,]<0) & all(dat[7,1]>0) ){ #facilitating 
-        rect(mudif[i-1], range[1]*1.2, mudif[i], range[2]*1.2, col="goldenrod", border=NA,
-             density=30, lty=3)
+        rect(mudif[i-1], range[1]*1.2, mudif[i], range[2]*1.2, col="darkgoldenrod2", border=NA,
+             density=40, lty=3)
       }
       
     }
@@ -60,22 +60,23 @@ dePlot2 <- function(mudif = seq(0, -0.8, -0.01), sigma, delta, legend=FALSE,...)
 #dePlot2(mudif, 4,0.5)
 #dePlot2(mudif, 5,0.5)
 #dePlot2(mudif, 6,0.5)
+mudif = seq(0.0, -2.5, length.out = 50)
 
-mudif_list <- list(mudif = seq(0.04, -0.01, length.out = 30),
-                   mudif = seq(-0.1, -0.2, length.out = 50),
-                   mudif = seq(-0.1, -0.5, length.out = 30),
+mudif_list <- list(mudif,
+                   mudif,
+                   mudif,
                    
[…]

### 2021-08-02 paper/Paper.Rnw
working on adding short lit revieew of environmental change and coexistence

--- 2021-07-21 paper/Paper.Rnw
@@
 in competition' and `storage effects' \citep{Chesson_1994} which will be familiar to many readers. The MCT has been
-applied to quite a few empirical systems, including annual plants \citep{Angert_2009, Hallett_2019}, perennial plants \citep{Adler_2010}, tropical trees \citep{Usinowicz_2012}, plankton \citep{Narawani_2013, Caceres_1997}, and corals \citep{}.
+applied to quite a few empirical systems, including annual plants \citep{Angert_2009, Hallett_2019}, perennial plants \citep{Adler_2010, Adler_2006}, tropical trees \citep{Usinowicz_2012}, plankton \citep{Narawani_2013, Caceres_1997}, and corals \citep{}.
 %***DAN: Jasmin, fill in Y and Y with names of systems, and accompanying reference(s), 
@@
 There are many studies on the influence of environmental extremes and variability on competition and coexistence \citep{Adler_2006,  Parepa_2013, Diez_2012}; however, the study of the influence of relationships and interactions between drivers in their extremes on competition and coexistence, is lacking in the scientific literature.
+%;we review these studies here  In one of the first empirical tests for storage effect and its requirements, \cite{Adler_2006} showed that climate variability - including extreme events - promoted the coexistence of grass species in Kansas prairie, specifically their ability to recover from low densities.
+Moreover, in studies on the invasibility of plants, the density and competitive balance of invasive Japanese knotweed increases as resource and thus climate variability increased.\citep{Parepa_2013}
 Interacting environmental drivers can be asymmetrically related in the extreme values (as shown in \cite{Gosh_2020_AER}). 

### 2021-08-02 paper/Paper.Rnw

--- 2021-07-30 paper/Paper.Rnw
@@
 There are many studies on the influence of environmental extremes and variability on competition and coexistence \citep{Adler_2006,  Parepa_2013, Diez_2012}; however, the study of the influence of relationships and interactions between drivers in their extremes on competition and coexistence, is lacking in the scientific literature.
-%;we review these studies here 
+%;we review these studies here  In one of the first empirical tests for storage effect and its requirements, \cite{Adler_2006} showed that climate variability - including extreme events - promoted the coexistence of grass species in Kansas prairie, specifically their ability to recover from low densities.
+Moreover, in studies on the invasibility of plants, the density and competitive balance of invasive Japanese knotweed increases as resource and thus climate variability increased.\citep{Parepa_2013}
 Interacting environmental drivers can be asymmetrically related in the extreme values (as shown in \cite{Gosh_2020_AER}). 

### 2021-08-02 paper/Paper.Rnw
more climate chnage refs and lit review

@@
 %***DAN: stopped here 2021 07 20
-[As we know], anthropogenic climate change threatens almost all aspects of life including via extreme climatic events, which have only become more extreme - increasing in magnitude and frequency - and have contributed to both the increase in global means as well as an increase in environmental variability (citation).
-There are many studies on the influence of environmental extremes and variability on competition and coexistence \citep{Adler_2006,  Parepa_2013, Diez_2012}; however, the study of the influence of relationships and interactions between drivers in their extremes on competition and coexistence, is lacking in the scientific literature.
+%[As we know], anthropogenic climate change threatens almost all aspects of life including via extreme climatic events, which have only become more extreme - increasing in magnitude and frequency - and have contributed to both the increase in global means as well as an increase in environmental variability (citation).
+%There are many studies on the influence of environmental extremes and variability on competition and coexistence \citep{Adler_2006,  Parepa_2013, Diez_2012}; however, the study of the influence of relationships and interactions between drivers in their extremes on competition and coexistence, is lacking in the scientific literature.
 %;we review these studies here  In one of the first empirical tests for storage effect and its requirements, \cite{Adler_2006} showed that climate variability - including extreme events - promoted the coexistence of grass species in Kansas prairie, specifically their ability to recover from low densities.
 Moreover, in studies on the invasibility of plants, the density and competitive balance of invasive Japanese knotweed increases as resource and thus climate variability increased.\citep{Parepa_2013}
+Environmental fluctuations are becoming more influenced by extreme climatic events (Hansen et al 2012) as extreme events increase in frequency, duration, and intensity \citep{Meehl_2004}.
+Extreme events (heat-waves, floods, hurricanes, etc.) alter the means and variation of environmental fluctuations, and thus have implications for coexistence and competition.
+Competitive output and community composition of grassland plants has been shown to be impacted by heat-waves and extreme flooding \citep{White_2000, White_2001, Hoover_2014}.
+Studies on the direct effect of extreme events on competing species are slim (but see above); however, the consequences of change in environmental means, which may be due to extreme events or gradual climate change in general, on coexistence is more commonly studied.
+\cite{Descombes_ 2020}, through experimental manipulation, found that range shifts, as a product of climate change, favored coexistence between alpine plants while \cite{Usinowicz_2021} predict increases in competitive pressures as the climate change induced range shifts lead populations to interact with new conspecifics.
+Similarly, \cite{diPaola_2013} predict the collapse of coexistence between Mediterranean tree species as rainfall levels are forecasted to decrease.
+Extreme events constitute large disturbances, quantitatively amounting to large fluctuations in environmental drivers (citation).
+More variable environmental fluctuations are even less so considered compared to changes in spatial means; although, environmental variability has been analytically shown to be important for classical coexistence \citep{Chesson_1981, Warner_1985}.
+Moreover, increased variability has been found to promote coexistence in plants communities \citep{Adler_2006,  Angert_2009}, but disrupt coexistence in birds \citep{Stenseth_2002, Stenseth_2015}.
+
+Extreme events can have important ecological impacts for controphic competing communities where species growth and the maintenance of biodiversity depends on distinct temporal events/dynamics.
+If extreme events are more temporally extensive (last longer, have higher duration) than moderate disturbances, then it is more likely that two species will be influenced by the same events.
+Both species may become more alike (similar, correlated), an important consideration for competition and coexistence.
+For example, an extreme flood event that lasts across the growth season of two different species and increases recruitment in both species will cause recruitment of the two species to become more correlated in the upper tail end of the distribution.
+This is a change in ATA not just from the event alone, but from the relationship between species.
+Though previous literature considers impacts of extreme events, change in means, and increased variability on species coexistence and competition, they do not directly consider ATA between species response caused by environment.
+
 Interacting environmental drivers can be asymmetrically related in the extreme values (as shown in \cite{Gosh_2020_AER}). 

### 2021-08-03 paper/Paper.Rnw
working on storage effect paragraph - more concise

@@
 in competition' and `storage effects' \citep{Chesson_1994} which will be familiar to many readers. The MCT has been
-applied to quite a few empirical systems, including annual plants \citep{Angert_2009, Hallett_2019}, perennial plants \citep{Adler_2010, Adler_2006}, tropical trees \citep{Usinowicz_2012}, plankton \citep{Narawani_2013, Caceres_1997}, and corals \citep{}.
+applied to quite a few empirical systems, including annual plants \citep{Angert_2009, Hallett_2019}, perennial plants \citep{Adler_2010, Adler_2006}, tropical trees \citep{Usinowicz_2012}, plankton \citep{Narawani_2013, Caceres_1997}, and corals \citep{},
 %***DAN: Jasmin, fill in Y and Y with names of systems, and accompanying reference(s), 
@@
 they are more strongly related to each
-other in the left (figure \ref{fig:noise}A, termed ``left-tail association'') or right (figure \ref{fig:noise}A,
+other in the left (figure \ref{fig:noise}A, termed ``left-tail association'') or right (figure \ref{fig:noise}C,
 termed `right-tail association') portions of the variables' distributions. Although ATAs and related ideas had previously been very seldom 
@@
 %***DAN: I have inserted some comments below that might help.
-
-Storage effect contributes as a stabilizing mechanism to coexistence when there is covariation between competitive pressures and environmental fluctuations.
-Because the contributive value of storage effects relies on the relationship between environment and competition \citep{Chesson_1994, Barabas_2018}, it reasonably follows to suspect the contribution of storage effectsto be altered by an association between environment and competition that is asymmetric, i.e., ATAs between environment and competition.
+Storage effects allow for species to fluctuate over time while escaping extinction, continously recovering from low densities. 
+In order for storage effects to contribute as a stabilizing mechanism to coexistence, there must first be covariation between competitive pressures and environmental fluctuations [EC covariance] \citep{Chesson_1994, Barabas_2018}.
+%Because the contributive value of storage effects relies on the relationship between environment and competition \citep{Chesson_1994, Barabas_2018}, it reasonably follows to suspect the contribution of storage effectsto be altered by an association between environment and competition that is asymmetric, i.e., ATAs between environment and competition.
+The realiance of storage effects on the relationship between environment and competition indicate that storage effects involve environment-coexistence associations, and hence there is potential for ATAs to influence storage effects.
+In addition to EC covariance, functioning storage effects also require that conspecific populations respond differently to environmental fluctuations, and different life stages must also vary in responses.
+In a simple two species system, differences in environmental response allow for one species to grow to high densities in its favorable environemnt while the other is very low. 
+Growth is stable for the high density species due to positive EC covariance -- high intraspecific competition limits its growth rate in a favorable environment. 
+Conversely, competition is more negatively related to environmental response for the low density species, leading to large fluctuations in population growth rate in both directions. 
+Although, this may appear like a scenario for extiction, it can, in fact, be at an advantage due to differences in between life stages, which buffer the species against extinction by limiting negative growth rates, but not positive ones.
+For example, insensitive adults or dormants seeds can still persist when in an unfavorable environment and overeall growth rate is negative \citep{Chesson_2012}; environment and competiton do not effect the buffered life stages.
+Thus, higher fluctuation in growth rate can lead to a higher long term mean growth rate. 
+In this paper, we investigate the consequences of having ATAs in the relationship between fluctuations environmental response and competitive pressure.
+More specifically, we answer: Does having higher association in a tail end of the EC relationship, while maintaining overall association, change the contribution of storage effect?
 %***DAN: You don't need to explain why it is reasonable to suspect something, that belabors the point too much. Just say indicate that storage effects
 %involve environment-coexistence associations, and hence there is the potential for ATAs to influence storage effects.
-Uniquely, storage effect is the only mechanism of modern coexistence theory which depends on the variance between two drivers rather than one.
-While we aimed to investigate the importance of ATAs on coexistence in general and all the mechanisms which compose the modern theory of coexistence, it was quickly revealed that ATAs are specifically important for storage effects and not for other mechanisms of coexistence.
+%Uniquely, storage effect is the only mechanism of modern coexistence theory which depends on the variance between two drivers rather than one.
+%While we aimed to investigate the importance of ATAs on coexistence in general and all the mechanisms which compose the modern theory of coexistence, it was quickly revealed that ATAs are specifically important for storage effects and not for other mechanisms of coexistence.
 %***DAN: Now that you have written this, I see it belabors the point too much. It sounds defensive to explain this early on that none of the 
@@
 %discover later that only storage effects are influenced. If a referee complains, I think we will be able to effectively respond after the fact. 
-
-For clarification, we explain the how storage effects operate in maintaining biodiversity and the conditions that are required for it to do so.
+%For clarification, we explain the how storage effects operate in maintaining biodiversity and the conditions that are required for it to do so.
 %***DAN: the above sentence will be superceded by the new topic sentence, above, of the combined paragraph, so you can cut this.
-We emphasize that although this paper focuses on storage effects, we do consider other mechanisms of modern coexistence theory, though we quickly realize the importance of ATAs is solely for storage effects; hence, it is still a paper generally about the importance of ATAs for coexistence because.
+%We emphasize that although this paper focuses on storage effects, we do consider other mechanisms of modern coexistence theory, though we quickly realize the importance of ATAs is solely for storage effects; hence, it is still a paper generally about the importance of ATAs for coexistence because.
 %***DAN: Protesting too much, sounds defensive. Maybe we can return to this in the Discussion, after we've made our points, but it 
@@
 %that ATAs do meaningfully influence species coexistence, and they do so through storage effects." Less defensive.
-Storage effects allow for species to fluctuate over time while escaping extinction, continously recovering from low densities. 
-In order for storage effect to be a functioning mechanism in coexistence, populations must fluctuate differently in response to environmental fluctuations, there must be a relationship between envrionmental responses and competition, and there must variation in responses at different life history stages.
-In a simple two species system, differences in environmental response allow for one species to grow to high densities while the other is very low. 
-Competition is positively related to environmental response for the high density species (species A), stabalizing its growth. 
-Conversely, competition is more negatively related to environmental response for the low density species (species B), leading to large fluctuations in population growth rate in both directions. 
-Seemingly, this may appear like a scenario for extiction of species B, but it can, in fact, be at an advantage due to differences in life history stages. 
-Different responses at different life stages buffer the species against extinction by limiting negative growth rates, but not positive ones. 
-Examples of life stages that can buffering against negative growth rates are insensitive adults or dormant seeds that persist even when environmental response amoung others in the population is low \citep{Chesson_2012}; environment and competiton do not effect the buffered life stages.
-Thus, higher fluctuation in growth rate can lead to a higher long term mean growth rate. 
-In this paper, we investigate the consequences of having ATAs in the relationship between fluctuations environmental response and competitive pressure.
-More specifically, we answer: Does having higher association in a tail end of the EC relationship, while maintaining overall association, change the contribution of storage effect?
+
 %***DAN: This explanation of storage effects is too long. See if you can cut it to approximately half its current length.

### 2021-08-03 paper/Paper.Rnw
more references and editing intro of paper

@@
 termed `right-tail association') portions of the variables' distributions. Although ATAs and related ideas had previously been very seldom 
-considered in ecology and related areas (but see \citep{Anderson_2018, deValpine_2014, Popovic_2019}), 
+considered in ecology and related areas (but see \citep{Anderson_2019, deValpine_2014, Popovic_2019}), 
 %***DAN: Jasmin, please insert bib entries for the references Anderson et al., 2018; de Valpine et al., 2014; Popovic et al., 2019
@@
 %Because the contributive value of storage effects relies on the relationship between environment and competition \citep{Chesson_1994, Barabas_2018}, it reasonably follows to suspect the contribution of storage effectsto be altered by an association between environment and competition that is asymmetric, i.e., ATAs between environment and competition.
-The realiance of storage effects on the relationship between environment and competition indicate that storage effects involve environment-coexistence associations, and hence there is potential for ATAs to influence storage effects.
-In addition to EC covariance, functioning storage effects also require that conspecific populations respond differently to environmental fluctuations, and different life stages must also vary in responses.
+The reliance of storage effect's magnitude on the degree of relationship between environment and competition indicate that storage effects involve environment-coexistence associations, and hence there is potential for ATAs to influence storage effects.
+In addition to EC covariance, functioning storage effects also require that heterospecific populations respond differently to environmental fluctuations, and different life stages must also vary in responses \citep{}.
 In a simple two species system, differences in environmental response allow for one species to grow to high densities in its favorable environemnt while the other is very low. 
 Growth is stable for the high density species due to positive EC covariance -- high intraspecific competition limits its growth rate in a favorable environment. 
-Conversely, competition is more negatively related to environmental response for the low density species, leading to large fluctuations in population growth rate in both directions. 
-Although, this may appear like a scenario for extiction, it can, in fact, be at an advantage due to differences in between life stages, which buffer the species against extinction by limiting negative growth rates, but not positive ones.
+Conversely, EC covariance is lesser for the low density species, leading to large fluctuations in population growth rate in both directions. 
+Although, this may appear like a scenario for extiction, it can, in fact, be advantageous due to differences between life stages, which buffer the species against extinction by limiting negative growth rates, but not positive ones.
 For example, insensitive adults or dormants seeds can still persist when in an unfavorable environment and overeall growth rate is negative \citep{Chesson_2012}; environment and competiton do not effect the buffered life stages.
-Thus, higher fluctuation in growth rate can lead to a higher long term mean growth rate. 
-In this paper, we investigate the consequences of having ATAs in the relationship between fluctuations environmental response and competitive pressure.
-More specifically, we answer: Does having higher association in a tail end of the EC relationship, while maintaining overall association, change the contribution of storage effect?
+Thus, averaged over multiple generations, higher fluctuation in growth rate can lead to a positive low density growth rate and coexistence is possible \citep{Chesson_2000}.
+Given that storage effect depends on the relationship between E and C, we question the consequences of EC relationships having ATAs, the presence of which may be increasing as climate change brings about more extreme climatic events.
+
 %***DAN: You don't need to explain why it is reasonable to suspect something, that belabors the point too much. Just say indicate that storage effects
@@
 Moreover, in studies on the invasibility of plants, the density and competitive balance of invasive Japanese knotweed increases as resource and thus climate variability increased.\citep{Parepa_2013}
-Environmental fluctuations are becoming more influenced by extreme climatic events (Hansen et al 2012) as extreme events increase in frequency, duration, and intensity \citep{Meehl_2004}.
+Environmental fluctuations are becoming more influenced by extreme climatic events (Hansen et al 2012) as extreme events increase in frequency, duration, and intensity \citep{Meehl_2004, Ummenhofer_2017}.
 Extreme events (heat-waves, floods, hurricanes, etc.) alter the means and variation of environmental fluctuations, and thus have implications for coexistence and competition.
-Competitive output and community composition of grassland plants has been shown to be impacted by heat-waves and extreme flooding \citep{White_2000, White_2001, Hoover_2014}.
+Competitive output and community composition of grassland plants has been shown to be impacted by heat-waves and extreme drought \citep{White_2000, White_2001, Hoover_2014}.
 Studies on the direct effect of extreme events on competing species are slim (but see above); however, the consequences of change in environmental means, which may be due to extreme events or gradual climate change in general, on coexistence is more commonly studied.
-\cite{Descombes_ 2020}, through experimental manipulation, found that range shifts, as a product of climate change, favored coexistence between alpine plants while \cite{Usinowicz_2021} predict increases in competitive pressures as the climate change induced range shifts lead populations to interact with new conspecifics.
-Similarly, \cite{diPaola_2013} predict the collapse of coexistence between Mediterranean tree species as rainfall levels are forecasted to decrease.
-Extreme events constitute large disturbances, quantitatively amounting to large fluctuations in environmental drivers (citation).
+\cite{Descombes_2020}, through experimental manipulation, found that range shifts, as a product of climate change, favored coexistence between alpine plants while \cite{Usinowicz_2021} predict increases in competitive pressures as the climate change induced range shifts lead populations to interact with new conspecifics.
+Similarly, \cite{diPaola_2012} predict the collapse of coexistence between Mediterranean tree species as rainfall levels are forecasted to decrease.
+Extreme events constitute large disturbances, quantitatively amounting to large fluctuations in environmental drivers and their ecological responses \citep{Smith_2011}.
 More variable environmental fluctuations are even less so considered compared to changes in spatial means; although, environmental variability has been analytically shown to be important for classical coexistence \citep{Chesson_1981, Warner_1985}.
@@
 
-Interacting environmental drivers can be asymmetrically related in the extreme values (as shown in \cite{Gosh_2020_AER}). 
+Interacting environmental drivers can be asymmetrically related in the extreme values (as shown in \cite{Ghosh_2020_AER}). 
 Environmental drivers directly influence competitive pressure experienced by populations through environmental response; competition is always a function of environmental response \citep{Chesson_2012, Ellner_2016}.

### 2021-08-04 paper/Paper.Rnw
small

--- 2021-08-02 paper/Paper.Rnw
@@
 %Coexistence and consequences of extreme environmental relationships
+%Coexistence and the consequences of extreme environmental relationships
 %Coexistence and consequences of environmental relationships in their extremes
@@
 %Extreme relationships (relationships in their extremes) and their deductions ([negative] contributions) to coexistence
+%Coexistence and relationships between environmental variables in their extremes
 
@@
 in competition' and `storage effects' \citep{Chesson_1994} which will be familiar to many readers. The MCT has been
-applied to quite a few empirical systems, including annual plants \citep{Angert_2009, Hallett_2019}, perennial plants \citep{Adler_2010, Adler_2006}, tropical trees \citep{Usinowicz_2012}, plankton \citep{Narawani_2013, Caceres_1997}, and corals \citep{}.
+applied to quite a few empirical systems, including annual plants \citep{Angert_2009, Hallett_2019}, perennial plants \citep{Adler_2010}, tropical trees \citep{Usinowicz_2012}, plankton \citep{Narawani_2013, Caceres_1997}, and corals \citep{}.
 %***DAN: Jasmin, fill in Y and Y with names of systems, and accompanying reference(s), 
@@
 %***DAN: stopped here 2021 07 20
-%[As we know], anthropogenic climate change threatens almost all aspects of life including via extreme climatic events, which have only become more extreme - increasing in magnitude and frequency - and have contributed to both the increase in global means as well as an increase in environmental variability (citation).
-%There are many studies on the influence of environmental extremes and variability on competition and coexistence \citep{Adler_2006,  Parepa_2013, Diez_2012}; however, the study of the influence of relationships and interactions between drivers in their extremes on competition and coexistence, is lacking in the scientific literature.
-%;we review these studies here  In one of the first empirical tests for storage effect and its requirements, \cite{Adler_2006} showed that climate variability - including extreme events - promoted the coexistence of grass species in Kansas prairie, specifically their ability to recover from low densities.
-Moreover, in studies on the invasibility of plants, the density and competitive balance of invasive Japanese knotweed increases as resource and thus climate variability increased.\citep{Parepa_2013}
-Environmental fluctuations are becoming more influenced by extreme climatic events (Hansen et al 2012) as extreme events increase in frequency, duration, and intensity \citep{Meehl_2004}.
-Extreme events (heat-waves, floods, hurricanes, etc.) alter the means and variation of environmental fluctuations, and thus have implications for coexistence and competition.
-Competitive output and community composition of grassland plants has been shown to be impacted by heat-waves and extreme flooding \citep{White_2000, White_2001, Hoover_2014}.
-Studies on the direct effect of extreme events on competing species are slim (but see above); however, the consequences of change in environmental means, which may be due to extreme events or gradual climate change in general, on coexistence is more commonly studied.
-\cite{Descombes_ 2020}, through experimental manipulation, found that range shifts, as a product of climate change, favored coexistence between alpine plants while \cite{Usinowicz_2021} predict increases in competitive pressures as the climate change induced range shifts lead populations to interact with new conspecifics.
-Similarly, \cite{diPaola_2013} predict the collapse of coexistence between Mediterranean tree species as rainfall levels are forecasted to decrease.
-Extreme events constitute large disturbances, quantitatively amounting to large fluctuations in environmental drivers (citation).
-More variable environmental fluctuations are even less so considered compared to changes in spatial means; although, environmental variability has been analytically shown to be important for classical coexistence \citep{Chesson_1981, Warner_1985}.
-Moreover, increased variability has been found to promote coexistence in plants communities \citep{Adler_2006,  Angert_2009}, but disrupt coexistence in birds \citep{Stenseth_2002, Stenseth_2015}.
-
-Extreme events can have important ecological impacts for controphic competing communities where species growth and the maintenance of biodiversity depends on distinct temporal events/dynamics.
-If extreme events are more temporally extensive (last longer, have higher duration) than moderate disturbances, then it is more likely that two species will be influenced by the same events.
-Both species may become more alike (similar, correlated), an important consideration for competition and coexistence.
-For example, an extreme flood event that lasts across the growth season of two different species and increases recruitment in both species will cause recruitment of the two species to become more correlated in the upper tail end of the distribution.
-This is a change in ATA not just from the event alone, but from the relationship between species.
-Though previous literature considers impacts of extreme events, change in means, and increased variability on species coexistence and competition, they do not directly consider ATA between species response caused by environment.
-
+%***DAN: Dan won't edit this para yet, Jasmin and I discussed and she will take another crack at it.
+[As we know], anthropogenic climate change threatens almost all aspects of life including via extreme climatic events, which have only become more extreme - increasing in magnitude and frequency - and have contributed to both the increase in global means as well as an increase in environmental variability (citation).
+There are many studies on the influence of environmental extremes and variability on competition and coexistence \citep{Adler_2006,  Parepa_2013, Diez_2012}; however, the study of the influence of relationships and interactions between drivers in their extremes on competition and coexistence, is lacking in the scientific literature.
 Interacting environmental drivers can be asymmetrically related in the extreme values (as shown in \cite{Gosh_2020_AER}). 
@@
 
-%If extreme events are more temporally extensive/expansive than regular environmental fluctuations, extreme events may cause the environmental response of two species who depend on temporally separated indicators of growth to become more correlated in the extremes.
-Extreme events can have important ecological impacts for controphic competing communities where species growth and the maintenance of biodiversity depends on distinct temporal events/dynamics.
-If extreme events are more temporally extensive (last longer, have higher duration) than moderate disturbances, then it is more likely that two species will be influenced by the same events.
-Both species may become more alike (similar, correlated), an important consideration for competition and coexistence.
-For example, an extreme flood event that lasts across the growth season of two different species and increases recruitment in both species will cause recruitment of the two species to become more correlated in the upper tail end of the distribution.
-This is a change in ATA not just from the event alone, but from the relationship between species.
-Though previous literature considers impacts of extreme events, change in means, and increased variability on species coexistence and competition, they do not directly consider ATA between species response caused by environment.
-To demonstrate plausibility of ATA altering coexistence between species, we carried out long time scale simulations using a simple lottery model. 
 To flesh out the idea of how ATAs in environmental response might manifest and modfiy the outcome of coexistence, we introduce an example of a two species grass-forb plant community (inspired by \cite{Chesson_2012}.
@@
 In simulating their populations,  ATAs can cause species to have longer periods of extreme dominance and lower growth rates overall.
-%move this paragraph to discussion
+
+%***DAN: I think we ought to put this in the Discussion
 Previously, storage effects were described as the contribution of covariance between the environment and competition to coexistence \citep{Chesson_2012, Ellner_2016, Ellner_2019}. 
-However, the description of storage effects as the covariance between drivers is a (potentially dangerous) misnomer in terminology, and we argue that storage effects, instead, should be described as the contribution of the relationship between drivers.
+However, the description of storage effects as the covariance between drivers is a (potentially dangerous/irresponsible?) misnomer in terminology, and we argue that storage effects, instead, should be described as the contribution of the relationship between drivers.
 All aspects and contributing values of the relationship between environmental response and competitive pressure are not fully captured by only the covariance between them because a measure of covariance does not capture the strucutre of associativity throuhgout the distribution of the relationship and any asymmetries there may be.
@@
 
+Maybe have a paragraph about how Chesson's weak noise and other assumptions make it impossible to detect this new effect
+via that framework. It's only the less-assumption-reliant simulation approach of Ellner et al that makes it possible to 
+study this new term in the decomp. Need to study a bit to substantiate this properly, or get Steve Ellner help.
+
 Need a para about multi-species models - how would things change? One difference is there are more different
@@
 \begin{figure}
-\caption{
-%Figure displaying the noise.
-Three panels, left-, symmetric, and right-tail assoc noise.
-Use standard normal marginals.Explain in caption that alternatives for $(b_1,b_2)$ are obtained by multipling by $\sigma$ and then adding $(\mu_1,\mu_2)$.Jasmin to make and insert the figure here using the includegraphics command, and then draft a caption along the lines described here.
-Simple example of bivariate noise with standard normal marginals.
-A) is left-tail associated, B) is symmetric, and C) is right-tail associated.
-All three cases have the same Pearson correlation of $\rho=$, up to sampling.., yet show different symmetry of tail associaton.
-Asymmetrically tail associated cases are perfectly correlated in the lower (A)  or upper (C) halfs of the distribution.
-The variables $(b_1,b_2)$ are transformed by multipling by $\sigma$ and then adding $(\mu_1,\mu_2)$.
-\label{fig:noise}}
+\caption{Figure displaying the noise. Three panels, left-, symmetric, and right-tail assoc noise. Use standard normal marginals.
+Explain in caption that alternatives for $(b_1,b_2)$ are obtained by multipling by $\sigma$ and then adding $(\mu_1,\mu_2)$.
+Jasmin to make and insert the figure here using the includegraphics command, and then draft a caption along the lines described here.\label{fig:noise}}
 \end{figure}
@@
 \begin{figure}
-\caption{%Pedagogical figure somehow displaying the different outcomes of competition in ATA and symmetric cases for an example.
+\caption{Pedagogical figure somehow displaying the different outcomes of competition in ATA and symmetric cases for an example.
 This will be cited from the Intro, so needs to be understandable to someone who does not know any of the ideas we introduce yet.
-So in particular you cannot use IGR as your metric of coexistence, you need to use something more intuitive.
-You will pick model parameter values and do two sims, one with ATA noise and one with symmtric noise, and show different outcomes.
-We should discuss and possibly take several iterations to get this figure right since it's a pedagogically important figure - if the reader is not convinced in the Intro that we are studying something that makes a difference, they are not going to slog through the Theory section, which is inevitably going to be tough going for many readers.
-Populations inlfuenced by asymmetric tail association...
-recover from low densities less often/ spend more time on average under one-species dominance.
-A) is a simulation over time of a population within a two species lottery model whose environmental response and competition is influenced by symmetric bivariate noise.
-B) is like A), but the population is influenced by bivariate noise with left-tailed asymmetric association.
-Both simulations were ran for total time of 100,000.
-Noticeable population density was defined at a threshold of 0.95 of total community space.
-Periods of noticable coexistence are marked in grey.
-While the population in B) had longer mean periods of coexistence, it had much longer mean periods of one-species dominace.
-The population in B) also recovered from low densities less often and thus has a lower ability to recover when rare and thus that community has lower coexistence.
-\label{fig:pedagog}}
+So in particular you cannot use IGR as your metric of coexistence, you need to use something more intuitive. You will pick
+model parameter values and do two sims, one with ATA noise and one with symmtric noise, and show different outcomes. We should
+discuss and possibly take several iterations to get this figure right since it's a pedagogically important figure - if the reader
+is not convinced in the Intro that we are studying something that makes a difference, they are not going to slog through
+the Theory section, which is inevitably going to be tough going for many readers.\label{fig:pedagog}}
 \end{figure}
+%possible examples:
+%population simulations - shows symmetric with faster fluctatuations to extremes of coexistence while asymmetric has longer periods of noticeable coexistence and dominace/exclusion; may be unclear
 

### 2021-08-04 paper/Paper.Rnw

--- 2021-08-03 paper/Paper.Rnw
@@
 %***DAN: stopped here 2021 07 20
+
 %[As we know], anthropogenic climate change threatens almost all aspects of life including via extreme climatic events, which have only become more extreme - increasing in magnitude and frequency - and have contributed to both the increase in global means as well as an increase in environmental variability (citation).
@@
 In simulating their populations,  ATAs can cause species to have longer periods of extreme dominance and lower growth rates overall.
+
 %move this paragraph to discussion

### 2021-08-04 paper/SupMat.Rnw
worked on editing the Intro

@@
 %The following is where you load in the numeric results that will be embedded in the text
-<<eval=T,echo=F,message=F,warning=F>>=
-system(paste("biber", sub("\\.Rnw$", "", current_input())))
-
-#place R code here for loading in necessary variables from Results
-#ordinarily this would be a readRDS command, but I'm just defining a variable here for now
-allregres<-matrix(rnorm(10),2,5)
-@
+%<<eval=T,echo=F,message=F,warning=F>>=
+%system(paste("biber", sub("\\.Rnw$", "", current_input())))
+
+%#place R code here for loading in necessary variables from Results
+%#ordinarily this would be a readRDS command, but I'm just defining a variable here for now
+%allregres<-matrix(rnorm(10),2,5)
+%@
 
@@
 \begin{enumerate}
-\item Use $\epsilon_i^0=0$, as in (\ref{eq:decomp:baseline_j}), and this is the exact value.
+\item Use $\epsilon_j^0=0$, as in (\ref{eq:decomp:baseline_j}), and this is the exact value.
 \item Estimate $\olep_j^E$, following (\ref{eq:decomp:contrib_of_var_E_j}), as

### 2021-08-06 Figure1.R
Figure 1 done!

@@ -1,21 +1,84 @@
 source("./makenoise.R")
 
-b <- makenoise(1000)
 
-x<-cor(b$b_l)[1,2]
-y<-cor(b$b_s)[1,2]
-z<-cor(b$b_r)[1,2]
-r <- c(x,y,z)
+
+
+P <- c(cor(b_tilde$l, method='pearson')[1,2], cor(b_tilde$s, method='pearson')[1,2], cor(b_tilde$r, method='pearson')[1,2])
+
+
+S <- c(cor(b_tilde$l, method='spearman')[1,2],cor(b_tilde$s, method='spearman')[1,2],cor(b_tilde$r, method='spearman')[1,2])
+
+col2rgb("darkblue")
+blue <- rgb(0,0,0.545,0.3)
+
+pdf("fig1.pdf")
+png("fig1.png", height=560, width=1600,pointsize=22)
+par(mar=c(0.5,0.5,0,0), oma=c(4,4,4,2))
+layout(matrix(c(4,10,10,6,11,11,8,12,
+                1,5 ,13,2,7 ,14,3,9 ),nrow=2,byrow=T),
+       heights=c(0.3,1), widths=c(1,0.25,0.25,1,0.25,0.25,1,0.25))
+#1-3
+plot(b_tilde$l[1:1000,1],b_tilde$l[1:1000,2], col=blue, ylab=NA, xlab=NA, xlim=c(-4,4), ylim=c(-4,4), pch=16, cex=1.5)
+text(x=4,y=-2,labels=paste("P = ", round(P[1],4)), adj=1)
+text(x=4,y=-2.5,labels=paste("S = ", round(S[1],4)), adj=1)
+plot(b_tilde$s[1:1000,1],b_tilde$s[1:1000,2], col=blue, xlab=NA, ylab=NA, xlim=c(-4,4), ylim=c(-4,4), pch=16, cex=1.5)
+text(x=4,y=-2,labels=paste("P = ", round(P[2],4)), adj=1)
+text(x=4,y=-2.5,labels=paste("S = ", round(S[2],4)), adj=1)
+plot(b_tilde$r[1:1000,1],b_tilde$r[1:1000,2], col=blue, xlab=NA, ylab=NA, xlim=c(-4,4), ylim=c(-4,4), pch=16, cex=1.5)
+text(x=4,y=-2,labels=paste("P = ", round(P[3],4)), adj=1)
+text(x=4,y=-2.5,labels=paste("S = ", round(S[3],4)), adj=1)
[…]

### 2021-08-06 MASTER.R
sourcing figure scripts to Master script

@@ -5,7 +5,7 @@ source("./makenoise.R") #make standard noise, change M here
 source("./parameters.R") #set parameters sigma, mu1, mu2, delta
 
 #FIGURE 1
-
+source("./Figure1.R")
 #FIGURE 2
 source("./Figure2.R")
 #FIGURE 3

### 2021-08-06 paper/Paper.Rnw
editing intro

@@
 %Now the new, edited version, with some additional changes requested, to be made by Jasmin, please:
-The maintenance of biodiversity in competitive ecosystems is a research topic at the absolute core of modern ecology (\citep{MacArthur_1958, Hutchinson_1961, Schmida_1984}.
+The maintenance of biodiversity in competitive ecosystems is a research topic at the absolute core of modern ecology \citep{MacArthur_1958, Hutchinson_1961, Schmida_1984}.
 %***DAN: Jasmin, add a few classic references. See the first sentence of Ellner et al 2019, read the abstracts of those papers
@@
 Storage effects allow for species to fluctuate over time while escaping extinction, continously recovering from low densities. 
-In order for storage effects to contribute as a stabilizing mechanism to coexistence, there must first be covariation between competitive pressures and environmental fluctuations [EC covariance] \citep{Chesson_1994, Barabas_2018}.
+In order for storage effects to contribute as a stabilizing mechanism to coexistence, there must first be covariation between competitive pressures and environmental fluctuations [EC covariance \citep{Chesson_1994, Barabas_2018} ].
 %Because the contributive value of storage effects relies on the relationship between environment and competition \citep{Chesson_1994, Barabas_2018}, it reasonably follows to suspect the contribution of storage effectsto be altered by an association between environment and competition that is asymmetric, i.e., ATAs between environment and competition.
@@
 Conversely, EC covariance is lesser for the low density species, leading to large fluctuations in population growth rate in both directions. 
-Although, this may appear like a scenario for extiction, it can, in fact, be advantageous due to differences between life stages, which buffer the species against extinction by limiting negative growth rates, but not positive ones.
+Although, this may appear like a scenario for extiction, it can, in fact, be advantageous due to differences between life stages, buffering species against extinction by limiting negative growth rates, but not positive ones.
 For example, insensitive adults or dormants seeds can still persist when in an unfavorable environment and overeall growth rate is negative \citep{Chesson_2012}; environment and competiton do not effect the buffered life stages.
@@
 Extreme events constitute large disturbances, quantitatively amounting to large fluctuations in environmental drivers and their ecological responses \citep{Smith_2011}.
-More variable environmental fluctuations are even less so considered compared to changes in spatial means; although, environmental variability has been analytically shown to be important for classical coexistence \citep{Chesson_1981, Warner_1985}.
+Increased variability of environmental fluctuations is a newer consideration for studying environmental changes in comparison to changes in means; however, environmental variability has been analytically shown to be important for classical coexistence \citep{Chesson_1981, Warner_1985}.
 Moreover, increased variability has been found to promote coexistence in plants communities \citep{Adler_2006,  Angert_2009}, but disrupt coexistence in birds \citep{Stenseth_2002, Stenseth_2015}.
+This present study has applications to extreme events and climate change as we consider relationships in their extremes, potentially caused by extreme events, and their implications for competition and coexistence.
 
@@
 This is a change in ATA not just from the event alone, but from the relationship between species.
-Though previous literature considers impacts of extreme events, change in means, and increased variability on species coexistence and competition, they do not directly consider ATA between species response caused by environment.
+Though previous literature considers impacts of extreme events, change in means, and increased variability on species coexistence and competition and community dynamics in gerneral (see above), they do not directly consider ATA between species response caused by environment (but see Ghosh for impact of ATA on community aggregate properties).
 
-Interacting environmental drivers can be asymmetrically related in the extreme values (as shown in \cite{Ghosh_2020_AER}). 
-Environmental drivers directly influence competitive pressure experienced by populations through environmental response; competition is always a function of environmental response \citep{Chesson_2012, Ellner_2016}.
-Thus we have implemented our tools for the study of ATAs to explore the importance of environmental relationships in their extremes on competition and coexistence. 
+To demonstrate plausibility of ATA altering coexistence between species, we carried out long time scale simulations using a simple two-species lottery model.
 Environmental variability has been classically shown to promote coexistence via storage mechanisms in a temporal lottery model of competition \citep{Chesson_1981, Warner_1985}, hence we continue/move forward with this simple model to examine the effect of extreme relationships between environmental responses on the contribution of storage effects to coexistence.
+We simulated two different communites up to 100,000 time steps to hopefully make it clear to the reader than ATA makes a difference (see figure 2).
+The only distinction between the two simulations is that one depends on asymmterically tail associated temporal dynamics and one does not.
+Markedly, the simulation with ATA has longer periods of one-species dominance than the simulation without, almost seven times as much....
+
+To contextualize idea of how ATAs in environmental response might manifest and modfiy the outcome of coexistence, we introduce an example of a two species grass-forb plant community (inspired by \cite{Chesson_2012}).
+The community follows the criteria required for temporal storage effects to operate.
+The forb dominates and has high fecundity when a sufficient threhold of rain is reached in the early season.
+The grass dominates and has high fecundity when there is sufficeint rain in the late season.
+Competition is influenced by resource comsumption of growing juvenile plants.
+Dormant seeds are unaffected by competiton and unfavorable environment and can persist in the community.
+When rain fall events temporally fluctuate from year to year, relative abundance and dominance of the grass-forb community also fluctuate year to year.
+The result is a relationship between the grass's and forb's environmental responses (fecundity) that is symmetrically associated.
+However, if environmental responses become asymmetrically associated by, say, extreme rainfall events coexistence can be altered.
+If the community is disturbed by frequent and long floods with some years of normalcy, the grass-forb environmental relationship may become more correlated in the higher values or right-tail associated.
+Conversely, if the community is disturbed by frequent and long drought with some years of normalcy, the grass-forb environmental response relationship may become more correlated in the lower values or left tail associated.
+We see the possible consequences of ATA in Figure 2.
+In simulating their populations,  ATAs can cause species to have longer periods of extreme dominance and lower growth rates overall.
+
+%Interacting environmental drivers can be asymmetrically related in the extreme values (as shown in \cite{Ghosh_2020_AER}). 
+%Environmental drivers directly influence competitive pressure experienced by populations through environmental response; competition is always a function of environmental response \citep{Chesson_2012, Ellner_2016}.
+%Thus we have implemented our tools for the study of ATAs to explore the importance of environmental relationships in their extremes on competition and coexistence. 
+%Environmental variability has been classically shown to promote coexistence via storage mechanisms in a temporal lottery model of competition \citep{Chesson_1981, Warner_1985}, hence we continue/move forward with this simple model to examine the effect of extreme relationships between environmental responses on the contribution of storage effects to coexistence.
 Moreover, the lottery model within the framework of coexistence theory provides a suitable opportunity to explore relationships in their extremes and the consequences to coexistence.
@@
 %If extreme events are more temporally extensive/expansive than regular environmental fluctuations, extreme events may cause the environmental response of two species who depend on temporally separated indicators of growth to become more correlated in the extremes.
-Extreme events can have important ecological impacts for controphic competing communities where species growth and the maintenance of biodiversity depends on distinct temporal events/dynamics.
+%Extreme events can have important ecological impacts for controphic competing communities where species growth and the maintenance of biodiversity depends on distinct temporal events/dynamics.
 If extreme events are more temporally extensive (last longer, have higher duration) than moderate disturbances, then it is more likely that two species will be influenced by the same events.
@@
 Though previous literature considers impacts of extreme events, change in means, and increased variability on species coexistence and competition, they do not directly consider ATA between species response caused by environment.
-To demonstrate plausibility of ATA altering coexistence between species, we carried out long time scale simulations using a simple lottery model. 
-To flesh out the idea of how ATAs in environmental response might manifest and modfiy the outcome of coexistence, we introduce an example of a two species grass-forb plant community (inspired by \cite{Chesson_2012}.
-The community follows the criteria required for temporal storage effects to operate.
-The forb dominates and has high fecundity when a sufficient threhold of rain is reached in the early season.
-The grass dominates and has high fecundity when there is sufficeint rain in the late season.
-Competition is influenced by the resource comsumption of growing juvenile plants.
-Dormant seeds are unaffected by competiton and unfavorable environment and can persist in the community.
-When rain conditions fluctuate from year to year, relative abundance and dominance of the grass-forb community also fluctuate year to year.
-This produces a relationship between the grass's and forb's environmental responses (fecundity) that is symmetrically associated.
-ATAs in environmental response can be caused by extreme climatic events.
-If the community is affected by frequent and long drought with some years of normalcy, the grass-forb environmental response relationship may become more correlated in the lower values or left tail associated.
-Conversely, if the community is effected by frequent and long floods with some years of normalcy, the grass-forb environmental relationship may become more correlated in the higher values or right-tail associated.
-In Figure 2 we see a possible outcome of ATAs between environmental responses.
-In simulating their populations,  ATAs can cause species to have longer periods of extreme dominance and lower growth rates overall.
 
 %move this paragraph to discussion
-Previously, storage effects were described as the contribution of covariance between the environment and competition to coexistence \citep{Chesson_2012, Ellner_2016, Ellner_2019}. 
-However, the description of storage effects as the covariance between drivers is a (potentially dangerous) misnomer in terminology, and we argue that storage effects, instead, should be described as the contribution of the relationship between drivers.
-All aspects and contributing values of the relationship between environmental response and competitive pressure are not fully captured by only the covariance between them because a measure of covariance does not capture the strucutre of associativity throuhgout the distribution of the relationship and any asymmetries there may be.
-Thus, aligning with the decomposition presented in \cite{Ellner_2019}, we propose storage effect can be broken down further into the contribution of ATA [to storage effect/coexistence] and the contribution of symmetric association [to storage effect/coexistence].
+%Previously, storage effects were described as the contribution of covariance between the environment and competition to coexistence \citep{Chesson_2012, Ellner_2016, Ellner_2019}. 
+%However, the description of storage effects as the covariance between drivers is a (potentially dangerous) misnomer in terminology, and we argue that storage effects, instead, should be described as the contribution of the relationship between drivers.
+%All aspects and contributing values of the relationship between environmental response and competitive pressure are not fully captured by only the covariance between them because a measure of covariance does not capture the strucutre of associativity throuhgout the distribution of the relationship and any asymmetries there may be.
+%Thus, aligning with the decomposition presented in \cite{Ellner_2019}, we propose storage effect can be broken down further into the contribution of ATA [to storage effect/coexistence] and the contribution of symmetric association [to storage effect/coexistence].
 
@@
 \section{Results}\label{section:results}
+%outline of main resutls
+Using our simple two-species lottery model, ATA made a difference for coexistence.
+Depending on death rate, variance, and difference between species, ATA can either impede or facilitate coexistence.
+In some cases, the contribution of ATA is the mechanism solely responsible for a positive (or negative) IGR, and the IGR without ATA is negative (or positive).
+In almost all cases we tested, ATA contributed negatively to coexistence expected for when delta = 1, annual species.
+Of the annual species cases, only when the species were the same or extremely close to being the same, did ATA have a rescue effect on storage effects and coexistence.
+For other annual species cases, as the species became moree different, IGR was more negative and the contribution of ATA and storage effect in general became more negligible.
+
+We found the range of species difference, where ATA either allowed the rescue of IGR or ATA imperiled IGR. %are we calling this 'differential coexistence'?
+
+In general, the direction of the ATA did not show any differences (other than small differences in very specific cases) for coexistence
 

### 2021-08-06 Figure1.R
revisions discussed in meeting

@@ -1,4 +1,4 @@
-source("./makenoise.R")
+#source("./makenoise.R")
 
 
 
@@ -13,44 +13,43 @@ blue <- rgb(0,0,0.545,0.3)
 
 pdf("fig1.pdf")
 png("fig1.png", height=560, width=1600,pointsize=22)
-par(mar=c(0.5,0.5,0,0), oma=c(4,4,4,2))
+par(mar=c(0.5,0.5,0,0), oma=c(5,5,2,2))
 layout(matrix(c(4,10,10,6,11,11,8,12,
                 1,5 ,13,2,7 ,14,3,9 ),nrow=2,byrow=T),
-       heights=c(0.3,1), widths=c(1,0.25,0.25,1,0.25,0.25,1,0.25))
+       heights=c(0.25,1), widths=c(1,0.25,0.25,1,0.25,0.25,1,0.25))
 #1-3
 plot(b_tilde$l[1:1000,1],b_tilde$l[1:1000,2], col=blue, ylab=NA, xlab=NA, xlim=c(-4,4), ylim=c(-4,4), pch=16, cex=1.5)
-text(x=4,y=-2,labels=paste("P = ", round(P[1],4)), adj=1)
-text(x=4,y=-2.5,labels=paste("S = ", round(S[1],4)), adj=1)
+title(main='Left tail association', line=-1.2, cex.main=1.5)
+text(x=4,y=-3.75,labels=paste("P = ", round(P[1],4)), adj=1)
+
 plot(b_tilde$s[1:1000,1],b_tilde$s[1:1000,2], col=blue, xlab=NA, ylab=NA, xlim=c(-4,4), ylim=c(-4,4), pch=16, cex=1.5)
-text(x=4,y=-2,labels=paste("P = ", round(P[2],4)), adj=1)
-text(x=4,y=-2.5,labels=paste("S = ", round(S[2],4)), adj=1)
+title(main='Symmetric association', line=-1.2, cex.main=1.5)
+text(x=4,y=-3.75,labels=paste("P = ", round(P[2],4)), adj=1)
+
 plot(b_tilde$r[1:1000,1],b_tilde$r[1:1000,2], col=blue, xlab=NA, ylab=NA, xlim=c(-4,4), ylim=c(-4,4), pch=16, cex=1.5)
-text(x=4,y=-2,labels=paste("P = ", round(P[3],4)), adj=1)
-text(x=4,y=-2.5,labels=paste("S = ", round(S[3],4)), adj=1)
[…]

### 2021-08-06 Figure4.R
some revisions to figure4

@@ -1,11 +1,11 @@
-source("./decomp2_plotting.R")
+#source("./decomp2_plotting.R")
 
 delta <- c(1, 0.8, 0.5)
 sigma <- c(2, 4 ,5,6)
 
 pdf("fig4_samexaxis50.pdf")
 
-par(mfrow=c(1,1), mar = c(1,2,1,1), oma=c(3,2,3,1))
+par(mfrow=c(1,1), mar = c(1,2,1,1), oma=c(3,2,3,3))
 
 layout(matrix(c(1,2,3,4,
                 5,9,10,11,	
@@ -40,13 +40,13 @@ m <- 1
 for (s in 1:length(sigma)){
   for (d in 1:length(delta)){
     res <- append(res,dePlot2(mudif_list[[m]],sigma[s], delta[d], xaxt="n"))
-    axis(1, labels=ifelse(m>1, yes=FALSE, no=TRUE), tick=TRUE)
+    axis(1, labels=ifelse(m%%3==0, yes=TRUE, no=FALSE), tick=TRUE)
     m <- m+1
   }
 }
 
-
-title(ylab=expression(sigma), outer=TRUE, line=-2.5, font.lab=2, cex.lab=1.5)
+title(ylab="contribution to coexistence", outer=TRUE, line=-2.5, font.lab=2, cex.lab=1.5)
+title(ylab=expression(sigma), outer=TRUE, line=-30, font.lab=2, cex.lab=1.5)
 title(xlab=expression(delta), outer=TRUE, line=-48, font.lab=2, cex.lab=1.5)
 title(xlab=expression(mu[1]-mu[2]), outer=TRUE, line=1.4, cex.lab=1.3)
 

### 2021-08-12 paper/Paper.Rnw
g3 methods

@@
 from ATAs compared to other mechanisms.
+G3) To define the parameter space for which species can coexist in our model. 
 Our examination and separation of the role that ATA play in contributing to coexistence is a novel approach to the study of coexistence...
@@
 SI section \ref{SI-sect:efficient_computation}.
-
 %What's next
@@
 %and sup mat results), put it above.
-%4) Think about this: Theoretical fig exploring why ATA matters to r_1\1 for the lottery model and hopefully more gernerally - try to get
+%4) Think about this: Theoretical fig exploring why , ATA matters to r_1\1 for the lottery model and hopefully more gernerally - try to get
 %at the intuition here. See if you can edge to more generality here, or not?
+
+To further explore the model and acheive G3 from the introduction i.e., exploring the parameter space for which coexistence is possible, we then plotted the invasion growth rate, $r_{i/i}-q_{12}r_{i/j}$ and the invasion growth rate [IGR] without ATA contributribution  $r_{i/i}-q_{12}r_{i/j} - \Delta_i^{[EC]}$ against the difference between species, $\mu_1-\mu_2$,
+
+Because $\mu$ is the only unique species parameter and the difference, $\mu_1-\mu_2$, is the only difference between the two species in this model, we henceforward call thus difference the trait difference between species.
+Letting the trait $\mu$ vary within a one dimensional space, this space is then a trait space, and the difference, $\mu_1-\mu_2$, is then a subset of that trait space.
+We then define $\psi_{ATA}$ to be the subset of trait space that allows a postive IGR when ATA is included and $\psi_{~ATA}$ to be the subset of trait space that allows a postive IGR when ATA is not included.
+We then use symmetric difference $\psi_{ATA} \Delta \psi_{~ATA}$ to make the following conslusions.
+If the length of $\psi_{ATA} > the length of $\psi_{~ATA}$ the $\psi_{ATA} \Delta \psi_{~ATA}$ consitutes ATA rescue and the degree of species trait difference that allows ATA rescue to act, hence, the ATA rescue trait space.
+If the length of $\psi_{ATA} < the length of $\psi_{~ATA}$ the $\psi_{ATA} \Delta \psi_{~ATA}$ consitutes ATA exclusion and the degree of species trait difference that allows ATA exclusion to act, hence, the ATA exclusion trait space.
+
+In order to find the ATA rescue trait space and ATA exclusion trait space, we plotted the trait space subset $\mu_1-\mu_2$ against IGR and IGR without ATA for different values of $\sigma$ and $\delta$, the shared, non-unique species parameters.
+We selected $\delta=0.5, 0.8, 1.0$ and $\sigma=2.0,4.0,5.0, 6.0$.
+The spaces in between the lines where they intersect the x-axis at y=0, indicate either an ATA rescue trait space or an ATA exclusion trait space.
+If the line plotting IGR is to the left of the line plotting IGR without ATA, then this space is an ATA rescue trait space.
+If the line plotting IGR is to the right of the line plotting IGR without ATA, then this space is an ATA exclusion trait space.
 

### 2021-08-12 paper/Paper.Rnw
starting G2 resujlts

@@
 \section{Results}\label{section:results}
+To explore the relative magnitudes of the contributions to coexistence that come from ATA versus other mechanisms in our simple two-species lottery model (G2), we plotted the decomposing mechanisms against increasing $\sigma$ for various values of $\mu_1 - \mu_2$ and $\delta$ (Figure 4).
+In almost all cases, aside from when $\delta=1$, the contribution of ATA is less than the contribution of storage effects without ATA, and the former decreases with $\sigma$ while the latter increases with $\sigma$.
+We see the opposite when $\delta=1$, where the contribution of ATA and greater than the contribution of storage effect without ATA and the trends are reversed.
 %outline of main resutls

### 2021-08-12 Figure3.R
matching configs of 3 and 4

@@ -1,19 +1,19 @@
 load("./params.RData")
 load("./noise_etc.RData")
 source("./decomp1_plotting.R")
-base <- par()
-
-#pdf("fig3.pdf")
-
-#par(mfrow=c(1,1), mar = c(1,2,1,1), oma=c(1.5,3,3,1))
 
 
-#layout(matrix(c(1,2,3,4,5,
-#                6,10,11,12,13,	
-#                7,14,15,16,17,
-#                8,18,19,20,21,
-#                9,22,23,24,25), ncol=5, byrow=TRUE), 
-#       heights=c(1,3,3,3,3), widths=c(1,2.5,2.5,2.5,2.5))
+pdf("fig3_legend.pdf")
+
+par(mgp=c(3,0.5,0), mar = c(1,1,1,1), oma=c(3,5,3,2), xpd=TRUE)
+
+
+layout(matrix(c(2,3,4,5,1,26,
+                10,11,12,13,6,26,	
+                14,15,16,17,7,26,
+                18,19,20,21,8,26,
+                22,23,24,25,9,26), ncol=6, byrow=TRUE), 
+       heights=c(1,3,3,3,3), widths=c(2.5,2.5,2.5,2.5,1.5,2))
 #layout.show(n=25)
 
 #plot(0, bty="n")
@@ -22,7 +22,7 @@ plot.new() #1
 #2-5
 for (d in 1:length(delta)){
   plot.new()
-  text(0.5,0.5,labels=paste(delta[d]),font=2)
+  text(0.5,0.3,labels=paste(delta[d]),font=2)
 }
 
 #6-9
@@ -33,28 +33,31 @@ for (m in 1:length(mudif)){
 
 #10-25
 res <- vector(mode='list',length=1)
+n <- 1
 for (m in 1:length(mudif)){
   for (d in 1:length(delta)){
     res <- append(res,dePlot1(mudif[m], delta[d], xaxt="n"))
-    axis(1, labels=FALSE, tick=TRUE)
+    axis(1, labels=ifelse(n>12, yes=TRUE, no=FALSE), tick=TRUE)
+    n <- n+1
   }
 }
 
-title(ylab=expression(mu[1]-mu[2]), outer=TRUE, line=0, font.lab=2, cex.lab=1.5)
-title(xlab=expression(delta), outer=TRUE, line=-50, font.lab=2, cex.lab=1.5)
-title(xlab=expression(sigma), outer=TRUE, line=0, cex.lab=1.3)
[…]

### 2021-08-12 Figure4.R

@@ -1,26 +1,22 @@
-#source("./decomp2_plotting.R")
+source("./decomp2_plotting.R")
 
 delta <- c(1, 0.8, 0.5)
 sigma <- c(2, 4 ,5,6)
 
-pdf("fig4_samexaxis50.pdf")
+pdf("fig4_samexaxis50legend.pdf")
 
-par(mfrow=c(1,1), mar = c(1,2,1,1), oma=c(3,2,3,3))
+par(mgp=c(3,0.5,0), mar = c(1,1,1,1), oma=c(3,5,3,2), xpd=TRUE)
 
-layout(matrix(c(1,2,3,4,
-                5,9,10,11,	
-                6,12,13,14,
-                7,15,16,17,
-                8,18,19,20), ncol=4, byrow=TRUE), 
-       heights=c(1,3,3,3,3), widths=c(1,2.5,2.5,2.5))
+layout(matrix(c(2,3,4,1,21,
+                9,10,11,5,21,	
+                12,13,14,6,21,
+                15,16,17,7,21,
+                18,19,20,8,21), ncol=5, byrow=TRUE), 
+       heights=c(1,3,3,3,3), widths=c(2.5,2.5,2.5,1,2))
 
 
 plot.new() #1
-legend("center", 
-       legend=c(expression(r), expression(Delta["[E||C]"])
-                ,expression(r-Delta["[E||C]"])), 
-       col = c("orange","red","navy"),
-       lty = c(1,1,5), bty="n", cex=0.6)
+
 
 #2-4
 for (d in 1:length(delta)){
@@ -31,7 +27,7 @@ for (d in 1:length(delta)){
 #5-8
 for (s in 1:length(sigma)){
   plot.new()
-  text(0.8,0.5,labels=paste(sigma[s]), font=2)
+  text(0.3,0.5,labels=paste(sigma[s]), font=2)
 }
 
 #9-20
@@ -40,16 +36,21 @@ m <- 1
 for (s in 1:length(sigma)){
   for (d in 1:length(delta)){
     res <- append(res,dePlot2(mudif_list[[m]],sigma[s], delta[d], xaxt="n"))
-    axis(1, labels=ifelse(m%%3==0, yes=TRUE, no=FALSE), tick=TRUE)
+    axis(1, labels=ifelse(m>9, yes=TRUE, no=FALSE), tick=TRUE)
     m <- m+1
   }
 }
 
[…]

### 2021-08-13 paper/Paper.Rnw
added to results and methods - explained trait space

--- 2021-08-06 paper/Paper.Rnw
@@
 from ATAs compared to other mechanisms.
+G3) To define the ATA rescue or ATA exlusion subset of trait space for which species will coexist or become competitively excluded in our model. 
+G4) To provide an empirical example to highlight the role that ATA plays in coexistence and its mechanistic decomposition.
 Our examination and separation of the role that ATA play in contributing to coexistence is a novel approach to the study of coexistence...
@@
 %models exhibiting solely temporal variation; spatial variation can be considered in future work.
-%G2) To explore, using a simple model, the relative magnitudes of the contributions to coexistence that come
+%G2) To explore, using a simple model, the relative magnitudes of the contributions to coexistence that come from ATAs compared to other mechanisms
+%G3) To define the ATA rescue or ATA exlusion subset of trait space for which species will coexist or become competitively excluded in our model. 
+%G4) To provide an empirical example to highlight the role that ATA plays in coexistence and its mechanistic decomposition.
 %from ATA compared to other mechanisms.
@@
 $\Delta_i^{(E\shp C)}$, $\Delta_i^{[E \nat C]}$ and $\Delta_i^{[EC]}$ [see (\ref{eq:full_decomp})] for the lottery model
+% and $r_1-q_{12}r_2$?
 against the independent parameters of the model, $\mu_1$, $\mu_2$, $\sigma$ and $\delta$; and then we compared their
@@
 the contributions of other mechanisms.
-%***DAN: If we select 4 values for mu1-mu2 and 4 for delta, then Fig. 3 is going to have 16 panels arranged in a 
-%4 by 4 grid, each of which will plot all the components against sigma. We'll need to use different colors and/or line
-%types for the components and have a legend. Sort it all out as one big multi-panel pdf fig, after running all the sims.
+
 The simulation methods of \cite{Ellner_2019} were used to compute the components of (\ref{eq:full_decomp}).
@@
 %and sup mat results), put it above.
-%4) Think about this: Theoretical fig exploring why ATA matters to r_1\1 for the lottery model and hopefully more gernerally - try to get
+%4) Think about this: Theoretical fig exploring why , ATA matters to r_1\1 for the lottery model and hopefully more gernerally - try to get
 %at the intuition here. See if you can edge to more generality here, or not?
+
+To further explore the model and acheive G3 from the introduction i.e., defining the ATA rescue or ATA exlusion subset of trait space for which species will coexist or become competitively excluded in our model, we plotted the invasion growth rate, $r_{i/i}-q_{12}r_{i/j}$ and the invasion growth rate [IGR] without ATA contributribution  $r_{i/i}-q_{12}r_{i/j} - \Delta_i^{[EC]}$ against the difference between species, $\mu_1-\mu_2$.
+The same mathematical simplifications described above still hold, such that $\mu1$ and $\mu2$ are the only ways that the species differ and computations for coexistence in the lottery model only depend on the difference, $\mu_1-\mu_2$.
+We henceforward call this difference the trait difference between species.
+Letting the trait $\mu$ vary within a one dimensional space, this space is then a trait space.
+Letting there me two points $\mu_1$ and $\mu_2$ within that trait space, then the difference, $\mu_1-\mu_2$, is a subset of that trait space.
+We then define $\psi_{ATA}$ to be the subset of trait space that allows a postive IGR when ATA is included and $\psi_{~ATA}$ to be the subset of trait space that allows a postive IGR when ATA is not included.
+We then use symmetric difference $\psi_{ATA} \Delta \psi_{~ATA}$ to make the following conslusions.
+If the length/range?
+of $\psi_{ATA}$ > the length/range?
+of $\psi_{~ATA}$ then $\psi_{ATA} \Delta \psi_{~ATA}$ consitutes ATA rescue and the degree of species trait difference that allows ATA rescue to act, hence, the ATA rescue trait space.
+If the length of $\psi_{ATA}$ < the length/range?
+of $\psi_{~ATA}$ then $\psi_{ATA} \Delta \psi_{~ATA}$ consitutes ATA exclusion and the degree of species trait difference that allows ATA exclusion to act, hence, the ATA exclusion trait space.
+We highlight and color code these areas in our plots.
+We made the plots for values of $\delta=0.5, 0.8, 1.0$ and $\sigma=2.0,4.0,5.0, 6.0$.
+The spaces in between the lines where they intersect the x-axis at y=0, indicate either an ATA rescue trait space or an ATA exclusion trait space.
+If the line plotting IGR is to the left of the line plotting IGR without ATA, then this space is an ATA rescue trait space.
+If the line plotting IGR is to the right of the line plotting IGR without ATA, then this space is an ATA exclusion trait space.
 
@@
 \section{Results}\label{section:results}
-%outline of main resutls
-Using our simple two-species lottery model, ATA made a difference for coexistence.
-Depending on death rate, variance, and difference between species, ATA can either impede or facilitate coexistence.
-In some cases, the contribution of ATA is the mechanism solely responsible for a positive (or negative) IGR, and the IGR without ATA is negative (or positive).
-In almost all cases we tested, ATA contributed negatively to coexistence expected for when delta = 1, annual species.
-Of the annual species cases, only when the species were the same or extremely close to being the same, did ATA have a rescue effect on storage effects and coexistence.
-For other annual species cases, as the species became moree different, IGR was more negative and the contribution of ATA and storage effect in general became more negligible.
+%from methods : Relative importance of contributions was judged from the plots to assess whether and to what extent ATA effects on coexistence were meaningfully large for the lottery model, compared to the contributions of other mechanisms.
+To explore the relative magnitudes of the contributions to coexistence that come from ATA versus other mechanisms in our simple two-species lottery model (G2), we plotted the decomposing mechanisms against increasing $\sigma$ for various values of $\mu_1 - \mu_2$ and $\delta$ (Figure 3).
+When species are the same, $\mu_1 - \mu_2=0$, only storage effect mechanisms contribute to IGR and other mechanisms (in black) stay at zero.
+When $\delta=1$, other mechanisms stay relatively more stable with $\Delta_i^0$ set at a negative value and other non-storage effect values being at zero or extremely close to zero.
+Otherwise, all mechanisms tend to increase in magnitude as $\sigma$ increases.
+The contribution of ATA captured in $\Delta_i^{[E \nat C]}$ (red) is not small.
+In almost all cases, aside from when $\delta=1$, ATA contributes negatively to coexistence while symmetric associations, $\Delta_i^{[EC]}$ (blue) contribute postively, and the former decreases with $\sigma$ while the latter increases with $\sigma$.
+We see the opposite when $\delta=1$, where the contribution of ATA are positve and the contribution of symmetric association are negative and the trends are reversed.
 
-We found the range of species difference, where ATA either allowed the rescue of IGR or ATA imperiled IGR. %are we calling this 'differential coexistence'?
+While the contribution of ATA is definitively not absent, in most cases it does not make a difference for coexistence or comepetitive exclusion.
+In most panels of Figure 3, the magnitude of negative contribution from ATA does not effect the sign of the IGR.
+The absense of ATA contribution will increase IGR in those cases, but the outcome of coexistence would likely not be different.
+%maybe move or expand more on this point in discussion
+%cite sources that say the sign of IGR matter more than the magnitude 
+However, there are cases where ATA makes a difference for coexistence or competitive exclusion where the maginitude of $\Delta_i^{[E \nat C]}$ is large and other mechanisms do not counteract resulting in a sign change for IGR.
+We call these cases ATA rescue or ATA exclusion, the exploration of which is at the core of G3.
 
-In general, the direction of the ATA did not show any differences (other than small differences in very specific cases) for coexistence
+To acheive G3, we assessed the effect of ATA by comparing and plotting (Figure 4) the IGR computed for G2 and the IGR if there were no ATA (IGR - $\Delta_i^{[E \nat C]}$).
+To find areas of ATA exclusion or ATA rescue we looked at where these line cross y=0 and highlighted and color coded ATA exclusion trait space and ATA rescue trait space (see Methods).
+In these spaces, contribution of ATA is the mechanism soley responsible for a positive or negative IGR.
+Occurence of ATA rescue is slim and only occurs when $\delta=1$ and species traits are extremely similar or identical.
+Other $\delta$ values are more indicative of ATA exclusion events.
+When $\delta<1$ ATA exlusion space increase in range?
+and magnitude as $\sigma$ increases.
+There is some indication that ATA exlcusion space also increases in magnitude as $\delta$ decreases, however a changes in range have a more complex trend.
+Range increases across decreasing $\delta$ for all $\sigma$ values except $\sigma=6$.
+Here, the range is limited by an increasing $\Delta_i^{[E \nat C]}$ as trait space increases and the two IGR's are close to converging.
 

### 2021-08-13 Figure3.R
change r to IGR in legend

@@ -53,7 +53,7 @@ legend("topright",
        legend=c(expression(Delta[0]), expression(Delta[E]),
                 expression(Delta[C]), expression(Delta[("E#C")]),
                 expression(Delta["[EC]"]), expression(Delta["[E||C]"])
-                ,expression(r)), 
+                ,expression(IGR)), 
        col = c("black","black","black","black","blue","red","orange"),
        lty = c(1,2,4,3,1,1,1), bty="n", cex=1.5, inset=c(0,0))
 

### 2021-08-16 paper/Paper.Rnw
figures and ATA exlcusion/rescue definition and trait space stuff - preparing to send to Dan

@@
 
-To further explore the model and acheive G3 from the introduction i.e., defining the ATA rescue or ATA exlusion subset of trait space for which species will coexist or become competitively excluded in our model, we plotted the invasion growth rate, $r_{i/i}-q_{12}r_{i/j}$ and the invasion growth rate [IGR] without ATA contributribution  $r_{i/i}-q_{12}r_{i/j} - \Delta_i^{[EC]}$ against the difference between species, $\mu_1-\mu_2$.
-The same mathematical simplifications described above still hold, such that $\mu1$ and $\mu2$ are the only ways that the species differ and computations for coexistence in the lottery model only depend on the difference, $\mu_1-\mu_2$.
-We henceforward call this difference the trait difference between species.
+To further explore the model and acheive G3 from the introduction i.e., identifying the ATA rescue or ATA exlusion subset of species trait space for which species will coexist or become competitively excluded in our model, we plotted the invasion growth rate [IGR], $r_{i/i}-q_{12}r_{i/j}$ and the IGR without ATA contributribution  $r_{i/i}-q_{12}r_{i/j} - \Delta_i^{[EC]}$ against the difference between species, $\mu_1-\mu_2$.
+The same mathematical simplifications described above hold, such that $\mu1$ and $\mu2$ are the only ways that the species differ and computations for coexistence in the lottery model only depend on the difference, $\mu_1-\mu_2$.
+We henceforward call this difference the trait difference between species or species trait difference or trait space subset range  or species trait range for short.
 Letting the trait $\mu$ vary within a one dimensional space, this space is then a trait space.
-Letting there me two points $\mu_1$ and $\mu_2$ within that trait space, then the difference, $\mu_1-\mu_2$, is a subset of that trait space.
+Letting there be two points $\mu_1$ and $\mu_2$ within that trait space, then the difference, $\mu_1-\mu_2$, is a range of a subset of that trait space.
 We then define $\psi_{ATA}$ to be the subset of trait space that allows a postive IGR when ATA is included and $\psi_{~ATA}$ to be the subset of trait space that allows a postive IGR when ATA is not included.
 We then use symmetric difference $\psi_{ATA} \Delta \psi_{~ATA}$ to make the following conslusions.
-If the length/range?
-of $\psi_{ATA}$ > the length/range?
+If the length/range?/rank?
+of $\psi_{ATA}$ > the length/range?/rank?
 of $\psi_{~ATA}$ then $\psi_{ATA} \Delta \psi_{~ATA}$ consitutes ATA rescue and the degree of species trait difference that allows ATA rescue to act, hence, the ATA rescue trait space.
-If the length of $\psi_{ATA}$ < the length/range?
+If the length/range/rank of $\psi_{ATA}$ < the length/range?/rank?
 of $\psi_{~ATA}$ then $\psi_{ATA} \Delta \psi_{~ATA}$ consitutes ATA exclusion and the degree of species trait difference that allows ATA exclusion to act, hence, the ATA exclusion trait space.
+In other words, ATA rescue is a subset of all $\mu_1-\mu_2$ that results in coexistence with ATA but failure to coexist without ATA and ATA exclusion is a subset of all $\mu_1-\mu_2$ that results in coexistence without ATA, but failure to coexist with ATA.
 We highlight and color code these areas in our plots.
@@
 \begin{figure}
+\includegraphics[width=\textwidth]{"../results_figs/fig1.png"}
 \caption{
@@
 \begin{figure}
+\includegraphics[width=\textwidth]{"../results_figs/figure2.PNG"}
 \caption{%Pedagogical figure somehow displaying the different outcomes of competition in ATA and symmetric cases for an example.
@@
 
+\begin{figure}
+\includegraphics[width=\textwidth]{"../results_figs/fig3_legend.pdf"}
+\caption{The contribution of ATA is non-negligible.
+All panels show the breakdown of contributing mechanisms of coexistence in accordance to \cite{Ellner_2019} new decomposition method, against increasing $\sigma$.
+Each panel is defined by a unique set of parameters $\mu_1-\mu_2$ and $\delta$.
+\label{fig:fig3}}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=\textwidth]{"../results_figs/fig4_samexaxis50legend.pdf"}
+\caption{ATA rescue and ATA exclusion can be defined within subsets species trait difference spaces.
+IGR (ATA included; yellow) and IGR without ATA (violet) are plotted against species trait difference $\mu_1-\mu_2$ in order to demonstrait ATA rescue and ATA exclusion.
+The red line, $\Delta_i^{[E \nat C]}$, is the difference between the two IGR's.
+Each panel is defined by a unique set of parameters $\sigma$ and $\delta$.
+ATA rescue is marked by regions in the panels where IGR wihtout ATA contribution is less than IGR at zero, denoted by regions of dotted yellow.
+ATA exclusion is marked by regions where IGR without ATA contribution is greater than IGR at zero, denoted by regions of dotted pink.
+ATA rescue is only seen in cases where $\delta=1$ (left column).
+All other cases shown here have subsets of species trait difference space where ATA exclusion occurs.
+Despite panels in the top row not containing the color coded region, these cases still contain ATA exlusion spaces; however, the space is too small to appropriately denote while still maintaining consistent x-axes across all panels.
+For $\delta<1$, ATA exclusion space increases down rows as $\sigma$ increases.
+ATA exclusion space also seems to increase across columns as $\delta$ decreases; however this trend is broken in the last row, indicating a more compex relationship between ATA exclusion space, $\delta$, and $\sigma$.
+\label{fig:fig4}}
+\end{figure}
+
 \end{document}

### 2021-08-19 paper/Paper.Rnw
draft sent to Dan

@@
 (e.g., various common forms of regression), but correlations are extremely common, and other common
-methods also provide only a small portion of the avaiable information \citep{Nelsen2006}.
+methods also provide only a small portion of the available information \citep{Nelsen2006}.
 The variables in figure \ref{fig:noise}A,C 
@@
 The reliance of storage effect's magnitude on the degree of relationship between environment and competition indicate that storage effects involve environment-coexistence associations, and hence there is potential for ATAs to influence storage effects.
-In addition to EC covariance, functioning storage effects also require that heterospecific populations respond differently to environmental fluctuations, and different life stages must also vary in responses \citep{}.
+In addition to EC covariance, functioning temporal storage effects also require that heterospecific populations respond differently to environmental fluctuations, and different life stages must also vary in responses \citep{}.
 In a simple two species system, differences in environmental response allow for one species to grow to high densities in its favorable environemnt while the other is very low. 
@@
 Thus, averaged over multiple generations, higher fluctuation in growth rate can lead to a positive low density growth rate and coexistence is possible \citep{Chesson_2000}.
-Given that storage effect depends on the relationship between E and C, we question the consequences of EC relationships having ATAs, the presence of which may be increasing as climate change brings about more extreme climatic events.
+Given that storage effect depends on the relationship between E and C, we question the consequences of EC relationships characterized by ATA, the presence of which may be increasing as climate change brings about more extreme climatic events.
 
@@
 
-To contextualize idea of how ATAs in environmental response might manifest and modfiy the outcome of coexistence, we introduce an example of a two species grass-forb plant community (inspired by \cite{Chesson_2012}).
+To contextualize idea of how ATA in environmental response might manifest and modfiy the outcome of coexistence, we introduce an example of a two species grass-forb plant community (inspired by \cite{Chesson_2012}).
 The community follows the criteria required for temporal storage effects to operate.
@@
 Competition is influenced by resource comsumption of growing juvenile plants.
-Dormant seeds are unaffected by competiton and unfavorable environment and can persist in the community.
-When rain fall events temporally fluctuate from year to year, relative abundance and dominance of the grass-forb community also fluctuate year to year.
+Dormant seeds are unaffected by competiton and can persist in the community in an unfavorable environment.
+When rain fall dynamics temporally fluctuate from year to year, relative abundance and dominance of the grass-forb community also fluctuate year to year.
 The result is a relationship between the grass's and forb's environmental responses (fecundity) that is symmetrically associated.
@@
 We see the possible consequences of ATA in Figure 2.
-In simulating their populations,  ATAs can cause species to have longer periods of extreme dominance and lower growth rates overall.
+In simulating their populations,  ATA can cause species to have longer periods of extreme dominance and lower growth rates overall.
 
@@
 from ATAs compared to other mechanisms.
-G3) To define the ATA rescue or ATA exlusion subset of trait space for which species will coexist or become competitively excluded in our model. 
-G4) To provide an empirical example to highlight the role that ATA plays in coexistence and its mechanistic decomposition.
+G3) To identify and define the ATA rescue or ATA exlusion set of trait space range for which species will coexist or become competitively excluded in our model. 
+G4) To provide an empirical example and highlight the role that ATA plays in coexistence and its mechanistic decomposition in a real system.
 Our examination and separation of the role that ATA play in contributing to coexistence is a novel approach to the study of coexistence...
@@
 \ref{SI-SIsect:other_mechanisms}, but we here provide a conceptual understanding. \cite{Ellner_2016} and
-\cite{Ellner_2019} provide a decomposition of the coexistence metric $\olr_{i \bs i} -q_{ij}\olr_{i \bs i}$ which is similar to and motivated by the 
+\cite{Ellner_2019} provide a decomposition of the coexistence metric $\olr_{i \bs i} -q_{ij}\olr_{j \bs i}$ which is similar to and motivated by the 
 original decomposition of \cite{Chesson_1994}, and which we use; the decomposition is how the mechanisms of 
@@
 \begin{equation}
-\olr_{i \bs i} -q_{ij}\olr_{i \bs i}=\Delta_i^0+\Delta_i^E+\Delta_i^C+\Delta_i^{(E\shp C)}+\Delta_i^{(EC)},\label{eq:previous_decomp}
+\olr_{i \bs i} -q_{ij}\olr_{j \bs i}=\Delta_i^0+\Delta_i^E+\Delta_i^C+\Delta_i^{(E\shp C)}+\Delta_i^{(EC)},\label{eq:previous_decomp}
 \end{equation}
@@
 \begin{equation}
-\olr_{i \bs i} -q_{ij}\olr_{i \bs i}=\Delta_i^0+\Delta_i^E+\Delta_i^C+\Delta_i^{(E\shp C)}+\Delta_i^{[E \nat C]} + \Delta_i^{[EC]},\label{eq:full_decomp}
+\olr_{i \bs i} -q_{ij}\olr_{j \bs i}=\Delta_i^0+\Delta_i^E+\Delta_i^C+\Delta_i^{(E\shp C)}+\Delta_i^{[E \nat C]} + \Delta_i^{[EC]},\label{eq:full_decomp}
 \end{equation}
@@
 
-To further explore the model and acheive G3 from the introduction i.e., identifying the ATA rescue or ATA exlusion subset of species trait space for which species will coexist or become competitively excluded in our model, we plotted the invasion growth rate [IGR], $r_{i/i}-q_{12}r_{i/j}$ and the IGR without ATA contributribution  $r_{i/i}-q_{12}r_{i/j} - \Delta_i^{[EC]}$ against the difference between species, $\mu_1-\mu_2$.
-The same mathematical simplifications described above hold, such that $\mu1$ and $\mu2$ are the only ways that the species differ and computations for coexistence in the lottery model only depend on the difference, $\mu_1-\mu_2$.
-We henceforward call this difference the trait difference between species or species trait difference or trait space subset range  or species trait range for short.
+To further explore the model and acheive G3 from the introduction i.e., to identify and define the ATA rescue or ATA exlusion set of trait space range for which species will coexist or become competitively excluded in our model, we plotted the invasion growth rate [IGR], $\olr_{i \bs i} -q_{ij}\olr_{j \bs i}$, and the IGR without ATA contributribution, $\olr_{i \bs i} -q_{ij}\olr_{j \bs i}- \Delta_i^{[E \nat C]}$, against the difference between species, $\mu_1-\mu_2$.
+The same mathematical simplifications described above hold, such that $\mu_1$ and $\mu_2$ are the only ways that the species differ and computations for coexistence in the lottery model only depend on the difference, $\mu_1-\mu_2$.
+We henceforward call this difference the species trait difference and sets of multiple $\mu_1-\mu_2$ are sets of trait space range.
 Letting the trait $\mu$ vary within a one dimensional space, this space is then a trait space.
-Letting there be two points $\mu_1$ and $\mu_2$ within that trait space, then the difference, $\mu_1-\mu_2$, is a range of a subset of that trait space.
-We then define $\psi_{ATA}$ to be the subset of trait space that allows a postive IGR when ATA is included and $\psi_{~ATA}$ to be the subset of trait space that allows a postive IGR when ATA is not included.
-We then use symmetric difference $\psi_{ATA} \Delta \psi_{~ATA}$ to make the following conslusions.
-If the length/range?/rank?
-of $\psi_{ATA}$ > the length/range?/rank?
-of $\psi_{~ATA}$ then $\psi_{ATA} \Delta \psi_{~ATA}$ consitutes ATA rescue and the degree of species trait difference that allows ATA rescue to act, hence, the ATA rescue trait space.
-If the length/range/rank of $\psi_{ATA}$ < the length/range?/rank?
-of $\psi_{~ATA}$ then $\psi_{ATA} \Delta \psi_{~ATA}$ consitutes ATA exclusion and the degree of species trait difference that allows ATA exclusion to act, hence, the ATA exclusion trait space.
-In other words, ATA rescue is a subset of all $\mu_1-\mu_2$ that results in coexistence with ATA but failure to coexist without ATA and ATA exclusion is a subset of all $\mu_1-\mu_2$ that results in coexistence without ATA, but failure to coexist with ATA.
-We highlight and color code these areas in our plots.
+Letting there be two points $\mu_1$ and $\mu_2$ within that trait space, then the set of points between $\mu_1$ and $\mu_2$ is a subset of trait space, and the difference, $\mu_1-\mu_2$, is the range of the subset of trait space.
+We then define $\psi_{ATA}$ to be the subset of trait space bound by a $\mu_1$ and $\mu_2$ whose range defines the maximal species trait difference that allows a postive IGR when ATA is included (coexistence with ATA) and $\psi_{\sim ATA}$ to be the subset of trait space bound by a $\mu_1$ and $\mu_2$ whose range defines the maximal species trait difference that allows a postive IGR when ATA is not included (coexistence without ATA).
+We then consider the difference between the range of $\psi_{ATA}$ and the range of $\psi_{\sim ATA}$ to make the following conclusions.
+If the range of $\psi_{ATA}$ > range of $\psi_{\sim ATA}$ then the continuous set of numbers in between the range of $\psi_{ATA}$ and the range of $\psi_{\sim ATA}$ defines the bounds of extensions to $\psi_{\sim ATA}$ that result coexistence with ATA but failure to coexist without ATA; hence ATA rescue.
+If the range of $\psi_{ATA}$ < the range of $\psi_{\sim ATA}$ then the continuous set of numbers in between the range of $\psi_{ATA}$ and the range of $\psi_{\sim ATA}$ defines the bounds of extensions to $\psi_{ATA}$ that result coexistence without ATA but failure to coexist with ATA; hence ATA exclusion.
+In other words, ATA rescue is a subset of all $\mu_1-\mu_2$ that results in coexistence with ATA but failure to coexist without ATA, and ATA exclusion is a subset of all $\mu_1-\mu_2$ that results in coexistence without ATA, but failure to coexist with ATA.
+We highlighted and color coded these areas in our plots.
 We made the plots for values of $\delta=0.5, 0.8, 1.0$ and $\sigma=2.0,4.0,5.0, 6.0$.
@@
 Otherwise, all mechanisms tend to increase in magnitude as $\sigma$ increases.
-The contribution of ATA captured in $\Delta_i^{[E \nat C]}$ (red) is not small.
+The contribution of ATA captured by $\Delta_i^{[E \nat C]}$ (red) is not small.
 In almost all cases, aside from when $\delta=1$, ATA contributes negatively to coexistence while symmetric associations, $\Delta_i^{[EC]}$ (blue) contribute postively, and the former decreases with $\sigma$ while the latter increases with $\sigma$.
@@
 Other $\delta$ values are more indicative of ATA exclusion events.
-When $\delta<1$ ATA exlusion space increase in range?
-and magnitude as $\sigma$ increases.
-There is some indication that ATA exlcusion space also increases in magnitude as $\delta$ decreases, however a changes in range have a more complex trend.
+When $\delta<1$ ATA exlusion space increase in range and magnitude as $\sigma$ increases.
+There is indication that ATA exlcusion space also increases in magnitude as $\delta$ decreases, however changes in range have a more complex trend.
 Range increases across decreasing $\delta$ for all $\sigma$ values except $\sigma=6$.
@@
 A) is left-tail associated, B) is symmetric, and C) is right-tail associated.
-All three cases have the same Pearson correlation of $\rho=$, up to sampling.., yet show different symmetry of tail associaton.
+All three cases have the same Pearson correlation up to sampling variation, yet show different symmetry of tail associaton.
 Asymmetrically tail associated cases are perfectly correlated in the lower (A)  or upper (C) halfs of the distribution.
@@
 We should discuss and possibly take several iterations to get this figure right since it's a pedagogically important figure - if the reader is not convinced in the Intro that we are studying something that makes a difference, they are not going to slog through the Theory section, which is inevitably going to be tough going for many readers.
-Populations inlfuenced by asymmetric tail association...
-recover from low densities less often/ spend more time on average under one-species dominance.
-A) is a simulation over time of a population within a two species lottery model whose environmental response and competition is influenced by symmetric bivariate noise.
-B) is like A), but the population is influenced by bivariate noise with left-tailed asymmetric association.
-Both simulations were ran for total time of 100,000.
+Populations inlfuenced by asymmetric tail association recover from low densities less often/ spend more time on average under one-species dominance.
+The top panel, A, is a simulation over time of a population within a two species lottery model whose environmental response and competition is influenced by symmetric bivariate noise.
+The bottom panel, B, is like A, but the population is influenced by bivariate noise with left-tailed asymmetric association.
+Both simulations were ran for total time of 100000.
 Noticeable population density was defined at a threshold of 0.95 of total community space.
 Periods of noticable coexistence are marked in grey.
-While the population in B) had longer mean periods of coexistence, it had much longer mean periods of one-species dominace.
-The population in B) also recovered from low densities less often and thus has a lower ability to recover when rare and thus that community has lower coexistence.
+While the population in B had longer mean periods of coexistence, it had much longer mean periods of one-species dominace.
+The population in B also recovered from low densities less often and thus has a lower ability to recover when rare and thus that community has lower coexistence.
 \label{fig:pedagog}}
@@
 Each panel is defined by a unique set of parameters $\mu_1-\mu_2$ and $\delta$.
+In most cases, the contribution of ATA (red) is negatively; however when $\delta=1$, ATA contribution is positive.
+The contribution of ATA is non-negligible and in special cases result in a sign change of IGR (orange).
 \label{fig:fig3}}

### 2021-08-20 Figure3.R
adapting code and making plots with q12 expression

@@ -60,4 +60,58 @@ legend("topright",
 dev.off()
 #res <- res[-1]
 
+pdf("fig3_qij_legend.pdf")
 
+par(mgp=c(3,0.5,0), mar = c(1,1,1,1), oma=c(3,5,3,2), xpd=TRUE)
+
+
+layout(matrix(c(2,3,4,5,1,26,
+                10,11,12,13,6,26,	
+                14,15,16,17,7,26,
+                18,19,20,21,8,26,
+                22,23,24,25,9,26), ncol=6, byrow=TRUE), 
+       heights=c(1,3,3,3,3), widths=c(2.5,2.5,2.5,2.5,1.5,2))
+#layout.show(n=25)
+
+#plot(0, bty="n")
+plot.new() #1
+
+#2-5
+for (d in 1:length(delta)){
+  plot.new()
+  text(0.5,0.3,labels=paste(delta[d]),font=2)
+}
+
+#6-9
+for (m in 1:length(mudif)){
+  plot.new()
+  text(0.3,0.5,labels=paste(mudif[m]), font=2)
+}
+
+#10-25
+res <- vector(mode='list',length=1)
+n <- 1
+for (m in 1:length(mudif)){
+  for (d in 1:length(delta)){
+    res <- append(res,dePlot1(mudif[m], delta[d], qij=TRUE, xaxt="n"))
+    axis(1, labels=ifelse(n>12, yes=TRUE, no=FALSE), tick=TRUE)
+    n <- n+1
+  }
+}
+
+mtext("contribution to coexistence", side=2, outer=TRUE, line=1.5, font=2, cex=1, at=0.46)
+mtext(expression(mu[1]-mu[2]), side=4, outer=TRUE, line=-8, font=2, cex=1.5, at=0.46)
+mtext(expression(delta), side=3, outer=TRUE, line=-0.5, font=2, cex=1.5, at=0.37)
+mtext(expression(sigma), outer=TRUE, side=1, line=0.5, cex.lab=1.3, at=0.37)
+
+#layout(matrix(c(1)))
+plot.new() #1
+legend("topright", 
+       legend=c(expression(Delta[0]), expression(Delta[E]),
+                expression(Delta[C]), expression(Delta[("E#C")]),
+                expression(Delta["[EC]"]), expression(Delta["[E||C]"])
+                ,expression(IGR)), 
+       col = c("black","black","black","black","blue","red","orange"),
[…]

### 2021-08-20 decomp1_plotting.R

@@ -2,7 +2,7 @@
 
 source("./decomposition_fxn.R")
 
-dePlot1 <- function(mudif, delta, legend=FALSE,...){
+dePlot1 <- function(mudif, delta, qij=FALSE, legend=FALSE,...){
   load("./noise_etc.RData")
   store <- vector(mode='list', length=7)
   sigma <- seq(0,7,1)
@@ -17,6 +17,9 @@ dePlot1 <- function(mudif, delta, legend=FALSE,...){
   for (i in 1:length(sigma)){
     if(i>=2){
       dat <- cbind(store[[i-1]]$D, store[[i]]$D)
+      if(qij==TRUE){
+        dat <- cbind(store[[i-1]]$Dq, store[[i]]$Dq)
+      }
       lines(c(sigma[i-1], sigma[i]), dat[1,1:2])#null
       lines(c(sigma[i-1], sigma[i]), dat[2,1:2], lty=2)#E
       lines(c(sigma[i-1], sigma[i]), dat[3,1:2], lty=4)#C
@@ -44,8 +47,8 @@ dePlot1 <- function(mudif, delta, legend=FALSE,...){
 }
 
 
-#dePlot(-0.3,0.8)
-
+dePlot1(-0.3,0.8)
+dePlot1(-0.3,1.0)
 
 
 

### 2021-08-20 Figure1.R
make pdfs

@@ -12,7 +12,7 @@ col2rgb("darkblue")
 blue <- rgb(0,0,0.545,0.3)
 
 pdf("fig1.pdf")
-png("fig1.png", height=560, width=1600,pointsize=22)
+#png("fig1.png", height=560, width=1600,pointsize=22)
 par(mar=c(0.5,0.5,0,0), oma=c(5,5,2,2))
 layout(matrix(c(4,10,10,6,11,11,8,12,
                 1,5 ,13,2,7 ,14,3,9 ),nrow=2,byrow=T),

### 2021-08-20 paper/Paper.Rnw
small
partial edits, got through Methods, next is Results

@@
 \begin{document}
-\SweaveOpts{concordance=TRUE}
+%\SweaveOpts{concordance=TRUE}
 %\SweaveOpts{concordance=TRUE}
@@
 G2) To explore, using a simple model, the relative magnitudes of the contributions to coexistence that come
-from ATAs compared to other mechanisms.
-G3) To identify and define the ATA rescue or ATA exlusion set of trait space range for which species will coexist or become competitively excluded in our model. 
-G4) To provide an empirical example and highlight the role that ATA plays in coexistence and its mechanistic decomposition in a real system.
-Our examination and separation of the role that ATA play in contributing to coexistence is a novel approach to the study of coexistence...
+from ATAs compared to other mechanisms. This is to help illuminate whether ATA contributions to coexistence 
+are likely to be negligible or meaningfully large. 
+%G3) To identify and define the ATA rescue or ATA exlusion set of trait space range for which species will coexist or become competitively excluded in our model. 
+%***DAN: This is a good additional goal. We will either have to have already defined the "ATA rescue and ATA exlusion sets of trait space"
+%or else readers won't understand this as it was written, above (now commented). These are complex concepts, so it may be better to 
+%save formal definitions for later and reword this. Here's a try at rewording: 
+G3) To explore, using our simple model, the likelihood that species can coexist with
+ATAs but not without, or vice versa. This is to begin to illuminate whether ATA contributions biodiversity maintenance 
+may be negligible or meaninfully large.
+G4) To provide an empirical example of species coexistence, and to highlight the role ATAs plays in coexistence for this example.
+Overall, our study presents a new mechanism of coexistence and biodiversity maintenance which may be influenced by climatic changes, 
+and a means for beginning to understand its importance for real systems.
 
@@
 for the lottery model, using an approach and notation
-similar to those of \cite{Ellner_2016} and \cite{Ellner_2019}; see those papers for additional details.
+similar to those of \cite{Ellner_2016;Ellner_2019}; see those papers for additional details.
 
@@
 \end{equation}
-where: $\olr_{i \bs i}$ refers to the mean ``invasion growth rate'' of species $i$, that is, the mean growth rate of 
+where: $\olr_{i \bs i}$ refers to the mean ``invasion growth rate'' (IGR) of species $i$, that is, the mean growth rate of 
 species $i$ when it is at negligible abundance and species $j$ (the other species) is at stochastic steady state;
@@
 We now specify how to decompose storage effects into
-contributions due to asymmetric tail associations and due
+contributions due to ATAs and due
 to $EC$ correlation \emph{per se}. 
@@
 The term $\Delta_i^{[E \nat C]}$ is the component of storage effects due to 
-asymmetries of tail association between $E_i$ and $C_{i \bs i}$, and between $E_j$ and $C_{j \bs i}$.
+ATAs between $E_i$ and $C_{i \bs i}$, and between $E_j$ and $C_{j \bs i}$.
 The notation $\nat$ was chosen because this new symbol is part of the symbol $\shp$ (i.e., two of its four lines), 
-corresponding to the removal of part of the relationship between $E$ and $C$ components (namely, ATA).
+corresponding to the removal of part of the relationship between $E$ and $C$ components (namely, ATAs).
 The new symbol should correspondingly be pronounced ``partial sharp''.
 The term $\Delta_i^{[EC]}$ is the component of storage effects due to $EC$ correlation \emph{per se}, remaining after 
-having removed the effects of asymmetric tail associations.
-%The brackets in the superscripts of $\Delta_i^{[E \nat C]}$ and $\Delta_i^{[EC]}$ are intended to indicate that 
-%this is a decompostion of storage effects, which were denoted by \cite{Ellner_2019} using a superscript $(EC)$,
-%i.e., $\Delta_i^{(EC)}$, to indicate that storage effects are, in turn, part of a decomposition of a difference between invader and
-%resident growth rates (see below).  
+having removed the effects of ATAs.
 The brackets in the superscripts of $\Delta_i^{[E \nat C]}$ and $\Delta_i^{[EC]}$ are intended to indicate that 
@@
 Only storage effects involve $EC$ covariation,
-so the only mechanism of coexistence influenced by ATA is storage effects. Combining our decomposition 
+so the only mechanism of coexistence influenced by ATA between $E$ and $C$ is storage effects. Combining our decomposition 
 (\ref{eq:storage_effects_further}) with the decomposition (\ref{eq:previous_decomp}) of \cite{Ellner_2019} gives 
@@
 \end{equation}
-which is the extension of the MTC which was goal G1 of the Introduction.
+which is the extension of the MTC which was goal G1 of the Introduction. To pursue goal G2, below, will quantify
+the magnitude of the new ATA term in this expression, $\Delta_i^{[E \nat C]}$, relative to the magnitudes of the other terms.
 
@@
 
-%Goals, for reference:
-%G1) To augment the modern theory of coexistence (MTC) to show how asymmetric tail association (ATA)
-%plays a role. The MTC decomposes a measure of coexistence into components due to each of several mechanisms. 
-%We address how ATA contributes to these components. We here examine mechanisms of coexistence that apply in 
-%models exhibiting solely temporal variation; spatial variation can be considered in future work.
-%G2) To explore, using a simple model, the relative magnitudes of the contributions to coexistence that come from ATAs compared to other mechanisms
-%G3) To define the ATA rescue or ATA exlusion subset of trait space for which species will coexist or become competitively excluded in our model. 
-%G4) To provide an empirical example to highlight the role that ATA plays in coexistence and its mechanistic decomposition.
-%from ATA compared to other mechanisms.
-
-%Figs, for reference:
-%1) pedagog fig showing what tail assoc is, also later used for introing the noise we use, in methods
-%2) pedgog fig showing a simple set of sims, somehow, the only difference being ATA, showing good things 
-%happening in one panel and bad in another - cited in Intro, only serves to prove to people that ATA matters for 
-%coexistence, so they keep reading
-%3) Fig with a grid of panels, each panel plots against sigma, and the grid controls a few values of 
-%delta and a few of mu1-mu1. Plotted is all the components of coexistence, for comparison.
-%4) Theoretical fig exploring why ATA matters to r_1\1 for the lottery model and hopefully more gernerally - try to get
-%at the intuition here. See if you can edge to more generality here, or not?
+%G2) To explore, using a simple model, the relative magnitudes of the contributions to coexistence that come
+%from ATAs compared to other mechanisms. This is to help illuminate whether ATA contributions to coexistence 
+%are likely to be negligible or meaningfully large.
 
@@
 the contributions of other mechanisms.
-
 The simulation methods of \cite{Ellner_2019} were used to compute the components of (\ref{eq:full_decomp}).
@@
 SI section \ref{SI-sect:efficient_computation}.
-%What's next
-%1) Write code implementing the above and making a figure - that'll be Fig. 3
-%2) Get Jasmin to work on Fig. 1 and Fig. 2 (Fig. 2 is largely done). Fig 1 should be facilitated by the work to make Fig. 3 in the previous item.
-%3) Decide what other results we will put in the main text and write a paragraph on methods for the other results (for both the main text
-%and sup mat results), put it above.
-%4) Think about this: Theoretical fig exploring why , ATA matters to r_1\1 for the lottery model and hopefully more gernerally - try to get
-%at the intuition here. See if you can edge to more generality here, or not?
 
-To further explore the model and acheive G3 from the introduction i.e., to identify and define the ATA rescue or ATA exlusion set of trait space range for which species will coexist or become competitively excluded in our model, we plotted the invasion growth rate [IGR], $\olr_{i \bs i} -q_{ij}\olr_{j \bs i}$, and the IGR without ATA contributribution, $\olr_{i \bs i} -q_{ij}\olr_{j \bs i}- \Delta_i^{[E \nat C]}$, against the difference between species, $\mu_1-\mu_2$.
+%G3) To explore, using our simple model, the likelihood that species can coexist with
+%ATAs but not without, or vice versa. This is to begin to illuminate whether ATA contributions biodiversity maintenance 
+%may be negligible or meaninfully large.
+
+To acheive goal G3 from the Introduction i.e., to explore, using the lottery model, the likelihood that species can coexist with
+ATAs but not without, or vice versa, we plotted the IGR $\olr_{i \bs i}=\olr_{i \bs i} -q_{ij}\olr_{j \bs i}$, 
+as well as the
+IGR minus its ATA contribution, $\olr_{i \bs i} -\Delta_i^{[E \nat C]}$. 
+For fixed values of $\sigma$ and $\delta$, we were interested in the values of $\mu_1-\mu_2$ for which one of these
+quantities is positive and the other is negative; this set of value characterizes the species which can coexist
+with ATAs, but not without, or vice versa. The one dimensional space of values $\mu$ is a species trait space,
+and $\mu_1$ and $\mu_2$ can be regarded as two points in it. But we instead define the ``trait-difference space'' of values $\mu_1-\mu_2$. 
+Points of the trait-difference space characterize possible combinations of traits of the two competing species in our model. For fixed values of 
+$\delta$ and $\sigma$, we defined the subset, $\psi_{\text{ATA}}(\delta,\sigma)$, of trait-difference space to be the set of 
+values $\mu_1 - \mu_2$ for which the the IGR, $\olr_{i \bs i}$, is positive. This 
+is the subspace for which the two species in the lottery model coexist in the presence of ATAs.
+We also defined the subset, $\psi_{\lnot\text{ATA}}(\delta,\sigma)$, of trait-difference space to be the set of values
+$\mu_1 - \mu_2$ for which the metric $\olr_{i \bs i} -\Delta_i^{[E \nat C]}$ is positive.
+This is the subspace for which the two species in the lottery model coexist in the absence of 
+ATAs. The symbol $\lnot$ denotes logical negation, so $\lnot \text{ATA}$ is intended to indicate the
+absence of ATAs. We defined the ``ATA rescue space" associated with given values of $\delta$ and $\sigma$ to be the 
+set difference $\psi_{\text{ATA}}(\delta,\sigma) \setminus \psi_{\lnot\text{ATA}}(\delta,\sigma)$, i.e., the
+values $\mu_1 - \mu_2$ for which model species coexist in the presence of ATAs, but not in the absence of ATAs. 
+Similarly, we define ``ATA exclusion space'' to be the set difference 
+$\psi_{\lnot\text{ATA}}(\delta,\sigma) \setminus \psi_{\text{ATA}}(\delta,\sigma)$,
+i.e., $\mu_1 - \mu_2$ for which coexistence occurs in the absence, but not the presence, of ATAs.
+On plots of $\olr_{i \bs i}$ and $\olr_{i \bs i} -\Delta_i^{[E \nat C]}$
+against $\mu_1 - \mu_2$, we highlighted the ATA rescue and exclusion spaces. We plotted against
+$\mu_1 - \mu_2$ for all combinations of $\delta=0.5, 0.8, 1.0$ and $\sigma=2.0,4.0,5.0, 6.0$.
+
+%***DAN: Jasmin, I edited the below to get the above. Kept the below so I could add one comment to it, and in case it is needed again in
+%the near future
+%to identify and define the ATA rescue or ATA exlusion set of trait space range for which species will coexist or become competitively excluded in our model, we plotted the invasion growth rate [IGR],
+%***DAN: Note that this is not the IGR. The IGR is just r_{1\1}. So I abandoned this terminology and used "coexistence metric" instead
+%$\olr_{i \bs i} -q_{ij}\olr_{j \bs i}$, and the IGR without ATA contributribution, $\olr_{i \bs i} -q_{ij}\olr_{j \bs i}- \Delta_i^{[E \nat C]}$, against the difference between species, $\mu_1-\mu_2$.
 The same mathematical simplifications described above hold, such that $\mu_1$ and $\mu_2$ are the only ways that the species differ and computations for coexistence in the lottery model only depend on the difference, $\mu_1-\mu_2$.
@@
 We then consider the difference between the range of $\psi_{ATA}$ and the range of $\psi_{\sim ATA}$ to make the following conclusions.
-If the range of $\psi_{ATA}$ > range of $\psi_{\sim ATA}$ then the continuous set of numbers in between the range of $\psi_{ATA}$ and the range of $\psi_{\sim ATA}$ defines the bounds of extensions to $\psi_{\sim ATA}$ that result coexistence with ATA but failure to coexist without ATA; hence ATA rescue.
-If the range of $\psi_{ATA}$ < the range of $\psi_{\sim ATA}$ then the continuous set of numbers in between the range of $\psi_{ATA}$ and the range of $\psi_{\sim ATA}$ defines the bounds of extensions to $\psi_{ATA}$ that result coexistence without ATA but failure to coexist with ATA; hence ATA exclusion.
-In other words, ATA rescue is a subset of all $\mu_1-\mu_2$ that results in coexistence with ATA but failure to coexist without ATA, and ATA exclusion is a subset of all $\mu_1-\mu_2$ that results in coexistence without ATA, but failure to coexist with ATA.
+%If the range of $\psi_{ATA}$ > range of $\psi_{\sim ATA}$ then the continuous set of numbers in between the range of $\psi_{ATA}$ and the range of $\psi_{\sim ATA}$ defines the bounds of extensions to $\psi_{\sim ATA}$ that result coexistence with ATA but failure to coexist without ATA; hence ATA rescue.
+%If the range of $\psi_{ATA}$ < the range of $\psi_{\sim ATA}$ then the continuous set of numbers in between the range of $\psi_{ATA}$ and the range of $\psi_{\sim ATA}$ defines the bounds of extensions to $\psi_{ATA}$ that result coexistence without ATA but failure to coexist with ATA; hence ATA exclusion.
+%In other words, ATA rescue is a subset of all $\mu_1-\mu_2$ that results in coexistence with ATA but failure to coexist without ATA, and ATA exclusion is a subset of all $\mu_1-\mu_2$ that results in coexistence without ATA, but failure to coexist with ATA.
 We highlighted and color coded these areas in our plots.
 We made the plots for values of $\delta=0.5, 0.8, 1.0$ and $\sigma=2.0,4.0,5.0, 6.0$.
-The spaces in between the lines where they intersect the x-axis at y=0, indicate either an ATA rescue trait space or an ATA exclusion trait space.
+
+%***DAN: Some version of the below text ought to go in the figure caption, do that later, Dan
+%The spaces in between the lines where they intersect the x-axis at y=0, indicate either an ATA rescue trait space or an ATA exclusion trait space.
 If the line plotting IGR is to the left of the line plotting IGR without ATA, then this space is an ATA rescue trait space.
@@
 %Ideas while driving to pick up Max:
-%1) I think in every case for the lottery model, making tail associations asymmetric reduced the potential for coexistence - I think that's what Jasmin told me, but if we present it we should make sure I'm not remembering wrong, of course.
-Is that generally the case, for other models?
-What are the features of a model, or of the growth rate of the invader, that make it so that more asymmetric tail associations means reduced coexistence?
-Does this have implications for climate change?
-%2) In this paper with Jasmin you need to cite the two papers you just got from the professor at University of Massachusetts, Lowell.
+%1) In this paper with Jasmin you need to cite the two papers you just got from the professor at University of Massachusetts, Lowell.
 I mean the papers about changes in the special extent of heat waves: https://agupubs.onlinelibrary.wiley.com/doi/full/10.1029/2020GL087097 and https://iopscience.iop.org/article/10.1088/1748-9326/ab4b41
-%3) Discussion content: We should (in future) study tail associations between environmental variables, and how these might be changing.
+%2) Discussion content: We should (in future) study tail associations between environmental variables, and how these might be changing.
 However, more than that, the variables E in the lottery model are not actual environmental variables but rather are the effects of environmental variables on the species of the model.
 Probably real environmental variables are actually related to vital rates in a non-linear and probably often threshold-like way.
-Threshold relationships can produce tail association, as we have showed elsewhere.
+Threshold relationships can produce tail association, as we showed elsewhere.
 Changes in the means of true environmental variables may change where those distributions of environmental variables are compared to thresholds, which may change the nature of tail associations between the effects of those environmental variables.
@@
 \begin{figure}
-\includegraphics[width=\textwidth]{"../results_figs/fig1.png"}
+%\includegraphics[width=\textwidth]{"../results_figs/fig1.png"}
 \caption{
@@
 \begin{figure}
-\includegraphics[width=\textwidth]{"../results_figs/figure2.PNG"}
+%\includegraphics[width=\textwidth]{"../results_figs/figure2.PNG"}
 \caption{%Pedagogical figure somehow displaying the different outcomes of competition in ATA and symmetric cases for an example.
@@
 \begin{figure}
-\includegraphics[width=\textwidth]{"../results_figs/fig3_legend.pdf"}
+%\includegraphics[width=\textwidth]{"../results_figs/fig3_legend.pdf"}
 \caption{The contribution of ATA is non-negligible.
@@
 \begin{figure}
-\includegraphics[width=\textwidth]{"../results_figs/fig4_samexaxis50legend.pdf"}
+%\includegraphics[width=\textwidth]{"../results_figs/fig4_samexaxis50legend.pdf"}
 \caption{ATA rescue and ATA exclusion can be defined within subsets species trait difference spaces. 
-IGR (ATA included; yellow) and IGR without ATA (violet) are plotted against species trait difference $\mu_1-\mu_2$ in order to demonstrait ATA rescue and ATA exclusion.
-The red line, $\Delta_i^{[E \nat C]}$, is the difference between the two IGR's.
-Each panel is defined by a unique set of parameters $\sigma$ and $\delta$.
-ATA rescue is marked by regions in the panels where IGR wihtout ATA contribution is less than IGR at zero, denoted by regions of dotted yellow.
-ATA exclusion is marked by regions where IGR without ATA contribution is greater than IGR at zero, denoted by regions of dotted pink.
-ATA rescue is only seen in cases where $\delta=1$ (left column).
-All other cases shown here have subsets of species trait difference space where ATA exclusion occurs.
-Despite panels in the top row not containing the color coded region, these cases still contain ATA exlusion spaces; however, the space is too small to appropriately denote while still maintaining consistent x-axes across all panels.
-For $\delta<1$, ATA exclusion space increases down rows as $\sigma$ increases.
-ATA exclusion space also seems to increase across columns as $\delta$ decreases; however this trend is broken in the last row, indicating a more compex relationship between ATA exclusion space, $\delta$, and $\sigma$.
-\label{fig:fig4}}
+IGR (ATA included; yellow) and IGR without ATA (violet) are plotted against species trait difference 
+$\mu_1-\mu_2$ in order to demonstrate ATA rescue and ATA exclusion. The red line, $\Delta_i^{[E || C]}$, 
+%***DAN: I would normally have used \nat instead of ||, above, but \nat seems to not work in figure captions for some reason,
+%it crashes the latex on my machine at least. If we can fix this, we should, but it's not the end of the world if we can't.
+%Jasmin, could you compile on your machine when || above was \nat?
+is the difference between the two IGR's. Each panel is defined by a unique set of parameters $\sigma$ 
+and $\delta$. ATA rescue is marked by regions in the panels where IGR wihtout ATA contribution is less 
+than IGR at zero, denoted by regions of dotted yellow. ATA exclusion is marked by regions where IGR 
+without ATA contribution is greater than IGR at zero, denoted by regions of dotted pink. ATA rescue 
+is only seen in cases where $\delta=1$ (left column). All other cases shown here have subsets of 
+species trait difference space where ATA exclusion occurs. Despite panels in the top row not 
+containing the color coded region, these cases still contain ATA exlusion spaces; however, the space 
+is too small to appropriately denote while still maintaining consistent x-axes across all panels. 
+For $\delta<1$, ATA exclusion space increases down rows as $\sigma$ increases. ATA exclusion space also 
+seems to increase across columns as $\delta$ decreases; however this trend is broken in the last row, 
+indicating a more compex relationship between ATA exclusion space, $\delta$, and $\sigma$.   
+}\label{fig:fig4}
 \end{figure}

### 2021-08-23 paper/Paper.Rnw
starting editing results, but just coments so far until we resolve an issue with fig 3 and standard errors

@@
 To acheive goal G3 from the Introduction i.e., to explore, using the lottery model, the likelihood that species can coexist with
-ATAs but not without, or vice versa, we plotted the IGR $\olr_{i \bs i}=\olr_{i \bs i} -q_{ij}\olr_{j \bs i}$, 
+ATAs but not without, or vice versa, we plotted the IGR, $\olr_{i \bs i}=\olr_{i \bs i} -q_{ij}\olr_{j \bs i}$, 
 as well as the
 IGR minus its ATA contribution, $\olr_{i \bs i} -\Delta_i^{[E \nat C]}$. 
-For fixed values of $\sigma$ and $\delta$, we were interested in the values of $\mu_1-\mu_2$ for which one of these
+For fixed values of $\sigma$ and $\delta$ and for left-tail associated noise, 
+we were interested in the values of $\mu_1-\mu_2$ for which one of these
 quantities is positive and the other is negative; this set of value characterizes the species which can coexist
@@
 $\psi_{\lnot\text{ATA}}(\delta,\sigma) \setminus \psi_{\text{ATA}}(\delta,\sigma)$,
-i.e., $\mu_1 - \mu_2$ for which coexistence occurs in the absence, but not the presence, of ATAs.
+i.e., $\mu_1 - \mu_2$ values for which coexistence occurs in the absence, but not the presence, of ATAs.
 On plots of $\olr_{i \bs i}$ and $\olr_{i \bs i} -\Delta_i^{[E \nat C]}$
@@
 \section{Results}\label{section:results}
+
+%G2
+
+%Para on the main goal, whether the ATA contribution is meaningfully large
+%ATA effects tend to oppose "the rest" of storage effects
+
+%Para on how it's usually negative, whereas \Delta_[EC] is positive, so except when \delta=1, ATA effects tend to weaken storage effects
+%Note that when mu1-mu2=2, we are in the case where only storage effects allow for coexistence
+
+%Para on when delta=1 (annual species) - in this case, coexistence is only possible when mu1-mu2 is close to 0, and in that case the only
+%thing allowing for coexistence at all is storage effects. Interetingly (and this is the topic sentence), "the rest" of storage effects are
+%negative, and it's the ATA component of storage effects which are allowing the species to coexist. Suggests the possibility that for 
+%annual species whcih are close competitors, ATAs might be especially important? Jasmin ought to check on this to make sure it's accurate.
+%What I'm suspicious about is that for annual species which are close competitors (say, mu1-mu2=0), and in the symmetric case, storage
+%effects in the lotery model actually don't opperate, and there is no coexistence. That sounds weird to me. Can we get some independent
+%verification of that? Why would annual species be different in this respect? There must be some delta between 0.8 and 1 where \Delta^[EC]
+%switches from positive to negative.
+
+%***DAN: Jasmin, you computed standard errors of everything, but they are not currently reported. The best way to handle this would
+%be if we could say in captions in appropriate places "standard errors were always less than xyz", where xyz is some extremely 
+%small number. Standard errors are smalled for larger
+
 %from methods : Relative importance of contributions was judged from the plots to assess whether and to what extent ATA effects on coexistence were meaningfully large for the lottery model, compared to the contributions of other mechanisms.
@@
 We see the opposite when $\delta=1$, where the contribution of ATA are positve and the contribution of symmetric association are negative and the trends are reversed.
+
+%G3
 

### 2021-08-25 parameters.R
delete delta=1

@@ -15,7 +15,7 @@ for (i in 1:length(mu1)){
 mu <- mu[-1,]
 #print(mu) #45x2
 
-delta <- c(0.2, 0.5, 0.8, 1.0)
+delta <- c(0.2, 0.5, 0.8)
 sigma <- c(0.8, 1.6, 3.2, 6.4)
 mudif <- c(0.0, -0.3, -0.6, -0.8)
 

### 2021-08-25 parameters.R
changed delta values

@@ -15,7 +15,7 @@ for (i in 1:length(mu1)){
 mu <- mu[-1,]
 #print(mu) #45x2
 
-delta <- c(0.2, 0.5, 0.8)
+delta <- c(0.2, 0.4, 0.6, 0.8)
 sigma <- c(0.8, 1.6, 3.2, 6.4)
 mudif <- c(0.0, -0.3, -0.6, -0.8)
 

### 2021-08-25 Figure3.R
change subscript to superscript of legend

@@ -107,9 +107,9 @@ mtext(expression(sigma), outer=TRUE, side=1, line=0.5, cex.lab=1.3, at=0.37)
 #layout(matrix(c(1)))
 plot.new() #1
 legend("topright", 
-       legend=c(expression(Delta[0]), expression(Delta[E]),
-                expression(Delta[C]), expression(Delta[("E#C")]),
-                expression(Delta["[EC]"]), expression(Delta["[E||C]"])
+       legend=c(expression(Delta^0), expression(Delta^E),
+                expression(Delta^C), expression(Delta^"(E#C)"),
+                expression(Delta^"[EC]"), expression(Delta^"[E||C]")
                 ,expression(IGR)), 
        col = c("black","black","black","black","blue","red","orange"),
        lty = c(1,2,4,3,1,1,1), bty="n", cex=1.5, inset=c(0,0))

### 2021-08-25 Figure4.R

@@ -99,7 +99,7 @@ mtext(expression(mu[1]-mu[2]), outer=TRUE, side=1, line=0.5, cex.lab=1.3, at=0.4
 
 plot.new() #21
 legend("topright", 
-       legend=c(expression(IGR),expression(IGR-Delta["[E||C]"]), expression(Delta["[E||C]"])),
+       legend=c(expression(IGR),expression(IGR-Delta^"[E||C]"), expression(Delta^"[E||C]")),
        col = c("orange","navy", "red"),
        lty = c(1,1,5), bty="n", cex=1.2, inset=c(0,0))
 

### 2021-09-01 paper/Paper.Rnw
minor changes to comments

@@
 
+%***DAN: drop the delta =1 case from here
 %Para on when delta=1 (annual species) - in this case, coexistence is only possible when mu1-mu2 is close to 0, and in that case the only
@@
 %switches from positive to negative.
+
+%***DAN: consider possile para on mu1-mu2=0 case
+%***DAN: For discussion: is it true that ATA always weakens storage effects?
+Same as saying the ATA terms is always negative?
+The actual question is, for what models are ATA effects negative, versus positive?
 
@@
 \section{Discussion}\label{section:discussion}
+
+I think ata specialists are a thing - if their competitor is ubiquitous, they can only live in places on the landscape 
+where there are atas. Could have a discussion para on it, if it does not make the results. Or both. BTW, ATA exclusion has a 
+similar thing going on - a species excluded by ATAs (and competition) can only live on the portions of the landscape where
+there are no ATAs. Could also describe the temporal way of it - the ATA specialist can only survive in a era/epoch for which there
+are ATAs in its local, otherwise it'll have to go dormant or move elsewhere.
 
